### 2015-01-13 paper.tex
Added trec notebook paper + working sigir abstract

% This is "sig-alternate.tex" V2.0 May 2012 This file should be compiled with
% V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls' V2.5
% LaTeX2e document class file. It is for those submitting articles to ACM
% Conference Proceedings WHO DO NOT WISH TO STRICTLY ADHERE TO THE SIGS
% (PUBS-BOARD-ENDORSED) STYLE.  The 'sig-alternate.cls' file will produce a
% similar-looking, albeit, 'tighter' paper resulting in, invariably, fewer
% pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces: 1) The Permission
% Statement 2) The Conference (location) Info information 3) The Copyright
% Line with ACM data 4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which DOES NOT produce 1) thru'
% 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within the source
% .tex file, over both the CopyrightYear (defaulted to 200X) and the ACM
% Copyright Data (defaulted to X-XXXXX-XX-X/XX/XX).  e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the
% copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use the .bib file (from which
% the .bbl file % is produced).  REMEMBER HOWEVER: After having produced the
% .bbl file, and prior to final submission, you *NEED* to 'insert' your .bbl
% file into your source .tex file so as to provide ONE 'self-contained' source
% file.
%
% ================= IF YOU HAVE QUESTIONS ======================= Questions
% regarding the SIGS styles, SIGS policies and procedures, Conferences etc.
% should be sent to Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

%\documentclass{sig-alternate} \usepackage{url} \usepackage{color}
\documentclass[10pt]{article} \usepackage{url} \usepackage{color}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[]{algorithm2e}
\DeclareMathOperator{\corpus}{\mathcal{C}}
\DeclareMathOperator{\doc}{\mathnormal{d}}
\DeclareMathOperator{\sent}{\mathnormal{s}}
\DeclareMathOperator{\order}{\pi}
\DeclareMathOperator{\dtime}{\mathnormal{t}}
\DeclareMathOperator{\hour}{\mathnormal{h}}
\DeclareMathOperator{\hours}{\mathcal{H}}
\DeclareMathOperator{\Sim}{\mathbf{K}}
\DeclareMathOperator{\SMat}{\mathbf{X}}
\DeclareMathOperator{\Pref}{\boldsymbol{\pi}}
\DeclareMathOperator{\Exemp}{\mathnormal{Exemplars}}
\DeclareMathOperator{\Updates}{\mathnormal{Updates}}
\usepackage{cleveref}
\crefname{section}{§}{§§}
\Crefname{section}{§}{§§}

\usepackage{authblk}
\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}

\begin{document}

%
% --- Author Metadata here ---
%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
%- IF NEED BE.  \crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.  --- End of Author
%Metadata ---

\title{Columbia University at TREC 2014: Temporal Summarization}

\author[1]{Chris Kedzie}
\author[1]{Kathleen McKeown}
\author[2]{Fernando Diaz}
\affil[1]{Columbia University, Department of Computer Science}
\affil[2]{Microsoft Research}
%    Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
%Format\titlenote{(Produces the permission block, and copyright information).
%For use with SIG-ALTERNATE.CLS. Supported by ACM.}} \subtitle{[Extended
%Abstract] \titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement and
% alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time' i.e. three
% 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of "name/affiliations" may
% appear. We just ask that you restrict the number of 'columns' to three.
%
% Because of the available 'opening page real-estate' we ask you to refrain
% from putting more than six authors (two rows with three columns) beneath the
% article title.  More than six makes the first-page appear very cluttered
% indeed.
%
% Use the \alignauthor commands to handle the names and affiliations for an
% 'aesthetic maximum' of six authors.  Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the \additionalauthors
% command.  These 'additional authors' will be output/set for you without
% further effort on your part as the last section in the body of your article
% BEFORE References or any Appendices.

%\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting reasons)
% and the remaining two appear in the \additionalauthors section.
%
%\author{
% You can go ahead and credit any number of authors here, e.g. one 'row of
% three' or two rows (consisting of one row of three and a second row of one,
% two or three).
%
% The command \alignauthor (no curly braces needed) should precede each author
% name, affiliation/snail-mail address and e-mail address. Additionally, tag
% each line of affiliation/address with \affaddr, and tag the e-mail address
% with \email.
%
% 1st. author

%\alignauthor Chris Kedzie\\ \affaddr{Columbia University}\\
%\affaddr{Department of Computer Science} \email{kedzie@cs.columbia.edu}

% 2nd. author
%\alignauthor Kathleen McKeown\\ \affaddr{Columbia University}\\
%\affaddr{Department of Computer Science}\\ \email{kathy@cs.columbia.edu}

% 3rd. author
%\alignauthor Fernando Diaz\\ \affaddr{Microsoft Research}\\
%\email{fdiaz@microsoft.com}

% There's nothing stopping you putting the seventh, eighth, etc.  author on
% the opening page (as the 'third row') but we ask, for aesthetic reasons that
% you place these 'additional authors' in the \additional authors block, viz.
%} 
\date{29 October 2014}
% Just remember to make sure that the TOTAL number of authors is the number
% that will appear on the first page PLUS the number that will appear in the
% \additionalauthors section.

\maketitle \begin{abstract} 
Information need is highest during times of crisis; both officials and 
civilians need effective access to make informed decisions about their 
safety and the safety of others. However, these situations pose many 
interesting challenges for the IR systems that support these decisions.
Specifically, we address event monitoring and updating, a.k.a. the
temporal summarization task (TS). Given an event query, we monitor streams
of online news and select important text to present to the user. Updates
are selected by predicting the salience of the text w.r.t. to the event.
Our model utilizes a variety of features at the query, document, and corpus
level. Additionally, these predictions are used to directly bias a clustering
algorithm for sentence selection, 
increasing the novelty of the updates. We evaluate our system
on a varied set of retrospective events using a combination of human 
judgements
and automatic evaluation measures, as compared to a baseline system.
We demonstrate the effect of the various feature groups, and the importance
of salience prediction w.r.t. update precision when compared to a clustering
only baseline.

\end{abstract}

% A category with the (minimum) three required fields
%\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
%performance measures]

%\terms{Summarization}

%\keywords{Extractive Summarization, Affinity Propagation}

\section{Introduction}

During crises, information is critical for first responders and those caught
in the event.  When the event is significant, as in the case of Hurricane
Sandy, the amount of information produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. Methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively  (e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
these experiments were not performed in the large and heterogeneous
environment of the modern web; there is still a need for robust and scalable 
methods for automatic summarization.

%Previous work on generating event descriptions and/or multi-document
%summarization has relied on clustering algorithms to find representative
%sentences appropriate for an event summary.  These methods impose a metric
%space on the text data that can make it difficult to incorporate external
%sources of information elegantly -- in this paper we argue that centroid
%sentences are not a priori the best candidates for inclusion in an event
%summary.

In this paper, we detail our participation in the second year of the TREC
Temporal Summarization (TS) track. Broadly, TS systems are required to monitor
a time ordered corpus for information about a specific event and ``emit'' 
updates about that event. Under the current TREC formulation of this task,
successful TS systems must synthesize techniques from information retrieval,
 extractive multi-document summarization, and update summarization.
Specifically, we must select updates to emit based on three primary criterion:
relevance, salience, and redundancy.

As this was our first time participating in TS track, we focused mostly on 
second and third criterion, relying on the TREC supplied resources for 
document retrieval and filtering heuristics for relevance.
Our system predicts sentence salience in the context of a
large-scale event and integrates these predictions into
a clustering based multi-document summarization system. We fit a Gaussian 
Process (GP) regression model to predict sentence salience and use 
these predictions to bias the
formation of sentence clusters around more salient regions in the input space
using affinity propagation (AP) clustering.  AP uses the salience predictions
as well as pairwise similarities among input sentences to identify
\emph{exemplar} sentences, which we use as our summary output.  Our approach
differs from other methods of summarization that compute salience by pairwise
comparisons alone, ignoring features of importance that are intrinsic to the
sentences themselves.

The remainder of the paper is organized as follows. In section
~\ref{sec:background} we give an overview of the semantic similarity, 
Gaussian process, and affinity propagation algorithms that make up the bulk of
our TS system. 
Next we describe the data with which we build our various models.
Then in section~\ref{sec:approach} we give a high level sketch
of our system, and then explain each component in detail. 
We finish with the presentation of our results and related works.

\section{Background}\label{sec:background}

\subsection{Affinity Propagation}

Affinity propagation (AP) is a message passing algorithm that identifies both
exemplar data points and assignments of each point to an exemplar.  This is
done iteratively by passing \emph{responsibility} and \emph{availability}
messages between data points that quantify the fitness of one data point to
represent another, and the fitness of a data point to be represented based on
the choices of other data points respectively \cite{dueck2007non}.

AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
our preference that the $i$-th data point can serve as an exemplar a priori of
other data points.

AP has several useful properties that comport well with the TS track 
requirements. First, the number of clusters $k$ is not a hyper-parameter
of the model. Since we will be running this algorithm many times in per run,
the usual methods of hyperparameter search become infeasible 
(?). The number of clusters falls out of the algorithm
organically-- lower overall preference values will result in fewer clusters.

Secondly, the arbitrary nature of the preferences
allow us to incorporate a variety of signals for
identifying the best exemplars, i.e. salience and redundancy signals. 
The preferences can be thought of as self-similarities (similarities that
pairwise-comparison based algorithms would ignore) that we can exploit to 
incorporate our prior beliefs about a data point.

Finally, cluster exemplars are guaranteed to be actual data points. Many 
clustering algorithms group data around mathematical objects (e.g., the
mean) that are not necessarily observed in the data. The extractive nature of
this TS task requires that we emit actual data points (i.e. sentences).
We are able to sidestep 
the additional requirement of selecting a most
representative cluster member, as this is computed explicitly in the AP 
algorithm.

%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
%the exemplar of the $i^{th}$ point relative to other potential exemplars.
%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
%serve as an exemplar of the $i^{th}$ point, taking into account other points'
%preference for the $k^{th}$ point as an exemplar.

%While, AP does not set the number of exemplars before-hand, a lower overall
%preference values will result in a smaller number of exemplars.

%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
%real-valued preference $\pi_i$ quantifying our belief a priori of the
%$i^{th}$ element's ability to serve as an exemplar.

\subsection{Semantic Similarity}\label{subsec:semsim}

Whenever we make a pairwise comparison between sentences, we use the weighted
textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
model can be thought of as a variant of latent semantic analysis (?), 
where words that are not present in a sentence are explicitly modeled.
More formally, we have a term-sentence matrix 
$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
$j$ contains word $i$. In the WTMF regime, we want to find an approximation
of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
space. These matrices are found by minimizing the objective function

$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
\mathbf{Q}_{\cdot,j} 
- \mathbf{X}_{i,j})^2 
 + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$

where $\mathbf{W}_{i,j} = 
\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
\end{cases}$
and $\lambda$ is a hyperparameter controlling the regularization terms.

The $w_m$ term is another model hyperparameter that is set to a small constant
($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
reconstruction error of missing terms (words that did not occur a sentence).

Given an unseen sentence $\hat{i}$ we can project its term vector into the
latent sentence vector space with

$$
\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
$$

where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.

The WTMF model is used extensively throughout our TS system. When making any 
pairwise comparison between sentence $i$ and $j$, we first construct
their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
$\displaystyle \operatorname{cos-sim}
(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
||\mathbf{Q}_{\cdot,j}||_2   }$.

Because the events for the TS task come from different domains, we construct
domain specific latent word vector spaces for each domain using in-domain 
Wikipedia pages (see~\cref{sec:data} for more details).

\subsection{Gaussian Processes}

A Gaussian process (GP) is a distribution over functions and is a 
generalization of the multi-variate Gaussian to the infinite dimensional
setting. That is, we use the observed data to define a distribution over 
possible functions that generated this data, without having to explicitly 
parameterize the function---in this sense GPs are considered 
a non-parametric model.

Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
of an input space $\mathcal{X}$ to the reals,

$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
 an $n$-dimensional vector whose elements are evaluations of the function $f$
at points $x_i \in \mathcal{X}$.
We say $p(f)$ is a Gaussian process if for any finite subset 
$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
covariance function $K(x,x^\prime)$. Generally, the mean function is simply
set to 0, leaving the distribution to be completely characterized by the
kernel function on the data.

In the regression setting, we typically have a response variable $y$ that
is the sum of our model prediction  and 
some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
two distributions
of principal interest are the marginal likelihood
$p(\mathbf{y}|\mathbf{X}) = 
\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
distribution,

$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$

where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
response, and
\begin{align*}
\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
\boldsymbol{\sigma}^2_* & 
= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
\mathbf{K}_*^T + \sigma^2\\
\end{align*}.

Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.

GP's are incredibly general, and are state of the art for many regression 
tasks (?). The reliance on the covariance matrix 
$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
for regression, and many varieties of similarity functions can be used.
In our experiments we used a radial basis function kernel 
$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
The $\ell_i$ are feature dependent scaling parameters; once learned, they not
only improve the accuracy of the model, but give us some introspection 
into which features are more important.

%the formation of sentence clusters around more salient sentences. We
%introduce the affinity propagation algorithm as as an elegant way to
%incorporate our salience predictions into a

%What follows is a description of our ongoing event summarization efforts.  We
%briefly situate our approach to summarization within the broader field of
%multi-document summarization, and then introduce the affinity propagation
%algorithm which we use for clustering. This algorithm allows us to elegantly
%address the salient sentence selection problem by incorporating our prior
%beliefs about sentence quality. Next, we describe out method for modeling
%summary sentence quality, and the features used in this model.  Finally, we
%address future features and system improvements that we are incorporating
%into our summarizer.

\section{Data}\label{sec:data}

Our primary dataset is the TREC 2014 Stream Corpus, a ?tb corpus of hourly 
web crawls from October 2011
through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
As per the TS specifications, all summary sentences come from processing this
corpus in a time aligned manner. Additionally, we use the query terms 
provided by the track organizers for document retrieval.

For our semantic similarity component (\cref{subsec:semsim}),
domain specific term-sentence matrices were built using subsections of
Wikipedia. For each event type we identified one or more related Wikipedia
categories and collected all pages that were listed under that category.
For all pages collected, we used the latest possible revision date before the
start of the Stream Corpus.

We use the same Wikipedia corpora to construct the lanuage models used
in our salience prediction model (\cref{subsubsec:lm}).

\begin{figure}
\begin{tabular}{| p{5cm} c c |}
\hline
Event & WP Categories & No. Docs/Sents/Words\\
\hline \hline
Boston\_Marathon\_bombings \newline
Christopher\_Dorner\_shootings\_and\_manhunt \newline 
 In\_Amenas\_hostage\_crisis
& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
\hline
 Cyclone\_Oswald \newline
Early\_2012\_European\_cold\_wave \newline
February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
\hline
Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
\hline
Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
\hline
Port\_Said\_Stadium\_riot \newline
2012\_Afghanistan\_Quran\_burning\_protests \newline
2011-13\_Russian\_protests \newline
2012\_Romanian\_protests \newline
2012-13\_Egyptian\_protests \newline
2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
\hline
\end{tabular}
\caption{Domain specific Wikipedia corpora for semantic similarity and 
language models. }
\end{figure}

We use the TREC 2013 Stream Corpus and TS 2013 event/gold nugget data to 
train our salience models. See section~\cref{subsec:Predict} 
for a detailed description of our sampling procedure to generate labeled
data.

%Our documents for summarization come from the online news portion of the TREC
%Stream Corpus, a 6.45tb corpus obtained by hourly web crawls from October 2011
%through mid February 2013
%
%Summary events come from the TREC Temporal Summarization track, and include
%natural disasters like Hurricane Sandy as well as man-made events like a 2012
%train accident in Buenos
%Aires.\footnote{\url{http://trec.nist.gov/data/tempsumm2013.html}} The track
%organizers also provide a search query for each event \cite{aslam2013trec}.
%For each event, we collect the documents that contain all query words and
%stratify them by the hour they were collected. 
%
%For evaluation purposes, the track organizers also provided gold nugget
%information (i.e. important pieces of information, usually the length of a
%short clause or sentence). These gold nuggets come from the event's related
%Wikipedia article and also include the timestamp of when they were added to
%the page.
%
%%KM _ Ask Chris: is this true?
%To create hourly gold summaries to evaluate our system, we simply take the set
%of gold nugget information from the start of the event up to the current hour. 
%

\section{Approach}\label{sec:approach}

We first begin with some notation. For a given event, let $\corpus$ be the set
of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
$\corpus_{\hour_i}$ be the set of retrieved documents such that 
$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.

%Our TS system involves ? phases involving sentence level classification, 
%regression, and clustering. 
Figure ? outlines our general temporal summarization algorithm. The description
of our approach is as follows.
For each event, we iterate over the retrieved 
documents in hourly chunks, emitting 0 or more updates at each hour.
At each hour $\hour$, we process each document
$\doc \in \corpus_{\hour}$. First, we identify where the document content
actually is; Second, we predict the salience of all content sentences.
To account for redundancy, the predicted salience is penalized based on the 
distance of the sentence in question to the previous summary updates.

Finally, we cluster all content sentences 
for the current hour using the penalized salience predictions to bias the 
formation
of clusters around the most salient sentences. For each cluster center, or 
exemplar, that results, we check that the salience is above a threshold and 
that it does not belong to a singleton cluster; exemplars that satisfy these
conditions are emitted as an update.
Additionally, we maintain the complete set of updates in order to penalize
salience predictions in the subsequent time steps.

\begin{figure}
\centering
\begin{algorithm}[H]
 \KwData{$Query$ --- the set of event query words\\ 
        $SC$ --- the stream corpus\\

}
 ~\\
 Initialize empty list $\mathbf{U}$ of updates\\
 Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
 $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
 \For{$i \gets 1,\ldots,t $}{

  Initialize empty lists $\SMat, \Pref$ \;

  \For{$\doc \in \corpus_{\hour_i}$}{
   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
     $\SMat.\operatorname{append}(\sent)$\;
     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
     $\Pref.\operatorname{append}(\sigma)$\;

   }        
  }
  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
  %$\operatorname{}$
  ~\\ 
  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
  ~\\
  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
%gets \Updates_{cache} \cup \Updates_{\hour}$\;

 } 
 \caption{Temporal Summarization Algorithm}
\end{algorithm}
\end{figure}

\subsection{Document Retrieval}\label{subsec:Document Retrieval}

The focus of our system was on salience prediction and clustering stages, and
so we relied heavily on the pre-filtered corpus provided by the track 
organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
a subset of the full TREC 2014 StreamCorpus and is
intended to be a high recall retrieval of documents related to all 15 of 
the 2014 TS events.

Track participants were provided with a set of query words for each event.
For example, the event ``Costa Concordia disaster and recovery'' had query
words [``costa'', ``concordia''].
In order to construct an event specific corpus,
we retrieve all documents whose timestamps fall within the event start/stop 
times and whose raw html content contains at least one keyword from the 
event's query words. We further restricted our document set to only those 
articles from the news domain.

\subsection{Content Detection}\label{subsec:Content Detection}

In early versions of our summarization system, we found that structural html
artifacts and sentence tokenization errors
were negatively effecting the performance of later stages. Examples of the 
former include strings of link text like ``World Politics Sports ...'', while
examples of the latter included concatenations of various article headlines,
i.e. headlines pertaining to the event in question, as well as headlines from
other non-related events.
Both types of ``sentences'' were problematic for computing sentence similarity
as they were more likely to have higher average similarity to all input 
sentences. In turn they would be more likely to appear as cluster exemplars
in our clustering stage. From the clustering algorithm's point of view, this
is the correct decision to make---such multi-topic sentences are more general 
than a single topic, and better able to represent all aspects contained in a 
cluster. They make poor choices as updates, however, as they contain 
irrelevant information.

In order to filter out these problematic inputs, we trained a classifier to 
identify which sentences came from inside a document's main article and which
came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
inside or outside the document's main article. We then trained a logistic 
regression classifier using the following features :

\begin{itemize}
 \item the position of the sentence
 \item word counts
 \item the last token in the sentence
 \item the last two tokens in the sentence
 \item the last three tokens in the sentence.
\end{itemize}

These features were sufficient to capture 
the main difference between content and non-content sentences,
which 
was that content sentences generally ended with sentence final punctuation, 
i.e. periods or a closing quotation mark.

Within our larger summarization framework, we process a document at a time,
identifying the subset of its sentences that are content sentences.
We then check to make sure \emph{all} event
query terms can be found within the document's content sentences. If so,
we send the content sentences on to the next stage of our pipeline; otherwise
we ignore all sentences in this document.

\subsection{Predicting Sentence Salience}\label{subsec:Predict}

In order to use AP clustering for summarization, we need to assign a
preference value to each input sentence.  In our approach, we equate a
sentence's salience with its preference.  A good model of sentence salience
should predict higher values for sentences that are more likely to appear in a
human generated summary of the event.

We do not have such human judgments, but we do have last year's gold nugget
sentences.
The response variable we try to predict is a sentence's semantic similarity
(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
to a gold nugget when the gold nugget is not known.

To build training data for this regression task, 
we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
sentences using the document retrieval and content selection steps outlined
in the previous selection. We then sample with replacement 200 sentences
from this collection, and extract non-lexical features (described in more 
detail below) for each sentence
to construct our design matrix $\mathbf{X}$. To build our response variables
$\mathbf{y}$ we compute the maximum semantic similarity of each input
sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
an RBF kernel to this data, optimizing kernel parameters with the scaled 
conjugate gradient method. This sampling procedure is repeated 100 times
for each of the ? 2013 TS events, yielding ? total models. In our salience 
predictions for the 2014 events, we take the mean prediction of all models.

%we take a subset of sentences
%relevant to the TREC events (approximately 1000) and match them to the gold
%nugget sentence with highest similarity as determined by the sentence
%similarity system of \cite{guo2012simple}. 
%\cite{} have used this system previously to correlate sentences to meaningful
%units of information in human generated summaries. 
%We use the real-valued similarity scores as our salience scores for the
%training sentences.

We want our model to be predictive across different kinds of events so we
avoid lexical features.  Instead, we extract a variety of features including
language model scores, geographic relevance, and temporal relevance from each
sentence.  These features are used to fit a Gaussian process regression model
that can predict the similarity of a sentence to a gold summary
\cite{preotiuc2013temporal}.  
%We use the model predicted salience of each
%sentence as it's preference value in the AP clustering.

\subsubsection{Basic Features}

We employ several basic features that have been used previously in supervised
models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
These include sentence length, the number of capitalized words normalized by
sentence length, and the number of query words present in the sentence.  Query
words include the event's type (e.g., \emph{earthquake}) and are expanded with
the event type's WordNet \cite{miller1995wordnet} synset, hypernyms, and
hyponyms.  For \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
``seism,'' ``aftershock,'' etc.

\subsubsection{Language Model Features}\label{subsubsec:lm}

%Because the data in our experiments is scraped from the web, it is common to
%find sentences that contain both salient informantion and two kinds of noise:
%noisey fragments of web page structure (e.g. section titles, \emph{News},
%\emph{Sports}, etc.) and references to other news not relevant to the topic
%summary.
%
We use two trigram language models, trained using the SRILM toolkit
\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
the sentence's total log probability normalized by sentence length) from each
model.  This first model is trained on 4 years (2005-2009) of articles from
the Gigaword corpus.  Specifically, we use articles from the Associated Press
and the New York Times. This model is intended to assess the general writing
quality (grammaticality, word usage) of an input sentence and helps us to
filter out text snippets which are not sentences (e.g., web page titles).  The
second model is a domain specific language model. We build a corpus of
Wikipedia articles for each event type, consisting of documents from a related
Wikipedia category. E.g. for earthquakes, we collect pages under the category
\emph{Category:Earthquakes}. This model assigns higher probability to
sentences that are focused on the given domain.

%For both models, we Finally, we extract the percentage of capitalized words,
%and sentence length as features. These last two features also help to
%identify sentences that are less likely to contain relevant content-- overly
%long and heavily capitalized sentences in our corpus were likely to be long
%strings of web-page headlines, section headers, and other irrelevant page
%structure.

\subsubsection{Geographic Relevance Features}

Locations are identified using a named entity tagger. For each location in a
sentence, we obtain its latitude and longitude using the Google Maps API.  We
then compute its distance to that of the event location.  It is possible for a
sentence and an event to have multiple locations so we take as features the
minimum, maximum, and average distance of all sentence-event location pairs.
Distances are calculated using the Vincenty distance.

\subsubsection{Temporal Relevance Features}

Our data consists of hourly crawls of online content and so we exploit the
temporality of corpus by capturing the burstiness of a sentence, i.e.  the
change in word frequency from one hour to the next.``Bursty'' sentences often
indicate new and important data.

Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
\end{align*}
% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
%\end{align*}

We similarly find the sentence's 5-hour burstiness.  In addition to
burstiness, we also include the sentence's average tf-idf and hours since the
event in question started as features.

\subsection{Sentence Selection}\label{subsec:SentenceSelection}

In the sentence selection stage, we use the salience predictions from our GP
model as preferences in the AP clustering algorithm. The AP algorithm is 
parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
parameters, and did not perform robustly on our range of inputs.
In order to improve the quality of the clusters and exemplar selection,
we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
The raw preferences are scaled to lie within the  range $(-3, -2)$

The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
similarity between input sentences. Self-similarities and similarities below 
a threshold $\lambda$ were masked and the remaining values scaled to the range
$(-3, -1)$.

\subsubsection{Preferences}

Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
$\mathbf{U}$. We call this the redundancy penalty

$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
previous update. We calculate our new penalized preferences 
$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
within the range $(-3,-2)$.

\subsubsection{Filtering}

We run AP clustering with our rescaled and penalized sentence similarities
and preferences. When the clustering has converged we emit all exemplar
sentences (cluster centers) whose preference is $\lambda$ standard deviations
above the mean preference for the current time period and who do not belong
to singleton clusters as updates. These updates and their preferences are 
retained for penalty calculations in subsequent time steps.

\section{Run Submissions}

We submitted three different runs for official evaluation. The first system,
(AP+) used AP clustering without salience predictions or redundancy penalty.
The clustering was run every hour, and all non-singleton exemplars were 
taken as the updates.

Our second submission (AP+Sal+) uses our salience prediction model, with the
AP clustering to select sentences. We do not penalize for redundancy.

Our final run (AP+Sal+Red-) is the same as the previous run but the 
redundancy penalty is applied.

\section{Results}

Overall it appears that (AP+Sal+) is our best performing model of the three
submissions. One of the notable differences is in the average number of 
updates per event. (AP+Sal+) was by far the most terse system, with 
383.2 updates/event compared to 5997.8 and 1070.7 for (AP+) and (AP+Sal+Red-)
respectively.
Despite not returning many updates, $\mathbb{E}$[Latency Gain] was high enough
on average to outperform on the F1 score compared to the other two systems.
The differences in F1 from the baseline run for both (AP+Sal+) and 
(AP+Sal+Red-) are statistically significant; 
this is empirical validation of our salience modeling efforts.

\begin{figure}
\centering
\begin{tabular}{| c | c | c | c |}
\hline
\textbf{Run} & \textbf{nE[Lat. Gain]} & \textbf{Lat. Comp} & \textbf{F1} \\
\hline
AP+ & 0.0222  & 0.5777 & 0.0403\\
\hline
AP+Sal+ & 0.0751 &  0.4139 & $\mathbf{0.1162}$\\
\hline
AP+Sal+Red- & 0.0375 & 0.3413 & 0.0602\\
\hline
\end{tabular}
\caption{Submission results, averaged over 2014 TREC TS events.}
\end{figure}

Unfortunately, there is still much work to be done with our redundancy 
component. (AP+Sal+Red-) returned an order of magnitude more updates on 
average than (AP+Sal+) but still achieved comparatively lower comprehensiveness
scores. The redundancy penalty is perhaps too aggressive, driving the 
clustering to focus on very novel input spaces---so novel as to be 
irrelevant.

A related issue with our current model is that our salience predictions
are relative from the point of view of the clustering algorithm. If all inputs
have low salience predictions, the AP clustering will still find exemplars, 
albeit from the most salient inputs. This becomes especially problematic 
with the redundancy component, as with every timestep, we are evaluating 
riskier and riskier sets of more novel sentences, and being forced to return
something. This behavior can be observed especially in the updates from the
Chelyabinsk meteor event, where our document retrieval stage was 
under-performing and the predicted salience of all updates was quite low.
This would suggest implementing some minimum threshold of salience. We
initially experimented with such a threshold but found tuning it across 
events of different magnitudes to be difficult.

\section{Related Work}

A principal concern in extractive multi-document summarization is the
selection of salient sentences for inclusion in summary output
\cite{nenkova2012survey}.  This has often been approached as a ranking
problem.
%We broadly conceptualize this decision as either an intrinsic or extrinsic
%sentence evaluation process. Intrinsic approaches evaluate sentences
%individually, possibly by predicting the impact on summary quality using
%sentece level features. 
Sentences have been ranked by the average word probability, average tf-idf
score, and the number of topically related words (topic-signatures in the
summarization literature)
\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
statistics are easily computable from the input sentences, while the third
only requires an additional, generic background corpus.  Another ranking
approach, centroid summarization, involves creating an average bag of words
(BOW) vector, the centroid, from the input sentences and ranking sentences by
their similarity to the centroid \cite{radev2004centroid}.  Graph
\cite{erkan2004lexrank} and clustering
\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
based approaches, on the other hand, make use of pair-wise similarity
comparisons amongst input sentences.  In these models, salient sentences are
more central to the input or cluster, respectively.

%identify salient regions of the input space while simultaneously coping with
%redundancy.  Graph-based algorithms have been used to rank sentences
%Clustering algorithms, e.g., are commonly used to exploit redundancy in
%input. Input sentences are clustered and summaries are generated by selecting
%the most representative sentence from each cluster.  Graph-based models have
%also been used for summarization.  E.g., the LexRank algorithm treats
%sentences as nodes in a graph, where edges are constructed by way of cosine
%similarity between sentence nodes; edges are either continuosly weighted by
%similarity or discrete, existing only when the similarity is above a
%threshold.  The PageRank algorithm is used on the graph to find the most
%important sentence nodes. In both clustering and graph-based approaches,
%sentence salience is largely determined by the pairwise relations between
%sentences.

Supervised learning has also been applied to this task. Model features are
usually derived from human generated summaries, and are non-lexical in nature
(e.g., sentence starting position, number of topic-signatures, number of
unique words, word frequencies). Seminal work in this area has employed naive
Bayes and logistic regression classifiers to identify sentences for summary
inclusion \cite{kupiec1995trainable,conroy2001using}.

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary.
%}

%This work uses the TREC Stream Corpus data set that is in use by the TREC
%Temporal Summarization track \cite{frank2012building,aslam2013trec} .
%Generally, last year's participants used a pipelined approach to build
%summaries, generally ranking sentences, filtering out all but the most
%relevant, and then performing some sort of deduplication/redundancy removal
%step.  Ranking approaches ranged from simple query word match to more
%sophisticated query expansion and query based language model scoring
%\cite{liu2013ictnet,xu2013trec,baruah2013univ}.  Perhaps most similar to our
%approach, the system of \cite{xu2013trec} uses a weighted combination of
%features (similarity to query, named entity frequency, predicate frequency,
%presence of numerical values, sentence novelty, etc.) to score sentences;
%sentences above a threshold are added to the summary.  Both the weights and
%the threshold are selected by hand.
%
%
%Our system seeks to combine the best of these approaches, using supervised
%learning to predict salience rankings, and directly incorporate this
%information in a clustering algorithm to bias the formation of sentence
%clusters around highly salient regions.

%Broadly, we conceive of the above

%Another popular graph-based model

%Larger clusters generally indicate more important For example

%\cite{} also explore the use of a hidden Markov model to predict summary
%inclusion based on previously selected sentences.

%ranking.
%
%The field of extractive summarization has long been focused on determing
%sentence importance or salience.  Many These efforts have largely involved
%unsupervised, data-driven approaches, where systems rank input sentences
%using features easily computable from the input, and sometimes an unrelated
%background corpus.  The most basic sentence rankings are by average word
%probability and average TF-IDF score, statistics that are computable from the
%input alone.  One of the more effective techniques, topic-sigantures,
%involves identifying the most topically related words in the input sentences.
%This is done using a log-likelihood ratio test ($\chi^2$) to determine which
%of two hypotheses hold for a word in the input: (1) the probability of the
%word in the input distribution and the background distribution are identical,
%or (2) the probability of the word is greater in the input distribution than
%in the background distribution. If (2) is true with some statistical
%significance, the word is said to be a topic signature. As with the previous
%features, a common approach is to rank sentences by the number of
%topic-signatures they contain.
%
%In mulit-document summarization, clustering algorthims are often employed to
%determine sentence importance, with larger clusters corresponding to more
%important topics. 
%
%Supervised learning approaches have also been explored. Features are usually
%derived from human generated summaries, and are non-lexical in nature (e.g.,
%sentence starting position, number of topic-signatures, number of unique
%words, word frequencies). Seminal work in this area has employed naive Bayes
%and logistic regression classifiers to identify sentences for summary
%inclusion. \cite{} also explore the use of a hidden Markov model to predict
%summary inclusion based on previously selected sentences.
%
%If this  
%
%tried to rank sentences by their overlap with query words. 
%

%and models are developed to predict the likelihood of a sentences inclusion
%in a summary. Naive Bayes, hidden Markov models, and logistic regression to
%predict a sentence's summary inclusion.
%
%Researches have also experimented with supervised learning for summarization,
%however, machine learning approaches have not yielded significant advances in
%generic summarization. One difficulty is that gold data is difficult to
%obtain.
%
%
%In  
%
%
%%which remain competetive to supervised machine-learning techniques.
%
%have ranked sentences by
%
%
%their average word probability, average TF-IDF weighting, and/or topic
%signatures; these representations are all easily computable from the input
%sentences, and, for topic signatures, an unrelated background corpus.  Topic
%signatures in
%
% 
%These efforts can be broadly split into two camps: feature based ranking and
%clustering.
%
%
%
%Determing sentence importance has been A large portion of automatic
%summarization research has focused on  
%
%

%\newcommand{corpus}{}

%A common approach to automatic summarization is to identify sentences with the
%highest centrality with respect to the input sentences.  Intuitively,
%sentences with a high degree of centrality are more semantically related to
%the entire set of input sentences.  A summary can thus be obtained by
%returning the $k$ most central sentences.  This generally implies the
%calculation of pairwise distances between all sentences
%\cite{radev2004centroid, erkan2004lexrank}.  In these approaches, sentences
%are evaluated extrinsically by their distance to other sentences, either
%directly \cite{erkan2004lexrank} or through an aggregate centroid object
%\cite{radev2004centroid}. The distance between sentences is most commonly the
%cosine distance of sentence term-vectors, but in general this can be an
%arbitrary real-valued similarity function.
%%; the distance function imposes a topology on the space of input sentences.
%
%For some domains, it is very likely that we will have additional background
%knowledge that could be predictive of sentence salience for the event being
%summarized.  For example, certain kinds of information extracted from the
%sentence text (e.g., temporal or geographic proximity) can indicate relevance
%to a given event.  It would be difficult to incorporate this kind of salience
%into the measure of centrality.
%
%
%For example, consider a cluster of three sentence vectors $s_1 = (1,1,0)$,
%$s_2 = (1,1,1)$, and $s_3 = (0,1,1)$. Without any other information, $s_2$ has
%the highest degree of centrality, i.e. it has the highest average cosine
%similarity and smallest average Euclidean distance to the other sentences.
%Now, if we believe that $s_1$ is $\alpha$ times more salient than the other
%sentences, we cannot simply scale $s_1$ by $\alpha$--the average cosine
%similarity will remain unchanged, since the vector magnitude does not affect
%the angle and $s_1$ will have an even greater the average Euclidean distance
%from the rest.  Worse still, $s_2$ will still be the most representative
%sentence of the three.
%
%

%In order to predict the salience score for a sentence,

 %\cite{guo2012simple} produces a real-valued score of the similarity We use
 %\cite{guo2012simple} produces real valued similarity scores s.
 %\cite{guo2012simple} determines the similarity of two sentences by deriving
 %low dimensional We have As a proxy measure for sentence salience, we use the
 %maximum similarity of an input sentence to a human generated summary
 %sentence, using the state-of-the-art sentence similarity module of
 %\cite{guo2012simple}. \cite{guo2012simple} derive a lower-dimensional vector
 %of latent features for a sentence. The similarity of two sentences is then
 %the cosine similarity of the latent vectors.  In order to predict this
 %similarity we extract a variety of features including language model scores,
 %geographic relevance, and temporal relevance.  We learn a predictive
 %function of sentence-summary similarity by fitting a Gaussian process model
 %on these features. \cite{preotiuc2013temporal}

%As training data for this model, we take a portion of the sentences found in
%the relevant documents and find their highest similarity to the gold nugget
%information seen thus far.

%When assigning preference scores

%After obtaining similarity scores for sentences in our training set, we fit a
%Gaussian process model to learn

%Our model

%Correllating automatic measures of sentence salience with human judgments is
%a difficult task. , we use the maximum cosine similarity to

%For example, if we scaled the BOW vector based on our belief in its salience,
%it would not effect the cosine distance since the angle between two vectors
%is not changed by the magnitude of either

%Simply scaling the sentence vector based on our belief in it's fitness does
%not with our prior beliefs-- When using the BOWs representation, the input
%space is lexically defined, but our salience metric may not be based on any
%specific lexical items.

%These approaches make it difficult to leverage prior knowledge to guide the
%selection of salient sentences. Sentences

%
%Centroid based algorthims are commonly employed in the summarization
%literature to compress a large number of sentences into a much smaller,
%representative set. 
%
%The $k$-medoids clustering algorithm is commonly used to partition input
%sentences into topically distinct clusters, where the most representative
%sentences (the cluster medoids) are taken to build the summary. The quality
%of the summary in this framework is dependent on the method of sentence
%representation and the ability of a similarity or distance metric to
%adequately capture pairwise comparisons between sentences. Previous work in
%event clustering has examined the use of different term weightings (tf-idf,
%named-entities)  and lower dimensional term representations (LSA, LDA) for
%use in summarization.
%
%
%A popular graph-based alternative, LexRank, treats each sentence as a node in
%a graph, and identifies the nodes that are most central by in-degree.  The
%existence of an edge between two sentences is determined by a similarity
%function (either by setting a threshold above which the edge exists, or by
%taking the similarity as a continuous edge weight). The sentences with the
%highest eigenvector centrality can be found using random walks (PageRank).
%The $k$ most central sentences are used as the summary.
%
%One drawback of these algorithms is that they do not provide an elegant way
%to incorporate prior information about sentence salience.

%Move this out!
%In our approach, the system generates clusters using an affinity propagation
%algorithm and from each cluster an exemplar sentence is selected that is added
%to the summary.  In the following sections, we show how  prior information
%representing salience can easily be incorporated into the affinity propagation
%algorithm.  We believe the incorporation of salience to be useful in noisy
%environments (e.g., a web crawl), and that it can help the formation of
%clusters around the most relevant inputs. Our current system  is trained using
%features derived from location, changes in wording across time and language
%models that characterize the language of disaster to generate summaries at
%regular intervals across time.  As we develop the system further, we will
%extend it to generate updates across time,  penalizing the salience of
%concepts already selected by the summarizer to encourage the discovery of
%novel sentences as the event unfolds. 
%

%\begin{table*}[ht] \begin{center} \begin{tabular}{| c p{7.0cm} || c p{7.0cm}
%|} \hline Preference & AP Clustering & Preference & $k$-Medoids Clustering\\
%\hline 9.010 & The magnitude-7.5 quake, about 20 miles deep, was centered off
%the town of Champerico .People fled buildings in Guatemala City , in Mexico
%City and in the capital of the Mexican state of Chiapas , across the border
%from Guatemala & 9.010 & The magnitude-7.5 quake, about 20 miles deep, was
%centered off the town of Champerico .People fled buildings in Guatemala City ,
%in Mexico City and in the capital of the Mexican state of Chiapas , across the
%border from Guatemala   \\ \hline 9.010 & A reporter in the town of San Marcos
%, about 80 miles north of the epicenter, told local radio station Emisoras
%Unidas that houses had collapsed onto residents and smashed televisions and
%other appliances had been scattered into the streets.  & 3.007 & ``Things fell
%in my kitchen .'' Perez said more than 2,000 soldiers were deployed from a
%base in San Marcos to help with disaster relief.\\ \hline 9.007 & The local
%fire department said on its Twitter account that a school had collapsed and
%eight injured people had been taken to a nearby hospital.  & 3.007 &  Ingrid
%Lopez , who went to the hospital with a 72-year-old aunt whose legs was
%crushed by a falling wall, said she had waited hours for an X-ray.\\ \hline
%7.008 & There are three confirmed dead and many missing after the strongest
%earthquake to hit Guatemala since a deadly 1976 quake that killed 23,000.  &
%1.007 & Hundreds of people crammed into the hallways of the small town
%hospital waiting for medical staff to help out hundreds of injured family
%members, some complaining they were not getting care quickly enough. \\ \hline
%\end{tabular} \end{center} \caption{Example summary using affinity propagation
%(left) and $k$-medoids (right)} \end{table*} \begin{table*}[t] \begin{center}
%\begin{tabular}{| c || c | c | c || c | c | c || c | c | c |} \hline Method &
%\multicolumn{3}{c}{ROUGE-1} & \multicolumn{3}{c}{ROUGE-2} &
%\multicolumn{3}{c}{ROUGE-3} \\ & Recall & Prec. & F-1 & Recall & Prec. & F-1 &
%Recall & Prec. & F-1 \\ \hline $k$-medoids & 0.127 & 0.414 & 0.181 & 0.025 &
%0.076 & 0.035 & 0.003 & 0.010 & 0.005\\ \hline AP & 0.117 & 0.440 & 0.173 &
%0.022 & 0.082 & 0.033 & 0.004 & 0.015 & 0.006\\ \hline \end{tabular}
%\end{center} \caption{ROUGE scores for $k$-medoids and affinity propagation
%methods} \end{table*}
%
%
%
%
%\section{Experiments}
%
%We carried out a small set of initial experiments on one event.  We collected
%a subset of pages from the TREC Stream Corpus that were relevant to a 2012
%earthquake off the coast of Guatemala, and further subdivided this collection
%by the hour they were created. For each hour we generated a summary using the
%AP clustering algorithm. 
%
% We also generated baseline summaries using the $k$-medoids (using the
% Partitioning Around Medoids algorithm), setting $k = |\mathcal{E}|$ , i.e.
% the number of exemplar sentences returned by the AP. Because $k$-medoids
% begins with a random initialization, we took the best (minimum average
% distance) result of 100 restarts.
%
%Table 1 shows example output of the AP and $k$-medoids generated summaries.
%Sentences are ranked by preference score, although preference has no effect on
%the $k$-medoids algorithm.  Quantitatively, AP exemplar sentences had higher
%predicted sentence quality scores (preferences) than the cluster medoids.
%Qualitatively, the AP method appears to select more general details about the
%earthquake. Looking at the third sentence selection in table 1, we can see
%that $k$-medoids selects a personal experience that was reported. This is
%perhaps less newsworthy or reportable compared to the third sentence in in the
%AP generated summary which reports a notable structure collapse and injuries
%related to the quake.  We believe AP results in a more readable and
%informative summary, although we have yet to perform a rigorous human
%evaluation of the summary output.
%
%
%We evaluated both algorithms with the ROUGE toolkit \cite{lin2004rouge}.
%N-ROUGE works by calculating the n-gram recall and precision of an
%automatically generated summary in reference to a model summary.  We created
%model summaries by taking the gold nugget sentences with timestamps up to and
%including the current system time as the gold summary for that hour.
%%\fdcomment{may be interesting to measure track metrics to measure latency,
%%redundancy, etc.} KM - Ask Chris
%
%Table 2 shows average recall, precision, and F-measure for various orders of
%ROUGE score. AP demonstrated consistently higher precision than our baseline.
%%scores over $k$-medoids (ROUGE evaluations are usually focused on recall).
%While not statistically significant, it is difficult to show significance with
%Rouge using a small test; we hope further tests will confirm this improvement. 
%%   , AP had higher precision on average.
%On average, the AP summaries were slightly shorter than the baseline, which
%would partially explain this difference. It is also possible that our language
%models are biased toward shorter sentences; we are more likely to have seen a
%shorter sentence in the language model input.
%%\fdcomment{is there a theory to why this is happening?}
%We are currently adapting our summarizer to add updates over time, and
%maintaining precision will be important to prevent topic drift.  
%
%
%
%%\begin{table}[h!] \tiny \begin{center} \begin{tabular}{| c p{3.5cm} |} \hline
%%Preference & Sentence\\ \hline 9.010 & The magnitude-7.5 quake, about 20
%%miles deep, was centered off the town of Champerico .People fled buildings in
%%Guatemala City , in Mexico City and in the capital of the Mexican state of
%%Chiapas , across the border from Guatemala\\ \hline 3.007 & "Things fell in
%%my kitchen ." Perez said more than 2,000 soldiers were deployed from a base
%%in San Marcos to help with disaster relief.\\ \hline 3.007 & Ingrid Lopez ,
%%who went to the hospital with a 72-year-old aunt whose legs was crushed by a
%%falling wall, said she had waited hours for an X-ray. \\ \hline 1.007 &
%%Hundreds of people crammed into the hallways of the small town hospital
%%waiting for medical staff to help out hundreds of injured family members,
%%some complaining they were not getting care quickly enough. \\ \hline
%%\end{tabular} \end{center} %  \caption{Example summary using $k$-medoids with
%%100 random restarts} \end{table}
%
%
%

%\section{Conclusions and Future Work}

%We have developed a summarizer that can generate summaries over time from web
%crawls on disasters. We show that our method of identifying exemplar sentences
%for a summary using AP clustering produces  summaries with higher precision
%    compared to those based on clustering with K-medoids. A key component of
%    our approach is the prediction of salient information using features based
%    on location, temporal changes in topic, and two different language models.
%
%Currently, we run each hour of summarization independently. In order to avoid
%repeating information, we would like to incorporate previously chosen
%exemplars in the preference computation. One possibility would be to
%down-weight a candidate exemplar's preference based on its similarity to
%previous exemplars.
%
%Secondly, we would like to do more intelligent inference of missing
%geographical information since not all sentences contain locations. Currently
%we are using mean values for missing data.
%
%Finally, we would like to experiment with non-symmetric similarity matrices,
%specifically using narrative chains\cite{chambers2009unsupervised}. Under this
%model $S(i,j)$ would express the likelihood that the events in sentence $j$
%precede the events in sentence $i$. We hope such a parameterization would
%promote more causally motivated sentences into exemplar positions, which would
%better describe the disaster event domain.
%
%
\bibliographystyle{abbrv}
%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this
%case
\bibliography{cites.bib}
% You must have a proper ".bib" file and remember to run: latex bibtex latex
% latex to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional \balancecolumns \balancecolumns % GM June 2007 That's
%all folks!
\end{document}

### 2015-01-13 introduction.tex
refactoring sections.

--- 2015-01-13 paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012 This file should be compiled with
-% V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls' V2.5
-% LaTeX2e document class file. It is for those submitting articles to ACM
-% Conference Proceedings WHO DO NOT WISH TO STRICTLY ADHERE TO THE SIGS
-% (PUBS-BOARD-ENDORSED) STYLE.  The 'sig-alternate.cls' file will produce a
-% similar-looking, albeit, 'tighter' paper resulting in, invariably, fewer
-% pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces: 1) The Permission
-% Statement 2) The Conference (location) Info information 3) The Copyright
-% Line with ACM data 4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which DOES NOT produce 1) thru'
-% 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within the source
-% .tex file, over both the CopyrightYear (defaulted to 200X) and the ACM
-% Copyright Data (defaulted to X-XXXXX-XX-X/XX/XX).  e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the
-% copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use the .bib file (from which
-% the .bbl file % is produced).  REMEMBER HOWEVER: After having produced the
-% .bbl file, and prior to final submission, you *NEED* to 'insert' your .bbl
-% file into your source .tex file so as to provide ONE 'self-contained' source
-% file.
-%
-% ================= IF YOU HAVE QUESTIONS ======================= Questions
-% regarding the SIGS styles, SIGS policies and procedures, Conferences etc.
-% should be sent to Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-%\documentclass{sig-alternate} \usepackage{url} \usepackage{color}
-\documentclass[10pt]{article} \usepackage{url} \usepackage{color}
-\usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage[]{algorithm2e}
-\DeclareMathOperator{\corpus}{\mathcal{C}}
-\DeclareMathOperator{\doc}{\mathnormal{d}}
-\DeclareMathOperator{\sent}{\mathnormal{s}}
-\DeclareMathOperator{\order}{\pi}
-\DeclareMathOperator{\dtime}{\mathnormal{t}}
-\DeclareMathOperator{\hour}{\mathnormal{h}}
-\DeclareMathOperator{\hours}{\mathcal{H}}
-\DeclareMathOperator{\Sim}{\mathbf{K}}
-\DeclareMathOperator{\SMat}{\mathbf{X}}
-\DeclareMathOperator{\Pref}{\boldsymbol{\pi}}
-\DeclareMathOperator{\Exemp}{\mathnormal{Exemplars}}
-\DeclareMathOperator{\Updates}{\mathnormal{Updates}}
-\usepackage{cleveref}
-\crefname{section}{§}{§§}
-\Crefname{section}{§}{§§}
-
-\usepackage{authblk}
-\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
-\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
-
-\begin{document}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
-%- IF NEED BE.  \crdata{0-12345-67-8/90/01}  % Allows default copyright data
-%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.  --- End of Author
-%Metadata ---
-
-\title{Columbia University at TREC 2014: Temporal Summarization}
-
-\author[1]{Chris Kedzie}
-\author[1]{Kathleen McKeown}
-\author[2]{Fernando Diaz}
-\affil[1]{Columbia University, Department of Computer Science}
-\affil[2]{Microsoft Research}
-%    Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
-%Format\titlenote{(Produces the permission block, and copyright information).
-%For use with SIG-ALTERNATE.CLS. Supported by ACM.}} \subtitle{[Extended
-%Abstract] \titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement and
-% alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time' i.e. three
-% 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of "name/affiliations" may
-% appear. We just ask that you restrict the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate' we ask you to refrain
-% from putting more than six authors (two rows with three columns) beneath the
-% article title.  More than six makes the first-page appear very cluttered
-% indeed.
-%
-% Use the \alignauthor commands to handle the names and affiliations for an
-% 'aesthetic maximum' of six authors.  Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the \additionalauthors
-% command.  These 'additional authors' will be output/set for you without
-% further effort on your part as the last section in the body of your article
-% BEFORE References or any Appendices.
-
-%\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting reasons)
-% and the remaining two appear in the \additionalauthors section.
-%
-%\author{
-% You can go ahead and credit any number of authors here, e.g. one 'row of
-% three' or two rows (consisting of one row of three and a second row of one,
-% two or three).
-%
-% The command \alignauthor (no curly braces needed) should precede each author
-% name, affiliation/snail-mail address and e-mail address. Additionally, tag
-% each line of affiliation/address with \affaddr, and tag the e-mail address
-% with \email.
-%
-% 1st. author
-
-%\alignauthor Chris Kedzie\\ \affaddr{Columbia University}\\
-%\affaddr{Department of Computer Science} \email{kedzie@cs.columbia.edu}
-
-% 2nd. author
-%\alignauthor Kathleen McKeown\\ \affaddr{Columbia University}\\
-%\affaddr{Department of Computer Science}\\ \email{kathy@cs.columbia.edu}
-
-% 3rd. author
-%\alignauthor Fernando Diaz\\ \affaddr{Microsoft Research}\\
-%\email{fdiaz@microsoft.com}
-
-% There's nothing stopping you putting the seventh, eighth, etc.  author on
-% the opening page (as the 'third row') but we ask, for aesthetic reasons that
-% you place these 'additional authors' in the \additional authors block, viz.
-%} 
-\date{29 October 2014}
-% Just remember to make sure that the TOTAL number of authors is the number
-% that will appear on the first page PLUS the number that will appear in the
-% \additionalauthors section.
-
-\maketitle \begin{abstract} 
-Information need is highest during times of crisis; both officials and 
-civilians need effective access to make informed decisions about their 
-safety and the safety of others. However, these situations pose many 
-interesting challenges for the IR systems that support these decisions.
-Specifically, we address event monitoring and updating, a.k.a. the
-temporal summarization task (TS). Given an event query, we monitor streams
-of online news and select important text to present to the user. Updates
-are selected by predicting the salience of the text w.r.t. to the event.
-Our model utilizes a variety of features at the query, document, and corpus
-level. Additionally, these predictions are used to directly bias a clustering
-algorithm for sentence selection, 
-increasing the novelty of the updates. We evaluate our system
-on a varied set of retrospective events using a combination of human 
-judgements
-and automatic evaluation measures, as compared to a baseline system.
-We demonstrate the effect of the various feature groups, and the importance
-of salience prediction w.r.t. update precision when compared to a clustering
-only baseline.
-
-\end{abstract}
-
-% A category with the (minimum) three required fields
-%\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
-%performance measures]
-
-%\terms{Summarization}
-
-%\keywords{Extractive Summarization, Affinity Propagation}
-
 \section{Introduction}
-
+\label{sec:introduction}
 During crises, information is critical for first responders and those caught
@@
 
-The remainder of the paper is organized as follows. In section
+The remainder of the paper is organized as follows. We begin with a review of related work in the information retrieval and multi-document summarization literature.  In section
 ~\ref{sec:background} we give an overview of the semantic similarity, 
@@
 of our system, and then explain each component in detail.
-We finish with the presentation of our results and related works.
-
-\section{Background}\label{sec:background}
-
-\subsection{Affinity Propagation}
-
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\section{Data}\label{sec:data}
-
-Our primary dataset is the TREC 2014 Stream Corpus, a ?tb corpus of hourly 
-web crawls from October 2011
-through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
-As per the TS specifications, all summary sentences come from processing this
-corpus in a time aligned manner. Additionally, we use the query terms 
-provided by the track organizers for document retrieval.
-
-For our semantic similarity component (\cref{subsec:semsim}),
-domain specific term-sentence matrices were built using subsections of
-Wikipedia. For each event type we identified one or more related Wikipedia
-categories and collected all pages that were listed under that category.
-For all pages collected, we used the latest possible revision date before the
-start of the Stream Corpus.
-
-We use the same Wikipedia corpora to construct the lanuage models used
-in our salience prediction model (\cref{subsubsec:lm}).
-
-\begin{figure}
-\begin{tabular}{| p{5cm} c c |}
-\hline
-Event & WP Categories & No. Docs/Sents/Words\\
-\hline \hline
-Boston\_Marathon\_bombings \newline
-Christopher\_Dorner\_shootings\_and\_manhunt \newline 
- In\_Amenas\_hostage\_crisis
-& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
-\hline
- Cyclone\_Oswald \newline
-Early\_2012\_European\_cold\_wave \newline
-February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
-\hline
-Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
-\hline
-Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
-\hline
-Port\_Said\_Stadium\_riot \newline
-2012\_Afghanistan\_Quran\_burning\_protests \newline
-2011-13\_Russian\_protests \newline
-2012\_Romanian\_protests \newline
-2012-13\_Egyptian\_protests \newline
-2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
-2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
-\hline
-\end{tabular}
-\caption{Domain specific Wikipedia corpora for semantic similarity and 
-language models. }
-\end{figure}
-
-We use the TREC 2013 Stream Corpus and TS 2013 event/gold nugget data to 
-train our salience models. See section~\cref{subsec:Predict} 
-for a detailed description of our sampling procedure to generate labeled
-data.
-
-%Our documents for summarization come from the online news portion of the TREC
-%Stream Corpus, a 6.45tb corpus obtained by hourly web crawls from October 2011
-%through mid February 2013
-%
-%Summary events come from the TREC Temporal Summarization track, and include
-%natural disasters like Hurricane Sandy as well as man-made events like a 2012
-%train accident in Buenos
-%Aires.\footnote{\url{http://trec.nist.gov/data/tempsumm2013.html}} The track
-%organizers also provide a search query for each event \cite{aslam2013trec}.
-%For each event, we collect the documents that contain all query words and
-%stratify them by the hour they were collected. 
-%
-%For evaluation purposes, the track organizers also provided gold nugget
-%information (i.e. important pieces of information, usually the length of a
-%short clause or sentence). These gold nuggets come from the event's related
-%Wikipedia article and also include the timestamp of when they were added to
-%the page.
-%
-%%KM _ Ask Chris: is this true?
-%To create hourly gold summaries to evaluate our system, we simply take the set
-%of gold nugget information from the start of the event up to the current hour. 
-%
-
-\section{Approach}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-\begin{figure}
-\centering
-\begin{algorithm}[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-\end{figure}
-
-\subsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, and the number of query words present in the sentence.  Query
-words include the event's type (e.g., \emph{earthquake}) and are expanded with
-the event type's WordNet \cite{miller1995wordnet} synset, hypernyms, and
-hyponyms.  For \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.
-
-\section{Run Submissions}
-
-We submitted three different runs for official evaluation. The first system,
-(AP+) used AP clustering without salience predictions or redundancy penalty.
-The clustering was run every hour, and all non-singleton exemplars were 
-taken as the updates.
-
-Our second submission (AP+Sal+) uses our salience prediction model, with the
-AP clustering to select sentences. We do not penalize for redundancy.
-
-Our final run (AP+Sal+Red-) is the same as the previous run but the 
-redundancy penalty is applied.
-
-\section{Results}
-
-Overall it appears that (AP+Sal+) is our best performing model of the three
-submissions. One of the notable differences is in the average number of 
-updates per event. (AP+Sal+) was by far the most terse system, with 
-383.2 updates/event compared to 5997.8 and 1070.7 for (AP+) and (AP+Sal+Red-)
-respectively.
-Despite not returning many updates, $\mathbb{E}$[Latency Gain] was high enough
-on average to outperform on the F1 score compared to the other two systems.
-The differences in F1 from the baseline run for both (AP+Sal+) and 
-(AP+Sal+Red-) are statistically significant; 
-this is empirical validation of our salience modeling efforts.
-
-\begin{figure}
-\centering
-\begin{tabular}{| c | c | c | c |}
-\hline
-\textbf{Run} & \textbf{nE[Lat. Gain]} & \textbf{Lat. Comp} & \textbf{F1} \\
-\hline
-AP+ & 0.0222  & 0.5777 & 0.0403\\
-\hline
-AP+Sal+ & 0.0751 &  0.4139 & $\mathbf{0.1162}$\\
-\hline
-AP+Sal+Red- & 0.0375 & 0.3413 & 0.0602\\
-\hline
-\end{tabular}
-\caption{Submission results, averaged over 2014 TREC TS events.}
-\end{figure}
-
-Unfortunately, there is still much work to be done with our redundancy 
-component. (AP+Sal+Red-) returned an order of magnitude more updates on 
-average than (AP+Sal+) but still achieved comparatively lower comprehensiveness
-scores. The redundancy penalty is perhaps too aggressive, driving the 
-clustering to focus on very novel input spaces---so novel as to be 
-irrelevant.
-
-A related issue with our current model is that our salience predictions
-are relative from the point of view of the clustering algorithm. If all inputs
-have low salience predictions, the AP clustering will still find exemplars, 
-albeit from the most salient inputs. This becomes especially problematic 
-with the redundancy component, as with every timestep, we are evaluating 
-riskier and riskier sets of more novel sentences, and being forced to return
-something. This behavior can be observed especially in the updates from the
-Chelyabinsk meteor event, where our document retrieval stage was 
-under-performing and the predicted salience of all updates was quite low.
-This would suggest implementing some minimum threshold of salience. We
-initially experimented with such a threshold but found tuning it across 
-events of different magnitudes to be difficult.
-
-\section{Related Work}
-
-A principal concern in extractive multi-document summarization is the
-selection of salient sentences for inclusion in summary output
-\cite{nenkova2012survey}.  This has often been approached as a ranking
-problem.
-%We broadly conceptualize this decision as either an intrinsic or extrinsic
-%sentence evaluation process. Intrinsic approaches evaluate sentences
-%individually, possibly by predicting the impact on summary quality using
-%sentece level features. 
-Sentences have been ranked by the average word probability, average tf-idf
-score, and the number of topically related words (topic-signatures in the
-summarization literature)
-\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
-statistics are easily computable from the input sentences, while the third
-only requires an additional, generic background corpus.  Another ranking
-approach, centroid summarization, involves creating an average bag of words
-(BOW) vector, the centroid, from the input sentences and ranking sentences by
-their similarity to the centroid \cite{radev2004centroid}.  Graph
-\cite{erkan2004lexrank} and clustering
-\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
-based approaches, on the other hand, make use of pair-wise similarity
-comparisons amongst input sentences.  In these models, salient sentences are
-more central to the input or cluster, respectively.
-
-%identify salient regions of the input space while simultaneously coping with
-%redundancy.  Graph-based algorithms have been used to rank sentences
-%Clustering algorithms, e.g., are commonly used to exploit redundancy in
-%input. Input sentences are clustered and summaries are generated by selecting
-%the most representative sentence from each cluster.  Graph-based models have
-%also been used for summarization.  E.g., the LexRank algorithm treats
-%sentences as nodes in a graph, where edges are constructed by way of cosine
-%similarity between sentence nodes; edges are either continuosly weighted by
-%similarity or discrete, existing only when the similarity is above a
-%threshold.  The PageRank algorithm is used on the graph to find the most
-%important sentence nodes. In both clustering and graph-based approaches,
-%sentence salience is largely determined by the pairwise relations between
-%sentences.
-
-Supervised learning has also been applied to this task. Model features are
-usually derived from human generated summaries, and are non-lexical in nature
-(e.g., sentence starting position, number of topic-signatures, number of
-unique words, word frequencies). Seminal work in this area has employed naive
-Bayes and logistic regression classifiers to identify sentences for summary
-inclusion \cite{kupiec1995trainable,conroy2001using}.
-
-%\fdadd{
-Several researchers have recognized the importance of summarization during
-natural disasters.  Guo \textit{et al.} developed a system for detecting
-novel, relevant, and comprehensive sentences immediately after a natural
-disaster \cite{qi:temporal-summarization}.  The method uses a model of
-sentence relevance and novelty in order to select appropriate updates.
-Training data for regression targets is automatically generated from
-retrospective Wikipedia data.  The system is evaluated on news documents
-related to 197 natural and human disasters from 2009 to 2011 using variants of
-Rouge modified to capture novelty, relevance, and comprehensiveness
-\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
-efficiency detect important updates during natural disasters
-\cite{wang:update-summarization}.  The algorithm works by hierarchically
-clustering sentences online, allowing the system to output a more expressive
-narrative structure than Guo \textit{et al.}.  The method is evaluated on
-official press releases related to Hurricane Wilma  in 2005 using Rouge score
-between the system summary and a manually generated target summary.
-%}
-
-%This work uses the TREC Stream Corpus data set that is in use by the TREC
-%Temporal Summarization track \cite{frank2012building,aslam2013trec} .
-%Generally, last year's participants used a pipelined approach to build
-%summaries, generally ranking sentences, filtering out all but the most
-%relevant, and then performing some sort of deduplication/redundancy removal
-%step.  Ranking approaches ranged from simple query word match to more
-%sophisticated query expansion and query based language model scoring
-%\cite{liu2013ictnet,xu2013trec,baruah2013univ}.  Perhaps most similar to our
-%approach, the system of \cite{xu2013trec} uses a weighted combination of
-%features (similarity to query, named entity frequency, predicate frequency,
-%presence of numerical values, sentence novelty, etc.) to score sentences;
-%sentences above a threshold are added to the summary.  Both the weights and
-%the threshold are selected by hand.
-%
-%
-%Our system seeks to combine the best of these approaches, using supervised
-%learning to predict salience rankings, and directly incorporate this
-%information in a clustering algorithm to bias the formation of sentence
-%clusters around highly salient regions.
-
-%Broadly, we conceive of the above
-
-%Another popular graph-based model
-
-%Larger clusters generally indicate more important For example
-
-%\cite{} also explore the use of a hidden Markov model to predict summary
-%inclusion based on previously selected sentences.
-
-%ranking.
-%
-%The field of extractive summarization has long been focused on determing
-%sentence importance or salience.  Many These efforts have largely involved
-%unsupervised, data-driven approaches, where systems rank input sentences
-%using features easily computable from the input, and sometimes an unrelated
-%background corpus.  The most basic sentence rankings are by average word
-%probability and average TF-IDF score, statistics that are computable from the
-%input alone.  One of the more effective techniques, topic-sigantures,
-%involves identifying the most topically related words in the input sentences.
-%This is done using a log-likelihood ratio test ($\chi^2$) to determine which
-%of two hypotheses hold for a word in the input: (1) the probability of the
-%word in the input distribution and the background distribution are identical,
-%or (2) the probability of the word is greater in the input distribution than
-%in the background distribution. If (2) is true with some statistical
-%significance, the word is said to be a topic signature. As with the previous
-%features, a common approach is to rank sentences by the number of
-%topic-signatures they contain.
-%
-%In mulit-document summarization, clustering algorthims are often employed to
-%determine sentence importance, with larger clusters corresponding to more
-%important topics. 
-%
-%Supervised learning approaches have also been explored. Features are usually
-%derived from human generated summaries, and are non-lexical in nature (e.g.,
-%sentence starting position, number of topic-signatures, number of unique
-%words, word frequencies). Seminal work in this area has employed naive Bayes
-%and logistic regression classifiers to identify sentences for summary
-%inclusion. \cite{} also explore the use of a hidden Markov model to predict
-%summary inclusion based on previously selected sentences.
-%
-%If this  
-%
-%tried to rank sentences by their overlap with query words. 
-%
-
-%and models are developed to predict the likelihood of a sentences inclusion
-%in a summary. Naive Bayes, hidden Markov models, and logistic regression to
-%predict a sentence's summary inclusion.
-%
-%Researches have also experimented with supervised learning for summarization,
-%however, machine learning approaches have not yielded significant advances in
-%generic summarization. One difficulty is that gold data is difficult to
-%obtain.
-%
-%
-%In  
-%
-%
-%%which remain competetive to supervised machine-learning techniques.
-%
-%have ranked sentences by
-%
-%
-%their average word probability, average TF-IDF weighting, and/or topic
-%signatures; these representations are all easily computable from the input
-%sentences, and, for topic signatures, an unrelated background corpus.  Topic
-%signatures in
-%
-% 
-%These efforts can be broadly split into two camps: feature based ranking and
-%clustering.
-%
-%
-%
-%Determing sentence importance has been A large portion of automatic
-%summarization research has focused on  
-%
-%
-
-%\newcommand{corpus}{}
-
-%A common approach to automatic summarization is to identify sentences with the
-%highest centrality with respect to the input sentences.  Intuitively,
-%sentences with a high degree of centrality are more semantically related to
-%the entire set of input sentences.  A summary can thus be obtained by
-%returning the $k$ most central sentences.  This generally implies the
-%calculation of pairwise distances between all sentences
-%\cite{radev2004centroid, erkan2004lexrank}.  In these approaches, sentences
-%are evaluated extrinsically by their distance to other sentences, either
-%directly \cite{erkan2004lexrank} or through an aggregate centroid object
-%\cite{radev2004centroid}. The distance between sentences is most commonly the
-%cosine distance of sentence term-vectors, but in general this can be an
-%arbitrary real-valued similarity function.
-%%; the distance function imposes a topology on the space of input sentences.
-%
-%For some domains, it is very likely that we will have additional background
-%knowledge that could be predictive of sentence salience for the event being
-%summarized.  For example, certain kinds of information extracted from the
-%sentence text (e.g., temporal or geographic proximity) can indicate relevance
-%to a given event.  It would be difficult to incorporate this kind of salience
-%into the measure of centrality.
-%
-%
-%For example, consider a cluster of three sentence vectors $s_1 = (1,1,0)$,
-%$s_2 = (1,1,1)$, and $s_3 = (0,1,1)$. Without any other information, $s_2$ has
-%the highest degree of centrality, i.e. it has the highest average cosine
-%similarity and smallest average Euclidean distance to the other sentences.
-%Now, if we believe that $s_1$ is $\alpha$ times more salient than the other
-%sentences, we cannot simply scale $s_1$ by $\alpha$--the average cosine
-%similarity will remain unchanged, since the vector magnitude does not affect
-%the angle and $s_1$ will have an even greater the average Euclidean distance
-%from the rest.  Worse still, $s_2$ will still be the most representative
-%sentence of the three.
-%
-%
-
-%In order to predict the salience score for a sentence,
-
- %\cite{guo2012simple} produces a real-valued score of the similarity We use
- %\cite{guo2012simple} produces real valued similarity scores s.
- %\cite{guo2012simple} determines the similarity of two sentences by deriving
- %low dimensional We have As a proxy measure for sentence salience, we use the
- %maximum similarity of an input sentence to a human generated summary
- %sentence, using the state-of-the-art sentence similarity module of
- %\cite{guo2012simple}. \cite{guo2012simple} derive a lower-dimensional vector
- %of latent features for a sentence. The similarity of two sentences is then
- %the cosine similarity of the latent vectors.  In order to predict this
- %similarity we extract a variety of features including language model scores,
- %geographic relevance, and temporal relevance.  We learn a predictive
- %function of sentence-summary similarity by fitting a Gaussian process model
- %on these features. \cite{preotiuc2013temporal}
-
-%As training data for this model, we take a portion of the sentences found in
-%the relevant documents and find their highest similarity to the gold nugget
-%information seen thus far.
-
-%When assigning preference scores
-
-%After obtaining similarity scores for sentences in our training set, we fit a
-%Gaussian process model to learn
-
-%Our model
-
-%Correllating automatic measures of sentence salience with human judgments is
-%a difficult task. , we use the maximum cosine similarity to
-
-%For example, if we scaled the BOW vector based on our belief in its salience,
-%it would not effect the cosine distance since the angle between two vectors
-%is not changed by the magnitude of either
-
-%Simply scaling the sentence vector based on our belief in it's fitness does
-%not with our prior beliefs-- When using the BOWs representation, the input
-%space is lexically defined, but our salience metric may not be based on any
-%specific lexical items.
-
-%These approaches make it difficult to leverage prior knowledge to guide the
-%selection of salient sentences. Sentences
-
-%
-%Centroid based algorthims are commonly employed in the summarization
-%literature to compress a large number of sentences into a much smaller,
-%representative set. 
-%
-%The $k$-medoids clustering algorithm is commonly used to partition input
-%sentences into topically distinct clusters, where the most representative
-%sentences (the cluster medoids) are taken to build the summary. The quality
-%of the summary in this framework is dependent on the method of sentence
-%representation and the ability of a similarity or distance metric to
-%adequately capture pairwise comparisons between sentences. Previous work in
-%event clustering has examined the use of different term weightings (tf-idf,
-%named-entities)  and lower dimensional term representations (LSA, LDA) for
-%use in summarization.
-%
-%
-%A popular graph-based alternative, LexRank, treats each sentence as a node in
-%a graph, and identifies the nodes that are most central by in-degree.  The
-%existence of an edge between two sentences is determined by a similarity
-%function (either by setting a threshold above which the edge exists, or by
-%taking the similarity as a continuous edge weight). The sentences with the
-%highest eigenvector centrality can be found using random walks (PageRank).
-%The $k$ most central sentences are used as the summary.
-%
-%One drawback of these algorithms is that they do not provide an elegant way
-%to incorporate prior information about sentence salience.
-
-%Move this out!
-%In our approach, the system generates clusters using an affinity propagation
-%algorithm and from each cluster an exemplar sentence is selected that is added
-%to the summary.  In the following sections, we show how  prior information
-%representing salience can easily be incorporated into the affinity propagation
-%algorithm.  We believe the incorporation of salience to be useful in noisy
-%environments (e.g., a web crawl), and that it can help the formation of
-%clusters around the most relevant inputs. Our current system  is trained using
-%features derived from location, changes in wording across time and language
-%models that characterize the language of disaster to generate summaries at
-%regular intervals across time.  As we develop the system further, we will
-%extend it to generate updates across time,  penalizing the salience of
-%concepts already selected by the summarizer to encourage the discovery of
-%novel sentences as the event unfolds. 
-%
-
-%\begin{table*}[ht] \begin{center} \begin{tabular}{| c p{7.0cm} || c p{7.0cm}
-%|} \hline Preference & AP Clustering & Preference & $k$-Medoids Clustering\\
-%\hline 9.010 & The magnitude-7.5 quake, about 20 miles deep, was centered off
-%the town of Champerico .People fled buildings in Guatemala City , in Mexico
-%City and in the capital of the Mexican state of Chiapas , across the border
-%from Guatemala & 9.010 & The magnitude-7.5 quake, about 20 miles deep, was
-%centered off the town of Champerico .People fled buildings in Guatemala City ,
-%in Mexico City and in the capital of the Mexican state of Chiapas , across the
-%border from Guatemala   \\ \hline 9.010 & A reporter in the town of San Marcos
-%, about 80 miles north of the epicenter, told local radio station Emisoras
-%Unidas that houses had collapsed onto residents and smashed televisions and
-%other appliances had been scattered into the streets.  & 3.007 & ``Things fell
-%in my kitchen .'' Perez said more than 2,000 soldiers were deployed from a
-%base in San Marcos to help with disaster relief.\\ \hline 9.007 & The local
-%fire department said on its Twitter account that a school had collapsed and
-%eight injured people had been taken to a nearby hospital.  & 3.007 &  Ingrid
-%Lopez , who went to the hospital with a 72-year-old aunt whose legs was
-%crushed by a falling wall, said she had waited hours for an X-ray.\\ \hline
-%7.008 & There are three confirmed dead and many missing after the strongest
-%earthquake to hit Guatemala since a deadly 1976 quake that killed 23,000.  &
-%1.007 & Hundreds of people crammed into the hallways of the small town
-%hospital waiting for medical staff to help out hundreds of injured family
-%members, some complaining they were not getting care quickly enough. \\ \hline
-%\end{tabular} \end{center} \caption{Example summary using affinity propagation
-%(left) and $k$-medoids (right)} \end{table*} \begin{table*}[t] \begin{center}
-%\begin{tabular}{| c || c | c | c || c | c | c || c | c | c |} \hline Method &
-%\multicolumn{3}{c}{ROUGE-1} & \multicolumn{3}{c}{ROUGE-2} &
-%\multicolumn{3}{c}{ROUGE-3} \\ & Recall & Prec. & F-1 & Recall & Prec. & F-1 &
-%Recall & Prec. & F-1 \\ \hline $k$-medoids & 0.127 & 0.414 & 0.181 & 0.025 &
-%0.076 & 0.035 & 0.003 & 0.010 & 0.005\\ \hline AP & 0.117 & 0.440 & 0.173 &
-%0.022 & 0.082 & 0.033 & 0.004 & 0.015 & 0.006\\ \hline \end{tabular}
-%\end{center} \caption{ROUGE scores for $k$-medoids and affinity propagation
-%methods} \end{table*}
-%
-%
-%
-%
-%\section{Experiments}
-%
-%We carried out a small set of initial experiments on one event.  We collected
-%a subset of pages from the TREC Stream Corpus that were relevant to a 2012
-%earthquake off the coast of Guatemala, and further subdivided this collection
-%by the hour they were created. For each hour we generated a summary using the
-%AP clustering algorithm. 
-%
-% We also generated baseline summaries using the $k$-medoids (using the
-% Partitioning Around Medoids algorithm), setting $k = |\mathcal{E}|$ , i.e.
-% the number of exemplar sentences returned by the AP. Because $k$-medoids
-% begins with a random initialization, we took the best (minimum average
-% distance) result of 100 restarts.
-%
-%Table 1 shows example output of the AP and $k$-medoids generated summaries.
-%Sentences are ranked by preference score, although preference has no effect on
-%the $k$-medoids algorithm.  Quantitatively, AP exemplar sentences had higher
-%predicted sentence quality scores (preferences) than the cluster medoids.
-%Qualitatively, the AP method appears to select more general details about the
-%earthquake. Looking at the third sentence selection in table 1, we can see
-%that $k$-medoids selects a personal experience that was reported. This is
-%perhaps less newsworthy or reportable compared to the third sentence in in the
-%AP generated summary which reports a notable structure collapse and injuries
-%related to the quake.  We believe AP results in a more readable and
-%informative summary, although we have yet to perform a rigorous human
-%evaluation of the summary output.
-%
-%
-%We evaluated both algorithms with the ROUGE toolkit \cite{lin2004rouge}.
-%N-ROUGE works by calculating the n-gram recall and precision of an
-%automatically generated summary in reference to a model summary.  We created
-%model summaries by taking the gold nugget sentences with timestamps up to and
-%including the current system time as the gold summary for that hour.
-%%\fdcomment{may be interesting to measure track metrics to measure latency,
-%%redundancy, etc.} KM - Ask Chris
-%
-%Table 2 shows average recall, precision, and F-measure for various orders of
-%ROUGE score. AP demonstrated consistently higher precision than our baseline.
-%%scores over $k$-medoids (ROUGE evaluations are usually focused on recall).
-%While not statistically significant, it is difficult to show significance with
-%Rouge using a small test; we hope further tests will confirm this improvement. 
-%%   , AP had higher precision on average.
-%On average, the AP summaries were slightly shorter than the baseline, which
-%would partially explain this difference. It is also possible that our language
-%models are biased toward shorter sentences; we are more likely to have seen a
-%shorter sentence in the language model input.
-%%\fdcomment{is there a theory to why this is happening?}
-%We are currently adapting our summarizer to add updates over time, and
-%maintaining precision will be important to prevent topic drift.  
-%
-%
-%
-%%\begin{table}[h!] \tiny \begin{center} \begin{tabular}{| c p{3.5cm} |} \hline
-%%Preference & Sentence\\ \hline 9.010 & The magnitude-7.5 quake, about 20
-%%miles deep, was centered off the town of Champerico .People fled buildings in
-%%Guatemala City , in Mexico City and in the capital of the Mexican state of
-%%Chiapas , across the border from Guatemala\\ \hline 3.007 & "Things fell in
-%%my kitchen ." Perez said more than 2,000 soldiers were deployed from a base
-%%in San Marcos to help with disaster relief.\\ \hline 3.007 & Ingrid Lopez ,
-%%who went to the hospital with a 72-year-old aunt whose legs was crushed by a
-%%falling wall, said she had waited hours for an X-ray. \\ \hline 1.007 &
-%%Hundreds of people crammed into the hallways of the small town hospital
-%%waiting for medical staff to help out hundreds of injured family members,
-%%some complaining they were not getting care quickly enough. \\ \hline
-%%\end{tabular} \end{center} %  \caption{Example summary using $k$-medoids with
-%%100 random restarts} \end{table}
-%
-%
-%
-
-%\section{Conclusions and Future Work}
-
-%We have developed a summarizer that can generate summaries over time from web
-%crawls on disasters. We show that our method of identifying exemplar sentences
-%for a summary using AP clustering produces  summaries with higher precision
-%    compared to those based on clustering with K-medoids. A key component of
-%    our approach is the prediction of salient information using features based
-%    on location, temporal changes in topic, and two different language models.
-%
-%Currently, we run each hour of summarization independently. In order to avoid
-%repeating information, we would like to incorporate previously chosen
-%exemplars in the preference computation. One possibility would be to
-%down-weight a candidate exemplar's preference based on its similarity to
-%previous exemplars.
-%
-%Secondly, we would like to do more intelligent inference of missing
-%geographical information since not all sentences contain locations. Currently
-%we are using mean values for missing data.
-%
-%Finally, we would like to experiment with non-symmetric similarity matrices,
-%specifically using narrative chains\cite{chambers2009unsupervised}. Under this
-%model $S(i,j)$ would express the likelihood that the events in sentence $j$
-%precede the events in sentence $i$. We hope such a parameterization would
-%promote more causally motivated sentences into exemplar positions, which would
-%better describe the disaster event domain.
-%
-%
-\bibliographystyle{abbrv}
-%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this
-%case
-\bibliography{cites.bib}
-% You must have a proper ".bib" file and remember to run: latex bibtex latex
-% latex to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional \balancecolumns \balancecolumns % GM June 2007 That's
-%all folks!
-\end{document}

### 2015-01-13 paper.tex
formatting and latex errors; added sig-alt
refactoring sections.

@@
 
-%\documentclass{sig-alternate} \usepackage{url} \usepackage{color}
-\documentclass[10pt]{article} \usepackage{url} \usepackage{color}
+\documentclass{sig-alternate} 
+\usepackage{url} 
+\usepackage{color}
+% \documentclass[10pt]{article} \usepackage{url} \usepackage{color}
 \usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
+% \usepackage[margin=1in]{geometry}
 \usepackage{amssymb}
@@
 
-\title{Columbia University at TREC 2014: Temporal Summarization}
-
-\author[1]{Chris Kedzie}
-\author[1]{Kathleen McKeown}
-\author[2]{Fernando Diaz}
-\affil[1]{Columbia University, Department of Computer Science}
-\affil[2]{Microsoft Research}
+\title{The Anatomy of a Temporal Summarization System}
+
+\numberofauthors{1} %  in this sample file, there are a *total*
+\author{}
+% \author[1]{Chris Kedzie}
+% \author[1]{Kathleen McKeown}
+% \author[2]{Fernando Diaz}
+% \affil[1]{Columbia University, Department of Computer Science}
+% \affil[2]{Microsoft Research}
 %    Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
@@
 
-\section{Introduction}
-
-During crises, information is critical for first responders and those caught
-in the event.  When the event is significant, as in the case of Hurricane
-Sandy, the amount of information produced by traditional news outlets,
-government agencies, relief organizations, and social media can vastly
-overwhelm those trying to monitor the situation. Methods for identifying,
-tracking, and summarizing events from text based input have been explored
-extensively  (e.g.,
-\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
-these experiments were not performed in the large and heterogeneous
-environment of the modern web; there is still a need for robust and scalable 
-methods for automatic summarization.
-
-%Previous work on generating event descriptions and/or multi-document
-%summarization has relied on clustering algorithms to find representative
-%sentences appropriate for an event summary.  These methods impose a metric
-%space on the text data that can make it difficult to incorporate external
-%sources of information elegantly -- in this paper we argue that centroid
-%sentences are not a priori the best candidates for inclusion in an event
-%summary.
-
-In this paper, we detail our participation in the second year of the TREC
-Temporal Summarization (TS) track. Broadly, TS systems are required to monitor
-a time ordered corpus for information about a specific event and ``emit'' 
-updates about that event. Under the current TREC formulation of this task,
-successful TS systems must synthesize techniques from information retrieval,
- extractive multi-document summarization, and update summarization.
-Specifically, we must select updates to emit based on three primary criterion:
-relevance, salience, and redundancy.
-
-As this was our first time participating in TS track, we focused mostly on 
-second and third criterion, relying on the TREC supplied resources for 
-document retrieval and filtering heuristics for relevance.
-Our system predicts sentence salience in the context of a
-large-scale event and integrates these predictions into
-a clustering based multi-document summarization system. We fit a Gaussian 
-Process (GP) regression model to predict sentence salience and use 
-these predictions to bias the
-formation of sentence clusters around more salient regions in the input space
-using affinity propagation (AP) clustering.  AP uses the salience predictions
-as well as pairwise similarities among input sentences to identify
-\emph{exemplar} sentences, which we use as our summary output.  Our approach
-differs from other methods of summarization that compute salience by pairwise
-comparisons alone, ignoring features of importance that are intrinsic to the
-sentences themselves.
-
-The remainder of the paper is organized as follows. In section
-~\ref{sec:background} we give an overview of the semantic similarity, 
-Gaussian process, and affinity propagation algorithms that make up the bulk of
-our TS system. 
-Next we describe the data with which we build our various models.
-Then in section~\ref{sec:approach} we give a high level sketch
-of our system, and then explain each component in detail. 
-We finish with the presentation of our results and related works.
+\input{introduction.tex}
+
+\input{relatedwork.tex}
 
@@
 
-\begin{figure}
+\begin{figure*}
+	\begin{center}
 \begin{tabular}{| p{5cm} c c |}
@@
 language models. }
-\end{figure}
+\end{center}
+\end{figure*}
 
@@
 
-\begin{figure}
-\centering
-\begin{algorithm}[H]
+% \begin{figure}
+% \centering
+\begin{algorithm}%[H]
  \KwData{$Query$ --- the set of event query words\\ 
@@
 \end{algorithm}
-\end{figure}
+% \end{figure}
 
@@
 
-\section{Related Work}
-
-A principal concern in extractive multi-document summarization is the
-selection of salient sentences for inclusion in summary output
-\cite{nenkova2012survey}.  This has often been approached as a ranking
-problem.
-%We broadly conceptualize this decision as either an intrinsic or extrinsic
-%sentence evaluation process. Intrinsic approaches evaluate sentences
-%individually, possibly by predicting the impact on summary quality using
-%sentece level features. 
-Sentences have been ranked by the average word probability, average tf-idf
-score, and the number of topically related words (topic-signatures in the
-summarization literature)
-\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
-statistics are easily computable from the input sentences, while the third
-only requires an additional, generic background corpus.  Another ranking
-approach, centroid summarization, involves creating an average bag of words
-(BOW) vector, the centroid, from the input sentences and ranking sentences by
-their similarity to the centroid \cite{radev2004centroid}.  Graph
-\cite{erkan2004lexrank} and clustering
-\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
-based approaches, on the other hand, make use of pair-wise similarity
-comparisons amongst input sentences.  In these models, salient sentences are
-more central to the input or cluster, respectively.
-
-%identify salient regions of the input space while simultaneously coping with
-%redundancy.  Graph-based algorithms have been used to rank sentences
-%Clustering algorithms, e.g., are commonly used to exploit redundancy in
-%input. Input sentences are clustered and summaries are generated by selecting
-%the most representative sentence from each cluster.  Graph-based models have
-%also been used for summarization.  E.g., the LexRank algorithm treats
-%sentences as nodes in a graph, where edges are constructed by way of cosine
-%similarity between sentence nodes; edges are either continuosly weighted by
-%similarity or discrete, existing only when the similarity is above a
-%threshold.  The PageRank algorithm is used on the graph to find the most
-%important sentence nodes. In both clustering and graph-based approaches,
-%sentence salience is largely determined by the pairwise relations between
-%sentences.
-
-Supervised learning has also been applied to this task. Model features are
-usually derived from human generated summaries, and are non-lexical in nature
-(e.g., sentence starting position, number of topic-signatures, number of
-unique words, word frequencies). Seminal work in this area has employed naive
-Bayes and logistic regression classifiers to identify sentences for summary
-inclusion \cite{kupiec1995trainable,conroy2001using}.
-
-%\fdadd{
-Several researchers have recognized the importance of summarization during
-natural disasters.  Guo \textit{et al.} developed a system for detecting
-novel, relevant, and comprehensive sentences immediately after a natural
-disaster \cite{qi:temporal-summarization}.  The method uses a model of
-sentence relevance and novelty in order to select appropriate updates.
-Training data for regression targets is automatically generated from
-retrospective Wikipedia data.  The system is evaluated on news documents
-related to 197 natural and human disasters from 2009 to 2011 using variants of
-Rouge modified to capture novelty, relevance, and comprehensiveness
-\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
-efficiency detect important updates during natural disasters
-\cite{wang:update-summarization}.  The algorithm works by hierarchically
-clustering sentences online, allowing the system to output a more expressive
-narrative structure than Guo \textit{et al.}.  The method is evaluated on
-official press releases related to Hurricane Wilma  in 2005 using Rouge score
-between the system summary and a manually generated target summary.
-%}
-
-%This work uses the TREC Stream Corpus data set that is in use by the TREC
-%Temporal Summarization track \cite{frank2012building,aslam2013trec} .
-%Generally, last year's participants used a pipelined approach to build
-%summaries, generally ranking sentences, filtering out all but the most
-%relevant, and then performing some sort of deduplication/redundancy removal
-%step.  Ranking approaches ranged from simple query word match to more
-%sophisticated query expansion and query based language model scoring
-%\cite{liu2013ictnet,xu2013trec,baruah2013univ}.  Perhaps most similar to our
-%approach, the system of \cite{xu2013trec} uses a weighted combination of
-%features (similarity to query, named entity frequency, predicate frequency,
-%presence of numerical values, sentence novelty, etc.) to score sentences;
-%sentences above a threshold are added to the summary.  Both the weights and
-%the threshold are selected by hand.
-%
-%
-%Our system seeks to combine the best of these approaches, using supervised
-%learning to predict salience rankings, and directly incorporate this
-%information in a clustering algorithm to bias the formation of sentence
-%clusters around highly salient regions.
-
-%Broadly, we conceive of the above
-
-%Another popular graph-based model
-
-%Larger clusters generally indicate more important For example
-
-%\cite{} also explore the use of a hidden Markov model to predict summary
-%inclusion based on previously selected sentences.
-
-%ranking.
-%
-%The field of extractive summarization has long been focused on determing
-%sentence importance or salience.  Many These efforts have largely involved
-%unsupervised, data-driven approaches, where systems rank input sentences
-%using features easily computable from the input, and sometimes an unrelated
-%background corpus.  The most basic sentence rankings are by average word
-%probability and average TF-IDF score, statistics that are computable from the
-%input alone.  One of the more effective techniques, topic-sigantures,
-%involves identifying the most topically related words in the input sentences.
-%This is done using a log-likelihood ratio test ($\chi^2$) to determine which
-%of two hypotheses hold for a word in the input: (1) the probability of the
-%word in the input distribution and the background distribution are identical,
-%or (2) the probability of the word is greater in the input distribution than
-%in the background distribution. If (2) is true with some statistical
-%significance, the word is said to be a topic signature. As with the previous
-%features, a common approach is to rank sentences by the number of
-%topic-signatures they contain.
-%
-%In mulit-document summarization, clustering algorthims are often employed to
-%determine sentence importance, with larger clusters corresponding to more
-%important topics. 
-%
-%Supervised learning approaches have also been explored. Features are usually
-%derived from human generated summaries, and are non-lexical in nature (e.g.,
-%sentence starting position, number of topic-signatures, number of unique
-%words, word frequencies). Seminal work in this area has employed naive Bayes
-%and logistic regression classifiers to identify sentences for summary
-%inclusion. \cite{} also explore the use of a hidden Markov model to predict
-%summary inclusion based on previously selected sentences.
-%
-%If this  
-%
-%tried to rank sentences by their overlap with query words. 
-%
-
-%and models are developed to predict the likelihood of a sentences inclusion
-%in a summary. Naive Bayes, hidden Markov models, and logistic regression to
-%predict a sentence's summary inclusion.
-%
-%Researches have also experimented with supervised learning for summarization,
-%however, machine learning approaches have not yielded significant advances in
-%generic summarization. One difficulty is that gold data is difficult to
-%obtain.
-%
-%
-%In  
-%
-%
-%%which remain competetive to supervised machine-learning techniques.
-%
-%have ranked sentences by
-%
-%
-%their average word probability, average TF-IDF weighting, and/or topic
-%signatures; these representations are all easily computable from the input
-%sentences, and, for topic signatures, an unrelated background corpus.  Topic
-%signatures in
-%
-% 
-%These efforts can be broadly split into two camps: feature based ranking and
-%clustering.
-%
-%
-%
-%Determing sentence importance has been A large portion of automatic
-%summarization research has focused on  
-%
-%
-
-%\newcommand{corpus}{}
-
-%A common approach to automatic summarization is to identify sentences with the
-%highest centrality with respect to the input sentences.  Intuitively,
-%sentences with a high degree of centrality are more semantically related to
-%the entire set of input sentences.  A summary can thus be obtained by
-%returning the $k$ most central sentences.  This generally implies the
-%calculation of pairwise distances between all sentences
-%\cite{radev2004centroid, erkan2004lexrank}.  In these approaches, sentences
-%are evaluated extrinsically by their distance to other sentences, either
-%directly \cite{erkan2004lexrank} or through an aggregate centroid object
-%\cite{radev2004centroid}. The distance between sentences is most commonly the
-%cosine distance of sentence term-vectors, but in general this can be an
-%arbitrary real-valued similarity function.
-%%; the distance function imposes a topology on the space of input sentences.
-%
-%For some domains, it is very likely that we will have additional background
-%knowledge that could be predictive of sentence salience for the event being
-%summarized.  For example, certain kinds of information extracted from the
-%sentence text (e.g., temporal or geographic proximity) can indicate relevance
-%to a given event.  It would be difficult to incorporate this kind of salience
-%into the measure of centrality.
-%
-%
-%For example, consider a cluster of three sentence vectors $s_1 = (1,1,0)$,
-%$s_2 = (1,1,1)$, and $s_3 = (0,1,1)$. Without any other information, $s_2$ has
-%the highest degree of centrality, i.e. it has the highest average cosine
-%similarity and smallest average Euclidean distance to the other sentences.
-%Now, if we believe that $s_1$ is $\alpha$ times more salient than the other
-%sentences, we cannot simply scale $s_1$ by $\alpha$--the average cosine
-%similarity will remain unchanged, since the vector magnitude does not affect
-%the angle and $s_1$ will have an even greater the average Euclidean distance
-%from the rest.  Worse still, $s_2$ will still be the most representative
-%sentence of the three.
-%
-%
-
-%In order to predict the salience score for a sentence,
-
- %\cite{guo2012simple} produces a real-valued score of the similarity We use
- %\cite{guo2012simple} produces real valued similarity scores s.
- %\cite{guo2012simple} determines the similarity of two sentences by deriving
- %low dimensional We have As a proxy measure for sentence salience, we use the
- %maximum similarity of an input sentence to a human generated summary
- %sentence, using the state-of-the-art sentence similarity module of
- %\cite{guo2012simple}. \cite{guo2012simple} derive a lower-dimensional vector
- %of latent features for a sentence. The similarity of two sentences is then
- %the cosine similarity of the latent vectors.  In order to predict this
- %similarity we extract a variety of features including language model scores,
- %geographic relevance, and temporal relevance.  We learn a predictive
- %function of sentence-summary similarity by fitting a Gaussian process model
- %on these features. \cite{preotiuc2013temporal}
-
-%As training data for this model, we take a portion of the sentences found in
-%the relevant documents and find their highest similarity to the gold nugget
-%information seen thus far.
-
-%When assigning preference scores
-
-%After obtaining similarity scores for sentences in our training set, we fit a
-%Gaussian process model to learn
-
-%Our model
-
-%Correllating automatic measures of sentence salience with human judgments is
-%a difficult task. , we use the maximum cosine similarity to
-
-%For example, if we scaled the BOW vector based on our belief in its salience,
-%it would not effect the cosine distance since the angle between two vectors
-%is not changed by the magnitude of either
-
-%Simply scaling the sentence vector based on our belief in it's fitness does
-%not with our prior beliefs-- When using the BOWs representation, the input
-%space is lexically defined, but our salience metric may not be based on any
-%specific lexical items.
-
-%These approaches make it difficult to leverage prior knowledge to guide the
-%selection of salient sentences. Sentences
-
-%
-%Centroid based algorthims are commonly employed in the summarization
-%literature to compress a large number of sentences into a much smaller,
-%representative set. 
-%
-%The $k$-medoids clustering algorithm is commonly used to partition input
-%sentences into topically distinct clusters, where the most representative
-%sentences (the cluster medoids) are taken to build the summary. The quality
-%of the summary in this framework is dependent on the method of sentence
-%representation and the ability of a similarity or distance metric to
-%adequately capture pairwise comparisons between sentences. Previous work in
-%event clustering has examined the use of different term weightings (tf-idf,
-%named-entities)  and lower dimensional term representations (LSA, LDA) for
-%use in summarization.
-%
-%
-%A popular graph-based alternative, LexRank, treats each sentence as a node in
-%a graph, and identifies the nodes that are most central by in-degree.  The
-%existence of an edge between two sentences is determined by a similarity
-%function (either by setting a threshold above which the edge exists, or by
-%taking the similarity as a continuous edge weight). The sentences with the
-%highest eigenvector centrality can be found using random walks (PageRank).
-%The $k$ most central sentences are used as the summary.
-%
-%One drawback of these algorithms is that they do not provide an elegant way
-%to incorporate prior information about sentence salience.
-
-%Move this out!
-%In our approach, the system generates clusters using an affinity propagation
-%algorithm and from each cluster an exemplar sentence is selected that is added
-%to the summary.  In the following sections, we show how  prior information
-%representing salience can easily be incorporated into the affinity propagation
-%algorithm.  We believe the incorporation of salience to be useful in noisy
-%environments (e.g., a web crawl), and that it can help the formation of
-%clusters around the most relevant inputs. Our current system  is trained using
-%features derived from location, changes in wording across time and language
-%models that characterize the language of disaster to generate summaries at
-%regular intervals across time.  As we develop the system further, we will
-%extend it to generate updates across time,  penalizing the salience of
-%concepts already selected by the summarizer to encourage the discovery of
-%novel sentences as the event unfolds. 
-%
-
-%\begin{table*}[ht] \begin{center} \begin{tabular}{| c p{7.0cm} || c p{7.0cm}
-%|} \hline Preference & AP Clustering & Preference & $k$-Medoids Clustering\\
-%\hline 9.010 & The magnitude-7.5 quake, about 20 miles deep, was centered off
-%the town of Champerico .People fled buildings in Guatemala City , in Mexico
-%City and in the capital of the Mexican state of Chiapas , across the border
-%from Guatemala & 9.010 & The magnitude-7.5 quake, about 20 miles deep, was
-%centered off the town of Champerico .People fled buildings in Guatemala City ,
-%in Mexico City and in the capital of the Mexican state of Chiapas , across the
-%border from Guatemala   \\ \hline 9.010 & A reporter in the town of San Marcos
-%, about 80 miles north of the epicenter, told local radio station Emisoras
-%Unidas that houses had collapsed onto residents and smashed televisions and
-%other appliances had been scattered into the streets.  & 3.007 & ``Things fell
-%in my kitchen .'' Perez said more than 2,000 soldiers were deployed from a
-%base in San Marcos to help with disaster relief.\\ \hline 9.007 & The local
-%fire department said on its Twitter account that a school had collapsed and
-%eight injured people had been taken to a nearby hospital.  & 3.007 &  Ingrid
-%Lopez , who went to the hospital with a 72-year-old aunt whose legs was
-%crushed by a falling wall, said she had waited hours for an X-ray.\\ \hline
-%7.008 & There are three confirmed dead and many missing after the strongest
-%earthquake to hit Guatemala since a deadly 1976 quake that killed 23,000.  &
-%1.007 & Hundreds of people crammed into the hallways of the small town
-%hospital waiting for medical staff to help out hundreds of injured family
-%members, some complaining they were not getting care quickly enough. \\ \hline
-%\end{tabular} \end{center} \caption{Example summary using affinity propagation
-%(left) and $k$-medoids (right)} \end{table*} \begin{table*}[t] \begin{center}
-%\begin{tabular}{| c || c | c | c || c | c | c || c | c | c |} \hline Method &
-%\multicolumn{3}{c}{ROUGE-1} & \multicolumn{3}{c}{ROUGE-2} &
-%\multicolumn{3}{c}{ROUGE-3} \\ & Recall & Prec. & F-1 & Recall & Prec. & F-1 &
-%Recall & Prec. & F-1 \\ \hline $k$-medoids & 0.127 & 0.414 & 0.181 & 0.025 &
-%0.076 & 0.035 & 0.003 & 0.010 & 0.005\\ \hline AP & 0.117 & 0.440 & 0.173 &
-%0.022 & 0.082 & 0.033 & 0.004 & 0.015 & 0.006\\ \hline \end{tabular}
-%\end{center} \caption{ROUGE scores for $k$-medoids and affinity propagation
-%methods} \end{table*}
-%
-%
-%
-%
-%\section{Experiments}
-%
-%We carried out a small set of initial experiments on one event.  We collected
-%a subset of pages from the TREC Stream Corpus that were relevant to a 2012
-%earthquake off the coast of Guatemala, and further subdivided this collection
-%by the hour they were created. For each hour we generated a summary using the
-%AP clustering algorithm. 
-%
-% We also generated baseline summaries using the $k$-medoids (using the
-% Partitioning Around Medoids algorithm), setting $k = |\mathcal{E}|$ , i.e.
-% the number of exemplar sentences returned by the AP. Because $k$-medoids
-% begins with a random initialization, we took the best (minimum average
-% distance) result of 100 restarts.
-%
-%Table 1 shows example output of the AP and $k$-medoids generated summaries.
-%Sentences are ranked by preference score, although preference has no effect on
-%the $k$-medoids algorithm.  Quantitatively, AP exemplar sentences had higher
-%predicted sentence quality scores (preferences) than the cluster medoids.
-%Qualitatively, the AP method appears to select more general details about the
-%earthquake. Looking at the third sentence selection in table 1, we can see
-%that $k$-medoids selects a personal experience that was reported. This is
-%perhaps less newsworthy or reportable compared to the third sentence in in the
-%AP generated summary which reports a notable structure collapse and injuries
-%related to the quake.  We believe AP results in a more readable and
-%informative summary, although we have yet to perform a rigorous human
-%evaluation of the summary output.
-%
-%
-%We evaluated both algorithms with the ROUGE toolkit \cite{lin2004rouge}.
-%N-ROUGE works by calculating the n-gram recall and precision of an
-%automatically generated summary in reference to a model summary.  We created
-%model summaries by taking the gold nugget sentences with timestamps up to and
-%including the current system time as the gold summary for that hour.
-%%\fdcomment{may be interesting to measure track metrics to measure latency,
-%%redundancy, etc.} KM - Ask Chris
-%
-%Table 2 shows average recall, precision, and F-measure for various orders of
-%ROUGE score. AP demonstrated consistently higher precision than our baseline.
-%%scores over $k$-medoids (ROUGE evaluations are usually focused on recall).
-%While not statistically significant, it is difficult to show significance with
-%Rouge using a small test; we hope further tests will confirm this improvement. 
-%%   , AP had higher precision on average.
-%On average, the AP summaries were slightly shorter than the baseline, which
-%would partially explain this difference. It is also possible that our language
-%models are biased toward shorter sentences; we are more likely to have seen a
-%shorter sentence in the language model input.
-%%\fdcomment{is there a theory to why this is happening?}
-%We are currently adapting our summarizer to add updates over time, and
-%maintaining precision will be important to prevent topic drift.  
-%
-%
-%
-%%\begin{table}[h!] \tiny \begin{center} \begin{tabular}{| c p{3.5cm} |} \hline
-%%Preference & Sentence\\ \hline 9.010 & The magnitude-7.5 quake, about 20
-%%miles deep, was centered off the town of Champerico .People fled buildings in
-%%Guatemala City , in Mexico City and in the capital of the Mexican state of
-%%Chiapas , across the border from Guatemala\\ \hline 3.007 & "Things fell in
-%%my kitchen ." Perez said more than 2,000 soldiers were deployed from a base
-%%in San Marcos to help with disaster relief.\\ \hline 3.007 & Ingrid Lopez ,
-%%who went to the hospital with a 72-year-old aunt whose legs was crushed by a
-%%falling wall, said she had waited hours for an X-ray. \\ \hline 1.007 &
-%%Hundreds of people crammed into the hallways of the small town hospital
-%%waiting for medical staff to help out hundreds of injured family members,
-%%some complaining they were not getting care quickly enough. \\ \hline
-%%\end{tabular} \end{center} %  \caption{Example summary using $k$-medoids with
-%%100 random restarts} \end{table}
-%
-%
-%
-
-%\section{Conclusions and Future Work}
-
-%We have developed a summarizer that can generate summaries over time from web
-%crawls on disasters. We show that our method of identifying exemplar sentences
-%for a summary using AP clustering produces  summaries with higher precision
-%    compared to those based on clustering with K-medoids. A key component of
-%    our approach is the prediction of salient information using features based
-%    on location, temporal changes in topic, and two different language models.
-%
-%Currently, we run each hour of summarization independently. In order to avoid
-%repeating information, we would like to incorporate previously chosen
-%exemplars in the preference computation. One possibility would be to
-%down-weight a candidate exemplar's preference based on its similarity to
-%previous exemplars.
-%
-%Secondly, we would like to do more intelligent inference of missing
-%geographical information since not all sentences contain locations. Currently
-%we are using mean values for missing data.
-%
-%Finally, we would like to experiment with non-symmetric similarity matrices,
-%specifically using narrative chains\cite{chambers2009unsupervised}. Under this
-%model $S(i,j)$ would express the likelihood that the events in sentence $j$
-%precede the events in sentence $i$. We hope such a parameterization would
-%promote more causally motivated sentences into exemplar positions, which would
-%better describe the disaster event domain.
-%
-%
 \bibliographystyle{abbrv}

### 2015-01-13 relatedwork.tex
refactoring sections.

\section{Related Work}
\label{sec:relatedwork}
A principal concern in extractive multi-document summarization is the
selection of salient sentences for inclusion in summary output
\cite{nenkova2012survey}.  This has often been approached as a ranking
problem.
%We broadly conceptualize this decision as either an intrinsic or extrinsic
%sentence evaluation process. Intrinsic approaches evaluate sentences
%individually, possibly by predicting the impact on summary quality using
%sentece level features. 
Sentences have been ranked by the average word probability, average tf-idf
score, and the number of topically related words (topic-signatures in the
summarization literature)
\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
statistics are easily computable from the input sentences, while the third
only requires an additional, generic background corpus.  Another ranking
approach, centroid summarization, involves creating an average bag of words
(BOW) vector, the centroid, from the input sentences and ranking sentences by
their similarity to the centroid \cite{radev2004centroid}.  Graph
\cite{erkan2004lexrank} and clustering
\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
based approaches, on the other hand, make use of pair-wise similarity
comparisons amongst input sentences.  In these models, salient sentences are
more central to the input or cluster, respectively.

%identify salient regions of the input space while simultaneously coping with
%redundancy.  Graph-based algorithms have been used to rank sentences
%Clustering algorithms, e.g., are commonly used to exploit redundancy in
%input. Input sentences are clustered and summaries are generated by selecting
%the most representative sentence from each cluster.  Graph-based models have
%also been used for summarization.  E.g., the LexRank algorithm treats
%sentences as nodes in a graph, where edges are constructed by way of cosine
%similarity between sentence nodes; edges are either continuosly weighted by
%similarity or discrete, existing only when the similarity is above a
%threshold.  The PageRank algorithm is used on the graph to find the most
%important sentence nodes. In both clustering and graph-based approaches,
%sentence salience is largely determined by the pairwise relations between
%sentences.

Supervised learning has also been applied to this task. Model features are
usually derived from human generated summaries, and are non-lexical in nature
(e.g., sentence starting position, number of topic-signatures, number of
unique words, word frequencies). Seminal work in this area has employed naive
Bayes and logistic regression classifiers to identify sentences for summary
inclusion \cite{kupiec1995trainable,conroy2001using}.

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary.

### 2015-01-15 algorithms.tex
Split out methods, algos, and motivation; added a temporary chart

--- 2015-01-13 paper.tex
@@
-% This is "sig-alternate.tex" V2.0 May 2012 This file should be compiled with
-% V2.5 of "sig-alternate.cls" May 2012
-%
-% This example file demonstrates the use of the 'sig-alternate.cls' V2.5
-% LaTeX2e document class file. It is for those submitting articles to ACM
-% Conference Proceedings WHO DO NOT WISH TO STRICTLY ADHERE TO THE SIGS
-% (PUBS-BOARD-ENDORSED) STYLE.  The 'sig-alternate.cls' file will produce a
-% similar-looking, albeit, 'tighter' paper resulting in, invariably, fewer
-% pages.
-%
-% ----------------------------------------------------------------------------------------------------------------
-% This .tex file (and associated .cls V2.5) produces: 1) The Permission
-% Statement 2) The Conference (location) Info information 3) The Copyright
-% Line with ACM data 4) NO page numbers
-%
-% as against the acm_proc_article-sp.cls file which DOES NOT produce 1) thru'
-% 3) above.
-%
-% Using 'sig-alternate.cls' you have control, however, from within the source
-% .tex file, over both the CopyrightYear (defaulted to 200X) and the ACM
-% Copyright Data (defaulted to X-XXXXX-XX-X/XX/XX).  e.g.
-% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
-% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the
-% copyright line.
-%
-% ---------------------------------------------------------------------------------------------------------------
-% This .tex source is an example which *does* use the .bib file (from which
-% the .bbl file % is produced).  REMEMBER HOWEVER: After having produced the
-% .bbl file, and prior to final submission, you *NEED* to 'insert' your .bbl
-% file into your source .tex file so as to provide ONE 'self-contained' source
-% file.
-%
-% ================= IF YOU HAVE QUESTIONS ======================= Questions
-% regarding the SIGS styles, SIGS policies and procedures, Conferences etc.
-% should be sent to Adrienne Griscti (griscti@acm.org)
-%
-% Technical questions _only_ to Gerald Murray (murray@hq.acm.org)
-% ===============================================================
-%
-% For tracking purposes - this is V2.0 - May 2012
-
-\documentclass{sig-alternate} 
-\usepackage{url} 
-\usepackage{color}
-% \documentclass[10pt]{article} \usepackage{url} \usepackage{color}
-\usepackage[utf8]{inputenc}
-% \usepackage[margin=1in]{geometry}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage[]{algorithm2e}
-\DeclareMathOperator{\corpus}{\mathcal{C}}
-\DeclareMathOperator{\doc}{\mathnormal{d}}
-\DeclareMathOperator{\sent}{\mathnormal{s}}
-\DeclareMathOperator{\order}{\pi}
-\DeclareMathOperator{\dtime}{\mathnormal{t}}
-\DeclareMathOperator{\hour}{\mathnormal{h}}
-\DeclareMathOperator{\hours}{\mathcal{H}}
-\DeclareMathOperator{\Sim}{\mathbf{K}}
-\DeclareMathOperator{\SMat}{\mathbf{X}}
-\DeclareMathOperator{\Pref}{\boldsymbol{\pi}}
-\DeclareMathOperator{\Exemp}{\mathnormal{Exemplars}}
-\DeclareMathOperator{\Updates}{\mathnormal{Updates}}
-\usepackage{cleveref}
-\crefname{section}{§}{§§}
-\Crefname{section}{§}{§§}
-
-\usepackage{authblk}
-\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
-\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
-
-\begin{document}
-
-%
-% --- Author Metadata here ---
-%\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
-%- IF NEED BE.  \crdata{0-12345-67-8/90/01}  % Allows default copyright data
-%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.  --- End of Author
-%Metadata ---
-
-\title{The Anatomy of a Temporal Summarization System}
-
-\numberofauthors{1} %  in this sample file, there are a *total*
-\author{}
-% \author[1]{Chris Kedzie}
-% \author[1]{Kathleen McKeown}
-% \author[2]{Fernando Diaz}
-% \affil[1]{Columbia University, Department of Computer Science}
-% \affil[2]{Microsoft Research}
-%    Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
-%Format\titlenote{(Produces the permission block, and copyright information).
-%For use with SIG-ALTERNATE.CLS. Supported by ACM.}} \subtitle{[Extended
-%Abstract] \titlenote{A full version of this paper is available as
-%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-%\LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}
-%
-% You need the command \numberofauthors to handle the 'placement and
-% alignment' of the authors beneath the title.
-%
-% For aesthetic reasons, we recommend 'three authors at a time' i.e. three
-% 'name/affiliation blocks' be placed beneath the title.
-%
-% NOTE: You are NOT restricted in how many 'rows' of "name/affiliations" may
-% appear. We just ask that you restrict the number of 'columns' to three.
-%
-% Because of the available 'opening page real-estate' we ask you to refrain
-% from putting more than six authors (two rows with three columns) beneath the
-% article title.  More than six makes the first-page appear very cluttered
-% indeed.
-%
-% Use the \alignauthor commands to handle the names and affiliations for an
-% 'aesthetic maximum' of six authors.  Add names, affiliations, addresses for
-% the seventh etc. author(s) as the argument for the \additionalauthors
-% command.  These 'additional authors' will be output/set for you without
-% further effort on your part as the last section in the body of your article
-% BEFORE References or any Appendices.
-
-%\numberofauthors{3} %  in this sample file, there are a *total*
-% of EIGHT authors. SIX appear on the 'first-page' (for formatting reasons)
-% and the remaining two appear in the \additionalauthors section.
-%
-%\author{
-% You can go ahead and credit any number of authors here, e.g. one 'row of
-% three' or two rows (consisting of one row of three and a second row of one,
-% two or three).
-%
-% The command \alignauthor (no curly braces needed) should precede each author
-% name, affiliation/snail-mail address and e-mail address. Additionally, tag
-% each line of affiliation/address with \affaddr, and tag the e-mail address
-% with \email.
-%
-% 1st. author
-
-%\alignauthor Chris Kedzie\\ \affaddr{Columbia University}\\
-%\affaddr{Department of Computer Science} \email{kedzie@cs.columbia.edu}
-
-% 2nd. author
-%\alignauthor Kathleen McKeown\\ \affaddr{Columbia University}\\
-%\affaddr{Department of Computer Science}\\ \email{kathy@cs.columbia.edu}
-
-% 3rd. author
-%\alignauthor Fernando Diaz\\ \affaddr{Microsoft Research}\\
-%\email{fdiaz@microsoft.com}
-
-% There's nothing stopping you putting the seventh, eighth, etc.  author on
-% the opening page (as the 'third row') but we ask, for aesthetic reasons that
-% you place these 'additional authors' in the \additional authors block, viz.
-%} 
-\date{29 October 2014}
-% Just remember to make sure that the TOTAL number of authors is the number
-% that will appear on the first page PLUS the number that will appear in the
-% \additionalauthors section.
-
-\maketitle \begin{abstract} 
-Information need is highest during times of crisis; both officials and 
-civilians need effective access to make informed decisions about their 
-safety and the safety of others. However, these situations pose many 
-interesting challenges for the IR systems that support these decisions.
-Specifically, we address event monitoring and updating, a.k.a. the
-temporal summarization task (TS). Given an event query, we monitor streams
-of online news and select important text to present to the user. Updates
-are selected by predicting the salience of the text w.r.t. to the event.
-Our model utilizes a variety of features at the query, document, and corpus
-level. Additionally, these predictions are used to directly bias a clustering
-algorithm for sentence selection, 
-increasing the novelty of the updates. We evaluate our system
-on a varied set of retrospective events using a combination of human 
-judgements
-and automatic evaluation measures, as compared to a baseline system.
-We demonstrate the effect of the various feature groups, and the importance
-of salience prediction w.r.t. update precision when compared to a clustering
-only baseline.
-
-\end{abstract}
-
-% A category with the (minimum) three required fields
-%\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
-%performance measures]
-
-%\terms{Summarization}
-
-%\keywords{Extractive Summarization, Affinity Propagation}
-
-\input{introduction.tex}
-
-\input{relatedwork.tex}
-
-\section{Background}\label{sec:background}
+\section{Algorithms}\label{sec:background}
 
@@
 
-\section{Data}\label{sec:data}
-
-Our primary dataset is the TREC 2014 Stream Corpus, a ?tb corpus of hourly 
-web crawls from October 2011
-through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
-As per the TS specifications, all summary sentences come from processing this
-corpus in a time aligned manner. Additionally, we use the query terms 
-provided by the track organizers for document retrieval.
-
-For our semantic similarity component (\cref{subsec:semsim}),
-domain specific term-sentence matrices were built using subsections of
-Wikipedia. For each event type we identified one or more related Wikipedia
-categories and collected all pages that were listed under that category.
-For all pages collected, we used the latest possible revision date before the
-start of the Stream Corpus.
-
-We use the same Wikipedia corpora to construct the lanuage models used
-in our salience prediction model (\cref{subsubsec:lm}).
-
-\begin{figure*}
-	\begin{center}
-\begin{tabular}{| p{5cm} c c |}
-\hline
-Event & WP Categories & No. Docs/Sents/Words\\
-\hline \hline
-Boston\_Marathon\_bombings \newline
-Christopher\_Dorner\_shootings\_and\_manhunt \newline 
- In\_Amenas\_hostage\_crisis
-& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
-\hline
- Cyclone\_Oswald \newline
-Early\_2012\_European\_cold\_wave \newline
-February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
-\hline
-Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
-\hline
-Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
-\hline
-Port\_Said\_Stadium\_riot \newline
-2012\_Afghanistan\_Quran\_burning\_protests \newline
-2011-13\_Russian\_protests \newline
-2012\_Romanian\_protests \newline
-2012-13\_Egyptian\_protests \newline
-2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
-2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
-\hline
-\end{tabular}
-\caption{Domain specific Wikipedia corpora for semantic similarity and 
-language models. }
-\end{center}
-\end{figure*}
-
-We use the TREC 2013 Stream Corpus and TS 2013 event/gold nugget data to 
-train our salience models. See section~\cref{subsec:Predict} 
-for a detailed description of our sampling procedure to generate labeled
-data.
-
-%Our documents for summarization come from the online news portion of the TREC
-%Stream Corpus, a 6.45tb corpus obtained by hourly web crawls from October 2011
-%through mid February 2013
-%
-%Summary events come from the TREC Temporal Summarization track, and include
-%natural disasters like Hurricane Sandy as well as man-made events like a 2012
-%train accident in Buenos
-%Aires.\footnote{\url{http://trec.nist.gov/data/tempsumm2013.html}} The track
-%organizers also provide a search query for each event \cite{aslam2013trec}.
-%For each event, we collect the documents that contain all query words and
-%stratify them by the hour they were collected. 
-%
-%For evaluation purposes, the track organizers also provided gold nugget
-%information (i.e. important pieces of information, usually the length of a
-%short clause or sentence). These gold nuggets come from the event's related
-%Wikipedia article and also include the timestamp of when they were added to
-%the page.
-%
-%%KM _ Ask Chris: is this true?
-%To create hourly gold summaries to evaluate our system, we simply take the set
-%of gold nugget information from the start of the event up to the current hour. 
-%
-
-\section{Approach}\label{sec:approach}
+\subsection{Temporal Summarization System}\label{sec:approach}
 
@@
 
-\subsection{Document Retrieval}\label{subsec:Document Retrieval}
+\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
 
@@
 
-\subsection{Content Detection}\label{subsec:Content Detection}
+\subsubsection{Content Detection}\label{subsec:Content Detection}
 
@@
 
-\subsection{Predicting Sentence Salience}\label{subsec:Predict}
+\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
 
@@
 These include sentence length, the number of capitalized words normalized by
-sentence length, and the number of query words present in the sentence.  Query
-words include the event's type (e.g., \emph{earthquake}) and are expanded with
-the event type's WordNet \cite{miller1995wordnet} synset, hypernyms, and
-hyponyms.  For \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
+sentence length, document position, number of named entities, etc.
+
+\subsubsection{Query Features}
+
+These features are derived from the event query and the event type.
+These include the number of query words present in the sentence
+in addition to the number of event type synonyms, hypernyms, and hyponyms
+as found in WordNet \cite{miller1995wordnet}.
+E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
 ``seism,'' ``aftershock,'' etc.
@@
 
-\subsection{Sentence Selection}\label{subsec:SentenceSelection}
+\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
 
@@
 retained for penalty calculations in subsequent time steps.
-
-\section{Run Submissions}
-
-We submitted three different runs for official evaluation. The first system,
-(AP+) used AP clustering without salience predictions or redundancy penalty.
-The clustering was run every hour, and all non-singleton exemplars were 
-taken as the updates.
-
-Our second submission (AP+Sal+) uses our salience prediction model, with the
-AP clustering to select sentences. We do not penalize for redundancy.
-
-Our final run (AP+Sal+Red-) is the same as the previous run but the 
-redundancy penalty is applied.
-
-\section{Results}
-
-Overall it appears that (AP+Sal+) is our best performing model of the three
-submissions. One of the notable differences is in the average number of 
-updates per event. (AP+Sal+) was by far the most terse system, with 
-383.2 updates/event compared to 5997.8 and 1070.7 for (AP+) and (AP+Sal+Red-)
-respectively.
-Despite not returning many updates, $\mathbb{E}$[Latency Gain] was high enough
-on average to outperform on the F1 score compared to the other two systems.
-The differences in F1 from the baseline run for both (AP+Sal+) and 
-(AP+Sal+Red-) are statistically significant; 
-this is empirical validation of our salience modeling efforts.
-
-\begin{figure}
-\centering
-\begin{tabular}{| c | c | c | c |}
-\hline
-\textbf{Run} & \textbf{nE[Lat. Gain]} & \textbf{Lat. Comp} & \textbf{F1} \\
-\hline
-AP+ & 0.0222  & 0.5777 & 0.0403\\
-\hline
-AP+Sal+ & 0.0751 &  0.4139 & $\mathbf{0.1162}$\\
-\hline
-AP+Sal+Red- & 0.0375 & 0.3413 & 0.0602\\
-\hline
-\end{tabular}
-\caption{Submission results, averaged over 2014 TREC TS events.}
-\end{figure}
-
-Unfortunately, there is still much work to be done with our redundancy 
-component. (AP+Sal+Red-) returned an order of magnitude more updates on 
-average than (AP+Sal+) but still achieved comparatively lower comprehensiveness
-scores. The redundancy penalty is perhaps too aggressive, driving the 
-clustering to focus on very novel input spaces---so novel as to be 
-irrelevant.
-
-A related issue with our current model is that our salience predictions
-are relative from the point of view of the clustering algorithm. If all inputs
-have low salience predictions, the AP clustering will still find exemplars, 
-albeit from the most salient inputs. This becomes especially problematic 
-with the redundancy component, as with every timestep, we are evaluating 
-riskier and riskier sets of more novel sentences, and being forced to return
-something. This behavior can be observed especially in the updates from the
-Chelyabinsk meteor event, where our document retrieval stage was 
-under-performing and the predicted salience of all updates was quite low.
-This would suggest implementing some minimum threshold of salience. We
-initially experimented with such a threshold but found tuning it across 
-events of different magnitudes to be difficult.
-
-\bibliographystyle{abbrv}
-%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this
-%case
-\bibliography{cites.bib}
-% You must have a proper ".bib" file and remember to run: latex bibtex latex
-% latex to resolve all references
-%
-% ACM needs 'a single self-contained file'!
-%
-%APPENDICES are optional \balancecolumns \balancecolumns % GM June 2007 That's
-%all folks!
-\end{document}

### 2015-01-15 methods.tex
Split out methods, algos, and motivation; added a temporary chart

\section{Methods}

\subsection{Data}\label{sec:data}

Our primary datasets are the TREC 2013/2014 Stream Corpus, 
a ?tb corpus of hourly 
web crawls from October 2011
through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
As per the TS specifications, all summary sentences come from processing this
corpus in a time aligned manner. While the 2014 corpus encompasses the 
2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
year because sentence ID's used in evaluation changed from year to year.

For our semantic similarity component (\cref{subsec:semsim}),
domain specific term-sentence matrices were built using subsections of
Wikipedia. For each event type we identified one or more related Wikipedia
categories and collected all pages that were listed under that category.
For all pages collected, we used the latest possible revision date before the
start of the Stream Corpus.

We use the same Wikipedia corpora to construct the lanuage models used
in our salience prediction model (\cref{subsubsec:lm}).

\begin{figure*}
	\begin{center}
\begin{tabular}{| p{5cm} c c |}
\hline
Event & WP Categories & No. Docs/Sents/Words\\
\hline \hline
Boston\_Marathon\_bombings \newline
Christopher\_Dorner\_shootings\_and\_manhunt \newline 
 In\_Amenas\_hostage\_crisis
& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
\hline
 Cyclone\_Oswald \newline
Early\_2012\_European\_cold\_wave \newline
February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
\hline
Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
\hline
Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
\hline
Port\_Said\_Stadium\_riot \newline
2012\_Afghanistan\_Quran\_burning\_protests \newline
2011-13\_Russian\_protests \newline
2012\_Romanian\_protests \newline
2012-13\_Egyptian\_protests \newline
2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
\hline
\end{tabular}
\caption{Domain specific Wikipedia corpora for semantic similarity and 
language models. }
\end{center}
\end{figure*}

\subsubsection{Events}

We use the event set developed for the TREC Temporal Summarization track 
for years 2013 and 2014. There were 10 and 15 events respectively. We omit
one event from the 2013 track because there were no news documents in the 
corpus during the duration of the event, yielding a final event set of 24 
events. Meta-data provided for each event includes the name of the event
as it appears on Wikipedia, a short text query for the event, 
the type of event, and the 
duration of the event (the duration for the system 
to monitor not necessarily the complete timespan of the event). 
All data except the Wikipedia name is exposed to the TS system to use.

\subsubsection{Nuggets}
For evaluation purposes, important pieces of information, hereafter 
referred to as nuggets, were extracted from each event's Wikipedia page.
Assessors graded nuggets by importance, and were also able to timestamp their
existance using the timestamp of the page's revision history. Nuggets are 
usually a short to medium length clause containing an atomic piece of
information.

\subsubsection{Matches}
TREC assessors provided judgements on a subset of participant updates each year
of the track. Judged updates are matched to one or more nuggets (a real 
sentence often contains more than one atomic piece of information) or marked
not relevant. There was not much overlap in the participant update sets so
overall coverage of relevant sentences in the corpus is small (percent?).
To expand the judgement corpora, we include all system updates as positive
nugget matches if they are $< .2$ close in Levenshtein distance to a human
assessed matched update.

\subsection{Experiment Design}

\subsubsection{Salience Models}
For each event (both in the complete system and feature ablation experiments)
we randomly sample 1000 sentences after relevant document filtering and 
over the duration of the event 
(as determined by 
TREC; avg. run length is ? days) of the event. Nugget similarities
are computed for each sentence and are standardized with respect to 
the sample mean and variance. After standardization, the maximum nugget similarities are computed for each sentence. 
We fit a Gaussian process for each event, using the maximum similarities 
as our regression target, yielding 24 models in total. Kernel lengthscale
parameters are also individually fit for each model.
When making event-specific predictions, we take the average prediction of the
23 other models.

\subsubsection{Redundancy Threshold}

For every time interval with input, a clustering algorithm is
guarranteed to produce at least one output; 
this can significantly hurt the performance of all clustering based 
temporal summarization systems. In order to avoid this, we set a threshold $t$
such that a new update must have a maximum similarity $< t$ to all previous 
updates.

We determined the value of $t$ by examining the distribution of similarities
of nuggets to their human-matched sentences compared to the distribution
of similarities of nuggets to relevant but non matching sentences. We fit
normal distributions for the matching and non-matching similarities and
used the point of intersection as the value of $t$.
Figure ? shows a normalized histogram of cosine similarities for both 
distributions, the fitted distributions, and the point of intersection.

\begin{figure}
\includegraphics[scale=.40]{match-dist.png} 
\caption{Distribution of nugget similarity for matching and non-matching 
updates.}
\end{figure}
\subsubsection{Effect of Salience Models}

In order to assess the effectiveness of the salience predictions on update
selection, we perform a baseline (AP-base) run using affinity propagation 
clustering
with uniform preference, i.e. all preferences are equal to the median sentence
similarity. This run is compared to our full system (AP-sal), using each
sentence's salience prediction as its preference.

We are also interested
in the effect of directly incorporating salience in the clustering algorithm.
A third run (HAC) is performed using hierarchical agglomerative 
clustering; we produce flattened clusters by setting a maximum cluster 
distance threshold. The sentence with the highest predicted salience from each
 cluster is selected as an update. 
The distance threshold was tuned using grid search. By comparing
HAV with AP-base we can empirically assess the trade off between preference
and neighbor similarity that affinity propagation is making.

\subsubsection{Feature Ablation}

We remove each feature group and train models models on these small faller 
feature susbstets. All other parameters are the same as in our full system 
(AP-sal).

### 2015-01-15 motivation.tex
Split out methods, algos, and motivation; added a temporary chart

\section{Motivation}

Most multi-document summarization approaches involve either sentence 
clustering or ranking or both to select summary sentences.
When the summarization task
is broad or underspecified, clustering is most appropriate in that the data
"speaks for itself." This is probably most true when summarization is 
interpretted to mean exploratory data analysis. 
When there are more constraints on the output, as in many text summarization
tasks,
it is often easier to design appropriate features for the ranking approach, 
either by rule or learning-to-rank. The primary contribution of this paper
is a framework for combining both of these approaches using affinity
propagation clustering. Additionally, we provide a task specific evaluation
of this clustering model on the TREC Temporal Summarization task.

\ckcomment{Not sure how to present this together, we also essentially evaluate
a salience model by feature ablation for the TS task, but this feels very specific compared to the general framework idea.}

### 2015-01-15 paper.tex
Split out methods, algos, and motivation; added a temporary chart

@@
 \Crefname{section}{§}{§§}
-
+\usepackage{graphicx}
 \usepackage{authblk}
@@
 \newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
+\newcommand{\ckcomment}[1]{\textbf{\textcolor{red}{[CK: #1]}}}
 
@@
 
-\section{Background}\label{sec:background}
+\input{motivation.tex}
 
-\subsection{Affinity Propagation}
+\input{algorithms.tex}
 
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\section{Data}\label{sec:data}
-
-Our primary dataset is the TREC 2014 Stream Corpus, a ?tb corpus of hourly 
-web crawls from October 2011
-through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
-As per the TS specifications, all summary sentences come from processing this
-corpus in a time aligned manner. Additionally, we use the query terms 
-provided by the track organizers for document retrieval.
-
-For our semantic similarity component (\cref{subsec:semsim}),
-domain specific term-sentence matrices were built using subsections of
-Wikipedia. For each event type we identified one or more related Wikipedia
-categories and collected all pages that were listed under that category.
-For all pages collected, we used the latest possible revision date before the
-start of the Stream Corpus.
-
-We use the same Wikipedia corpora to construct the lanuage models used
-in our salience prediction model (\cref{subsubsec:lm}).
-
-\begin{figure*}
-	\begin{center}
-\begin{tabular}{| p{5cm} c c |}
-\hline
-Event & WP Categories & No. Docs/Sents/Words\\
-\hline \hline
-Boston\_Marathon\_bombings \newline
-Christopher\_Dorner\_shootings\_and\_manhunt \newline 
- In\_Amenas\_hostage\_crisis
-& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
-\hline
- Cyclone\_Oswald \newline
-Early\_2012\_European\_cold\_wave \newline
-February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
-\hline
-Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
-\hline
-Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
-\hline
-Port\_Said\_Stadium\_riot \newline
-2012\_Afghanistan\_Quran\_burning\_protests \newline
-2011-13\_Russian\_protests \newline
-2012\_Romanian\_protests \newline
-2012-13\_Egyptian\_protests \newline
-2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
-2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
-\hline
-\end{tabular}
-\caption{Domain specific Wikipedia corpora for semantic similarity and 
-language models. }
-\end{center}
-\end{figure*}
-
-We use the TREC 2013 Stream Corpus and TS 2013 event/gold nugget data to 
-train our salience models. See section~\cref{subsec:Predict} 
-for a detailed description of our sampling procedure to generate labeled
-data.
-
-%Our documents for summarization come from the online news portion of the TREC
-%Stream Corpus, a 6.45tb corpus obtained by hourly web crawls from October 2011
-%through mid February 2013
-%
-%Summary events come from the TREC Temporal Summarization track, and include
-%natural disasters like Hurricane Sandy as well as man-made events like a 2012
-%train accident in Buenos
-%Aires.\footnote{\url{http://trec.nist.gov/data/tempsumm2013.html}} The track
-%organizers also provide a search query for each event \cite{aslam2013trec}.
-%For each event, we collect the documents that contain all query words and
-%stratify them by the hour they were collected. 
-%
-%For evaluation purposes, the track organizers also provided gold nugget
-%information (i.e. important pieces of information, usually the length of a
-%short clause or sentence). These gold nuggets come from the event's related
-%Wikipedia article and also include the timestamp of when they were added to
-%the page.
-%
-%%KM _ Ask Chris: is this true?
-%To create hourly gold summaries to evaluate our system, we simply take the set
-%of gold nugget information from the start of the event up to the current hour. 
-%
-
-\section{Approach}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, and the number of query words present in the sentence.  Query
-words include the event's type (e.g., \emph{earthquake}) and are expanded with
-the event type's WordNet \cite{miller1995wordnet} synset, hypernyms, and
-hyponyms.  For \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.
+\input{methods.tex}
 

### 2015-01-19 algorithm/ap.tex
restructuring

--- 2015-01-15 algorithms.tex
@@
-\section{Algorithms}\label{sec:background}
+\subsection{Selecting Sentences}
+\label{sec:ap}
+In the sentence selection stage, we use the salience predictions from our GP
+model as preferences in the AP clustering algorithm. The AP algorithm is 
+parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
+preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
+parameters, and did not perform robustly on our range of inputs.
+In order to improve the quality of the clusters and exemplar selection,
+we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
+The raw preferences are scaled to lie within the  range $(-3, -2)$
 
-\subsection{Affinity Propagation}
+The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
+similarity between input sentences. Self-similarities and similarities below 
+a threshold $\lambda$ were masked and the remaining values scaled to the range
+$(-3, -1)$.
 
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
+\subsubsection{Semantic Similarity}\label{subsec:semsim}
 
@@
 
-\subsection{Gaussian Processes}
+\subsubsection{Affinity Propagation}
+Affinity propagation (AP) is a message passing algorithm that identifies both
+exemplar data points and assignments of each point to an exemplar.  This is
+done iteratively by passing \emph{responsibility} and \emph{availability}
+messages between data points that quantify the fitness of one data point to
+represent another, and the fitness of a data point to be represented based on
+the choices of other data points respectively \cite{dueck2007non}.
 
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
+AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
+preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
+similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
+need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
+our preference that the $i$-th data point can serve as an exemplar a priori of
+other data points.
 
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
+AP has several useful properties that comport well with the TS track 
+requirements. First, the number of clusters $k$ is not a hyper-parameter
+of the model. Since we will be running this algorithm many times in per run,
+the usual methods of hyperparameter search become infeasible 
+(?). The number of clusters falls out of the algorithm
+organically-- lower overall preference values will result in fewer clusters.
 
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
+Secondly, the arbitrary nature of the preferences
+allow us to incorporate a variety of signals for
+identifying the best exemplars, i.e. salience and redundancy signals. 
+The preferences can be thought of as self-similarities (similarities that
+pairwise-comparison based algorithms would ignore) that we can exploit to 
+incorporate our prior beliefs about a data point.
 
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\subsection{Temporal Summarization System}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
+Finally, cluster exemplars are guaranteed to be actual data points. Many 
+clustering algorithms group data around mathematical objects (e.g., the
+mean) that are not necessarily observed in the data. The extractive nature of
+this TS task requires that we emit actual data points (i.e. sentences).
+We are able to sidestep 
+the additional requirement of selecting a most
+representative cluster member, as this is computed explicitly in the AP 
+algorithm.
 
@@
 within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.

### 2015-01-19 algorithm/filtering.tex
restructuring

--- 2015-01-15 algorithms.tex
@@
-\section{Algorithms}\label{sec:background}
+\subsection{Filtering Content}
+\label{sec:filtering}
+We simulate our system using realistic data insofar as the document stream consists of relevant and non-relevant content.
+As with many information retrieval systems, we crude, first phase of high-recall document filtering.
+That is, we would like to remove most of the content which we can easily detect as unlikely to contain useful updates.
 
-\subsection{Affinity Propagation}
+\subsubsection{Filtering Documents}
+\label{sec:filtering:documents}
+As a first filter, we retained only those documents whose raw html content contains at least one keyword from the event's query words.
+We further restricted our document set to only those articles from the news domain.\footnote{This filtered set was distributed by track organizers.}
 
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\subsection{Temporal Summarization System}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
+\subsubsection{Filtering Sentences}
+\label{sec:filtering:sentences}
 In early versions of our summarization system, we found that structural html
@@
 we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.

### 2015-01-19 algorithm/introduction.tex
restructuring

--- 2015-01-15 algorithms.tex
@@
-\section{Algorithms}\label{sec:background}
-
-\subsection{Affinity Propagation}
-
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\subsection{Temporal Summarization System}\label{sec:approach}
-
 We first begin with some notation. For a given event, let $\corpus$ be the set
@@
  Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
+ $\corpus \gets \operatorname{FilterDocuments}(Query, SC)$ (See \cref{sec:filtering:documents})\\
  \For{$i \gets 1,\ldots,t $}{
@@
   \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
+   \For{$\sent \in \operatorname{FilterSentences}(\doc, Query)$ (See \cref{sec:filtering:sentences})\\}{
      $\SMat.\operatorname{append}(\sent)$\;
@@
   $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
+    (\SMat, \Pref)$ (See \cref{sec:ap})\\
   ~\\
@@
 % \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.

### 2015-01-19 algorithm/novelty.tex
restructuring

--- 2015-01-15 algorithms.tex
@@
-\section{Algorithms}\label{sec:background}
-
-\subsection{Affinity Propagation}
-
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
-
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
-
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
-
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\subsection{Temporal Summarization System}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
+\subsection{Ensuring Novelty}
 We run AP clustering with our rescaled and penalized sentence similarities

### 2015-01-19 algorithm/salience.tex
restructuring

--- 2015-01-15 algorithms.tex
@@
-\section{Algorithms}\label{sec:background}
+\subsection{Predicting Sentence Salience}
+\label{subsec:Predict}
 
-\subsection{Affinity Propagation}
+In order to use AP clustering for summarization, we need to assign a
+preference value to each input sentence.  In our approach, we equate a
+sentence's salience with its preference.  A good model of sentence salience
+should predict higher values for sentences that are more likely to appear in a
+human generated summary of the event.
 
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
+We do not have such human judgments, but we do have last year's gold nugget
+sentences.
+The response variable we try to predict is a sentence's semantic similarity
+(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
+to a gold nugget when the gold nugget is not known.
 
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
+To build training data for this regression task, 
+we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
+sentences using the document retrieval and content selection steps outlined
+in the previous selection. We then sample with replacement 200 sentences
+from this collection, and extract non-lexical features (described in more 
+detail below) for each sentence
+to construct our design matrix $\mathbf{X}$. To build our response variables
+$\mathbf{y}$ we compute the maximum semantic similarity of each input
+sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
+an RBF kernel to this data, optimizing kernel parameters with the scaled 
+conjugate gradient method. This sampling procedure is repeated 100 times
+for each of the ? 2013 TS events, yielding ? total models. In our salience 
+predictions for the 2014 events, we take the mean prediction of all models.
 
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
+%we take a subset of sentences
+%relevant to the TREC events (approximately 1000) and match them to the gold
+%nugget sentence with highest similarity as determined by the sentence
+%similarity system of \cite{guo2012simple}. 
+%\cite{} have used this system previously to correlate sentences to meaningful
+%units of information in human generated summaries. 
+%We use the real-valued similarity scores as our salience scores for the
+%training sentences.
 
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
+We want our model to be predictive across different kinds of events so we
+avoid lexical features.  Instead, we extract a variety of features including
+language model scores, geographic relevance, and temporal relevance from each
+sentence.  These features are used to fit a Gaussian process regression model
+that can predict the similarity of a sentence to a gold summary
+\cite{preotiuc2013temporal}.  
+%We use the model predicted salience of each
+%sentence as it's preference value in the AP clustering.
 
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
+\subsubsection{Features}
+\paragraph{Basic Features}
 
-%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
-%the exemplar of the $i^{th}$ point relative to other potential exemplars.
-%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
-%serve as an exemplar of the $i^{th}$ point, taking into account other points'
-%preference for the $k^{th}$ point as an exemplar.
+We employ several basic features that have been used previously in supervised
+models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
+These include sentence length, the number of capitalized words normalized by
+sentence length, document position, number of named entities, etc.
 
-%While, AP does not set the number of exemplars before-hand, a lower overall
-%preference values will result in a smaller number of exemplars.
+\paragraph{Query Features}
 
-%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
-%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
-%real-valued preference $\pi_i$ quantifying our belief a priori of the
-%$i^{th}$ element's ability to serve as an exemplar.
+These features are derived from the event query and the event type.
+These include the number of query words present in the sentence
+in addition to the number of event type synonyms, hypernyms, and hyponyms
+as found in WordNet \cite{miller1995wordnet}.
+E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
+``seism,'' ``aftershock,'' etc.
 
-\subsection{Semantic Similarity}\label{subsec:semsim}
+\paragraph{Language Model Features}\label{subsubsec:lm}
 
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
+%Because the data in our experiments is scraped from the web, it is common to
+%find sentences that contain both salient informantion and two kinds of noise:
+%noisey fragments of web page structure (e.g. section titles, \emph{News},
+%\emph{Sports}, etc.) and references to other news not relevant to the topic
+%summary.
+%
+We use two trigram language models, trained using the SRILM toolkit
+\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
+the sentence's total log probability normalized by sentence length) from each
+model.  This first model is trained on 4 years (2005-2009) of articles from
+the Gigaword corpus.  Specifically, we use articles from the Associated Press
+and the New York Times. This model is intended to assess the general writing
+quality (grammaticality, word usage) of an input sentence and helps us to
+filter out text snippets which are not sentences (e.g., web page titles).  The
+second model is a domain specific language model. We build a corpus of
+Wikipedia articles for each event type, consisting of documents from a related
+Wikipedia category. E.g. for earthquakes, we collect pages under the category
+\emph{Category:Earthquakes}. This model assigns higher probability to
+sentences that are focused on the given domain.
 
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
+%For both models, we Finally, we extract the percentage of capitalized words,
+%and sentence length as features. These last two features also help to
+%identify sentences that are less likely to contain relevant content-- overly
+%long and heavily capitalized sentences in our corpus were likely to be long
+%strings of web-page headlines, section headers, and other irrelevant page
+%structure.
 
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
+\paragraph{Geographic Relevance Features}
 
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
+Locations are identified using a named entity tagger. For each location in a
+sentence, we obtain its latitude and longitude using the Google Maps API.  We
+then compute its distance to that of the event location.  It is possible for a
+sentence and an event to have multiple locations so we take as features the
+minimum, maximum, and average distance of all sentence-event location pairs.
+Distances are calculated using the Vincenty distance.
 
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
+\paragraph{Temporal Relevance Features}
 
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
+Our data consists of hourly crawls of online content and so we exploit the
+temporality of corpus by capturing the burstiness of a sentence, i.e.  the
+change in word frequency from one hour to the next.``Bursty'' sentences often
+indicate new and important data.
 
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
+Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
+w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
+burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
+\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
+\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
+w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
+D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
 
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
+where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
+\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
+\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
+\end{align*}
+% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
+%\end{align*}
 
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
+We similarly find the sentence's 5-hour burstiness.  In addition to
+burstiness, we also include the sentence's average tf-idf and hours since the
+event in question started as features.
 
-\subsection{Gaussian Processes}
+\subsubsection{Gaussian Process Regression}
+\fdcomment{transition to gp description}
 
@@
 into which features are more important.
-
-%the formation of sentence clusters around more salient sentences. We
-%introduce the affinity propagation algorithm as as an elegant way to
-%incorporate our salience predictions into a
-
-%What follows is a description of our ongoing event summarization efforts.  We
-%briefly situate our approach to summarization within the broader field of
-%multi-document summarization, and then introduce the affinity propagation
-%algorithm which we use for clustering. This algorithm allows us to elegantly
-%address the salient sentence selection problem by incorporating our prior
-%beliefs about sentence quality. Next, we describe out method for modeling
-%summary sentence quality, and the features used in this model.  Finally, we
-%address future features and system improvements that we are incorporating
-%into our summarizer.
-
-\subsection{Temporal Summarization System}\label{sec:approach}
-
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.

### 2015-01-19 algorithms.tex
restructuring

@@
-\section{Algorithms}\label{sec:background}
-
-\subsection{Affinity Propagation}
-
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
+\section{Algorithm}
+\label{sec:background}
+\input{algorithm/introduction.tex}
+\input{algorithm/filtering.tex}
+\input{algorithm/salience.tex}
+\input{algorithm/ap.tex}
+\input{algorithm/novelty.tex}
 
@@
 
-\subsection{Semantic Similarity}\label{subsec:semsim}
-
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
-
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
-
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
-
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
-
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
-
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
-
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
-
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
-
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsection{Gaussian Processes}
-
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
-
 %the formation of sentence clusters around more salient sentences. We
@@
 
-\subsection{Temporal Summarization System}\label{sec:approach}
+%
+% \subsection{Temporal Summarization System}
+% \label{sec:approach}
+%
+%
+% We first begin with some notation. For a given event, let $\corpus$ be the set
+% of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
+% of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$.
+% Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let
+% $\corpus_{\hour_i}$ be the set of retrieved documents such that
+% $\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
+%
+% %Our TS system involves ? phases involving sentence level classification,
+% %regression, and clustering.
+% Figure ? outlines our general temporal summarization algorithm. The description
+% of our approach is as follows.
+% For each event, we iterate over the retrieved
+% documents in hourly chunks, emitting 0 or more updates at each hour.
+% At each hour $\hour$, we process each document
+% $\doc \in \corpus_{\hour}$. First, we identify where the document content
+% actually is; Second, we predict the salience of all content sentences.
+% To account for redundancy, the predicted salience is penalized based on the
+% distance of the sentence in question to the previous summary updates.
+%
+% Finally, we cluster all content sentences
+% for the current hour using the penalized salience predictions to bias the
+% formation
+% of clusters around the most salient sentences. For each cluster center, or
+% exemplar, that results, we check that the salience is above a threshold and
+% that it does not belong to a singleton cluster; exemplars that satisfy these
+% conditions are emitted as an update.
+% Additionally, we maintain the complete set of updates in order to penalize
+% salience predictions in the subsequent time steps.
+%
+% % \begin{figure}
+% % \centering
+% \begin{algorithm}%[H]
+%  \KwData{$Query$ --- the set of event query words\\
+%         $SC$ --- the stream corpus\\
+%
+%
+% }
+%  ~\\
+%  Initialize empty list $\mathbf{U}$ of updates\\
+%  Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
+%  $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
+%  \For{$i \gets 1,\ldots,t $}{
+%
+%   Initialize empty lists $\SMat, \Pref$ \;
+%
+%
+%   \For{$\doc \in \corpus_{\hour_i}$}{
+%    \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
+%      $\SMat.\operatorname{append}(\sent)$\;
+%      $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\
+%      $\Pref.\operatorname{append}(\sigma)$\;
+%
+%    }
+%   }
+%   %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
+%   %$\operatorname{}$
+%   ~\\
+%   $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
+%     (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
+%   ~\\
+%   $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\
+%   $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
+%   $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
+% %gets \Updates_{cache} \cup \Updates_{\hour}$\;
+%
+%  }
+%  \caption{Temporal Summarization Algorithm}
+% \end{algorithm}
+% % \end{figure}
 
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
+% \subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
+%
+% The focus of our system was on salience prediction and clustering stages, and
+% so we relied heavily on the pre-filtered corpus provided by the track
+% organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
+% a subset of the full TREC 2014 StreamCorpus and is
+% intended to be a high recall retrieval of documents related to all 15 of
+% the 2014 TS events.
+%
+% Track participants were provided with a set of query words for each event.
+% For example, the event ``Costa Concordia disaster and recovery'' had query
+% words [``costa'', ``concordia''].
+% In order to construct an event specific corpus,
+% we retrieve all documents whose timestamps fall within the event start/stop
+% times and whose raw html content contains at least one keyword from the
+% event's query words. We further restricted our document set to only those
+% articles from the news domain.
 
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
-
-% \begin{figure}
-% \centering
-\begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
-        $SC$ --- the stream corpus\\
-
-}
- ~\\
- Initialize empty list $\mathbf{U}$ of updates\\
- Initialize empty list $\boldsymbol{\Pref}^{(U)}$ of update preferences\\
- $\corpus \gets $ RetrieveDocuments($Query$, $SC$) (See \cref{subsec:Document Retrieval})\\
- \For{$i \gets 1,\ldots,t $}{
-
-  Initialize empty lists $\SMat, \Pref$ \;
-
-  \For{$\doc \in \corpus_{\hour_i}$}{
-   \For{$\sent \in \operatorname{getContent}(\doc, Query)$ (See \cref{subsec:Content Detection})\\}{
-     $\SMat.\operatorname{append}(\sent)$\;
-     $\sigma \gets \operatorname{PredictSalience}(\sent)$ (See \cref{subsec:Predict})\\     
-     $\Pref.\operatorname{append}(\sigma)$\;
-
-   }        
-  }
-  %$\Sim \gets \operatorname{ComputeSimilarityMatrix}(X)$\;
-  %$\operatorname{}$
-  ~\\ 
-  $\mathbf{U}_{h_i}, \Pref^{(U)}_{h_i} \gets \operatorname{SentenceSelection}
-    (\SMat, \Pref)$ (See \cref{subsec:SentenceSelection})\\
-  ~\\
-  $\operatorname{Emit}(\mathbf{U}_{h_i})$ \\ 
-  $\mathbf{U}\operatorname{.append}(\mathbf{U}_{h_i})$ \\
-  $\mathbf{\Pref}^{(U)}\operatorname{.append}(\Pref^{(U)}_{h_i})$ \\
-%gets \Updates_{cache} \cup \Updates_{\hour}$\;
-
- } 
- \caption{Temporal Summarization Algorithm}
-\end{algorithm}
-% \end{figure}
-
-\subsubsection{Document Retrieval}\label{subsec:Document Retrieval}
-
-The focus of our system was on salience prediction and clustering stages, and
-so we relied heavily on the pre-filtered corpus provided by the track 
-organizers. The TREC Temporal Summarization 2014 (TREC-TS-2014F) corpus is
-a subset of the full TREC 2014 StreamCorpus and is
-intended to be a high recall retrieval of documents related to all 15 of 
-the 2014 TS events.
-
-Track participants were provided with a set of query words for each event.
-For example, the event ``Costa Concordia disaster and recovery'' had query
-words [``costa'', ``concordia''].
-In order to construct an event specific corpus,
-we retrieve all documents whose timestamps fall within the event start/stop 
-times and whose raw html content contains at least one keyword from the 
-event's query words. We further restricted our document set to only those 
-articles from the news domain.
-
-\subsubsection{Content Detection}\label{subsec:Content Detection}
-
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
-
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
-
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
-
-\subsubsection{Predicting Sentence Salience}\label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
-
-\subsubsection{Basic Features}
-
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
-
-\subsubsection{Query Features}
-
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
-
-\subsubsection{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\subsubsection{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-
-\subsubsection{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
-\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-\end{align*}
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
-
-\subsubsection{Sentence Selection}\label{subsec:SentenceSelection}
-
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
-
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
-
-\subsubsection{Filtering}
-
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.
+% \subsubsection{Sentence Selection}\label{subsec:SentenceSelection}

### 2015-01-19 motivation.tex
restructuring

@@
 is broad or underspecified, clustering is most appropriate in that the data
-"speaks for itself." This is probably most true when summarization is 
-interpretted to mean exploratory data analysis. 
+``speaks for itself.'' This is probably most true when summarization is 
+interpreted to mean exploratory data analysis. 
 When there are more constraints on the output, as in many text summarization
@@
 a salience model by feature ablation for the TS task, but this feels very specific compared to the general framework idea.}
+
+\fdcomment{I guess I think we need to two things in this section (and/or introduction).  1. explain why the problem matters.  2. explain why our approach is compelling.  }
+
+\fdcomment{if easy to generate, would be good to have a single figure for an event that includes 1.
+the query, 2.
+the distribution of (filtered) document volume over time, and 3.
+a sample of the gold nuggets.
+this provides the reader with a visual explanation of the problem, the solution, and the difficulty}

### 2015-01-19 problemdefinition.tex
restructuring

\section{Problem Definition}
\label{sec:problemdefinition}
In order to evaluate a temporal summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization Track.  We provide a brief overview of the problem.  Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.

A temporal summarization system takes as input 
\begin{enumerate*}[label=\itshape\alph*\upshape)]
  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$, presented in temporal order, and
	\item an evaluation time period of interest, $(\stime,\etime)$.  
\end{enumerate*}
Throughout the time period of interest, the system must output sentences, known as \emph{updates}, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.  We present the evaluation metrics which measure these properties in Section \ref{sec:methods}.

In the ideal version of this problem, the system makes decisions every time it consumes a document $\doc_i$.  In practice, we assume that that we receive batches of new sentence-segmented documents every hour.

### 2015-01-23 kdd_paper/root.tex
Added kdd paper

% This is "sig-alternate.tex" V2.0 May 2012 This file should be compiled with
% V2.5 of "sig-alternate.cls" May 2012
%
% This example file demonstrates the use of the 'sig-alternate.cls' V2.5
% LaTeX2e document class file. It is for those submitting articles to ACM
% Conference Proceedings WHO DO NOT WISH TO STRICTLY ADHERE TO THE SIGS
% (PUBS-BOARD-ENDORSED) STYLE.  The 'sig-alternate.cls' file will produce a
% similar-looking, albeit, 'tighter' paper resulting in, invariably, fewer
% pages.
%
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V2.5) produces: 1) The Permission
% Statement 2) The Conference (location) Info information 3) The Copyright
% Line with ACM data 4) NO page numbers
%
% as against the acm_proc_article-sp.cls file which DOES NOT produce 1) thru'
% 3) above.
%
% Using 'sig-alternate.cls' you have control, however, from within the source
% .tex file, over both the CopyrightYear (defaulted to 200X) and the ACM
% Copyright Data (defaulted to X-XXXXX-XX-X/XX/XX).  e.g.
% \CopyrightYear{2007} will cause 2007 to appear in the copyright line.
% \crdata{0-12345-67-8/90/12} will cause 0-12345-67-8/90/12 to appear in the
% copyright line.
%
% ---------------------------------------------------------------------------------------------------------------
% This .tex source is an example which *does* use the .bib file (from which
% the .bbl file % is produced).  REMEMBER HOWEVER: After having produced the
% .bbl file, and prior to final submission, you *NEED* to 'insert' your .bbl
% file into your source .tex file so as to provide ONE 'self-contained' source
% file.
%
% ================= IF YOU HAVE QUESTIONS ======================= Questions
% regarding the SIGS styles, SIGS policies and procedures, Conferences etc.
% should be sent to Adrienne Griscti (griscti@acm.org)
%
% Technical questions _only_ to Gerald Murray (murray@hq.acm.org)
% ===============================================================
%
% For tracking purposes - this is V2.0 - May 2012

\documentclass{sig-alternate} \usepackage{url} \usepackage{color}
\newcommand{\fdadd}[1]{\textcolor{red}{#1}}
\newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}

\begin{document}

%
% --- Author Metadata here ---
\conferenceinfo{WOODSTOCK}{'97 El Paso, Texas USA}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
%- IF NEED BE.  \crdata{0-12345-67-8/90/01}  % Allows default copyright data
%(0-89791-88-6/97/05) to be over-ridden - IF NEED BE.  --- End of Author
%Metadata ---

\title{Summarizing Disasters Over Time}
%    Alternate {\ttlit ACM} SIG Proceedings Paper in LaTeX
%Format\titlenote{(Produces the permission block, and copyright information).
%For use with SIG-ALTERNATE.CLS. Supported by ACM.}} \subtitle{[Extended
%Abstract] \titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at \texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement and
% alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time' i.e. three
% 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of "name/affiliations" may
% appear. We just ask that you restrict the number of 'columns' to three.
%
% Because of the available 'opening page real-estate' we ask you to refrain
% from putting more than six authors (two rows with three columns) beneath the
% article title.  More than six makes the first-page appear very cluttered
% indeed.
%
% Use the \alignauthor commands to handle the names and affiliations for an
% 'aesthetic maximum' of six authors.  Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the \additionalauthors
% command.  These 'additional authors' will be output/set for you without
% further effort on your part as the last section in the body of your article
% BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting reasons)
% and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here, e.g. one 'row of
% three' or two rows (consisting of one row of three and a second row of one,
% two or three).
%
% The command \alignauthor (no curly braces needed) should precede each author
% name, affiliation/snail-mail address and e-mail address. Additionally, tag
% each line of affiliation/address with \affaddr, and tag the e-mail address
% with \email.
%
% 1st. author
\alignauthor Chris Kedzie\\ \affaddr{Columbia University}\\
\affaddr{Department of Computer Science} \email{kedzie@cs.columbia.edu}
% 2nd. author
\alignauthor Kathleen McKeown\\ \affaddr{Columbia University}\\
\affaddr{Department of Computer Science}\\ \email{kathy@cs.columbia.edu}
% 3rd. author
\alignauthor Fernando Diaz\\ \affaddr{Microsoft Research}\\
\email{fdiaz@microsoft.com}
% There's nothing stopping you putting the seventh, eighth, etc.  author on
% the opening page (as the 'third row') but we ask, for aesthetic reasons that
% you place these 'additional authors' in the \additional authors block, viz.
} \date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors is the number
% that will appear on the first page PLUS the number that will appear in the
% \additionalauthors section.

\maketitle \begin{abstract} We have developed a text summarization system that
can generate summaries over time from web crawls on disasters. We show that
our method of identifying exemplar sentences for a summary using affinity
propagation clustering produces better summaries than clustering based on
K-medoids as measured using Rouge on a small set of examples. A key component
of our approach is the prediction of salient information using event related
features based on location, temporal changes in topic, and two different
language models.  \end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures,
performance measures]

\terms{Summarization}

\keywords{Extractive Summarization, Affinity Propagation}

\section{Introduction}

During crises, information is critical for first responders and those caught
in the event.  When the event is significant, as in the case of Hurricane
Sandy, the amount of information produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. Methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively  (e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
these experiments were not performed in the large and heterogeneous
environment of the modern web.

%Previous work on generating event descriptions and/or multi-document
%summarization has relied on clustering algorithms to find representative
%sentences appropriate for an event summary.  These methods impose a metric
%space on the text data that can make it difficult to incorporate external
%sources of information elegantly -- in this paper we argue that centroid
%sentences are not a priori the best candidates for inclusion in an event
%summary.

In this paper, we present an update summarization system to track events
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster,  and integrates these predictions into
a clustering based multi-document summarization system. We train a regression
model to predict sentence salience and use these predictions to bias the
formation of sentence clusters around more salient regions in the input space
using affinity propagation (AP) clustering.  AP uses the salience predictions
as well as pairwise similarities among input sentences to identify
\emph{exemplar} sentences, which we use as our summary output.  Our approach
differs from other methods of summarization that compute salience by pairwise
comparisons alone, ignoring features of importance that are intrinsic to the
sentences themselves.

%the formation of sentence clusters around more salient sentences. We
%introduce the affinity propagation algorithm as as an elegant way to
%incorporate our salience predictions into a

%What follows is a description of our ongoing event summarization efforts.  We
%briefly situate our approach to summarization within the broader field of
%multi-document summarization, and then introduce the affinity propagation
%algorithm which we use for clustering. This algorithm allows us to elegantly
%address the salient sentence selection problem by incorporating our prior
%beliefs about sentence quality. Next, we describe out method for modeling
%summary sentence quality, and the features used in this model.  Finally, we
%address future features and system improvements that we are incorporating
%into our summarizer.

\section{Related Work}

A principal concern in extractive multi-document summarization is the
selection of salient sentences for inclusion in summary output
\cite{nenkova2012survey}.  This has often been approached as a ranking
problem.
%We broadly conceptualize this decision as either an intrinsic or extrinsic
%sentence evaluation process. Intrinsic approaches evaluate sentences
%individually, possibly by predicting the impact on summary quality using
%sentece level features. 
Sentences have been ranked by the average word probability, average tf-idf
score, and the number of topically related words (topic-signatures in the
summarization literature)
\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
statistics are easily computable from the input sentences, while the third
only requires an additional, generic background corpus.  Another ranking
approach, centroid summarization, involves creating an average bag of words
(BOW) vector, the centroid, from the input sentences and ranking sentences by
their similarity to the centroid \cite{radev2004centroid}.  Graph
\cite{erkan2004lexrank} and clustering
\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
based approaches, on the other hand, make use of pair-wise similarity
comparisons amongst input sentences.  In these models, salient sentences are
more central to the input or cluster, respectively.

%identify salient regions of the input space while simultaneously coping with
%redundancy.  Graph-based algorithms have been used to rank sentences
%Clustering algorithms, e.g., are commonly used to exploit redundancy in
%input. Input sentences are clustered and summaries are generated by selecting
%the most representative sentence from each cluster.  Graph-based models have
%also been used for summarization.  E.g., the LexRank algorithm treats
%sentences as nodes in a graph, where edges are constructed by way of cosine
%similarity between sentence nodes; edges are either continuosly weighted by
%similarity or discrete, existing only when the similarity is above a
%threshold.  The PageRank algorithm is used on the graph to find the most
%important sentence nodes. In both clustering and graph-based approaches,
%sentence salience is largely determined by the pairwise relations between
%sentences.

Supervised learning has also been applied to this task. Model features are
usually derived from human generated summaries, and are non-lexical in nature
(e.g., sentence starting position, number of topic-signatures, number of
unique words, word frequencies). Seminal work in this area has employed naive
Bayes and logistic regression classifiers to identify sentences for summary
inclusion \cite{kupiec1995trainable,conroy2001using}.

%\fdadd{
Several researchers have recognized the importance of summarization during
natural disasters.  Guo \textit{et al.} developed a system for detecting
novel, relevant, and comprehensive sentences immediately after a natural
disaster \cite{qi:temporal-summarization}.  The method uses a model of
sentence relevance and novelty in order to select appropriate updates.
Training data for regression targets is automatically generated from
retrospective Wikipedia data.  The system is evaluated on news documents
related to 197 natural and human disasters from 2009 to 2011 using variants of
Rouge modified to capture novelty, relevance, and comprehensiveness
\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
efficiency detect important updates during natural disasters
\cite{wang:update-summarization}.  The algorithm works by hierarchically
clustering sentences online, allowing the system to output a more expressive
narrative structure than Guo \textit{et al.}.  The method is evaluated on
official press releases related to Hurricane Wilma  in 2005 using Rouge score
between the system summary and a manually generated target summary.
%}

This work uses the TREC Stream Corpus data set that is in use by the TREC
Temporal Summarization track \cite{frank2012building,aslam2013trec} .
Generally, last year's participants used a pipelined approach to build
summaries, generally ranking sentences, filtering out all but the most
relevant, and then performing some sort of deduplication/redundancy removal
step.  Ranking approaches ranged from simple query word match to more
sophisticated query expansion and and query based language model scoring
\cite{liu2013ictnet,xu2013trec,baruah2013univ}.  Perhaps most similar to our
approach, the system of \cite{xu2013trec} uses a weighted combination of
features (similarity to query, named entity frequency, predicate frequency,
presence of numerical values, sentence novelty, etc.) to score sentences;
sentences above a threshold are added to the summary.  Both the weights and
the threshold are selected by hand.

Our system seeks to combine the best of these approaches, using supervised
learning to predict salience rankings, and directly incorporate this
information in a clustering algorithm to bias the formation of sentence
clusters around highly salient regions.

%Broadly, we conceive of the above

%Another popular graph-based model

%Larger clusters generally indicate more important For example

%\cite{} also explore the use of a hidden Markov model to predict summary
%inclusion based on previously selected sentences.

%ranking.
%
%The field of extractive summarization has long been focused on determing
%sentence importance or salience.  Many These efforts have largely involved
%unsupervised, data-driven approaches, where systems rank input sentences
%using features easily computable from the input, and sometimes an unrelated
%background corpus.  The most basic sentence rankings are by average word
%probability and average TF-IDF score, statistics that are computable from the
%input alone.  One of the more effective techniques, topic-sigantures,
%involves identifying the most topically related words in the input sentences.
%This is done using a log-likelihood ratio test ($\chi^2$) to determine which
%of two hypotheses hold for a word in the input: (1) the probability of the
%word in the input distribution and the background distribution are identical,
%or (2) the probability of the word is greater in the input distribution than
%in the background distribution. If (2) is true with some statistical
%significance, the word is said to be a topic signature. As with the previous
%features, a common approach is to rank sentences by the number of
%topic-signatures they contain.
%
%In mulit-document summarization, clustering algorthims are often employed to
%determine sentence importance, with larger clusters corresponding to more
%important topics. 
%
%Supervised learning approaches have also been explored. Features are usually
%derived from human generated summaries, and are non-lexical in nature (e.g.,
%sentence starting position, number of topic-signatures, number of unique
%words, word frequencies). Seminal work in this area has employed naive Bayes
%and logistic regression classifiers to identify sentences for summary
%inclusion. \cite{} also explore the use of a hidden Markov model to predict
%summary inclusion based on previously selected sentences.
%
%If this  
%
%tried to rank sentences by their overlap with query words. 
%

%and models are developed to predict the likelihood of a sentences inclusion
%in a summary. Naive Bayes, hidden Markov models, and logistic regression to
%predict a sentence's summary inclusion.
%
%Researches have also experimented with supervised learning for summarization,
%however, machine learning approaches have not yielded significant advances in
%generic summarization. One difficulty is that gold data is difficult to
%obtain.
%
%
%In  
%
%
%%which remain competetive to supervised machine-learning techniques.
%
%have ranked sentences by
%
%
%their average word probability, average TF-IDF weighting, and/or topic
%signatures; these representations are all easily computable from the input
%sentences, and, for topic signatures, an unrelated background corpus.  Topic
%signatures in
%
% 
%These efforts can be broadly split into two camps: feature based ranking and
%clustering.
%
%
%
%Determing sentence importance has been A large portion of automatic
%summarization research has focused on  
%
%

\section{Multi-Document Summarization Framework}

%To generate event summaries, we identify the most salient sentences from a
%collection of related documents and present these sentences as the event
%summary. This approach, commonly referred to as extractive summarization,
%avoids the complexity of parsing a semantic representation from the input
%sentences and subsequently generating a summary\cite{nenkova2012survey}.

A common approach to automatic summarization is to identify sentences with the
highest centrality with respect to the input sentences.  Intuitively,
sentences with a high degree of centrality are more semantically related to
the entire set of input sentences.  A summary can thus be obtained by
returning the $k$ most central sentences.  This generally implies the
calculation of pairwise distances between all sentences
\cite{radev2004centroid, erkan2004lexrank}.  In these approaches, sentences
are evaluated extrinsically by their distance to other sentences, either
directly \cite{erkan2004lexrank} or through an aggregate centroid object
\cite{radev2004centroid}. The distance between sentences is most commonly the
cosine distance of sentence term-vectors, but in general this can be an
arbitrary real-valued similarity function.
%; the distance function imposes a topology on the space of input sentences.

For some domains, it is very likely that we will have additional background
knowledge that could be predictive of sentence salience for the event being
summarized.  For example, certain kinds of information extracted from the
sentence text (e.g., temporal or geographic proximity) can indicate relevance
to a given event.  It would be difficult to incorporate this kind of salience
into the measure of centrality.

For example, consider a cluster of three sentence vectors $s_1 = (1,1,0)$,
$s_2 = (1,1,1)$, and $s_3 = (0,1,1)$. Without any other information, $s_2$ has
the highest degree of centrality, i.e. it has the highest average cosine
similarity and smallest average Euclidean distance to the other sentences.
Now, if we believe that $s_1$ is $\alpha$ times more salient than the other
sentences, we cannot simply scale $s_1$ by $\alpha$--the average cosine
similarity will remain unchanged, since the vector magnitude does not affect
the angle and $s_1$ will have an even greater the average Euclidean distance
from the rest.  Worse still, $s_2$ will still be the most representative
sentence of the three.

%For example, if we scaled the BOW vector based on our belief in its salience,
%it would not effect the cosine distance since the angle between two vectors
%is not changed by the magnitude of either

%Simply scaling the sentence vector based on our belief in it's fitness does
%not with our prior beliefs-- When using the BOWs representation, the input
%space is lexically defined, but our salience metric may not be based on any
%specific lexical items.

%These approaches make it difficult to leverage prior knowledge to guide the
%selection of salient sentences. Sentences

%
%Centroid based algorthims are commonly employed in the summarization
%literature to compress a large number of sentences into a much smaller,
%representative set. 
%
%The $k$-medoids clustering algorithm is commonly used to partition input
%sentences into topically distinct clusters, where the most representative
%sentences (the cluster medoids) are taken to build the summary. The quality
%of the summary in this framework is dependent on the method of sentence
%representation and the ability of a similarity or distance metric to
%adequately capture pairwise comparisons between sentences. Previous work in
%event clustering has examined the use of different term weightings (tf-idf,
%named-entities)  and lower dimensional term representations (LSA, LDA) for
%use in summarization.
%
%
%A popular graph-based alternative, LexRank, treats each sentence as a node in
%a graph, and identifies the nodes that are most central by in-degree.  The
%existence of an edge between two sentences is determined by a similarity
%function (either by setting a threshold above which the edge exists, or by
%taking the similarity as a continuous edge weight). The sentences with the
%highest eigenvector centrality can be found using random walks (PageRank).
%The $k$ most central sentences are used as the summary.
%
%One drawback of these algorithms is that they do not provide an elegant way
%to incorporate prior information about sentence salience.

In our approach, the system generates clusters using an affinity propagation
algorithm and from each cluster an exemplar sentence is selected that is added
to the summary.  In the following sections, we show how  prior information
representing salience can easily be incorporated into the affinity propagation
algorithm.  We believe the incorporation of salience to be useful in noisy
environments (e.g., a web crawl), and that it can help the formation of
clusters around the most relevant inputs. Our current system  is trained using
features derived from location, changes in wording across time and language
models that characterize the language of disaster to generate summaries at
regular intervals across time.  As we develop the system further, we will
extend it to generate updates across time,  penalizing the salience of
concepts already selected by the summarizer to encourage the discovery of
novel sentences as the event unfolds.

\subsection{Data}

Our documents for summarization come from the online news portion of the TREC
Stream Corpus, a 6.45tb corpus obtained by hourly web crawls from October 2011
through mid February 2013
\cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
Summary events come from the TREC Temporal Summarization track, and include
natural disasters like Hurricane Sandy as well as man-made events like a 2012
train accident in Buenos
Aires.\footnote{\url{http://trec.nist.gov/data/tempsumm2013.html}} The track
organizers also provide a search query for each event \cite{aslam2013trec}.
For each event, we collect the documents that contain all query words and
stratify them by the hour they were collected.

For evaluation purposes, the track organizers also provided gold nugget
information (i.e. important pieces of information, usually the length of a
short clause or sentence). These gold nuggets come from the event's related
Wikipedia article and also include the timestamp of when they were added to
the page.

%KM _ Ask Chris: is this true?
To create hourly gold summaries to evaluate our system, we simply take the set
of gold nugget information from the start of the event up to the current hour.

\subsection{Affinity Propagation}

Affinity propagation (AP) is a message passing algorithm that identifies both
exemplar data points and assignments of each point to an exemplar.  This is
done iteratively by passing \emph{responsibility} and \emph{availability}
messages between data points that quantify the fitness of one data point to
represent another, and the fitness of a data point to be represented based on
the choices of other data points respectively \cite{dueck2007non}.

%Each element $R(i,k)$ expresses the fitness of the $k^{th}$ point to serve as
%the exemplar of the $i^{th}$ point relative to other potential exemplars.
%Each element $A(i, k)$ represents the $k^{th}$ point's ``availability'' to
%serve as an exemplar of the $i^{th}$ point, taking into account other points'
%preference for the $k^{th}$ point as an exemplar.

AP is parameterized by a $n\times n$ similarity matrix $S$ and a $n\times 1$
preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
our preference that the $i$-th data point can serve as an exemplar a priori of
other data points.

In our experiments $\pi(i)$ is set to the salience prediction from the
Gaussian process regression for the $i$-th sentence minus an offset. This
offset drives most of the preferences negative and reduces the number of
returned exemplars to a handful of sentences (around 4-5).  For the similarity
matrix, we use $S(i,j) = -\operatorname{dist}(i,j)^2$, where
$\operatorname{dist}$ is the Euclidean distance between the BOW vectors for
sentences $i$ and $j$.
%we used the negative squared cosine distance of sentence bag-of-words. 
The summary output is the set $\mathcal{E}$ of returned exemplars found after
convergence.

%While, AP does not set the number of exemplars before-hand, a lower overall
%preference values will result in a smaller number of exemplars.

%arbitrary pair-wise similarity function $S: \mathcal{R}^d \rightarrow
%\mathcal{R}$, where $d$ is the dimension of the data being clustered and a
%real-valued preference $\pi_i$ quantifying our belief a priori of the
%$i^{th}$ element's ability to serve as an exemplar.

AP has two useful properties for summarization. First, the number of clusters
identified is determined by the preferences -- lower overall preference
values  will result in fewer clusters. Unlike $k$-means, we do not have to
specify how many clusters we would like to find.  Determining the number of
clusters in a principled way each time we run the clustering algorithm would
be difficult in our setup.  Secondly, the arbitrary nature of the preferences
and similarity function allow us to incorporate a variety of signals for
identifying the best exemplars.

\subsection{Predicting Sentence Salience}

In order to use AP clustering for summarization, we need to assign a
preference value to each input sentence.  In our approach, we equate a
sentence's salience with its preference.  A good model of sentence salience
should predict higher values for sentences that are more likely to appear in a
human generated summary of the event.

To build training data for this regression task, we take a subset of sentences
relevant to the TREC events (approximately 1000) and match them to the gold
nugget sentence with highest similarity as determined by the sentence
similarity system of \cite{guo2012simple}. 
%\cite{} have used this system previously to correlate sentences to meaningful
%units of information in human generated summaries. 
We use the real-valued similarity scores as our salience scores for the
training sentences.

We want our model to be predictive across different kinds of events so we
avoid lexical features.  Instead, we extract a variety of features including
language model scores, geographic relevance, and temporal relevance from each
sentence.  These features are used to fit a Gaussian process regression model
that can predict the similarity of a sentence to a gold summary
\cite{preotiuc2013temporal}.  We use the model predicted salience of each
sentence as it's preference value in the AP clustering. 
%In order to predict the salience score for a sentence,

 %\cite{guo2012simple} produces a real-valued score of the similarity We use
 %\cite{guo2012simple} produces real valued similarity scores s.
 %\cite{guo2012simple} determines the similarity of two sentences by deriving
 %low dimensional We have As a proxy measure for sentence salience, we use the
 %maximum similarity of an input sentence to a human generated summary
 %sentence, using the state-of-the-art sentence similarity module of
 %\cite{guo2012simple}. \cite{guo2012simple} derive a lower-dimensional vector
 %of latent features for a sentence. The similarity of two sentences is then
 %the cosine similarity of the latent vectors.  In order to predict this
 %similarity we extract a variety of features including language model scores,
 %geographic relevance, and temporal relevance.  We learn a predictive
 %function of sentence-summary similarity by fitting a Gaussian process model
 %on these features. \cite{preotiuc2013temporal}

%As training data for this model, we take a portion of the sentences found in
%the relevant documents and find their highest similarity to the gold nugget
%information seen thus far.

%When assigning preference scores

%After obtaining similarity scores for sentences in our training set, we fit a
%Gaussian process model to learn

%Our model

%Correllating automatic measures of sentence salience with human judgments is
%a difficult task. , we use the maximum cosine similarity to

\begin{table*}[ht] \begin{center} \begin{tabular}{| c p{7.0cm} || c p{7.0cm}
|} \hline Preference & AP Clustering & Preference & $k$-Medoids Clustering\\
\hline 9.010 & The magnitude-7.5 quake, about 20 miles deep, was centered off
the town of Champerico .People fled buildings in Guatemala City , in Mexico
City and in the capital of the Mexican state of Chiapas , across the border
from Guatemala & 9.010 & The magnitude-7.5 quake, about 20 miles deep, was
centered off the town of Champerico .People fled buildings in Guatemala City ,
in Mexico City and in the capital of the Mexican state of Chiapas , across the
border from Guatemala   \\ \hline 9.010 & A reporter in the town of San Marcos
, about 80 miles north of the epicenter, told local radio station Emisoras
Unidas that houses had collapsed onto residents and smashed televisions and
other appliances had been scattered into the streets.  & 3.007 & ``Things fell
in my kitchen .'' Perez said more than 2,000 soldiers were deployed from a
base in San Marcos to help with disaster relief.\\ \hline 9.007 & The local
fire department said on its Twitter account that a school had collapsed and
eight injured people had been taken to a nearby hospital.  & 3.007 &  Ingrid
Lopez , who went to the hospital with a 72-year-old aunt whose legs was
crushed by a falling wall, said she had waited hours for an X-ray.\\ \hline
7.008 & There are three confirmed dead and many missing after the strongest
earthquake to hit Guatemala since a deadly 1976 quake that killed 23,000.  &
1.007 & Hundreds of people crammed into the hallways of the small town
hospital waiting for medical staff to help out hundreds of injured family
members, some complaining they were not getting care quickly enough. \\ \hline
\end{tabular} \end{center} \caption{Example summary using affinity propagation
(left) and $k$-medoids (right)} \end{table*} \begin{table*}[t] \begin{center}
\begin{tabular}{| c || c | c | c || c | c | c || c | c | c |} \hline Method &
\multicolumn{3}{c}{ROUGE-1} & \multicolumn{3}{c}{ROUGE-2} &
\multicolumn{3}{c}{ROUGE-3} \\ & Recall & Prec. & F-1 & Recall & Prec. & F-1 &
Recall & Prec. & F-1 \\ \hline $k$-medoids & 0.127 & 0.414 & 0.181 & 0.025 &
0.076 & 0.035 & 0.003 & 0.010 & 0.005\\ \hline AP & 0.117 & 0.440 & 0.173 &
0.022 & 0.082 & 0.033 & 0.004 & 0.015 & 0.006\\ \hline \end{tabular}
\end{center} \caption{ROUGE scores for $k$-medoids and affinity propagation
methods} \end{table*}

\subsection{Basic Features}

We employ several basic features that have been used previously in supervised
models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
These include sentence length, the number of capitalized words normalized by
sentence length, and the number of query words present in the sentence.  Query
words include the event's type (e.g., \emph{earthquake}) and are expanded with
the event type's WordNet \cite{miller1995wordnet} synset, hypernyms, and
hyponyms.  For \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
``seism,'' ``aftershock,'' etc.

\subsection{Language Model Features}

%Because the data in our experiments is scraped from the web, it is common to
%find sentences that contain both salient informantion and two kinds of noise:
%noisey fragments of web page structure (e.g. section titles, \emph{News},
%\emph{Sports}, etc.) and references to other news not relevant to the topic
%summary.
%
We use two trigram language models, trained using the SRILM toolkit
\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
the sentence's total log probability normalized by sentence length) from each
model.  This first model is trained on 4 years (2005-2009) of articles from
the Gigaword corpus.  Specifically, we use articles from the Associated Press
and the New York Times. This model is intended to assess the general writing
quality (grammaticality, word usage) of an input sentence and helps us to
filter out text snippets which are not sentences (e.g., web page titles).  The
second model is a domain specific language model. We build a corpus of
Wikipedia articles for each event type, consisting of documents from a related
Wikipedia category. E.g. for earthquakes, we collect pages under the category
\emph{Category:Earthquakes}. This model assigns higher probability to
sentences that are focused on the given domain.

%For both models, we Finally, we extract the percentage of capitalized words,
%and sentence length as features. These last two features also help to
%identify sentences that are less likely to contain relevant content-- overly
%long and heavily capitalized sentences in our corpus were likely to be long
%strings of web-page headlines, section headers, and other irrelevant page
%structure.

\subsection{Geographic Relevance Features}

Locations are identified using a named entity tagger. For each location in a
sentence, we obtain its latitude and longitude using the Google Maps API.  We
then compute its distance to that of the event location.  It is possible for a
sentence and an event to have multiple locations so we take as features the
minimum, maximum, and average distance of all sentence-event location pairs.
Distances are calculated using the Vincenty distance.

\subsection{Temporal Relevance Features}

Our data consists of hourly crawls of online content and so we exploit the
temporality of corpus by capturing the burstiness of a sentence, i.e.  the
change in word frequency from one hour to the next.``Bursty'' sentences often
indicate new and important data.

Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
\end{align*}
% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
%\end{align*}

We similarly find the sentence's 5-hour burstiness.  In addition to
burstiness, we also include the sentence's average tf-idf and hours since the
event in question started as features.

\section{Experiments}

We carried out a small set of initial experiments on one event.  We collected
a subset of pages from the TREC Stream Corpus that were relevant to a 2012
earthquake off the coast of Guatemala, and further subdivided this collection
by the hour they were created. For each hour we generated a summary using the
AP clustering algorithm.

 We also generated baseline summaries using the $k$-medoids (using the
 Partitioning Around Medoids algorithm), setting $k = |\mathcal{E}|$ , i.e.
 the number of exemplar sentences returned by the AP. Because $k$-medoids
 begins with a random initialization, we took the best (minimum average
 distance) result of 100 restarts.

Table 1 shows example output of the AP and $k$-medoids generated summaries.
Sentences are ranked by preference score, although preference has no effect on
the $k$-medoids algorithm.  Quantitatively, AP exemplar sentences had higher
predicted sentence quality scores (preferences) than the cluster medoids.
Qualitatively, the AP method appears to select more general details about the
earthquake. Looking at the third sentence selection in table 1, we can see
that $k$-medoids selects a personal experience that was reported. This is
perhaps less newsworthy or reportable compared to the third sentence in in the
AP generated summary which reports a notable structure collapse and injuries
related to the quake.  We believe AP results in a more readable and
informative summary, although we have yet to perform a rigorous human
evaluation of the summary output.

We evaluated both algorithms with the ROUGE toolkit \cite{lin2004rouge}.
N-ROUGE works by calculating the n-gram recall and precision of an
automatically generated summary in reference to a model summary.  We created
model summaries by taking the gold nugget sentences with timestamps up to and
including the current system time as the gold summary for that hour.
%\fdcomment{may be interesting to measure track metrics to measure latency,
%redundancy, etc.} KM - Ask Chris

Table 2 shows average recall, precision, and F-measure for various orders of
ROUGE score. AP demonstrated consistently higher precision than our baseline.
%scores over $k$-medoids (ROUGE evaluations are usually focused on recall).
While not statistically significant, it is difficult to show significance with
Rouge using a small test; we hope further tests will confirm this improvement. 
%   , AP had higher precision on average.
On average, the AP summaries were slightly shorter than the baseline, which
would partially explain this difference. It is also possible that our language
models are biased toward shorter sentences; we are more likely to have seen a
shorter sentence in the language model input.
%\fdcomment{is there a theory to why this is happening?}
We are currently adapting our summarizer to add updates over time, and
maintaining precision will be important to prevent topic drift.

%\begin{table}[h!] \tiny \begin{center} \begin{tabular}{| c p{3.5cm} |} \hline
%Preference & Sentence\\ \hline 9.010 & The magnitude-7.5 quake, about 20
%miles deep, was centered off the town of Champerico .People fled buildings in
%Guatemala City , in Mexico City and in the capital of the Mexican state of
%Chiapas , across the border from Guatemala\\ \hline 3.007 & "Things fell in
%my kitchen ." Perez said more than 2,000 soldiers were deployed from a base
%in San Marcos to help with disaster relief.\\ \hline 3.007 & Ingrid Lopez ,
%who went to the hospital with a 72-year-old aunt whose legs was crushed by a
%falling wall, said she had waited hours for an X-ray. \\ \hline 1.007 &
%Hundreds of people crammed into the hallways of the small town hospital
%waiting for medical staff to help out hundreds of injured family members,
%some complaining they were not getting care quickly enough. \\ \hline
%\end{tabular} \end{center} %  \caption{Example summary using $k$-medoids with
%100 random restarts} \end{table}

\section{Conclusions and Future Work}

We have developed a summarizer that can generate summaries over time from web
crawls on disasters. We show that our method of identifying exemplar sentences
for a summary using AP clustering produces  summaries with higher precision
    compared to those based on clustering with K-medoids. A key component of
    our approach is the prediction of salient information using features based
    on location, temporal changes in topic, and two different language models.

Currently, we run each hour of summarization independently. In order to avoid
repeating information, we would like to incorporate previously chosen
exemplars in the preference computation. One possibility would be to
down-weight a candidate exemplar's preference based on its similarity to
previous exemplars.

Secondly, we would like to do more intelligent inference of missing
geographical information since not all sentences contain locations. Currently
we are using mean values for missing data.

Finally, we would like to experiment with non-symmetric similarity matrices,
specifically using narrative chains\cite{chambers2009unsupervised}. Under this
model $S(i,j)$ would express the likelihood that the events in sentence $j$
precede the events in sentence $i$. We hope such a parameterization would
promote more causally motivated sentences into exemplar positions, which would
better describe the disaster event domain.

%
\bibliographystyle{abbrv}
%\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this
%case
\bibliography{kdd.bib}
% You must have a proper ".bib" file and remember to run: latex bibtex latex
% latex to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional \balancecolumns \balancecolumns % GM June 2007 That's
%all folks!
\end{document}

### 2015-01-24 algorithm/filtering.tex
edits to algo and problem definition mostly

@@
 We simulate our system using realistic data insofar as the document stream consists of relevant and non-relevant content.
-As with many information retrieval systems, we crude, first phase of high-recall document filtering.
-That is, we would like to remove most of the content which we can easily detect as unlikely to contain useful updates.
+As with many information retrieval systems, we use a crude, first phase of high-recall content filtering for both documents and sentences.
 
@@
 As a first filter, we retained only those documents whose raw html content contains at least one keyword from the event's query words.
-We further restricted our document set to only those articles from the news domain.\footnote{This filtered set was distributed by track organizers.}
+We further restricted our document set to only those articles from the news domain.\footnote{This filtered set was distributed by track organizers.} \fdcomment{not sure how much we have to describe this.}
 
@@
 \label{sec:filtering:sentences}
-In early versions of our summarization system, we found that structural html
-artifacts and sentence tokenization errors
-were negatively effecting the performance of later stages. Examples of the 
-former include strings of link text like ``World Politics Sports ...'', while
-examples of the latter included concatenations of various article headlines,
-i.e. headlines pertaining to the event in question, as well as headlines from
-other non-related events.
-Both types of ``sentences'' were problematic for computing sentence similarity
-as they were more likely to have higher average similarity to all input 
-sentences. In turn they would be more likely to appear as cluster exemplars
-in our clustering stage. From the clustering algorithm's point of view, this
-is the correct decision to make---such multi-topic sentences are more general 
-than a single topic, and better able to represent all aspects contained in a 
-cluster. They make poor choices as updates, however, as they contain 
-irrelevant information.
+Structural html artifacts and sentence tokenization errors can negatively effect the performance of a temporal summarization system.
+Examples of the former include strings of link text like ``World Politics Sports ...'', while examples of the latter included concatenations of unrelated content  (e.g.
+segmentation errors treating two headlines as a single sentence).
 
-In order to filter out these problematic inputs, we trained a classifier to 
-identify which sentences came from inside a document's main article and which
-came from various headers, titles, menus, and links to other content. We collected ? random sentences and manually labeled whether the sentence came from
-inside or outside the document's main article. We then trained a logistic 
-regression classifier using the following features :
+In order to filter out these problematic inputs, we trained a classifier to identify which sentences came from inside a document's main article and which came from various headers, titles, menus, and links to other content.
+We collected random sentences and manually labeled whether the sentence came from inside or outside the document's main article.
+We then trained a logistic  regression classifier using:
+\begin{enumerate*}[label=\itshape\alph*\upshape)]
+	\item the position of the sentence,
+	\item word counts,
+	\item the last token in the sentence,
+	\item the last two tokens in the sentence, and
+	\item the last three tokens in the sentence.
+\end{enumerate*}
+These features were sufficient to capture the main difference between content and non-content sentences, which  was that content sentences generally ended with sentence final punctuation, i.e.
+periods or a closing quotation mark.
+\fdcomment{include summary performance numbers here; also a reference to prior work.}
 
-\begin{itemize}
- \item the position of the sentence
- \item word counts
- \item the last token in the sentence
- \item the last two tokens in the sentence
- \item the last three tokens in the sentence.
-\end{itemize}
-
-These features were sufficient to capture 
-the main difference between content and non-content sentences,
-which 
-was that content sentences generally ended with sentence final punctuation, 
-i.e. periods or a closing quotation mark.
-
-Within our larger summarization framework, we process a document at a time,
-identifying the subset of its sentences that are content sentences.
-We then check to make sure \emph{all} event
-query terms can be found within the document's content sentences. If so,
-we send the content sentences on to the next stage of our pipeline; otherwise
-we ignore all sentences in this document.
+After identifying the subset of a document's sentences that are content sentences, we check to make sure \emph{all} event query terms can be found within the document's content sentences.
+If so, we send the content sentences on to the next stage of our pipeline; otherwise we ignore all sentences in this document.

### 2015-01-24 algorithm/introduction.tex
edits to algo and problem definition mostly

@@
-We first begin with some notation. For a given event, let $\corpus$ be the set
-of retrieved documents. A document $\doc \in \corpus$ is an ordered sequence
-of sentences $\{\sent_{1,\doc},\ldots,\sent_{|\doc|,\doc} \}$. 
-Additionally, each document has a timestamp $\dtime(\doc)$. Finally, let 
-$\corpus_{\hour_i}$ be the set of retrieved documents such that 
-$\hour_i \le \dtime(\doc) < \hour_{i+1}$ for all $\doc \in \corpus_{\hour_i}$.
-
-%Our TS system involves ? phases involving sentence level classification, 
-%regression, and clustering. 
-Figure ? outlines our general temporal summarization algorithm. The description
-of our approach is as follows.
-For each event, we iterate over the retrieved 
-documents in hourly chunks, emitting 0 or more updates at each hour.
-At each hour $\hour$, we process each document
-$\doc \in \corpus_{\hour}$. First, we identify where the document content
-actually is; Second, we predict the salience of all content sentences.
-To account for redundancy, the predicted salience is penalized based on the 
-distance of the sentence in question to the previous summary updates.
-
-Finally, we cluster all content sentences 
-for the current hour using the penalized salience predictions to bias the 
-formation
-of clusters around the most salient sentences. For each cluster center, or 
-exemplar, that results, we check that the salience is above a threshold and 
-that it does not belong to a singleton cluster; exemplars that satisfy these
-conditions are emitted as an update.
-Additionally, we maintain the complete set of updates in order to penalize
-salience predictions in the subsequent time steps.
+Figure \ref{alg:temporal-summarization} outlines our general temporal summarization algorithm.
+At each hour, the system processes each document in the batch.
+It first discards any documents below a predefined relevance threshold (Section \ref{sec:filtering:documents}).
+It then discards any sentences from the remaining documents which are unlikely to carry content (Section \ref{sec:filtering:sentences}).
+Following this filtering process, the system must decide which of the remaining sentences should be considered an update.
+We refer to this as salience prediction and describe a feature-based model for predicting salience (Section \ref{subsec:Predict}).
+Because a single sub-event may produce several high-quality-but-redundant sentences within an hour, we employ a cluster-based method for selecting exemplars of high salience sub-events (Section \ref{sec:ap}).
+Finally, we employ a final deduplication phase to ensure that we only select novel updates (Section \ref{sec:novelty}).
 
@@
 \begin{algorithm}%[H]
- \KwData{$Query$ --- the set of event query words\\ 
+ \KwData{\\$Query$ --- the set of event query words\\ 
         $SC$ --- the stream corpus\\
@@
  } 
- \caption{Temporal Summarization Algorithm}
+ \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
 \end{algorithm}

### 2015-01-24 algorithm/salience.tex
edits to algo and problem definition mostly

@@
 \label{subsec:Predict}
-
-In order to use AP clustering for summarization, we need to assign a
-preference value to each input sentence.  In our approach, we equate a
-sentence's salience with its preference.  A good model of sentence salience
-should predict higher values for sentences that are more likely to appear in a
-human generated summary of the event.
-
-We do not have such human judgments, but we do have last year's gold nugget
-sentences.
-The response variable we try to predict is a sentence's semantic similarity
-(see ?) to the gold nugget sentences, i.e. we want to predict the similarity
-to a gold nugget when the gold nugget is not known.
-
-To build training data for this regression task, 
-we use the TREC 2013 event/gold nuggets. For each event, we retrieve all 
-sentences using the document retrieval and content selection steps outlined
-in the previous selection. We then sample with replacement 200 sentences
-from this collection, and extract non-lexical features (described in more 
-detail below) for each sentence
-to construct our design matrix $\mathbf{X}$. To build our response variables
-$\mathbf{y}$ we compute the maximum semantic similarity of each input
-sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with
-an RBF kernel to this data, optimizing kernel parameters with the scaled 
-conjugate gradient method. This sampling procedure is repeated 100 times
-for each of the ? 2013 TS events, yielding ? total models. In our salience 
-predictions for the 2014 events, we take the mean prediction of all models.
-
-%we take a subset of sentences
-%relevant to the TREC events (approximately 1000) and match them to the gold
-%nugget sentence with highest similarity as determined by the sentence
-%similarity system of \cite{guo2012simple}. 
-%\cite{} have used this system previously to correlate sentences to meaningful
-%units of information in human generated summaries. 
-%We use the real-valued similarity scores as our salience scores for the
-%training sentences.
-
-We want our model to be predictive across different kinds of events so we
-avoid lexical features.  Instead, we extract a variety of features including
-language model scores, geographic relevance, and temporal relevance from each
-sentence.  These features are used to fit a Gaussian process regression model
-that can predict the similarity of a sentence to a gold summary
-\cite{preotiuc2013temporal}.  
-%We use the model predicted salience of each
-%sentence as it's preference value in the AP clustering.
+Once the input has been filtered down to a higher quality set of candidate updates, we can rank sentences according to their salience.
+We adopt a model-based approach to predicting salience: a good model should predict higher values for sentences that are more likely to appear in a human generated summary of the event.
+The response variable we try to predict is a sentence's semantic similarity \cite{guo:wtmf} to the gold nugget sentences, i.e.
+we want to predict the similarity to a gold nugget when the gold nugget is not known.
+If sentence is very similar to a nugget, then it is a better candidate for emission as an update compared to a sentence that is not similar to any nuggets.
 
 \subsubsection{Features}
+We want our model to be predictive across different kinds of events so we avoid lexical features.
+Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
+These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
+\fdcomment{can we enumerate all of these?}
+
 \paragraph{Basic Features}
 
-We employ several basic features that have been used previously in supervised
-models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by
-sentence length, document position, number of named entities, etc.
+We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
+These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
 
@@
 
-These features are derived from the event query and the event type.
-These include the number of query words present in the sentence
-in addition to the number of event type synonyms, hypernyms, and hyponyms
-as found in WordNet \cite{miller1995wordnet}.
-E.g., for event type \emph{earthquake}, e.g., we obtain ``quake,'' ``temblor,''
-``seism,'' ``aftershock,'' etc.
+Query features measure the relationship between the sentence and the event query and type.
+These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
+For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
 
 \paragraph{Language Model Features}\label{subsubsec:lm}
-
-%Because the data in our experiments is scraped from the web, it is common to
-%find sentences that contain both salient informantion and two kinds of noise:
-%noisey fragments of web page structure (e.g. section titles, \emph{News},
-%\emph{Sports}, etc.) and references to other news not relevant to the topic
-%summary.
-%
-We use two trigram language models, trained using the SRILM toolkit
-\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each
-model.  This first model is trained on 4 years (2005-2009) of articles from
-the Gigaword corpus.  Specifically, we use articles from the Associated Press
-and the New York Times. This model is intended to assess the general writing
-quality (grammaticality, word usage) of an input sentence and helps us to
-filter out text snippets which are not sentences (e.g., web page titles).  The
-second model is a domain specific language model. We build a corpus of
-Wikipedia articles for each event type, consisting of documents from a related
-Wikipedia category. E.g. for earthquakes, we collect pages under the category
-\emph{Category:Earthquakes}. This model assigns higher probability to
-sentences that are focused on the given domain.
+Language models allow us to measure the likelihood of a sentence having been produced from a particular source.
+We consider two types of language model features.
+The first model is estimated from a corpus of generic news articles.
+This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).
+The second model is estimated from text specific to our event types.
+For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.
+These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g.
+most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
 
@@
 
-Locations are identified using a named entity tagger. For each location in a
-sentence, we obtain its latitude and longitude using the Google Maps API.  We
-then compute its distance to that of the event location.  It is possible for a
-sentence and an event to have multiple locations so we take as features the
-minimum, maximum, and average distance of all sentence-event location pairs.
+Locations are identified using a named entity tagger.
+For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.
+We then compute its distance to that of the event location.
+It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.
 Distances are calculated using the Vincenty distance.
@@
 
-Our data consists of hourly crawls of online content and so we exploit the
-temporality of corpus by capturing the burstiness of a sentence, i.e.  the
-change in word frequency from one hour to the next.``Bursty'' sentences often
-indicate new and important data.
+Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
+the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.
 
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,
-w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour
-burstiness of sentence $s$ from document $d$ at hour $t$  as \begin{align*}
+Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
+We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
+\begin{align*}
 \operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
@@
 
-We similarly find the sentence's 5-hour burstiness.  In addition to
-burstiness, we also include the sentence's average tf-idf and hours since the
-event in question started as features.
+We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.
 
 \subsubsection{Gaussian Process Regression}
+
 \fdcomment{transition to gp description}

### 2015-01-24 methods.tex
restructuring
edits to algo and problem definition mostly

@@
 \section{Methods}
-
+\label{sec:methods}
 \subsection{Data}\label{sec:data}
@@
 
+\subsubsection{Estimating Language Models}
+We use two trigram language models, trained using the SRILM toolkit \cite{stolcke2002srilm}, taking as features the average log probability (i.e.
+the sentence's total log probability normalized by sentence length) from each model.
+This first model is trained on 4 years (2005-2009) of articles from the Gigaword corpus.
+Specifically, we use articles from the Associated Press and the New York Times.
+The second model is a domain specific language model.
+We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category.
+E.g.
+for earthquakes, we collect pages under the category \emph{Category:Earthquakes}.
+This model assigns higher probability to sentences that are focused on the given domain.
+
+% For each event, we retrieve all sentences using the document retrieval and content selection steps outlined in the previous selection.
+We then sample with replacement 200 sentences from this collection, and extract non-lexical features for each sentence to construct our design matrix $\mathbf{X}$.
+To build our response variables $\mathbf{y}$ we compute the maximum semantic similarity of each input sentence to the gold nugget sentences.
+We fit a $\operatorname{GP}$ with an RBF kernel to this data, optimizing kernel parameters with the scaled  conjugate gradient method.
+This sampling procedure is repeated 100 times for each of the ?
+2013 TS events, yielding \fdcomment{?} total models.
+In our salience predictions for the 2014 events, we take the mean prediction of all models.
+
 \subsubsection{Redundancy Threshold}
@@
 \caption{Distribution of nugget similarity for matching and non-matching 
-updates.}
+updates.\fdcomment{possible to output this as a vector graphic (pdf pref).  it will scale better.}}
 \end{figure}

### 2015-01-24 paper.tex
restructuring
edits to algo and problem definition mostly

@@
 \usepackage{amsmath}
+\usepackage[inline]{enumitem}
 \usepackage{amsfonts}
 \usepackage[]{algorithm2e}
-\DeclareMathOperator{\corpus}{\mathcal{C}}
-\DeclareMathOperator{\doc}{\mathnormal{d}}
-\DeclareMathOperator{\sent}{\mathnormal{s}}
-\DeclareMathOperator{\order}{\pi}
-\DeclareMathOperator{\dtime}{\mathnormal{t}}
-\DeclareMathOperator{\hour}{\mathnormal{h}}
-\DeclareMathOperator{\hours}{\mathcal{H}}
-\DeclareMathOperator{\Sim}{\mathbf{K}}
-\DeclareMathOperator{\SMat}{\mathbf{X}}
-\DeclareMathOperator{\Pref}{\boldsymbol{\pi}}
-\DeclareMathOperator{\Exemp}{\mathnormal{Exemplars}}
-\DeclareMathOperator{\Updates}{\mathnormal{Updates}}
 \usepackage{cleveref}
@@
 \newcommand{\fdcomment}[1]{\textbf{\textcolor{red}{[FD: #1]}}}
-\newcommand{\ckcomment}[1]{\textbf{\textcolor{red}{[CK: #1]}}}
+\newcommand{\ckcomment}[1]{\textbf{\textcolor{blue}{[CK: #1]}}}
+\newcommand{\kmcomment}[1]{\textbf{\textcolor{green}{[KM: #1]}}}
+\input{notation.tex}
 
@@
 \maketitle \begin{abstract} 
-Information need is highest during times of crisis; both officials and 
-civilians need effective access to make informed decisions about their 
-safety and the safety of others. However, these situations pose many 
-interesting challenges for the IR systems that support these decisions.
-Specifically, we address event monitoring and updating, a.k.a. the
-temporal summarization task (TS). Given an event query, we monitor streams
-of online news and select important text to present to the user. Updates
-are selected by predicting the salience of the text w.r.t. to the event.
-Our model utilizes a variety of features at the query, document, and corpus
-level. Additionally, these predictions are used to directly bias a clustering
-algorithm for sentence selection, 
-increasing the novelty of the updates. We evaluate our system
-on a varied set of retrospective events using a combination of human 
-judgements
-and automatic evaluation measures, as compared to a baseline system.
-We demonstrate the effect of the various feature groups, and the importance
-of salience prediction w.r.t. update precision when compared to a clustering
-only baseline.
+During crises such as natural disasters or other human tragedies, information needs of both civilians and responders often require urgent, specialized treatment.
+Slow or ineffective information access can significantly impact the safety and health of individuals.
+Monitoring and summarizing a text stream during such an event remains a difficult problem.
+The temporal summarization task refers to extracting short, sentence-length updates about a seed event from a stream of text data.
+We present a system which makes substantial improvements over the state of the art in temporal summarization.
+Our approach predicts the salience of sentences with respect to an event and then uses these predictions to directly bias a clustering algorithm for sentence selection, increasing the novelty of the updates.
+We use novel, disaster-specific features for salience prediction, including geo-locations and language models representing the language typically used to describe different disaster types.
+We evaluate our system on a standard set of retrospective events using a combination of human judgements and automatic evaluation measures.
+We demonstrate the effect of different feature groups, and the importance of salience prediction for update precision.
 
@@
 
+\fdcomment{We should try to pull out as much of the TREC-specific language as possible (e.g. `run submissions', `participation in the TS track') to make the results more general.}
 \input{introduction.tex}
@@
 \input{motivation.tex}
+
+\input{problemdefinition.tex}
 

### 2015-01-24 problemdefinition.tex
edits to algo and problem definition mostly

@@
   \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
+  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
   \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$, presented in temporal order, and
@@
 \end{enumerate*}
-Throughout the time period of interest, the system must output sentences, known as \emph{updates}, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.
-We present the evaluation metrics which measure these properties in Section \ref{sec:methods}.
+While processing documents throughout the time period of interest, the system must output sentences, known as updates, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.
+Precisely how to measure these properties will be discussed in Section \ref{sec:methods}.
+In our version of this problem, we assume that that the system receives one batch of new sentence-segmented documents every hour throughout the period of interest.
 
-In the ideal version of this problem, the system makes decisions every time it consumes a document $\doc_i$.
-In practice, we assume that that we receive batches of new sentence-segmented documents every hour.
+\fdcomment{include description of nuggets here.}

### 2015-01-24 algorithm/novelty.tex
edits to algo and problem definition mostly
finished section 5

@@
-\subsection{Ensuring Novelty}
-We run AP clustering with our rescaled and penalized sentence similarities
-and preferences. When the clustering has converged we emit all exemplar
-sentences (cluster centers) whose preference is $\lambda$ standard deviations
-above the mean preference for the current time period and who do not belong
-to singleton clusters as updates. These updates and their preferences are 
-retained for penalty calculations in subsequent time steps.
+\subsubsection{Ensuring Novelty with Respect to Previous Updates}
+\label{sec:novelty}
+After we run AP clustering, we have a set of candidates for the current hour.
+In order to avoid outputting redundant updates, we computed the semantic similarity between the exemplar sentences and previously output sentences.
+We then removed those sentences which were above some similarity threshold.
+We selected the threshold by inspecting the distribution of similarities between sentences associated with the same sub-event in the training set and the distribution of similarities between random pairs of sentences.
+
+% and preferences. When the clustering has converged we emit all exemplar
+% sentences (cluster centers) whose preference is $\lambda$ standard deviations
+% above the mean preference for the current time period and who do not belong
+% to singleton clusters as updates. These updates and their preferences are
+% retained for penalty calculations in subsequent time steps.
+%

### 2015-01-24 introduction.tex
updated intro and motivation Not yet done with motivation

@@
 tracking, and summarizing events from text based input have been explored
-extensively  (e.g.,
+extensively
+%KM - I think we should consider adding other event summarizers. 
+(e.g.,
 \cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
-these experiments were not performed in the large and heterogeneous
-environment of the modern web; there is still a need for robust and scalable 
-methods for automatic summarization.
+these experiments were not developed to handle streaming data from  the large and heterogeneous
+environment of the modern web. Neither were they developed to address the
+information needs that arise during  crisis
+situations. 
+%KM - this last sentence sounds somewhat negative.
+%there is still a need for robust and scalable 
+%methods for automatic summarization.
+
+In this paper, we present an update summarization system to track disasters
+across time. Our system predicts sentence salience in the context of a
+large-scale event, such as a disaster,  and integrates these predictions into
+a clustering based multi-document summarization system. We train a regression
+model to predict sentence salience and use these predictions to bias the
+formation of sentence clusters around more salient regions in the input space
+using affinity propagation (AP) clustering.  AP uses the salience predictions
+as well as pairwise similarities among input sentences to identify
+\emph{exemplar} sentences, which we use as our summary output.  Our approach
+differs from other methods of summarization that compute salience by pairwise
+comparisons alone, ignoring features of importance that are intrinsic to the
+sentences themselves.
+
+In addition to the tight integration between clustering and salience
+prediction, our approach also exploits knowledge about disaster to determine
+salience. Thus, salience does not just represent importance within a set of
+documents; it also represents both how typical this sentence is of the input event
+type (i.e., disaster, hurricane, tornado) and whether it specifies information
+about this particular disaster. We use a set of language models, one for each
+disaster type, to measure typicality of the sentence for the current event type. We
+use a feature that measures distance of mentioned location from the center of
+the disaster to represent its likelihood of referring to the input disaster.
 
@@
 
-In this paper, we detail our participation in the second year of the TREC
-Temporal Summarization (TS) track. Broadly, TS systems are required to monitor
-a time ordered corpus for information about a specific event and ``emit'' 
-updates about that event. Under the current TREC formulation of this task,
-successful TS systems must synthesize techniques from information retrieval,
- extractive multi-document summarization, and update summarization.
-Specifically, we must select updates to emit based on three primary criterion:
-relevance, salience, and redundancy.
-
-As this was our first time participating in TS track, we focused mostly on 
-second and third criterion, relying on the TREC supplied resources for 
-document retrieval and filtering heuristics for relevance.
-Our system predicts sentence salience in the context of a
-large-scale event and integrates these predictions into
-a clustering based multi-document summarization system. We fit a Gaussian 
-Process (GP) regression model to predict sentence salience and use 
-these predictions to bias the
-formation of sentence clusters around more salient regions in the input space
-using affinity propagation (AP) clustering.  AP uses the salience predictions
-as well as pairwise similarities among input sentences to identify
-\emph{exemplar} sentences, which we use as our summary output.  Our approach
-differs from other methods of summarization that compute salience by pairwise
-comparisons alone, ignoring features of importance that are intrinsic to the
-sentences themselves.
-
 The remainder of the paper is organized as follows. We begin with a review of related work in the information retrieval and multi-document summarization literature.  In section
 ~\ref{sec:background} we give an overview of the semantic similarity,
+%KM - Chris can you update this so it reflects the contributions. We don't
+%mention semantic similarity. We do mention salience and the disaster specific
+%features. Perhaps we should mention the use of a new approach for semantic
+%similarity? 
 Gaussian process, and affinity propagation algorithms that make up the bulk of

### 2015-01-24 motivation.tex
updated intro and motivation Not yet done with motivation

@@
 \section{Motivation}
+
+%KM If we are short on space, not sure how much this section is needed as it's
+%somewhat repetitive of intro.
+
+%KM - Chris, could you get some good references to work in the crisis
+%informatics space.
+Crisis informatics\cite{?} is dedicated to finding methods for sharing the
+right information in a timely fashion with relief organizations during a major crisis. With the
+increasing impact of climate change, the world is seeing an increase in
+disasters such as hurricanes, tornadoes, flooding and typhoons, all of which
+cause major damage and wreak havoc with food supplies, housing, and health
+issues. Health epidemics, such as the ebola crisis in West Africa, also create
+a need for timely information about where problems are greatest. Social and
+political 
+crises, such as the current situation in Syria, create similar needs for
+humanitarian assistance.
+
+At times of crisis, however, social media can overwhelm current information
+systems with the quantity of information, much of which is irrelevant,
+unnecessary detail, or out of date. Many approaches have focused on
+
+%KM - PRobably should make a little more inclusive, there are graph based
+%models that use methods such lexical similarity to build the graph,
+%probabilistic methods (e.g., based on word probabilities, regression or
+%topical signatures). There are also methods that use topic segmentation or
+%modeling and select sentences for each topic.
 

### 2015-01-24 motivation.tex
updated intro and motivation. Both complete

@@
 unnecessary detail, or out of date. Many approaches have focused on
+human-in-the-loop approaches \cite{?} to enable people on the ground to update
+crisis interfaces with information about needs. Others have ..
+%KM - Chris - fill this out a bit.
 
@@
 
-Most multi-document summarization approaches involve either sentence 
-clustering or ranking or both to select summary sentences.
+Multi-document summarization offers potential for enabling automatic updates of
+relevant, salient information at regular intervals. It would provide
+information even when human volunteers are unable to and would filter out
+unnecessary and irrelevant detail. 
+Most generic multi-document summarization approaches involve either sentence 
+clustering or ranking sentences according to metrics that measure the salience
+of their words. 
 When the summarization task
-is broad or underspecified, clustering is most appropriate in that the data
-``speaks for itself.'' This is probably most true when summarization is 
-interpreted to mean exploratory data analysis. 
-When there are more constraints on the output, as in many text summarization
-tasks,
-it is often easier to design appropriate features for the ranking approach, 
-either by rule or learning-to-rank. The primary contribution of this paper
+is broad or underspecified,
+%clustering is most appropriate in that the data
+these methods are appropriate as they let the data
+``to speak for itself.''
+%This is probably most true when summarization is 
+%interpreted to mean exploratory data analysis. 
+When there are more constraints on the output, as in query-focused
+summarization or the crisis update task we present here,
+results are more accurate when
+%often easier to design
+appropriate features for the ranking approach are used (e.g., amount of word
+overlap for the query-focused approach).
+%either by rule or learning-to-rank.
+The primary contribution of this paper
 is a framework for combining both of these approaches using affinity
-propagation clustering. Additionally, we provide a task specific evaluation
-of this clustering model on the TREC Temporal Summarization task.
+propagation clustering as well as disaster-specific features.
 

### 2015-01-24 algorithm/ap.tex

@@
 \label{sec:ap}
-In the sentence selection stage, we use the salience predictions from our GP
-model as preferences in the AP clustering algorithm. The AP algorithm is 
-parameterized by a similarity matrix $\mathbf{S}$ and a vector of 
-preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
-parameters, and did not perform robustly on our range of inputs.
-In order to improve the quality of the clusters and exemplar selection,
-we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$. 
-The raw preferences are scaled to lie within the  range $(-3, -2)$
+Because we are analyzing multiple text streams, there is likely to be a large amount of redundancy of content.
+That is, when a sub-event occurs, several news providers will publish stories reporting the occurrence.
+Predicting the salience of each sentence in isolation, as was described in the previous section, provides us with a measure of the usefulness of a sentence independent of other decisions.
+In order to reduce the redundancy in our updates, we need to select sentences that  are both salient and novel.
 
-The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
-similarity between input sentences. Self-similarities and similarities below 
-a threshold $\lambda$ were masked and the remaining values scaled to the range
-$(-3, -1)$.
+We take a two phase approach to dealing with redundancy.
+First, we select a set of the diverse, highly-salient sentences within our set of candidates from the current hour.
+This ensures that we focus only on the most recent sub-events and, as a result, indirectly helps preserve low latency.
+Second, we compare those exemplars to update decisions made in previous hours.
+This ensures that we do not repeat sub-events previously output to the user.
 
-\subsubsection{Semantic Similarity}\label{subsec:semsim}
+\subsubsection{Selecting Exemplars within the Hour}
+\fdcomment{subsection title needs to be changed}
 
-Whenever we make a pairwise comparison between sentences, we use the weighted
-textual matrix factorization (WTMF) model of \cite{guo2012simple}. This 
-model can be thought of as a variant of latent semantic analysis (?), 
-where words that are not present in a sentence are explicitly modeled.
-More formally, we have a term-sentence matrix 
-$\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a 
-vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
-$j$ contains word $i$. In the WTMF regime, we want to find an approximation
-of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where 
-$\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
-$\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
-space. These matrices are found by minimizing the objective function
+Many existing multi-document summarization systems operate by clustering a set of previously-deemed salient sentences.
+Because our salience assessments are based on a model, we cannot assume that values are perfectly reliable.
+Therefore using a clustering technique that ignored the uncertainty in the salience prediction may have undesirable outcomes.
 
-$$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
-\mathbf{Q}_{\cdot,j} 
-- \mathbf{X}_{i,j})^2 
- + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
+To address the uncertain in salience predictions, we adopted an algorithm known as affinity propagation (AP) \cite{dueck2007non}.
+AP is a message passing algorithm that identifies both exemplar data points and assignments of each point to an exemplar.
+This is done iteratively by passing \emph{responsibility} and \emph{availability} messages between data points that quantify the fitness of one data point to represent another, and the fitness of a data point to be represented based on the choices of other data points respectively.
 
-where $\mathbf{W}_{i,j} = 
-\begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
-w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
-\end{cases}$
-and $\lambda$ is a hyperparameter controlling the regularization terms.
+AP is parameterized by an $\numcandidates\times \numcandidates$ similarity matrix $\candidateSimMat$ and an $\numcandidates\times 1$ preference vector $\preferences$, derived from our salience predictions.
+$\candidateSimMat$ is a real-valued matrix where $\candidateSimMatij{i}{j}$ is the \emph{semantic} similarity of the $i$-th data point to the $j$-th candidate sentence.
+Note that this matrix, whose construction is described below, is different from our feature-based kernel matrix $\kernelMatrix$, in that it measures the similarity of the content as opposed to the properties of the content.
+The vector $\pi$ is a real-valued vector where $\preferences$ expresses our preference that the $i$-th data point can serve as an exemplar a priori of other data points.
 
-The $w_m$ term is another model hyperparameter that is set to a small constant
-($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
-reconstruction error of missing terms (words that did not occur a sentence).
+AP has several useful properties that comport well with the temporal summarization problem.
+First, the number of clusters $k$ is not a hyper-parameter of the model.
+Since we will be running this algorithm every hour, the usual methods of hyperparameter search become infeasible.
+This is especially critical in our situation, where the number of sub-events is likely to change dramatically between hours.
+The number of clusters falls out of the algorithm organically--lower overall preference values will result in fewer clusters.
+Secondly, the arbitrary nature of the preferences allow us to incorporate a variety of signals for identifying the best exemplars, i.e.
+salience and redundancy signals.
+The preferences can be thought of as self-similarities (similarities that pairwise-comparison based algorithms would ignore) that we can exploit to  incorporate our prior beliefs about a data point.
+Finally, cluster exemplars are guaranteed to be actual data points.
+Many  clustering algorithms group data around mathematical objects (e.g., the mean) that are not necessarily observed in the data.
+The extractive nature of this TS task requires that we emit actual data points (i.e.
+sentences).
+We are able to sidestep  the additional requirement of selecting a most representative cluster member, as this is computed explicitly in the AP  algorithm.
 
-Given an unseen sentence $\hat{i}$ we can project its term vector into the
-latent sentence vector space with
+\fdcomment{probably need more content explaining AP.}
 
-$$
-\mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
-\mathbf{P}^T  + \lambda\mathbf{I} )^{-1} 
-\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
-$$
+% In the sentence selection stage, we use the salience predictions from our GP
+% model as preferences in the AP clustering algorithm. The AP algorithm is
+% parameterized by a similarity matrix $\mathbf{S}$ and a vector of
+% preferences $\boldsymbol{\pi}$; we found AP to be very sensitive to these
+% parameters, and did not perform robustly on our range of inputs.
+% In order to improve the quality of the clusters and exemplar selection,
+% we re-scaled both the raw inputs $\mathbf{S}$ and $\boldsymbol{\pi}$.
+% The raw preferences are scaled to lie within the  range $(-3, -2)$
+%
+% The initial matrix $\mathbf{S}$ is computed by finding the pairwise semantic
+% similarity between input sentences. Self-similarities and similarities below
+% a threshold $\lambda$ were masked and the remaining values scaled to the range
+% $(-3, -1)$.
 
-where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
-where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$ 
-depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
+% \subsubsubsection{Semantic Similarity}
+% \label{subsec:semsim}
+The AP algorithm requires two data structures: the semantic similarity matrix $\candidateSimMat$ and the preference vector $\preferences$.
+In order to construct $\candidateSimMat$, we use the weighted textual matrix factorization (WTMF) model of \cite{guo2012simple}.
+This  model can be thought of as a variant of latent semantic analysis \cite{deerwester:lsa},  where strings, sentences in our case, are projected into a lower dimensional space.
+We then use the cosine similarity of embedded vectors in this space to define entries in $\candidateSimMat$.
+The preference vector $\preferences$ is constructed by using salience scores rescaled to lie within the  range $(-3, -2)$.
+\fdcomment{not sure how much we have to justify this.}
 
-The WTMF model is used extensively throughout our TS system. When making any 
-pairwise comparison between sentence $i$ and $j$, we first construct
-their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
-$\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity 
-$\displaystyle \operatorname{cos-sim}
-(\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) = 
-\frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
-||\mathbf{Q}_{\cdot,j}||_2   }$.
+% More formally, we have a term-sentence matrix
+% $\mathbf{X}\in\mathcal{R}^{v \times n}$ representing $n$ sentences with a
+% vocabulary of $v$ words; $\mathbf{X}_{i,j}$ indicates is non-zero if sentence
+% $j$ contains word $i$. In the WTMF regime, we want to find an approximation
+% of $\mathbf{X} \approx \mathbf{P}^T\mathbf{Q}$, where
+% $\mathbf{P} \in \mathcal{R}^{k \times v}$ is a latent word vector space and
+% $\mathbf{Q} \in \mathcal{R}^{k \times n}$ is a latent sentence vector
+% space. These matrices are found by minimizing the objective function
+%
+% $$\sum_i^v \sum_j^n \mathbf{W}_{i,j}(\mathbf{P}_{\cdot,i}^T
+% \mathbf{Q}_{\cdot,j}
+% - \mathbf{X}_{i,j})^2
+%  + \lambda ||\mathbf{P}||_2^2 + \lambda ||\mathbf{Q}||_2^2$$
+%
+% where $\mathbf{W}_{i,j} =
+% \begin{cases} 1, & \textrm{if $\mathbf{X}_{i,j} \ne 0$ } \\
+% w_m, & \textrm{if $\mathbf{X}_{i,j} = 0$ }\\
+% \end{cases}$
+% and $\lambda$ is a hyperparameter controlling the regularization terms.
+%
+% The $w_m$ term is another model hyperparameter that is set to a small constant
+% ($\le .01$). The weight matrix $\mathbf{W}$ has the effect of discounting the
+% reconstruction error of missing terms (words that did not occur a sentence).
+%
+% Given an unseen sentence $\hat{i}$ we can project its term vector into the
+% latent sentence vector space with
+%
+% $$
+% \mathbf{Q}_{\cdot,\hat{i}} = (\mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})}
+% \mathbf{P}^T  + \lambda\mathbf{I} )^{-1}
+% \mathbf{P}\mathbf{\tilde{W}}^{(\hat{i})} \mathbf{X}_{\cdot, \hat{i}}
+% $$
+%
+% where $\mathbf{\tilde{W}}^{(\hat{i})}$ is an $v\times v$ diagonal matrix
+% where $\mathbf{\tilde{W}}^{(\hat{i})}_{j,j}$ is equal to $1$ or $w_m$
+% depending on whether or not the $j$-th term occurs in sentence $\hat{i}$.
+%
+%
+% The WTMF model is used extensively throughout our TS system. When making any
+% pairwise comparison between sentence $i$ and $j$, we first construct
+% their latent sentence vectors $\mathbf{Q}_{\cdot,i}$ and
+% $\mathbf{Q}_{\cdot,j}$ and then find the cosine similarity
+% $\displaystyle \operatorname{cos-sim}
+% (\mathbf{Q}_{\cdot,i}, \mathbf{Q}_{\cdot,j}) =
+% \frac{\mathbf{Q}_{\cdot,i}^T\mathbf{Q}_{\cdot,j}}{||\mathbf{Q}_{\cdot,i}||_2
+% ||\mathbf{Q}_{\cdot,j}||_2   }$.
+%
+%
+% Because the events for the TS task come from different domains, we construct
+% domain specific latent word vector spaces for each domain using in-domain
+% Wikipedia pages (see~\cref{sec:data} for more details).
 
-Because the events for the TS task come from different domains, we construct
-domain specific latent word vector spaces for each domain using in-domain 
-Wikipedia pages (see~\cref{sec:data} for more details).
-
-\subsubsection{Affinity Propagation}
-Affinity propagation (AP) is a message passing algorithm that identifies both
-exemplar data points and assignments of each point to an exemplar.  This is
-done iteratively by passing \emph{responsibility} and \emph{availability}
-messages between data points that quantify the fitness of one data point to
-represent another, and the fitness of a data point to be represented based on
-the choices of other data points respectively \cite{dueck2007non}.
-
-AP is parameterized by an $n\times n$ similarity matrix $S$ and an $n\times 1$
-preference vector $\pi$.  $S$ is a real-valued matrix where $S(i,j)$ is the
-similarity of the $i$-th data point to the $j$-th data point.  $S$ does not
-need to be symmetric.  $\pi$ is a real-valued vector where $\pi(i)$ expresses
-our preference that the $i$-th data point can serve as an exemplar a priori of
-other data points.
-
-AP has several useful properties that comport well with the TS track 
-requirements. First, the number of clusters $k$ is not a hyper-parameter
-of the model. Since we will be running this algorithm many times in per run,
-the usual methods of hyperparameter search become infeasible 
-(?). The number of clusters falls out of the algorithm
-organically-- lower overall preference values will result in fewer clusters.
-
-Secondly, the arbitrary nature of the preferences
-allow us to incorporate a variety of signals for
-identifying the best exemplars, i.e. salience and redundancy signals. 
-The preferences can be thought of as self-similarities (similarities that
-pairwise-comparison based algorithms would ignore) that we can exploit to 
-incorporate our prior beliefs about a data point.
-
-Finally, cluster exemplars are guaranteed to be actual data points. Many 
-clustering algorithms group data around mathematical objects (e.g., the
-mean) that are not necessarily observed in the data. The extractive nature of
-this TS task requires that we emit actual data points (i.e. sentences).
-We are able to sidestep 
-the additional requirement of selecting a most
-representative cluster member, as this is computed explicitly in the AP 
-algorithm.
-
-\subsubsection{Preferences}
-
-Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
-the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
-$\mathbf{U}$. We call this the redundancy penalty
-
-$$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
-{\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$   
-where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$ 
-previous update. We calculate our new penalized preferences 
-$\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
-Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie 
-within the range $(-3,-2)$.
+%
+% \subsubsection{Preferences}
+% The second data structure we need for AP is the preference vector $\preferences$.
+% Before rescaling, we first penalize each $\boldsymbol{\pi}_i$ based on
+% the aggregate similarity of input sentence $\mathbf{s}_i$ to the set of previous updates
+% $\mathbf{U}$. We call this the redundancy penalty
+%
+% $$\rho_i = \sum_j \frac{\boldsymbol{\pi}^{(U)}_{j}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_j)}
+% {\sum_{j^\prime}\operatorname{cos-sim}(\mathbf{s}_i, \mathbf{U}_{j^\prime})} $$
+% where $\boldsymbol{\pi}^{(U)}_{j}$ is the salience prediction of the $j$
+% previous update. We calculate our new penalized preferences
+% $\boldsymbol{\pi}^{(\rho)}$ where $\boldsymbol{\pi}^{(\rho)}_i = \boldsymbol{\pi}_i - \rho_i$.
+% Finally, we rescale $\boldsymbol{\pi}^{(\rho)}$ such that all values lie
+% within the range $(-3,-2)$.
+%

### 2015-01-24 algorithm/salience.tex

@@
 
-\subsubsection{Gaussian Process Regression}
+\subsubsection{Model}
+We adopt Gaussian process regression in order to predict sentence sentence \cite{rasmussen:gaussian-process-book}.
+Gaussian process regressors are a class of data-driven, non-parametric model generalizing the multi-variate Gaussian to the infinite dimensional setting.
+Gaussian processes are  general and are state of the art for many regression tasks.
+A full treatment of Gaussian process regression is beyond the scope of this article and can be found in standard textbooks.
 
-\fdcomment{transition to gp description}
+Being non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.
+In our experiments, we used a radial basis function (RBF) kernel.
+Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
+\begin{align*}
+	\kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
+\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
+\end{align*}
+where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
+The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
 
-A Gaussian process (GP) is a distribution over functions and is a 
-generalization of the multi-variate Gaussian to the infinite dimensional
-setting. That is, we use the observed data to define a distribution over 
-possible functions that generated this data, without having to explicitly 
-parameterize the function---in this sense GPs are considered 
-a non-parametric model.
-
-Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-of an input space $\mathcal{X}$ to the reals,
-
-$$f: \mathcal{X} \rightarrow \mathcal{R}.$$ 
-Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
- an $n$-dimensional vector whose elements are evaluations of the function $f$
-at points $x_i \in \mathcal{X}$.
-We say $p(f)$ is a Gaussian process if for any finite subset 
-$\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over 
-that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a 
-covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-set to 0, leaving the distribution to be completely characterized by the
-kernel function on the data.
-
-In the regression setting, we typically have a response variable $y$ that
-is the sum of our model prediction  and 
-some Gaussian noise, i.e. $y = f(x) + \epsilon$ with 
-$\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-$f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-two distributions
-of principal interest are the marginal likelihood
-$p(\mathbf{y}|\mathbf{X}) = 
-\mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-distribution,
-
-$$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-\mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-
-where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-response, and
-\begin{align*}
-\boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-\boldsymbol{\sigma}^2_* & 
-= \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-\mathbf{K}_*^T + \sigma^2\\
-\end{align*}.
-
-Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and 
-$\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-
-GP's are incredibly general, and are state of the art for many regression 
-tasks (?). The reliance on the covariance matrix 
-$\mathbf{K}$ for parameterization opens up the wide world of kernel methods
-for regression, and many varieties of similarity functions can be used.
-In our experiments we used a radial basis function kernel 
-$$K(\mathbf{x},\mathbf{x}^\prime) = \sigma^2 \exp\bigg(- \frac{1}{2} 
-\sum_{i=1}^d \frac{ (x_i-x^\prime_i)^2}{\ell_i^2} \bigg)$$ where 
-$\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not
-only improve the accuracy of the model, but give us some introspection 
-into which features are more important.
+% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
+% of an input space $\mathcal{X}$ to the reals,
+%
+% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
+% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
+%  an $n$-dimensional vector whose elements are evaluations of the function $f$
+% at points $x_i \in \mathcal{X}$.
+% We say $p(f)$ is a Gaussian process if for any finite subset
+% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
+% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
+% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
+% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
+% set to 0, leaving the distribution to be completely characterized by the
+% kernel function on the data.
+%
+% In the regression setting, we typically have a response variable $y$ that
+% is the sum of our model prediction  and
+% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
+% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
+% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
+% two distributions
+% of principal interest are the marginal likelihood
+% $p(\mathbf{y}|\mathbf{X}) =
+% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
+% distribution,
+%
+% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
+% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
+%
+% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
+% response, and
+% \begin{align*}
+% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
+% \boldsymbol{\sigma}^2_* &
+% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
+% \mathbf{K}_*^T + \sigma^2\\
+% \end{align*}.
+%
+% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
+% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
+%
+%

### 2015-01-24 algorithm/introduction.tex
Updated intro, motivation, intro to algorithms and methods

@@
-Figure \ref{alg:temporal-summarization} outlines our general temporal summarization algorithm.
-At each hour, the system processes each document in the batch.
-It first discards any documents below a predefined relevance threshold (Section \ref{sec:filtering:documents}).
-It then discards any sentences from the remaining documents which are unlikely to carry content (Section \ref{sec:filtering:sentences}).
-Following this filtering process, the system must decide which of the remaining sentences should be considered an update.
-We refer to this as salience prediction and describe a feature-based model for predicting salience (Section \ref{subsec:Predict}).
-Because a single sub-event may produce several high-quality-but-redundant sentences within an hour, we employ a cluster-based method for selecting exemplars of high salience sub-events (Section \ref{sec:ap}).
+Figure \ref{alg:temporal-summarization} outlines our general temporal
+summarization algorithm.  At each hour, the system processes each document in
+the batch. It first discards any documents below a predefined relevance
+threshold (Section \ref{sec:filtering:documents}).  It then discards any
+sentences from the remaining documents which are unlikely to carry content
+(Section \ref{sec:filtering:sentences}).  Following this filtering process, the
+system must decide which of the remaining sentences should be considered an
+update.  We refer to this as salience prediction and describe a feature-based
+model for predicting salience (Section \ref{subsec:Predict}).  Because a single
+sub-event may produce several high-quality but redundant sentences within an hour, we employ a cluster-based method for selecting exemplars of high salience sub-events (Section \ref{sec:ap}).
 Finally, we employ a final deduplication phase to ensure that we only select novel updates (Section \ref{sec:novelty}).

### 2015-01-24 methods.tex
Updated intro, motivation, intro to algorithms and methods

@@
 web crawls from October 2011
-through mid February 2013 \cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
-As per the TS specifications, all summary sentences come from processing this
-corpus in a time aligned manner. While the 2014 corpus encompasses the 
-2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
-year because sentence ID's used in evaluation changed from year to year.
+through mid February 2013
+\cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
+%KM - Keep neutral from Trec
+%As per the TS specifications,
+All summary sentences come from processing this
+corpus in a time aligned manner.
+%KM - Are you using both 2013 and 2014 corpora.
+%KM - Mentioning sentence ids at this point is too low level.
+%KM - I would probably just drop this sentence. since you say above that you
+%use both.
+%While the 2014 corpus encompasses the 
+%2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
+%year because sentence ID's used in evaluation changed from year to year.
 
+%KM - As far as I can see this is the first time you've mentioned this.
+%KM - This paragraph is not clearly understandable. Is this the training data? 
 For our semantic similarity component (\cref{subsec:semsim}),
@@
 
-We use the same Wikipedia corpora to construct the lanuage models used
+We use the same Wikipedia corpora to construct the language models used
 in our salience prediction model (\cref{subsubsec:lm}).
@@
 
-We use the event set developed for the TREC Temporal Summarization track 
-for years 2013 and 2014. There were 10 and 15 events respectively. We omit
+%KM - You said this already so I think you can make it simpler. 
+%We use the event set developed for the TREC Temporal Summarization track 
+%for years 2013 and 2014.
+The TREC 2013 and 2014 Stream corpus contains  10 and 15 events respectively. We omit
 one event from the 2013 track because there were no news documents in the 
@@
 the type of event, and the 
-duration of the event (the duration for the system 
-to monitor not necessarily the complete timespan of the event). 
-All data except the Wikipedia name is exposed to the TS system to use.
+duration of the event
+%KM - the below parens is not clear. is ``to monitor'' missing from the above
+%sentence? I think get rid of the parens below.
+%(the duration for the system 
+%to monitor is not necessarily the complete timespan of the event).
+%KM - Let's get rid of all references to ``the TS system'' If you want to have
+%a system name, choose a better one. 
+All data except the Wikipedia name is exposed to
+systems that participate in TREC.
+%the TS system to use.
 
@@
 referred to as nuggets, were extracted from each event's Wikipedia page.
-Assessors graded nuggets by importance, and were also able to timestamp their
-existance using the timestamp of the page's revision history. Nuggets are 
+%KM - What assessors? Was this done by TREC? Make clear. I added ``TREC'' after
+%reading ahead.
+%KM - What do you mean ``graded''? Do you mean they graded your nuggets? Or do
+%you mean they ``ranked'' the nuggets by importance. Please make clear.
+TREC assessors graded nuggets by importance, and were also able to timestamp
+%KM - wrong wording. I've changed.
+%their
+%existance
+when the event represented by each nugget occurred 
+using the timestamp of the page's revision history. Nuggets are 
 usually a short to medium length clause containing an atomic piece of
@@
 
-\subsubsection{Matches}
+%KM - I changed title to make clearer
+\subsubsection{Gold Standard Matches}
 TREC assessors provided judgements on a subset of participant updates each year
@@
 sentence often contains more than one atomic piece of information) or marked
-not relevant. There was not much overlap in the participant update sets so
-overall coverage of relevant sentences in the corpus is small (percent?).
-To expand the judgement corpora, we include all system updates as positive
+not relevant.
+
+%KM - I don't see why overlap in particpant update sets would give you low
+%coverage. In fact, it might you give you broader coverage than if all
+%participants extracted the same sentences.  I removed first portion of the
+%sentence, but you can put back and clarify if you think I interpreted
+%incorrectly. 
+%There was not much overlap in the participant update sets so
+Overall coverage of
+%KM - Shouldn't this be nuggets?
+relevant sentences
+relevant nuggets in the corpus  by participant systems' generated sentences
+is small (percent?).
+%KM - This is complicated. More explanation is needed.
+%KM - The sentence below sounds like how you evaluate your system and not like
+%the judgment corpus.
+%KM not sure this owuld go here. Shouldn't it go in evaluation? Here you would
+%just talk about the corpus
+To evaluate new system output on the same event, we compare it against the
+corpus of matched sentences. Comparing it against the nuggets would not yield
+meaningful results as the nuggets are much smaller than the sentences of the
+corpus.
+Because there are so few matched sentences, we score all new system updates as positive
+%To expand the judgement corpora, we include all system updates as positive
 nugget matches if they are $< .2$ close in Levenshtein distance to a human
@@
 
+%KM - Got this far.
 \subsection{Experiment Design}

### 2015-01-25 methods.tex
finished review of methods

@@
 %KM - This paragraph is not clearly understandable. Is this the training data?
+%KM - Also confusing here is that you alternate between what TREC provides and
+%what you do. Perhaps use different sections for TREC and for you.
 For our semantic similarity component (\cref{subsec:semsim}),
@@
 the type of event, and the 
-duration of the event
+duration of the event.
 %KM - the below parens is not clear. is ``to monitor'' missing from the above
@@
 %existance
-when the event represented by each nugget occurred 
+the occurrence time of the events represented by each nugget
 using the timestamp of the page's revision history. Nuggets are 
@@
 information.
+%KM - I think an example of a nugget here could help.
 
@@
 %KM - Shouldn't this be nuggets?
-relevant sentences
+%relevant sentences
 relevant nuggets in the corpus  by participant systems' generated sentences
@@
 corpus.
+%KM - Let's discuss where these different things go. I think scoring should
+%probably go in the evaluation section.
 Because there are so few matched sentences, we score all new system updates as positive
@@
 
-%KM - Got this far.
 \subsection{Experiment Design}
@@
 over the duration of the event 
-(as determined by 
-TREC; avg. run length is ? days) of the event. Nugget similarities
+as determined by 
+TREC; avgerage run length is ? days.
+%KM - Should you say here using the sentence similarity software? 
+Nugget similarities
 are computed for each sentence and are standardized with respect to 
-the sample mean and variance. After standardization, the maximum nugget similarities are computed for each sentence. 
+the sample mean and variance. After standardization, the maximum nugget
+similarities are computed for each sentence.
+%KM - I think it would be helpful to mention that you use the features below
+%here. 
 We fit a Gaussian process for each event, using the maximum similarities 
@@
 parameters are also individually fit for each model.
+%KM - Do you mean for new events? Would help if you say so. 
 When making event-specific predictions, we take the average prediction of the
@@
 \subsubsection{Estimating Language Models}
-We use two trigram language models, trained using the SRILM toolkit \cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each model.
-This first model is trained on 4 years (2005-2009) of articles from the Gigaword corpus.
-Specifically, we use articles from the Associated Press and the New York Times.
-The second model is a domain specific language model.
+We use two trigram language models, trained using the SRILM toolkit
+\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
+the sentence's total log probability normalized by sentence length) from each
+model.  This first model is trained on 4 years (2005-2009) of articles from the
+Gigaword corpus.  Specifically, we use articles from the Associated Press and
+the New York Times. The second model is a domain specific language model.
+%KM - Remind us how many event types there are. Do you list all at some point?
 We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category.
@@
 used the point of intersection as the value of $t$.
+%KM _ I think you need to say here what sentence similarity method you use.
+%Just saying ``cosine similarity'' implies word overlap.
 Figure ? shows a normalized histogram of cosine similarities for both 
@@
 \end{figure}
+
+%KM _ What happened to the geo-codes? Aren't they in the salience models?
 \subsubsection{Effect of Salience Models}
@@
 
-We are also interested
-in the effect of directly incorporating salience in the clustering algorithm.
-A third run (HAC) is performed using hierarchical agglomerative 
-clustering; we produce flattened clusters by setting a maximum cluster 
-distance threshold. The sentence with the highest predicted salience from each
- cluster is selected as an update. 
-The distance threshold was tuned using grid search. By comparing
-HAV with AP-base we can empirically assess the trade off between preference
-and neighbor similarity that affinity propagation is making.
+%KM - I think this should be rewritten to say you compare to a usual
+%summarization technique, using hierarchical agglomerative clustering with
+%selection of sentences at the centroid of the cluster.
+%We are also interested
+%in the effect of directly incorporating salience in the clustering algorithm.
+%A third run (HAC) is performed using hierarchical agglomerative 
+%clustering; we produce flattened clusters by setting a maximum cluster 
+%distance threshold. The sentence with the highest predicted salience from each
+% cluster is selected as an update. 
+%The distance threshold was tuned using grid search. By comparing
+%HAV with AP-base we can empirically assess the trade off between preference
+%and neighbor similarity that affinity propagation is making.
+%KM - I now see why you were running experiments this way. It would be an
+%interesting test.
+
+%KM - Here is a paragraph you could use. I commented out above but you can put
+%back in if there are pieces you prefer. Of course, edit below any way you see
+%fit. Probably cite Radev and Jing here. They did this fairly early. There may
+%be others who did this later.
+
+We also compared our system against a typical summarization approach using
+hierarchical agglomerative clustering (HAC); we produce flattened clusters by setting a maximum cluster 
+distance threshold. We  select the centroid of each
+cluster as a summary sentence  \cite{??}.
 
@@
 
-We remove each feature group and train models models on these small faller 
-feature susbstets. All other parameters are the same as in our full system 
+%KM - Probably should describe this as either backward search over feature
+%groups. There's a word for this. I will find it. Also should start with the
+%purpose of doing it. 
+We remove each feature group and train models models on these small 
+feature subsets. All other parameters are the same as in our full system 
 (AP-sal).

### 2015-01-25 algorithm/ap.tex
Read and commented on algorithm sections

@@
 
-To address the uncertain in salience predictions, we adopted an algorithm known as affinity propagation (AP) \cite{dueck2007non}.
+To address the uncertainty in salience predictions, we adopted an algorithm known as affinity propagation (AP) \cite{dueck2007non}.
 AP is a message passing algorithm that identifies both exemplar data points and assignments of each point to an exemplar.
@@
 % \label{subsec:semsim}
-The AP algorithm requires two data structures: the semantic similarity matrix $\candidateSimMat$ and the preference vector $\preferences$.
-In order to construct $\candidateSimMat$, we use the weighted textual matrix factorization (WTMF) model of \cite{guo2012simple}.
-This  model can be thought of as a variant of latent semantic analysis \cite{deerwester:lsa},  where strings, sentences in our case, are projected into a lower dimensional space.
-We then use the cosine similarity of embedded vectors in this space to define entries in $\candidateSimMat$.
-The preference vector $\preferences$ is constructed by using salience scores rescaled to lie within the  range $(-3, -2)$.
-\fdcomment{not sure how much we have to justify this.}
+The AP algorithm requires two data structures: the semantic similarity matrix
+$\candidateSimMat$ and the preference vector $\preferences$.  In order to
+construct $\candidateSimMat$, we use the weighted textual matrix factorization
+(WTMF) model of \cite{guo2012simple}. This  model can be thought of as a
+variant of latent semantic analysis \cite{deerwester:lsa},  where strings,
+sentences in our case, are projected into a lower dimensional space.  We then
+use the cosine similarity of embedded vectors in this space to define entries
+in $\candidateSimMat$.
+%KM - Include advantage. Tried below
+This approach allows us to measure semantic similarity as opposed to simple
+word overlap.
+The preference vector $\preferences$ is constructed by using salience scores rescaled to lie within the  range $(-3, -2)$.  \fdcomment{not sure how much we have to justify this.}
 

### 2015-01-25 algorithm/novelty.tex
Read and commented on algorithm sections

@@
 \label{sec:novelty}
-After we run AP clustering, we have a set of candidates for the current hour.
-In order to avoid outputting redundant updates, we computed the semantic similarity between the exemplar sentences and previously output sentences.
+After we run AP clustering, we have a set of candidates for the current hour.  In order to avoid
+%KM - word change
+%outputting
+producing
+redundant updates, we computed the semantic similarity,
+%KM - You could give the semantic similarity algorithm a name.
+THat woudl make it easier to refer to.
+In any case, you need to specify the method you use here.
+I added clause below.
+Coudl remove it if you name the algorithm.
+using the same method as used in salience prediction,
+between the exemplar sentences and previously output sentences.
 We then removed those sentences which were above some similarity threshold.

### 2015-01-27 algorithm/salience.tex
Added features figure, event figure, and updated wikipedia lm figure

@@
 If sentence is very similar to a nugget, then it is a better candidate for emission as an update compared to a sentence that is not similar to any nuggets.
+
+\begin{figure}[t!]
+\begin{tabular}{| l |} 
+\hline
+\textbf{Basic Features}\\
+$\cdot$ Sent. position (normalized by doc. length) \\
+$\cdot$ Sent. length \\
+$\cdot$ Ratio of punc. to non-punc. chars. \\
+$\cdot$ Ratio of caps. to non-caps. chars. \\
+$\cdot$ Ratio of lowercase to other chars. \\
+$\cdot$ Ratio of uppercase to other chars. \\
+$\cdot$ \# of caps. words (normalized by \# of words)\\
+$\cdot$ \# of Person, Location, Org. Date, Number,\\
+$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. tags \\
+$\;\;$ (normalized by \# of words)\\
+\hline
+\textbf{Query Features}\\
+$\cdot$ \% of query words covered by sent.\\
+$\cdot$ Total query matches.\\
+$\cdot$ Total event-type synonyms/hypernyms/hyponyms\\
+$\;\;$ coverage.\\
+$\cdot$ Total event-type synonyms/hypernyms/hyponyms\\
+$\;\;$ matches.\\
+\hline
+\textbf{Language Model Features}\\
+$\cdot$ Avg. token log probability (domain lang. model)\\
+$\cdot$ Avg. token log probability (background lang. model)\\
+\hline
+\textbf{Geo-tag Features}\\
+$\cdot$ Median document distance to nearest location \\
+$\;\;$ cluster (current hour).\\
+$\cdot$ Median document distance to nearest location \\
+$\;\;$ clusters (previous hour).\\
+$\cdot$ Distance of first location in doc. to nearest location \\
+$\;\;$ cluster (current hour).\\
+$\cdot$ Distance of first location in doc. to nearest location \\
+$\;\;$ cluster (previous hour).\\
+\hline
+\textbf{Temporal Features}\\
+$\cdot$ Avg. tf-idf at current time.\\
+$\cdot$ Change in avg. tf-idf since previous hour (up to\\
+$\;\;$  24 hours).\\
+$\cdot$ Time since query/event start.\\
+\hline
+\end{tabular}
+\caption{Salience model features.}
+\end{figure}
 

### 2015-01-27 methods.tex
Added features figure, event figure, and updated wikipedia lm figure

@@
 \hline \hline
+2012 Aurora shooting \newline
+Wisconsin Sikh temple shooting\newline 
+2012 Tel Aviv bus bombing \newline
 Boston\_Marathon\_bombings \newline
@@
 \hline
+Hurricane Isaac (2012) \newline
+Hurricane Sandy \newline
+Typhoon Bopha \newline
+2012 Guatemala earthquake \newline
  Cyclone\_Oswald \newline
 Early\_2012\_European\_cold\_wave \newline
-February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
+February\_2013\_noreaster \newline & Natural Disasters & 35,554/591,850/12,794,438  \\
 \hline
+2012 Buenos Aires Rail Disaster
+2012 Pakistan garment factory fires
 Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
 \hline
-Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
+Chelyabinsk\_meteor & Astronomy & 14,515/283,509/6,135,803  \\
 \hline

### 2015-01-27 algorithm/salience.tex
Added comments and made a few wording changes to salience

@@
 \paragraph{Basic Features}
+%KM - Would be good to have quick justification of these features. I added a
+%sentence. Feel free to edit or remove.
 
@@
 These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
+Since training is one on grammatical English, some of these features help
+to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
+Others help to more heavily weight important sentences (e.g., that appear in
+prominent positions such as paragraph initial or article initial).
 
@@
 For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
-
 \paragraph{Language Model Features}\label{subsubsec:lm}
@@
 most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
+%KM - Note: Someplace the exact list of event types should appear. Probably not
+%here.
+%KM - I note you have it in a later section but it is labeled as data you use
+%to train language models and semantic similarity. I think it would be good to
+%have up front in definition of task.
 
@@
 Distances are calculated using the Vincenty distance.
+%KM - Probably should say how you determine event location. Some events move.
+%KM - In your figure you only include one distance.
 
@@
 \subsubsection{Model}
-We adopt Gaussian process regression in order to predict sentence sentence \cite{rasmussen:gaussian-process-book}.
-Gaussian process regressors are a class of data-driven, non-parametric model generalizing the multi-variate Gaussian to the infinite dimensional setting.
-Gaussian processes are  general and are state of the art for many regression tasks.
-A full treatment of Gaussian process regression is beyond the scope of this article and can be found in standard textbooks.
+%KM - I think one of the words ``sentence'' below should be something else, btu
+%not sure what you meant. Ah.. I think ``salience''?
+%We adopt Gaussian process regression in order to predict sentence sentence
+We adopt Gaussian process regression in order to predict sentence salience
+\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
+class of data-driven, non-parametric model generalizing the multi-variate
+Gaussian to the infinite dimensional setting.  Gaussian processes are general
+and are state of the art for many regression tasks.  A full treatment of
+Gaussian process regression is beyond the scope of this article and can be
+found in standard textbooks.
 

### 2015-01-27 problemdefinition.tex
Added event types to problemdefinition

@@
 \end{enumerate*}
-While processing documents throughout the time period of interest, the system must output sentences, known as updates, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.
-Precisely how to measure these properties will be discussed in Section \ref{sec:methods}.
-In our version of this problem, we assume that that the system receives one batch of new sentence-segmented documents every hour throughout the period of interest.
+While processing documents throughout the time period of interest, the system
+must output sentences, known as updates, which are \emph{relevant} to the
+query, \emph{comprehensive} with respect to the event, \emph{novel} to the
+user, and \emph{timely} with respect to when the update occurred.  Precisely
+how to measure these properties will be discussed in Section \ref{sec:methods}.
+In our version of this problem, we assume that that the system receives one
+batch of new sentence-segmented documents every hour throughout the period of
+interest.
+
+We address  disaster types such as terrorism and mass shootings (e.g.,
+the 2012 shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy),
+accidents (e.g., the 2012 Pakistan garment factory fire), astronomical
+disasters (e.g,. the Chelyabinsk comet in Russia), and social activism (e.g.,
+the Arab spring).
+\kmcomment{This was as best as I could do on this. These are the example
+  categories from the wikipedia page. However you refer in the language model
+  description to the event type earthquake which is more specific. Also, in the
+  TREC data you refer to event types that are specified in the metadata. So I'm
+  wondering how many different categories you have and where they come from.}
 
 \fdcomment{include description of nuggets here.}
+\kmcomment{We should jointly discuss what goes in data and what goes in problem
+  definition. I have moved event types here because I think knowing how many
+  types and examples of them would be helpful upfront. I think where nuggets
+  and wikipedia pages go is questionable.}

### 2015-02-22 introduction.tex
updated intro and motivation. Both complete
Moving to acl style

@@
-\section{Introduction}
 \label{sec:introduction}
-During crises, information is critical for first responders and those caught
-in the event.  When the event is significant, as in the case of Hurricane
-Sandy, the amount of information produced by traditional news outlets,
+During crises, information is critical for first responders,
+ crisis management organizations, and those caught in the event.  
+When the event is significant, as in the case of Hurricane
+Sandy, the amount of content produced by traditional news outlets,
 government agencies, relief organizations, and social media can vastly
-overwhelm those trying to monitor the situation. Methods for identifying,
+overwhelm those trying to monitor the situation. 
+Crisis informatics\cite{?} is dedicated to finding methods for sharing the
+right information in a timely fashion with relief organizations during a major crisis. 
+Research in this field has focused on
+human-in-the-loop approaches ranging from on the ground information gathering to
+crowdsourced reporting and disaster management.
+
+Multi-document summarization offers potential for enabling automatic updates of
+relevant, salient information at regular intervals. It would provide
+information even when human volunteers are unable to and would filter out
+unnecessary and irrelevant detail. 
+While methods for identifying,
 tracking, and summarizing events from text based input have been explored
@@
 (e.g.,
-\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
-these experiments were not developed to handle streaming data from  the large and heterogeneous
-environment of the modern web. Neither were they developed to address the
-information needs that arise during  crisis
-situations. 
+\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
+these experiments were not developed to handle streaming data from the large
+and heterogeneous environment of the modern web. 
+These methods also rely heavily on redundancy which is suboptimal for time
+sensitive domains where there is a high cost in delaying information.
+
+%\cite{?} to enable people on the ground to update
+%crisis interfaces with information about needs. Others have used crowd
+%sourcing \cite{?},
+
+%At times of crisis, however, social media can overwhelm current information
+%systems with the quantity of information, much of which is irrelevant,
+%unnecessary detail, or out of date.
+
+%Neither were they developed to address the
+%information needs that arise during  crisis
+%situations. 
 %KM - this last sentence sounds somewhat negative.
@@
 
-In this paper, we present an update summarization system to track disasters
+In this paper, we present an update summarization system to track events
 across time. Our system predicts sentence salience in the context of a
 large-scale event, such as a disaster, and integrates these predictions into
-a clustering based multi-document summarization system. We train a regression
-model to predict sentence salience and use these predictions to bias the
-formation of sentence clusters around more salient regions in the input space
-using affinity propagation (AP) clustering.  AP uses the salience predictions
-as well as pairwise similarities among input sentences to identify
-\emph{exemplar} sentences, which we use as our summary output.  Our approach
-differs from other methods of summarization that compute salience by pairwise
-comparisons alone, ignoring features of importance that are intrinsic to the
-sentences themselves.
+a clustering based multi-document summarization system. 
+Our system is better able to adapt to
+dynamic changes in input volume which adversely 
+effect methods of summarization that use either pairwise
+comparisons as a proxy for salience or predict salience but
+ignore redundancy in the input alone.
+
+%We train a regression
+%model to predict sentence salience and use these predictions to bias the
+%formation of sentence clusters around more salient regions in the input space
+%using affinity propagation (AP) clustering.  We adapt AP to incorporate
+%the salience predictions
+%as well as pairwise similarities among input sentences to identify
+%\emph{exemplar} sentences, which we use as our summary output.
+
+%ignoring features of importance that are intrinsic to the
+%sentences themselves.
+
+%KM - Chris - add a short overview on the kinds of things we measure and
+%results so people are encouraged to read ahead. Keep it general.
 
@@
 salience. Thus, salience does not just represent importance within a set of
-documents; it also represents both how typical this sentence is of the input event
+documents; it also represents both how typical this sentence is of the input 
+event
 type (i.e., disaster, hurricane, tornado) and whether it specifies information
-about this particular disaster. We use a set of language models, one for each
-disaster type, to measure typicality of the sentence for the current event type. We
-use a feature that measures distance of mentioned location from the center of
-the disaster to represent its likelihood of referring to the input disaster.
+about this particular disaster. 
+Our feature representation includes a set of language models, one for each
+disaster type, to measure typicality of the sentence for the current event 
+type, the distance of mentioned locations from the center of
+the disaster, and the change in word frequencies over the time of the event.
+% to represent its likelihood of referring to the input disaster. 
+%We also utilize the change in word frequencies as we process the stream 
+%to measure the changing importance of updates over time.
 
-%Previous work on generating event descriptions and/or multi-document
-%summarization has relied on clustering algorithms to find representative
-%sentences appropriate for an event summary.  These methods impose a metric
-%space on the text data that can make it difficult to incorporate external
-%sources of information elegantly -- in this paper we argue that centroid
-%sentences are not a priori the best candidates for inclusion in an event
-%summary.
+Our approach achieves suprerior ROUGE scores compared to multiple baselines.
+Additionally, we introduce a novel method for determining the 
+coverage of atomic facts in a summary, and show our system's
+superiority on this metric.
 
-The remainder of the paper is organized as follows. We begin with a review of related work in the information retrieval and multi-document summarization literature.  In section
-~\ref{sec:background} we give an overview of the semantic similarity,
-%KM - Chris can you update this so it reflects the contributions. We don't
-%mention semantic similarity. We do mention salience and the disaster specific
-%features. Perhaps we should mention the use of a new approach for semantic
-%similarity? 
-Gaussian process, and affinity propagation algorithms that make up the bulk of
-our TS system. 
-Next we describe the data with which we build our various models.
-Then in section~\ref{sec:approach} we give a high level sketch
-of our system, and then explain each component in detail.
+The remainder of the paper is organized as follows.
+We begin with a review of related work
+in the information retrieval and multi-document
+summarization literature. (Rest of outline goes here...)
+
+%Our main contributions are three fold. First, we present a straightforward 
+%framework for 
+%combining prior beliefs with a clustering algorithm. Second, we demonstrate
+%the empirical effectiveness of this approach on the update summarization
+%task. Finally, we present a novel feature represention for sentence salience
+%in the crisis/disaster domain.

### 2015-02-22 motivation.tex
Added features figure, event figure, and updated wikipedia lm figure
Moving to acl style

@@
-\section{Motivation}
-
 %KM If we are short on space, not sure how much this section is needed as it's
@@
 unnecessary and irrelevant detail.
-Most generic multi-document summarization approaches involve either sentence 
-clustering or ranking sentences according to metrics that measure the salience
-of their words. 
-When the summarization task
-is broad or underspecified,
-%clustering is most appropriate in that the data
-these methods are appropriate as they let the data
-``to speak for itself.''
-%This is probably most true when summarization is 
-%interpreted to mean exploratory data analysis. 
-When there are more constraints on the output, as in query-focused
-summarization or the crisis update task we present here,
-results are more accurate when
-%often easier to design
-appropriate features for the ranking approach are used (e.g., amount of word
-overlap for the query-focused approach).
-%either by rule or learning-to-rank.
-The primary contribution of this paper
-is a framework for combining both of these approaches using affinity
-propagation clustering as well as disaster-specific features.
 
-\ckcomment{Not sure how to present this together, we also essentially evaluate
-a salience model by feature ablation for the TS task, but this feels very specific compared to the general framework idea.}
+We conceptualize the space of existing summarization approaches as falling into
+one of three categories, each with specific tradeoffs with respect to update 
+summarization. First, centrality-focused approaches (including graph,
+cluster, and centroid methods) are very natural for retrospective analysis in
+the sense that they let the data "speak for itself." However, they rely
+chiefly on
+redundancy. When applied to an unfolding event, there may not exist enough
+redundant content at the event onset for these methods to exploit.
+Once the event onset has passed, however, the redundancy reduction of these 
+methods is quite beneficial.
 
-\fdcomment{I guess I think we need to two things in this section (and/or introduction).  1. explain why the problem matters.  2. explain why our approach is compelling.  }
+The second category, predictive approaches,
+includes ranking and classification based methods.
+These methods do not control for redundancy as well as the first group, 
+and will most likely be overwhelmed by web-scale summarization.
+On the other hand, these methods are not dependent on redundancy.
 
-\fdcomment{if easy to generate, would be good to have a single figure for an event that includes 1.
-the query, 2.
-the distribution of (filtered) document volume over time, and 3.
-a sample of the gold nuggets.
-this provides the reader with a visual explanation of the problem, the solution, and the difficulty}
+The last category includes probabilistic, information theoritic, and set cover
+approaches. While these methods are generally focused on producing diverse
+summaries, they are difficult to adapt to the streaming setting, where 
+we do not necessarily have a fixed summary length and the corpus to be
+summarized contains many irrelevent sentences, i.e. there are large
+portions of the corpora that we specificly want to avoid.
+
+Our main contribution is a framework for combining the strengths of the 
+centrality-based methods and predictive methods using salience predictions (we 
+prefer the term salience over rank) to bias cluster formation around high 
+salience sentences.% by combining regression with AP clustering.
+By algorithmically making the trade off between redundancy
+signals and salience signals, we mitigate the effect of 
+different input sizes. 
+When input is small, individual characteristics
+of the input determine centrality; when input is large, neighbor similarity 
+becomes more important.
+
+Second, we validate this method with the application to the update
+summarization
+task. We experiment on events of dramatically different scope and duration,
+and demonstrate improved average performance on ROUGE. Even within events
+the document streams can have widely varying levels of volume, corresponding
+to the day/night cycle (see fig. ?? for an example).
+
+Finally, we present an analysis of our feature representation for salience
+disaster domain. While we focus on a particular domain, the feature templates
+themselves are generally applicable to streaming/update summarization.
+
+%Most generic multi-document summarization approaches involve either sentence 
+%clustering or ranking sentences according to metrics that measure the salience
+%of their words. 
+%When the summarization task
+%is broad or underspecified,
+%these methods are appropriate as they let the data
+%``to speak for itself.''
+%When there are more constraints on the output, as in query-focused
+%summarization or the crisis update task we present here,
+%results are more accurate when
+%appropriate features for the ranking approach are used (e.g., amount of word
+%overlap for the query-focused approach).
+%
+%The primary contribution of this paper
+%is a framework for combining both of these approaches using affinity
+%propagation clustering as well as disaster-specific features.
+
+%\begin{figure}
+%\begin{tabular}{| p{8cm} |}
+%\hline
+%\textbf{Event: Hurricane Sandy}\\
+%\includesvg[width=\columnwidth]{docfreq}\\
+%\textbf{Example Nuggets:}\\
+%$\cdot$ October 22nd Sandy strengthened from a tropical depression into a tropical storm\\
+%$\cdot$ could turn towards Bermuda or go straight to the Eastern U.S.\\
+%$\cdot$ 11 pm Oct 24 winds 50 mph\\
+%%$\cdot$ 11 pm Oct 24 moving NNE at 9 knots\\
+%$\cdot$ More than 1,000 people went to shelters in Jamaica\\
+%\hline
+%\end{tabular}
+%\end{figure}

### 2015-02-22 problemdefinition.tex
Moving to acl style

--- 2015-01-24 problemdefinition.tex
@@
-\section{Problem Definition}
 \label{sec:problemdefinition}
-In order to evaluate a temporal summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization Track.
+In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization Track.
 We provide a brief overview of the problem.
@@
 
-A temporal summarization system takes as input 
+An update summarization system takes as input 
 \begin{enumerate*}[label=\itshape\alph*\upshape)]
@@
   \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
-  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$, presented in temporal order, and
-	\item an evaluation time period of interest, $(\stime,\etime)$.  
-\end{enumerate*}
-While processing documents throughout the time period of interest, the system must output sentences, known as updates, which are \emph{relevant} to the query, \emph{comprehensive} with respect to the event, \emph{novel} to the user, and \emph{timely} with respect to when the update occurred.
-Precisely how to measure these properties will be discussed in Section \ref{sec:methods}.
-In our version of this problem, we assume that that the system receives one batch of new sentence-segmented documents every hour throughout the period of interest.
+  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
+  presented in temporal order, and \item an evaluation time period of interest,
+      $(\stime,\etime)$.  \end{enumerate*} While processing documents
+      throughout the time period of interest, the system must output sentences,
+      known as updates, which are \emph{relevant} to the query,
+      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
+      and \emph{timely} with respect to when the update occurred.  Precisely
+      how to measure these properties will be discussed in Section
+      \ref{sec:methods}.  In our version of this problem, we assume that that
+      the system receives one batch of new sentence-segmented documents every
+      hour throughout the period of interest.
 
-\fdcomment{include description of nuggets here.}
+We address disaster types such as terrorism and mass shootings (e.g., the 2012
+shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
+(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
+Arab spring). The size and scope of the events varies considerably. Hurricane
+Sandy, for example, effected multiple countries over weeks while the
+2012 Aurora shooting was contained to a single location spanning several hours.
+
+%\kmcomment{This was as best as I could do on this. These are the example
+%    categories from the wikipedia page. However you refer in the language model
+%    description to the event type earthquake which is more specific. Also, in
+%    the TREC data you refer to event types that are specified in the metadata.
+%So I'm wondering how many different categories you have and where they come
+%from.}
+
+%\fdcomment{include description of nuggets here.} 
+%%\kmcomment{We should jointly
+%    discuss what goes in data and what goes in problem definition. I have moved
+%    event types here because I think knowing how many types and examples of
+%    them would be helpful upfront. I think where nuggets and wikipedia pages go
+%is questionable.}
+
+Figure \ref{alg:temporal-summarization} outlines our general update
+summarization algorithm.  At each hour, the system processes each input 
+sentence batch $S_t$.
+We predict the salience $P$ for all input sentences $s\in S_t$.
+Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
+obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
+to be updates and add those to our set of summary updates $U$.
+
+\begin{algorithm}%[H]
+ \KwData{ 
+     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
+     batches\\
+ \KwResult{U --- a list of updates, i.e. the update summary} 
+}
+ ~\\
+ Initialize empty list $U$ of updates\\
+    \For{$t \gets \stime,\ldots,\etime $}{
+        $P \gets \operatorname{PredictSalience}(S_t)$\\
+        $E \gets \operatorname{APCluster}(P, S_t)$ \\
+        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
+        $U \gets U \cup U_t$
+}
+ \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
+\end{algorithm}
+
+\subsection{Salience Prediction}
+%
+\begin{figure}[t!]
+\begin{tabular}{| l |} 
+\hline
+\textbf{Basic Features}\\
+$\cdot$ Sent. position (normalized by doc. length) \\
+$\cdot$ Sent. length \\
+$\cdot$ Ratio of punc. to non-punc. chars. \\
+$\cdot$ Ratio of caps. to non-caps. chars. \\
+$\cdot$ Ratio of lowercase to other chars. \\
+$\cdot$ Ratio of uppercase to other chars. \\
+$\cdot$ \# of caps. words (normalized by \# of words)\\
+$\cdot$ \# of Person, Location, Org. Date, Number,\\
+$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. tags \\
+$\;\;$ (normalized by \# of words)\\
+\hline
+\textbf{Query Features}\\
+$\cdot$ \% of query words covered by sent.\\
+$\cdot$ Total query matches.\\
+$\cdot$ Total event-type synonyms/hypernyms/hyponyms\\
+$\;\;$ coverage.\\
+$\cdot$ Total event-type synonyms/hypernyms/hyponyms\\
+$\;\;$ matches.\\
+\hline
+\textbf{Language Model Features}\\
+$\cdot$ Avg. token log probability (domain lang. model)\\
+$\cdot$ Avg. token log probability (background lang. model)\\
+\hline
+\textbf{Geo-tag Features}\\
+$\cdot$ Median document distance to nearest location \\
+$\;\;$ cluster (current hour).\\
+$\cdot$ Median document distance to nearest location \\
+$\;\;$ clusters (previous hour).\\
+$\cdot$ Distance of first location in doc. to nearest location \\
+$\;\;$ cluster (current hour).\\
+$\cdot$ Distance of first location in doc. to nearest location \\
+$\;\;$ cluster (previous hour).\\
+\hline
+\textbf{Temporal Features}\\
+$\cdot$ Avg. tf-idf at current time.\\
+$\cdot$ Change in avg. tf-idf since previous hour (up to\\
+$\;\;$  24 hours).\\
+$\cdot$ Time since query/event start.\\
+\hline
+\end{tabular}
+\caption{Salience model features.}
+\end{figure}
+
+\subsubsection{Features}
+We want our model to be predictive across different kinds of events so we avoid lexical features.
+Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
+These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
+\fdcomment{can we enumerate all of these?}
+
+\paragraph{Basic Features}
+%KM - Would be good to have quick justification of these features. I added a
+%sentence. Feel free to edit or remove.
+
+We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
+These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
+Since training is one on grammatical English, some of these features help
+to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
+Others help to more heavily weight important sentences (e.g., that appear in
+prominent positions such as paragraph initial or article initial).
+
+\paragraph{Query Features}
+
+Query features measure the relationship between the sentence and the event query and type.
+These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
+For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
+\paragraph{Language Model Features}\label{subsubsec:lm}
+Language models allow us to measure the likelihood of a sentence having been produced from a particular source.
+We consider two types of language model features.
+The first model is estimated from a corpus of generic news articles.
+This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).
+The second model is estimated from text specific to our event types.
+For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.
+These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g.
+most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
+%KM - Note: Someplace the exact list of event types should appear. Probably not
+%here.
+%KM - I note you have it in a later section but it is labeled as data you use
+%to train language models and semantic similarity. I think it would be good to
+%have up front in definition of task.
+
+%For both models, we Finally, we extract the percentage of capitalized words,
+%and sentence length as features. These last two features also help to
+%identify sentences that are less likely to contain relevant content-- overly
+%long and heavily capitalized sentences in our corpus were likely to be long
+%strings of web-page headlines, section headers, and other irrelevant page
+%structure.
+
+\paragraph{Geographic Relevance Features}
+
+Locations are identified using a named entity tagger.
+For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.
+We then compute its distance to that of the event location.
+It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.
+Distances are calculated using the Vincenty distance.
+%KM - Probably should say how you determine event location. Some events move.
+%KM - In your figure you only include one distance.
+
+\paragraph{Temporal Relevance Features}
+
+Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
+the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.
+
+Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
+We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
+\begin{align*}
+\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
+\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
+w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
+D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
+
+where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
+\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
+\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
+\end{align*}
+% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
+%\end{align*}
+
+We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.
+
+\subsubsection{Model}
+%KM - I think one of the words ``sentence'' below should be something else, btu
+%not sure what you meant. Ah.. I think ``salience''?
+%We adopt Gaussian process regression in order to predict sentence sentence
+We adopt Gaussian process regression in order to predict sentence salience
+\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
+class of data-driven, non-parametric model generalizing the multi-variate
+Gaussian to the infinite dimensional setting.  Gaussian processes are general
+and are state of the art for many regression tasks.  A full treatment of
+Gaussian process regression is beyond the scope of this article and can be
+found in standard textbooks.
+
+Being non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.
+In our experiments, we used a radial basis function (RBF) kernel.
+Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
+\begin{align*}
+        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
+\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
+\end{align*}
+where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
+The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
+
+% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
+% of an input space $\mathcal{X}$ to the reals,
+%
+% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
+% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
+%  an $n$-dimensional vector whose elements are evaluations of the function $f$
+% at points $x_i \in \mathcal{X}$.
+% We say $p(f)$ is a Gaussian process if for any finite subset
+% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
+% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
+% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
+% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
+% set to 0, leaving the distribution to be completely characterized by the
+% kernel function on the data.
+%
+% In the regression setting, we typically have a response variable $y$ that
+% is the sum of our model prediction  and
+% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
+% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
+% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
+% two distributions
+% of principal interest are the marginal likelihood
+% $p(\mathbf{y}|\mathbf{X}) =
+% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
+% distribution,
+%
+% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
+% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
+%
+% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
+% response, and
+% \begin{align*}
+% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
+% \boldsymbol{\sigma}^2_* &
+% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
+% \mathbf{K}_*^T + \sigma^2\\
+% \end{align*}.
+%
+% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
+% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
+%
+%

### 2015-02-22 relatedwork.tex
added deps to makefile; added glasgow reference;
Moving to acl style

@@
-\section{Related Work}
 \label{sec:relatedwork}
+
 A principal concern in extractive multi-document summarization is the
 selection of salient sentences for inclusion in summary output
-\cite{nenkova2012survey}.  This has often been approached as a ranking
-problem.
-%We broadly conceptualize this decision as either an intrinsic or extrinsic
-%sentence evaluation process. Intrinsic approaches evaluate sentences
-%individually, possibly by predicting the impact on summary quality using
-%sentece level features. 
+\cite{nenkova2012survey}. 
+%We conceptualize 
+Existing approaches generally fall into  %falling into
+one of three categories, each with specific tradeoffs with respect to update 
+summarization.
+
+First, centrality-focused approaches (including graph,
+cluster, and centroid methods) are very natural for retrospective analysis in
+the sense that they let the data "speak for itself." 
+These methods equate salience with centrality, either to the input or some other
+aggregate object (i.e. a cluster center or input centroid).
+However, they rely
+chiefly on
+redundancy. When applied to an unfolding event, there may not exist enough
+redundant content at the event onset for these methods to exploit.
+Once the event onset has passed, however, the redundancy reduction of these 
+methods is quite beneficial.
+
+The second category, predictive approaches,
+includes ranking and classification based methods.
 Sentences have been ranked by the average word probability, average tf-idf
@@
 statistics are easily computable from the input sentences, while the third
-only requires an additional, generic background corpus.  Another ranking
-approach, centroid summarization, involves creating an average bag of words
-(BOW) vector, the centroid, from the input sentences and ranking sentences by
-their similarity to the centroid \cite{radev2004centroid}.  Graph
-\cite{erkan2004lexrank} and clustering
-\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
-based approaches, on the other hand, make use of pair-wise similarity
-comparisons amongst input sentences.  In these models, salient sentences are
-more central to the input or cluster, respectively.
+only requires an additional, generic background corpus.  
+In classification based methods, model features are
+usually derived from human generated summaries, and are non-lexical in nature
+(e.g., sentence starting position, number of topic-signatures, number of
+unique words). Seminal work in this area has employed naive
+Bayes and logistic regression classifiers to identify sentences for summary
+inclusion \cite{kupiec1995trainable,conroy2001using}. 
+While these methods are less dependent on redundancy, the expressiveness of
+their features is limited. Our model expands on these basic features to 
+account for geographic, temporal, and language model features.
+
+The last category includes probabilistic, information theoritic, and set cover
+approaches. While these methods are generally focused on producing diverse
+summaries, they are difficult to adapt to the streaming setting, where 
+we do not necessarily have a fixed summary length and the corpus to be
+summarized contains many irrelevent sentences, i.e. there are large
+portions of the corpora that we specificly want to avoid.
+
+% This has often been approached as a ranking
+%problem.
+%We broadly conceptualize this decision as either an intrinsic or extrinsic
+%sentence evaluation process. Intrinsic approaches evaluate sentences
+%individually, possibly by predicting the impact on summary quality using
+%sentece level features. 
+%Sentences have been ranked by the average word probability, average tf-idf
+%score, and the number of topically related words (topic-signatures in the
+%summarization literature)
+%\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
+%statistics are easily computable from the input sentences, while the third
+%only requires an additional, generic background corpus.  Another ranking
+%approach, centroid summarization, involves creating an average bag of words
+%(BOW) vector, the centroid, from the input sentences and ranking sentences by
+%their similarity to the centroid \cite{radev2004centroid}.  Graph
+%\cite{erkan2004lexrank} and clustering
+%\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
+%based approaches, on the other hand, make use of pair-wise similarity
+%comparisons amongst input sentences.  In these models, salient sentences are
+%more central to the input or cluster, respectively.
 
@@
 %sentences.
-
-Supervised learning has also been applied to this task. Model features are
-usually derived from human generated summaries, and are non-lexical in nature
-(e.g., sentence starting position, number of topic-signatures, number of
-unique words, word frequencies). Seminal work in this area has employed naive
-Bayes and logistic regression classifiers to identify sentences for summary
-inclusion \cite{kupiec1995trainable,conroy2001using}.
 
@@
 between the system summary and a manually generated target summary.
+
+\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

### 2015-02-23 experiments.tex
Added experiments and results

We evaluate our system on two metrics: ROUGE \cite{?}, the most common automatic
summarization method and a modified method adapted 
from the Text Retrieval Conference Temporal Summarization (TREC-TS) track \cite{?}.

\subsection{Dataset}

Both evaluations make use of the TREC TS data from 2013 and 2014. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event time frame).
Additionally, TREC assessors constructed a set of ground truth facts, or 
nuggets, for each event. The events took place from 20??-20?? and had 
an associated Wikipedia (WP) page. The WP pages were edited and updated 
concurrently with the events, so the diffs between revisions produce a similar
ouput to the ideal update summarization system. Assessors analyzed the diffs
and selected phrases that were important enough to include in an update 
summary.  
For the document stream, we use the news portion of the
 2014 TREC Stream Corpus.
The documents from this corpus come from hourly crawls of the web covering 
 ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}

\subsection{Training and Testing}

Of the 25 events in the TREC TS data, 24 are covered by the Stream Corpus.
From these 24, we set aside three events to use as a development set.
All system salience and similarity threshold parameters are tuned on the
development set.

We train a salience model for each event using 1000 sentences randomly sampled
from the event's document stream. Kernel length scale parameters  for each
feature group are fit to maximize
the model's log likelihood using ???.
At test time, we predict a sentence's salience using the average predictions
of the 23 other models---essentially 24 fold cross validation approach.

\subsection{ROUGE Evaluation}

ROUGE measures the ngram overlap between a model (i.e. human) summary 
and an automatically generated system summary. 
Model summaries for each event were constructed by concatenating the event's 
nuggets. 
See figure~\ref{fig:test} for an example model summary excerpt.
Generally, ROUGE evaluation assumes both model and system summaries
are of a bounded length. Since our systems are summarizing events over a span
of two weeks time, the total length of our system output is much longer than
the model. To address this, for each system/event pair, we sample with replacement
1000 random summaries of length less than or equal to the model summary 
(truncating the last sentence when neccessary). The final ROUGE score for the 
system are the average scores from these 1000 samples.

Because we are interested in system performance over time, we also evaluate 
systems at 12 hour intervals using the same regime as above. 
The model summaries in this case are retrospective, and so we can see how quickly
each system takes to achieve its maximum performance.

\subsection{Nugget Evaluation}
We also modify the evaluation used in the TREC TS track, 
where systems were evaluated on the containment of event nuggets. 
We match each system update to 0 or more nuggets, where an update matches
a nugget if their semantic similarity is above a threshold.  
Summaries are then evaluated on the precision and recall of the matches.
More formally,
\[ \mathrm{Prec.} = \frac{|S_n|}{|S|} \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}\] 
where S is the set of system updates, $S_n$ is the set of nuggets mapped updates in $S$,
$N$ is the set of event nuggets, and $|\cdot|$ is the size of the set.
A precision of 1 indicates that every sentence contains one unique nugget.
Because an update can map to more than one nugget, it is possible to receive a precision greater than 1. 
A recall of 1 indicates that the summary has cover all nugget information for the event; the maximum
attainable recall is 1.
Determining 
an optimal threshold is difficult so we evaluate at threshold values ranging
from .5 to 1, where values closer to 1 are more conservative estimates of
performance.

\subsection{Model Comparisons}

Our proposed model using AP+Salience is compared against several variants and baselines. All thresholds
are tuned on the development set.

\textbf{AP+Salience Time Ranked} : This model is everywhere the same as AP+Salience except at the update selection stage, 
where the algorithm selects exemplars in temporal order rather than salience order.

\textbf{AP} : In this model we use the AP clustering algorithm without the salience predictions.
The preferences are chosen as the median value of the ? matrix, as is commonly used with AP \cite{?}.
The update selection procedes in time order. As in the AP+Salience model we tune a similarity threshold,
and ignore exemplar sentences whose maximum similarity to a previous update is above the threshold.

\textbf{Rank by Salience} : In this model we rank the sentences by salience and select the updates at each hour that are
above a salience threshold and are below a similarity threshold (similar to the AP+Salience model).

\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We also provide another clustering baseline, HAC using ?-linkage. 
This method was chosen over other clustering approaches because the number of clusters 
is not an explicit hyper parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold
(like the previous models).

### 2015-02-23 methods.tex
Added experiments and results

--- 2015-01-24 methods.tex
@@
-\section{Methods}
-\label{sec:methods}
-\subsection{Data}\label{sec:data}
+In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
+We provide a brief overview of the problem.
+Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.
 
-Our primary datasets are the TREC 2013/2014 Stream Corpus, 
-a ?tb corpus of hourly 
-web crawls from October 2011
-through mid February 2013
-\cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
-%KM - Keep neutral from Trec
-%As per the TS specifications,
-All summary sentences come from processing this
-corpus in a time aligned manner.
-%KM - Are you using both 2013 and 2014 corpora.
-%KM - Mentioning sentence ids at this point is too low level.
-%KM - I would probably just drop this sentence. since you say above that you
-%use both.
-%While the 2014 corpus encompasses the 
-%2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
-%year because sentence ID's used in evaluation changed from year to year.
+An update summarization system takes as input 
+\begin{enumerate*}[label=\itshape\alph*\upshape)]
+  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
+  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
+  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
+  presented in temporal order, and \item an evaluation time period of interest,
+      $(\stime,\etime)$.  \end{enumerate*} While processing documents
+      throughout the time period of interest, the system must output sentences,
+      known as updates, which are \emph{relevant} to the query,
+      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
+      and \emph{timely} with respect to when the update occurred.  Precisely
+      how to measure these properties will be discussed in Section
+      \ref{sec:methods}.  In our version of this problem, we assume that that
+      the system receives one batch of new sentence-segmented documents every
+      hour throughout the period of interest.
 
-%KM - As far as I can see this is the first time you've mentioned this.
-%KM - This paragraph is not clearly understandable. Is this the training data? 
-For our semantic similarity component (\cref{subsec:semsim}),
-domain specific term-sentence matrices were built using subsections of
-Wikipedia. For each event type we identified one or more related Wikipedia
-categories and collected all pages that were listed under that category.
-For all pages collected, we used the latest possible revision date before the
-start of the Stream Corpus.
+We address disaster types such as terrorism and mass shootings (e.g., the 2012
+shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
+(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
+Arab spring). The size and scope of the events varies considerably. Hurricane
+Sandy, for example, effected multiple countries over weeks while the
+2012 Aurora shooting was contained to a single location spanning several hours.
 
-We use the same Wikipedia corpora to construct the language models used
-in our salience prediction model (\cref{subsubsec:lm}).
+%\kmcomment{This was as best as I could do on this. These are the example
+%    categories from the wikipedia page. However you refer in the language model
+%    description to the event type earthquake which is more specific. Also, in
+%    the TREC data you refer to event types that are specified in the metadata.
+%So I'm wondering how many different categories you have and where they come
+%from.}
 
-\begin{figure*}
-        \begin{center}
-\begin{tabular}{| p{5cm} c c |}
+%\fdcomment{include description of nuggets here.} 
+%%\kmcomment{We should jointly
+%    discuss what goes in data and what goes in problem definition. I have moved
+%    event types here because I think knowing how many types and examples of
+%    them would be helpful upfront. I think where nuggets and wikipedia pages go
+%is questionable.}
+
+Figure \ref{alg:temporal-summarization} outlines our general update
+summarization algorithm.  At each hour, the system processes each input 
+sentence batch $S_t$.
+We predict the salience $P$ for all input sentences $s\in S_t$.
+Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
+obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
+to be updates and add those to our set of summary updates $U$.
+
+\begin{algorithm}%[H]
+ \KwData{ 
+     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
+     batches\\
+ \KwResult{U --- a list of updates, i.e. the update summary} 
+}
+ ~\\
+ Initialize empty list $U$ of updates\\
+    \For{$t \gets \stime,\ldots,\etime $}{
+        $P \gets \operatorname{PredictSalience}(S_t)$\\
+        $E \gets \operatorname{APCluster}(P, S_t)$ \\
+        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
+        $U \gets U \cup U_t$
+}
+ \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
+\end{algorithm}
+
+\subsection{Salience Prediction}
+
+\subsubsection{Model}
+%KM - I think one of the words ``sentence'' below should be something else, btu
+%not sure what you meant. Ah.. I think ``salience''?
+%We adopt Gaussian process regression in order to predict sentence sentence
+We adopt Gaussian process regression in order to predict sentence salience
+\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
+class of data-driven, non-parametric model generalizing the multi-variate
+Gaussian to the infinite dimensional setting.  Gaussian processes are general
+and are state of the art for many regression tasks.  A full treatment of
+Gaussian process regression is beyond the scope of this article and can be
+found in standard textbooks.
+
+As a non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.
+In our experiments, we used a radial basis function (RBF) kernel.
+Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
+\begin{align*}
+        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
+\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
+\end{align*}
+where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
+The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
+
+\subsubsection{Training Data}
+\ckcomment{Explain how we train the model here...}
+%In order to train the salience model, we need to quantify the notion of salience. We assume for each event we have a set of atomic facts, or nuggets.
+
+%
+\begin{figure}[t!]
+\begin{tabular}{| l |} 
 \hline
-Event & WP Categories & No. Docs/Sents/Words\\
-\hline \hline
-Boston\_Marathon\_bombings \newline
-Christopher\_Dorner\_shootings\_and\_manhunt \newline 
- In\_Amenas\_hostage\_crisis
-& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
+\textbf{Basic Features}\\
+$\cdot$ Sent. position (normalized by doc. length) \\
+$\cdot$ Sent. length \\
+$\cdot$ Ratio of punc. to non-punc. chars. \\
+$\cdot$ Ratio of caps. to non-caps. chars. \\
+$\cdot$ Ratio of lowercase to other chars. \\
+$\cdot$ Ratio of uppercase to other chars. \\
+$\cdot$ \# of caps. words (normalized by \# of words)\\
+$\cdot$ \# of Person, Location, Org. Date, Number,\\
+$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
+$\;\;$ tags (normalized by \# of words)\\
 \hline
- Cyclone\_Oswald \newline
-Early\_2012\_European\_cold\_wave \newline
-February\_2013\_noreaster \newline & Weather events & 35,554/591,850/12,794,438  \\
+\textbf{Query Features}\\
+$\cdot$ \% of query words covered by sent.\\
+$\cdot$ Total query matches.\\
+$\cdot$ Total event-type synonyms/hypernyms/ \\
+$\;\;$ hyponyms coverage.\\
+$\cdot$ Total event-type synonyms/hypernyms/ \\
+$\;\;$ hyponyms matches.\\
 \hline
-Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
+\textbf{Language Model Features}\\
+$\cdot$ Avg. token log probability (domain lang. \\
+$\;\;$ model)\\
+$\cdot$ Avg. token log probability (background \\
+$\;\;$ lang. model)\\
 \hline
-Chelyabinsk\_meteor & Earthquakes & 14,515/283,509/6,135,803  \\
+\textbf{Geo-tag Features}\\
+$\cdot$ Median document distance to nearest  \\
+$\;\;$ location cluster (current hour).\\
+$\cdot$ Median document distance to nearest  \\
+$\;\;$ location clusters (previous hour).\\
+$\cdot$ Distance of first location in doc. to nearest \\
+$\;\;$ location cluster (current hour).\\
+$\cdot$ Distance of first location in doc. to nearest \\
+$\;\;$ location cluster (previous hour).\\
 \hline
-Port\_Said\_Stadium\_riot \newline
-2012\_Afghanistan\_Quran\_burning\_protests \newline
-2011-13\_Russian\_protests \newline
-2012\_Romanian\_protests \newline
-2012-13\_Egyptian\_protests \newline
-2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
-2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
+\textbf{Temporal Features}\\
+$\cdot$ Avg. tf-idf at current time.\\
+$\cdot$ Change in avg. tf-idf since previous hour \\
+$\;\;$  (up to 24 hours).\\
+$\cdot$ Time since query/event start.\\
 \hline
 \end{tabular}
-\caption{Domain specific Wikipedia corpora for semantic similarity and 
-language models. }
-\end{center}
-\end{figure*}
+\caption{Salience model features.}
+\end{figure}
 
-\subsubsection{Events}
+\subsubsection{Features}
+We want our model to be predictive across different kinds of events so we avoid lexical features.
+Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
+These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
 
-%KM - You said this already so I think you can make it simpler. 
-%We use the event set developed for the TREC Temporal Summarization track 
-%for years 2013 and 2014.
-The TREC 2013 and 2014 Stream corpus contains  10 and 15 events respectively. We omit
-one event from the 2013 track because there were no news documents in the 
-corpus during the duration of the event, yielding a final event set of 24 
-events. Meta-data provided for each event includes the name of the event
-as it appears on Wikipedia, a short text query for the event, 
-the type of event, and the 
-duration of the event
-%KM - the below parens is not clear. is ``to monitor'' missing from the above
-%sentence? I think get rid of the parens below.
-%(the duration for the system 
-%to monitor is not necessarily the complete timespan of the event).
-%KM - Let's get rid of all references to ``the TS system'' If you want to have
-%a system name, choose a better one. 
-All data except the Wikipedia name is exposed to
-systems that participate in TREC.
-%the TS system to use.
+\paragraph{Basic Features}
+%KM - Would be good to have quick justification of these features. I added a
+%sentence. Feel free to edit or remove.
 
-\subsubsection{Nuggets}
-For evaluation purposes, important pieces of information, hereafter 
-referred to as nuggets, were extracted from each event's Wikipedia page.
-%KM - What assessors? Was this done by TREC? Make clear. I added ``TREC'' after
-%reading ahead.
-%KM - What do you mean ``graded''? Do you mean they graded your nuggets? Or do
-%you mean they ``ranked'' the nuggets by importance. Please make clear.
-TREC assessors graded nuggets by importance, and were also able to timestamp
-%KM - wrong wording. I've changed.
-%their
-%existance
-when the event represented by each nugget occurred 
-using the timestamp of the page's revision history. Nuggets are 
-usually a short to medium length clause containing an atomic piece of
-information.
+We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
+These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
+Since training is one on grammatical English, some of these features help
+to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
+Others help to more heavily weight important sentences (e.g., that appear in
+prominent positions such as paragraph initial or article initial).
 
-%KM - I changed title to make clearer
-\subsubsection{Gold Standard Matches}
-TREC assessors provided judgements on a subset of participant updates each year
-of the track. Judged updates are matched to one or more nuggets (a real 
-sentence often contains more than one atomic piece of information) or marked
-not relevant.
+\paragraph{Query Features}
 
-%KM - I don't see why overlap in particpant update sets would give you low
-%coverage. In fact, it might you give you broader coverage than if all
-%participants extracted the same sentences.  I removed first portion of the
-%sentence, but you can put back and clarify if you think I interpreted
-%incorrectly. 
-%There was not much overlap in the participant update sets so
-Overall coverage of
-%KM - Shouldn't this be nuggets?
-relevant sentences
-relevant nuggets in the corpus  by participant systems' generated sentences
-is small (percent?).
-%KM - This is complicated. More explanation is needed.
-%KM - The sentence below sounds like how you evaluate your system and not like
-%the judgment corpus.
-%KM not sure this owuld go here. Shouldn't it go in evaluation? Here you would
-%just talk about the corpus
-To evaluate new system output on the same event, we compare it against the
-corpus of matched sentences. Comparing it against the nuggets would not yield
-meaningful results as the nuggets are much smaller than the sentences of the
-corpus.
-Because there are so few matched sentences, we score all new system updates as positive
-%To expand the judgement corpora, we include all system updates as positive
-nugget matches if they are $< .2$ close in Levenshtein distance to a human
-assessed matched update.
+Query features measure the relationship between the sentence and the event query and type.
+These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
+For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
+\paragraph{Language Model Features}\label{subsubsec:lm}
+Language models allow us to measure the likelihood of a sentence having been produced from a particular source.
+We consider two types of language model features.
+The first model is estimated from a corpus of generic news articles.
+This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).
+The second model is estimated from text specific to our event types.
+For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.
+These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g.
+most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
+%KM - Note: Someplace the exact list of event types should appear. Probably not
+%here.
+%KM - I note you have it in a later section but it is labeled as data you use
+%to train language models and semantic similarity. I think it would be good to
+%have up front in definition of task.
 
-%KM - Got this far.
-\subsection{Experiment Design}
+%For both models, we Finally, we extract the percentage of capitalized words,
+%and sentence length as features. These last two features also help to
+%identify sentences that are less likely to contain relevant content-- overly
+%long and heavily capitalized sentences in our corpus were likely to be long
+%strings of web-page headlines, section headers, and other irrelevant page
+%structure.
 
-\subsubsection{Salience Models}
-For each event (both in the complete system and feature ablation experiments)
-we randomly sample 1000 sentences after relevant document filtering and 
-over the duration of the event 
-(as determined by 
-TREC; avg. run length is ? days) of the event. Nugget similarities
-are computed for each sentence and are standardized with respect to 
-the sample mean and variance. After standardization, the maximum nugget similarities are computed for each sentence. 
-We fit a Gaussian process for each event, using the maximum similarities 
-as our regression target, yielding 24 models in total. Kernel lengthscale
-parameters are also individually fit for each model.
-When making event-specific predictions, we take the average prediction of the
-23 other models.
+\paragraph{Geographic Relevance Features}
 
-\subsubsection{Estimating Language Models}
-We use two trigram language models, trained using the SRILM toolkit \cite{stolcke2002srilm}, taking as features the average log probability (i.e.
-the sentence's total log probability normalized by sentence length) from each model.
-This first model is trained on 4 years (2005-2009) of articles from the Gigaword corpus.
-Specifically, we use articles from the Associated Press and the New York Times.
-The second model is a domain specific language model.
-We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category.
-E.g.
-for earthquakes, we collect pages under the category \emph{Category:Earthquakes}.
-This model assigns higher probability to sentences that are focused on the given domain.
+Locations are identified using a named entity tagger.
+For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.
+We then compute its distance to that of the event location.
+It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.
+Distances are calculated using the Vincenty distance.
+%KM - Probably should say how you determine event location. Some events move.
+%KM - In your figure you only include one distance.
 
-% For each event, we retrieve all sentences using the document retrieval and content selection steps outlined in the previous selection.
-We then sample with replacement 200 sentences from this collection, and extract non-lexical features for each sentence to construct our design matrix $\mathbf{X}$.
-To build our response variables $\mathbf{y}$ we compute the maximum semantic similarity of each input sentence to the gold nugget sentences.
-We fit a $\operatorname{GP}$ with an RBF kernel to this data, optimizing kernel parameters with the scaled  conjugate gradient method.
-This sampling procedure is repeated 100 times for each of the ?
-2013 TS events, yielding \fdcomment{?} total models.
-In our salience predictions for the 2014 events, we take the mean prediction of all models.
+\paragraph{Temporal Relevance Features}
 
-\subsubsection{Redundancy Threshold}
+Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
+the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.
 
-For every time interval with input, a clustering algorithm is
-guarranteed to produce at least one output; 
-this can significantly hurt the performance of all clustering based 
-temporal summarization systems. In order to avoid this, we set a threshold $t$
-such that a new update must have a maximum similarity $< t$ to all previous 
-updates.
+Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
+We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
+%\begin{align*}
+%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
+%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
+%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
+%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
 
-We determined the value of $t$ by examining the distribution of similarities
-of nuggets to their human-matched sentences compared to the distribution
-of similarities of nuggets to relevant but non matching sentences. We fit
-normal distributions for the matching and non-matching similarities and
-used the point of intersection as the value of $t$.
-Figure ? shows a normalized histogram of cosine similarities for both 
-distributions, the fitted distributions, and the point of intersection.
+%where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
+%\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
+%\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
+%\end{align*}
 
-\begin{figure}
-\includegraphics[scale=.40]{match-dist.png} 
-\caption{Distribution of nugget similarity for matching and non-matching 
-updates.\fdcomment{possible to output this as a vector graphic (pdf pref).  it will scale better.}}
-\end{figure}
-\subsubsection{Effect of Salience Models}
+[\textrm{Simpler explanation goes here}]
+% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
+%\end{align*}
 
-In order to assess the effectiveness of the salience predictions on update
-selection, we perform a baseline (AP-base) run using affinity propagation 
-clustering
-with uniform preference, i.e. all preferences are equal to the median sentence
-similarity. This run is compared to our full system (AP-sal), using each
-sentence's salience prediction as its preference.
+We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.
 
-We are also interested
-in the effect of directly incorporating salience in the clustering algorithm.
-A third run (HAC) is performed using hierarchical agglomerative 
-clustering; we produce flattened clusters by setting a maximum cluster 
-distance threshold. The sentence with the highest predicted salience from each
- cluster is selected as an update. 
-The distance threshold was tuned using grid search. By comparing
-HAV with AP-base we can empirically assess the trade off between preference
-and neighbor similarity that affinity propagation is making.
+% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
+% of an input space $\mathcal{X}$ to the reals,
+%
+% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
+% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
+%  an $n$-dimensional vector whose elements are evaluations of the function $f$
+% at points $x_i \in \mathcal{X}$.
+% We say $p(f)$ is a Gaussian process if for any finite subset
+% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
+% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
+% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
+% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
+% set to 0, leaving the distribution to be completely characterized by the
+% kernel function on the data.
+%
+% In the regression setting, we typically have a response variable $y$ that
+% is the sum of our model prediction  and
+% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
+% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
+% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
+% two distributions
+% of principal interest are the marginal likelihood
+% $p(\mathbf{y}|\mathbf{X}) =
+% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
+% distribution,
+%
+% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
+% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
+%
+% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
+% response, and
+% \begin{align*}
+% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
+% \boldsymbol{\sigma}^2_* &
+% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
+% \mathbf{K}_*^T + \sigma^2\\
+% \end{align*}.
+%
+% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
+% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
+%
+%
 
-\subsubsection{Feature Ablation}
-
-We remove each feature group and train models models on these small faller 
-feature susbstets. All other parameters are the same as in our full system 
-(AP-sal).
+\subsection{Affinity Propagation Clustering}
+\ckcomment{Explain AP and how salience gets plugged in here}
+\subsection{Update Selection}
+\ckcomment{Explain thresholds}

### 2015-02-23 methods_old.tex
Added experiments and results

\section{Methods}
\label{sec:methods}
\subsection{Data}\label{sec:data}

Our primary datasets are the TREC 2013/2014 Stream Corpus, 
a ?tb corpus of hourly 
web crawls from October 2011
through mid February 2013
\cite{frank2012building}.\footnote{\url{http://trec-kba.org/kba-stream-corpus-2014.shtml}}
%KM - Keep neutral from Trec
%As per the TS specifications,
All summary sentences come from processing this
corpus in a time aligned manner.
%KM - Are you using both 2013 and 2014 corpora.
%KM - Mentioning sentence ids at this point is too low level.
%KM - I would probably just drop this sentence. since you say above that you
%use both.
%While the 2014 corpus encompasses the 
%2013 corpus, we use the 2013 corpus for evaluating events from the 2013 track
%year because sentence ID's used in evaluation changed from year to year.

\subsubsection{Events}

%KM - You said this already so I think you can make it simpler. 
%We use the event set developed for the TREC Temporal Summarization track 
%for years 2013 and 2014.
The TREC 2013 and 2014 Stream corpus contains  10 and 15 events respectively. We omit
one event from the 2013 track because there were no news documents in the 
corpus during the duration of the event, yielding a final event set of 24 
events. Meta-data provided for each event includes the name of the event
as it appears on Wikipedia, a short text query for the event, 
the type of event, and the 
duration of the event.
%KM - the below parens is not clear. is ``to monitor'' missing from the above
%sentence? I think get rid of the parens below.
%(the duration for the system 
%to monitor is not necessarily the complete timespan of the event).
%KM - Let's get rid of all references to ``the TS system'' If you want to have
%a system name, choose a better one. 
All data except the Wikipedia name is exposed to
systems that participate in TREC.
%the TS system to use.

\subsubsection{Wikipedia}

\kmcomment{
 As far as I can see this is the first time you've mentioned this.
 This paragraph is not clearly understandable. Is this the training data?
 Also confusing here is that you alternate between what TREC provides and
 what you do. This absolutely must be clear. Are these wikipedia pages provided
 by TREC or do you create a wikipedia corpus on your own?}

For each event type we identified one or more related Wikipedia
categories and collected all pages that were listed under that category for use
as training data.
For all pages collected, we used the latest possible revision date before the
start of the Stream Corpus.

We use the  Wikipedia corpora to construct the language models used
in our salience prediction model (\cref{subsubsec:lm}). For each of the five
event types shown in Fig.~\ref{fig:wikipedia}, we used all data in the pages
shown to build a language model. The amount of words per event type varied from
6 million to 250 million.
\kmcomment{moved this to the end since it is less of a priority}
We used the same corpus to train our semantic similarity component
(\cref{subsec:semsim}. 
Disaster specific term-sentence matrices were built using subsections of
Wikipedia.
\kmcomment{I don't think you ever mention term-sentence matrices earlier.}
\begin{figure*}
        \begin{center}
\begin{tabular}{| p{5cm} c c |}
\hline
Event & WP Categories & No. Docs/Sents/Words\\
\hline \hline
2012 Aurora shooting \newline
Wisconsin Sikh temple shooting\newline 
2012 Tel Aviv bus bombing \newline
Boston\_Marathon\_bombings \newline
Christopher\_Dorner\_shootings\_and\_manhunt \newline 
 In\_Amenas\_hostage\_crisis
& Terrorism, Mass Shootings & 33,732/1,139,588/26,201,659  \\
\hline
Hurricane Isaac (2012) \newline
Hurricane Sandy \newline
Typhoon Bopha \newline
2012 Guatemala earthquake \newline
 Cyclone\_Oswald \newline
Early\_2012\_European\_cold\_wave \newline
February\_2013\_noreaster \newline & Natural Disasters & 35,554/591,850/12,794,438  \\
\hline
2012 Buenos Aires Rail Disaster
2012 Pakistan garment factory fires
Costa\_Concordia\_disaster & Accidents & 22,874/732,945/16,520,242 \\
\hline
Chelyabinsk\_meteor & Astronomy & 14,515/283,509/6,135,803  \\
\hline
Port\_Said\_Stadium\_riot \newline
2012\_Afghanistan\_Quran\_burning\_protests \newline
2011-13\_Russian\_protests \newline
2012\_Romanian\_protests \newline
2012-13\_Egyptian\_protests \newline
2013\_Bulgarian\_protests\_against\_the\_Borisov\_cabinet \newline
2013\_Shahbag\_protests & Activism\_by\_type & 464,657/11,254,122/250,172,896  \\
\hline
\end{tabular}
\caption{Domain specific Wikipedia corpora for semantic similarity and 
language models. }
\label{fig:wikipedia}
\end{center}
\end{figure*}

\subsubsection{Nuggets}
For evaluation purposes, important pieces of information, hereafter 
referred to as nuggets, were extracted from each event's Wikipedia page.
%KM - What assessors? Was this done by TREC? Make clear. I added ``TREC'' after
%reading ahead.
%KM - What do you mean ``graded''? Do you mean they graded your nuggets? Or do
%you mean they ``ranked'' the nuggets by importance. Please make clear.
TREC assessors graded nuggets by importance, and were also able to timestamp
%KM - wrong wording. I've changed.
%their
%existance
the occurrence time of the events represented by each nugget
using the timestamp of the page's revision history. Nuggets are 
usually a short to medium length clause containing an atomic piece of
information.
%KM - I think an example of a nugget here could help.

\kmcomment{ I changed title to make clearer}
\subsubsection{Gold Standard} %Matches}

A gold standard dataset consisting of full sentences from the TREC datastream
was constructed to use for scoring of the system we describe here. It would be
difficult to automatically match full sentences against the nuggets identified
by TREC assessors given the substantial difference in length; each nugget
typically appears as a clause or phrase in the sentence.

TREC assessors provided judgements on a subset of participant updates (i.e.,
full sentences)  each year
of the track. Judged updates are matched to one or more nuggets (a real 
sentence often contains more than one atomic piece of information) or marked
not relevant.

%KM - I don't see why overlap in particpant update sets would give you low
%coverage. In fact, it might you give you broader coverage than if all
%participants extracted the same sentences.  I removed first portion of the
%sentence, but you can put back and clarify if you think I interpreted
%incorrectly. 
%There was not much overlap in the participant update sets so
Overall coverage of
%KM - Shouldn't this be nuggets?
%relevant sentences
relevant nuggets in the corpus  by participant systems' generated sentences
is small (percent?). Thus we augmented the pool of previous participant system
output with judgments on new system output using Amazon Mechanical Turk. For
each sentence produced by our system and by our baselines, we asked turkers to
indicate which of the nuggets were conveyed by the sentence. For each hit, we
provided one sentence and five nuggets. For each nugget, they had to answer a
yes-no question indicating whether it was conveyed by the full sentence.
%KM - This is complicated. More explanation is needed.
%KM - The sentence below sounds like how you evaluate your system and not like
%the judgment corpus.
%KM not sure this owuld go here. Shouldn't it go in evaluation? Here you would
%just talk about the corpus
To evaluate new system output on the same event, we compare it against the
corpus of matched sentences. Comparing it against the nuggets would not yield
meaningful results as the nuggets are much smaller than the sentences of the
corpus.
%KM - Let's discuss where these different things go. I think scoring should
%probably go in the evaluation section.
\kmcomment{I think this sentence should go in the evaluation section. This
  subsection is about data.}
Because there are so few matched sentences, we score all new system updates as positive
%To expand the judgement corpora, we include all system updates as positive
nugget matches if they are $< .2$ close in Levenshtein distance to a human
assessed matched update.

\subsection{Experiment Design}

\subsubsection{Salience Models}
For each event (both in the complete system and feature ablation experiments)
we randomly sample 1000 sentences after relevant document filtering and 
over the duration of the event 
as determined by 
TREC; avgerage run length is ? days.
%KM - Should you say here using the sentence similarity software? 
Nugget similarities
are computed for each sentence and are standardized with respect to 
the sample mean and variance. After standardization, the maximum nugget
similarities are computed for each sentence.
%KM - I think it would be helpful to mention that you use the features below
%here. 
We fit a Gaussian process for each event, using the maximum similarities 
as our regression target, yielding 24 models in total. Kernel lengthscale
parameters are also individually fit for each model.
%KM - Do you mean for new events? Would help if you say so. 
When making event-specific predictions, we take the average prediction of the
23 other models.

\subsubsection{Estimating Language Models}
We use two trigram language models, trained using the SRILM toolkit
\cite{stolcke2002srilm}, taking as features the average log probability (i.e.
the sentence's total log probability normalized by sentence length) from each
model.  This first model is trained on 4 years (2005-2009) of articles from the
Gigaword corpus.  Specifically, we use articles from the Associated Press and
the New York Times. The second model is a domain specific language model.
%KM - Remind us how many event types there are. Do you list all at some point?
We build a corpus of Wikipedia articles for each event type, consisting of documents from a related Wikipedia category. E.g. for earthquakes, we collect pages under the category \emph{Category:Earthquakes}. This model assigns higher probability to sentences that are focused on the given domain.

% For each event, we retrieve all sentences using the document retrieval and content selection steps outlined in the previous selection. We then sample with replacement 200 sentences from this collection, and extract non-lexical features for each sentence to construct our design matrix $\mathbf{X}$. To build our response variables $\mathbf{y}$ we compute the maximum semantic similarity of each input sentence to the gold nugget sentences. We fit a $\operatorname{GP}$ with an RBF kernel to this data, optimizing kernel parameters with the scaled  conjugate gradient method. This sampling procedure is repeated 100 times for each of the ? 2013 TS events, yielding \fdcomment{?} total models. In our salience predictions for the 2014 events, we take the mean prediction of all models.

\subsubsection{Redundancy Threshold}

For every time interval with input, a clustering algorithm is
guarranteed to produce at least one output; 
this can significantly hurt the performance of all clustering based 
temporal summarization systems. In order to avoid this, we set a threshold $t$
such that a new update must have a maximum similarity $< t$ to all previous 
updates.

We determined the value of $t$ by examining the distribution of similarities
of nuggets to their human-matched sentences compared to the distribution
of similarities of nuggets to relevant but non matching sentences. We fit
normal distributions for the matching and non-matching similarities and
used the point of intersection as the value of $t$.
%KM _ I think you need to say here what sentence similarity method you use.
%Just saying ``cosine similarity'' implies word overlap.
Figure ? shows a normalized histogram of cosine similarities for both 
distributions, the fitted distributions, and the point of intersection.

\begin{figure}
\includegraphics[scale=.40]{match-dist.png} 
\caption{Distribution of nugget similarity for matching and non-matching 
updates.\fdcomment{possible to output this as a vector graphic (pdf pref).  it will scale better.}}
\end{figure}

%KM _ What happened to the geo-codes? Aren't they in the salience models?
\subsubsection{Effect of Salience Models}

In order to assess the effectiveness of the salience predictions on update
selection, we perform a baseline (AP-base) run using affinity propagation 
clustering
with uniform preference, i.e. all preferences are equal to the median sentence
similarity. This run is compared to our full system (AP-sal), using each
sentence's salience prediction as its preference.

%KM - I think this should be rewritten to say you compare to a usual
%summarization technique, using hierarchical agglomerative clustering with
%selection of sentences at the centroid of the cluster.
%We are also interested
%in the effect of directly incorporating salience in the clustering algorithm.
%A third run (HAC) is performed using hierarchical agglomerative 
%clustering; we produce flattened clusters by setting a maximum cluster 
%distance threshold. The sentence with the highest predicted salience from each
% cluster is selected as an update. 
%The distance threshold was tuned using grid search. By comparing
%HAV with AP-base we can empirically assess the trade off between preference
%and neighbor similarity that affinity propagation is making.
%KM - I now see why you were running experiments this way. It would be an
%interesting test.

%KM - Here is a paragraph you could use. I commented out above but you can put
%back in if there are pieces you prefer. Of course, edit below any way you see
%fit. Probably cite Radev and Jing here. They did this fairly early. There may
%be others who did this later.

We also compared our system against a typical summarization approach using
hierarchical agglomerative clustering (HAC); we produce flattened clusters by setting a maximum cluster 
distance threshold. We  select the centroid of each
cluster as a summary sentence  \cite{??}.

\subsubsection{Feature Ablation}

%KM - Probably should describe this as either backward search over feature
%groups. There's a word for this. I will find it. Also should start with the
%purpose of doing it. 
We remove each feature group and train models models on these small 
feature subsets. All other parameters are the same as in our full system 
(AP-sal).

### 2015-02-23 results.tex
Added experiments and results

\subsection{ROUGE}

Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
Both versions of AP+Salience show improvement on ngram precision and recall over the 
baselines. This improvement is statistically significant at the $\alpha = .01$ level
using the Wilcoxon signed-rank test. Additionally, AP+Salience's improvement over
AP+Salience TR in recall and F-measure is also statistically significant at the same level.

AP+Salience maintains its performance above the baselines over time as well. Table~\ref{fig:trouge}
shows the ROUGE-1 performance over time. Within the initial days of the event, recall is 
competitive amongst system. However, both AP+Salience variants are able to achieve higher
precision early on.

\subsection{Nugget Evaluation}

\subsection{Feature Ablation}

\begin{table}[h]
\centering
% centering table
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-1}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\
[0.5ex]
\hline
AP+Salience &          0.289   &    0.344  &  0.312  \\
AP+Salience TR &       0.278   &    0.342  &  0.303 \\
AP    &         0.246   &    0.285  &  0.263 \\
HAC     &        0.238   &    0.276  &  0.255 \\
Rank    &        0.229   &    0.271  &  0.247 \\
% inserts single-line
% Entering 1st row
%& &soft &1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
%[-1ex]
%\raisebox{1.5ex}{Police} &\raisebox{1.5ex}{5}&hard& 2 & $-4$ & 4 & 4 & $-2$ & $-4$ & 4\\
%[1ex]
% Entering 2nd row
%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
%[-1ex]
%\raisebox{1.5ex}{Beethoven}&\raisebox{1.5ex}{5}& hard&8 & $-8$ & 2 & 8 & $-8$ & $-8$ & 6
%\\
%[1ex]
% Entering 3rd row
%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1
%\\
%[-1ex]
%\raisebox{1.5ex}{Metallica} &\raisebox{1.5ex}{5}& hard&4 & $-8$ & 8 & 4 & $-8$ & $-8$ & 8\\
%[1ex]
% [1ex]adds vertical space
\hline % inserts single-line
\end{tabular}
~\\[1ex]
~\\
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-2}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\[0.5ex]
\hline
AP+Salience & 0.047 &     0.056 &   0.051 \\
AP+Salience TR & 0.044  &    0.054 &  0.047 \\
AP & 0.033  &    0.038 &  0.035 \\
HAC & 0.031  &    0.035 &  0.034 \\
Rank & 0.031  &    0.037 &  0.034 \\
\hline % inserts single-line
\end{tabular}
\caption{System ROUGE performance.} % title name of the table
\label{tab:rouge}
\end{table}

\begin{figure*}
  \includegraphics[width=\textwidth]{rouge-time.png}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure*}

### 2015-02-26 experiments.tex
latest revision changes

@@
 At test time, we predict a sentence's salience using the average predictions
-of the 23 other models---essentially 24 fold cross validation approach.
+of the 23 other models---essentially a leave-one-out approach.
 
@@
 
-\subsection{Nugget Evaluation}
-We also modify the evaluation used in the TREC TS track, 
-where systems were evaluated on the containment of event nuggets. 
-We match each system update to 0 or more nuggets, where an update matches
-a nugget if their semantic similarity is above a threshold.  
-Summaries are then evaluated on the precision and recall of the matches.
-More formally,
-\[ \mathrm{Prec.} = \frac{|S_n|}{|S|} \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}\] 
-where S is the set of system updates, $S_n$ is the set of nuggets mapped updates in $S$,
-$N$ is the set of event nuggets, and $|\cdot|$ is the size of the set.
-A precision of 1 indicates that every sentence contains one unique nugget.
-Because an update can map to more than one nugget, it is possible to receive a precision greater than 1. 
-A recall of 1 indicates that the summary has cover all nugget information for the event; the maximum
-attainable recall is 1.
-Determining 
-an optimal threshold is difficult so we evaluate at threshold values ranging
+\subsection{Average Gain and Comprehensiveness}
+
+We treat the event's nuggets as unique units of information.
+When a system adds an update to its summary, it is potentially adding some
+of this nugget information. It would be instructive to know how much unique
+and novel information each update is adding on average to the summary.
+To that end, we define
+
+\[ \mathrm{Avg. Gain} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
+\] 
+where $S$ is the set of system updates, 
+$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
+elements in the set.
+To compute the set $S_n$ we match each system update to 0 or more nuggets, 
+where an update matches a nugget if their semantic similarity is above 
+a threshold. $S_n$ results from the unique set of nuggets matched.
+Because an update can map to more than one nugget, it is possible to receive an
+average gain greater than 1. An average gain of 1 would indicate that every
+sentence was both relevant and contained a unique piece of information.
+
+Additionally, we can use the nuggets to measure the completeness of an update 
+summary. We define
+
+\[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]
+
+where $N$ is the set of event nuggets.
+A comprehensiveness of 1 indicates that the summary has cover 
+all nugget information for the event; the maximum
+attainable comprehensiveness is 1.
+
+Determining an optimal threshold to count matches is difficult so 
+we evaluate at 
+threshold values ranging
 from .5 to 1, where values closer to 1 are more conservative estimates of
 performance.
+A qualitative assessment of matches 
+suggests that semantic similarity values around .7 produce reasonable matches.
+For example, the nugget 
+\begin{blockquote}
+watches and warnings in effect for much of the Mid-Atlantic states and 
+southern New England 
+\end{blockquote}
+\noindent
+matches with the update 
+\begin{blockquote}
+A tropical storm warning remains in effect for much of the coasts 
+of North and South Carolina\end{blockquote}
+\noindent
+with a similarity of .7. 
+The average semantic similarity of manual matches performed 
+by TREC assessors was considerably lower at approximately .25, increasing
+our confidence in the automatic matches in the .5-1 range.
+
+An assessment of manual matchings done by TREC 
+assessors revealed that the average cosine similarity
 

### 2015-02-26 introduction.tex
latest revision changes

@@
 overwhelm those trying to monitor the situation. 
-Crisis informatics\cite{?} is dedicated to finding methods for sharing the
-right information in a timely fashion with relief organizations during a major crisis. 
+Crisis informatics \cite{?} is dedicated to finding methods for sharing the
+right information in a timely fashion during such an event.
 Research in this field has focused on
-human-in-the-loop approaches ranging from on the ground information gathering to
-crowdsourced reporting and disaster management.
+human-in-the-loop approaches ranging from on the ground information 
+gathering to crowdsourced reporting and disaster management.
 
-Multi-document summarization offers potential for enabling automatic updates of
-relevant, salient information at regular intervals. It would provide
-information even when human volunteers are unable to and would filter out
-unnecessary and irrelevant detail. 
+Multi-document summarization has the potential to assist the crisis 
+informatics community. 
+Automatic summarization could deliver
+relevant and salient information at regular intervals, 
+even when human volunteers are unable to. 
+Perhaps more importantly it could help filter out
+unnecessary and irrelevant detail when the volume of incoming information
+is large. 
 While methods for identifying,
@@
 \cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
-these experiments were not developed to handle streaming data from the large
-and heterogeneous environment of the modern web. 
+these experiments were not developed to handle streaming data from a
+ heterogeneous environment at web scale. 
 These methods also rely heavily on redundancy which is suboptimal for time
 sensitive domains where there is a high cost in delaying information.
-
-%\cite{?} to enable people on the ground to update
-%crisis interfaces with information about needs. Others have used crowd
-%sourcing \cite{?},
-
-%At times of crisis, however, social media can overwhelm current information
-%systems with the quantity of information, much of which is irrelevant,
-%unnecessary detail, or out of date.
-
-%Neither were they developed to address the
-%information needs that arise during  crisis
-%situations. 
-%KM - this last sentence sounds somewhat negative.
-%there is still a need for robust and scalable 
-%methods for automatic summarization.
 
@@
 a clustering based multi-document summarization system. 
-Our system is better able to adapt to
+Not only does incorporating salience produce more relevant summaries,
+our system is better able to adapt to
 dynamic changes in input volume which adversely 
-effect methods of summarization that use either pairwise
-comparisons as a proxy for salience or predict salience but
-ignore redundancy in the input alone.
-
-%We train a regression
-%model to predict sentence salience and use these predictions to bias the
-%formation of sentence clusters around more salient regions in the input space
-%using affinity propagation (AP) clustering.  We adapt AP to incorporate
-%the salience predictions
-%as well as pairwise similarities among input sentences to identify
-%\emph{exemplar} sentences, which we use as our summary output.
-
-%ignoring features of importance that are intrinsic to the
-%sentences themselves.
-
-%KM - Chris - add a short overview on the kinds of things we measure and
-%results so people are encouraged to read ahead. Keep it general.
+effect methods that use redundancy as a proxy for salience. 
+%or predict salience but
+%ignore redundancy in the input alone.
 
 In addition to the tight integration between clustering and salience
-prediction, our approach also exploits knowledge about disaster to determine
+prediction, our approach also exploits knowledge about the event to determine
 salience. Thus, salience does not just represent importance within a set of
-documents; it also represents both how typical this sentence is of the input 
+documents; it also represents both how typical a sentence is of the  
 event
-type (i.e., disaster, hurricane, tornado) and whether it specifies information
-about this particular disaster. 
+type (i.e., industrial accident, hurricane, riot) and whether it specifies 
+information
+about this particular event. 
 Our feature representation includes a set of language models, one for each
-disaster type, to measure typicality of the sentence for the current event 
-type, the distance of mentioned locations from the center of
-the disaster, and the change in word frequencies over the time of the event.
+event type, to measure the typicality of the sentence with regard to the 
+current event, the distance of mentioned locations from the center of
+the event, and the change in word frequencies over the time of the event.
+While we ground these features in the domain of disaster, their templates
+are generally applicable to many update summarization domains.
 % to represent its likelihood of referring to the input disaster. 
@@
 
-Our approach achieves suprerior ROUGE scores compared to multiple baselines.
-Additionally, we introduce a novel method for determining the 
-coverage of atomic facts in a summary, and show our system's
-superiority on this metric.
+Our approach achieves a statistically significant improvement in 
+ ROUGE scores compared to multiple baselines.
+Additionally, we introduce novel methods for estimating the average information
+gain each update provides and how completely the update summary covers 
+the event it is tracking; our system's updates contain more relevant
+information on average than the competeing baselines.
 

### 2015-02-26 methods.tex
latest revision changes

@@
-In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
+%In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
 We provide a brief overview of the problem.
@@
 \begin{enumerate*}[label=\itshape\alph*\upshape)]
-  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
-  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
-  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
-  presented in temporal order, and \item an evaluation time period of interest,
-      $(\stime,\etime)$.  \end{enumerate*} While processing documents
+  \item a short query defining the event to be tracked (e.g. `Hurricane Sandy'), 
+  \item an event category defining the type of event to be tracked (e.g. `hurricane'), 
+  \item a stream of time-stamped documents %, $(\doc_0, \doc_1,\ldots,\doc_T)$,
+  presented in temporal order, and \item an evaluation time period of interest.
+    %  $(\stime,\etime)$.  
+      \end{enumerate*} While processing documents
       throughout the time period of interest, the system must output sentences,
-      known as updates, which are \emph{relevant} to the query,
-      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
-      and \emph{timely} with respect to when the update occurred.  Precisely
-      how to measure these properties will be discussed in Section
-      \ref{sec:methods}.  In our version of this problem, we assume that that
-      the system receives one batch of new sentence-segmented documents every
-      hour throughout the period of interest.
+      known as updates.
+      Updates should not only be relevant to the event but also salient,
+      i.e. worthy of
+      reporting.
+      For most domains, updates also need to be 
+      timely and novel, i.e. users need to be informed of changes in the 
+      event as quickly as possible.
+      Finally, in aggregate the updates must comprehensively summarize the 
+      event.
 
-We address disaster types such as terrorism and mass shootings (e.g., the 2012
-shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
-(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
-Arab spring). The size and scope of the events varies considerably. Hurricane
-Sandy, for example, effected multiple countries over weeks while the
-2012 Aurora shooting was contained to a single location spanning several hours.
+In order to quantify the salience of a sentence with respect to an event,
+we assume the event can be represented as a set of atomic facts, or nuggets.
+Nuggets are brief and important text snippets that should be conveyed
+by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
+\setlength{\fboxsep}{10pt}
+\begin{figure}
+    \fbox{\parbox{6.9cm}{
+        $\cdot$ hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay\\
+
+$\cdot$ Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief\\
+
+        $\cdot$ over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
+ }}
+ \caption{Example nuggets from Hurricane Sandy.\label{fig:nuggets}}
+\end{figure}
+
+Given an update $u$ and a set of nuggets $N$ we define the salience of 
+$u$ as,
+\[\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
+\operatorname{sim}(u, n) \]
+where $\operatorname{sim}(\cdot)$ is the semantic similarity \cite{?} (i.e.
+the cosine similarity of latent vectors associated with the update and 
+nugget text). Since we will not have nuggets for new events, we train
+a regression model to predict salience using features derived from 
+the updates themselves
+and the document stream.
+
+     % \ref{sec:methods}.  In our version of this problem, we assume that that
+     % the system receives one batch of new sentence-segmented documents every
+     % hour throughout the period of interest.
+\subsection{Update Summarization}
+
+Our system archictecture follows a simple pipeline design where each
+stage provides an additional level of processing or filtering of the input
+sentences.
+At the highest level, we begin with an empty update summary $U$.
+At each hour we receive a new batch of sentences $b_t$ from the stream
+and perform the following actions:
+\begin{enumerate}
+    \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
+  \item select a set of exemplar sentences (by combining clustering with 
+      salience predictions) (Section~\ref{sec:exsel}),
+  \item add the most novel and salient exemplars to $U$ (Section~\ref{sec:upsel}).
+\end{enumerate}
+
+The resultant list of updates $U$ is our summary of the event.
+Step 2 uses the affinity propagation algorithm to incorporate the salience
+predictions from step 1. For this reason we dub our summarization 
+system the \textbf{AP+Salience} model.
 
@@
 
-Figure \ref{alg:temporal-summarization} outlines our general update
-summarization algorithm.  At each hour, the system processes each input 
-sentence batch $S_t$.
-We predict the salience $P$ for all input sentences $s\in S_t$.
-Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
-obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
-to be updates and add those to our set of summary updates $U$.
-
-\begin{algorithm}%[H]
- \KwData{ 
-     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
-     batches\\
- \KwResult{U --- a list of updates, i.e. the update summary} 
-}
- ~\\
- Initialize empty list $U$ of updates\\
-    \For{$t \gets \stime,\ldots,\etime $}{
-        $P \gets \operatorname{PredictSalience}(S_t)$\\
-        $E \gets \operatorname{APCluster}(P, S_t)$ \\
-        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
-        $U \gets U \cup U_t$
-}
- \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
-\end{algorithm}
+%Figure \ref{alg:temporal-summarization} outlines our general update
+%summarization algorithm.  At each hour, the system processes each input 
+%sentence batch $S_t$.
+%We predict the salience $P$ for all input sentences $s\in S_t$.
+%Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
+%obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
+%to be updates and add those to our set of summary updates $U$.
+%
+%\begin{algorithm}%[H]
+% \KwData{ 
+%     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
+%     batches\\
+% \KwResult{U --- a list of updates, i.e. the update summary} 
+%}
+% ~\\
+% Initialize empty list $U$ of updates\\
+%    \For{$t \gets \stime,\ldots,\etime $}{
+%        $P \gets \operatorname{PredictSalience}(S_t)$\\
+%        $E \gets \operatorname{APCluster}(P, S_t)$ \\
+%        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
+%        $U \gets U \cup U_t$
+%}
+% \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
+%\end{algorithm}
 
 \subsection{Salience Prediction}
+\label{sec:salpred}
 
 \subsubsection{Model}
-%KM - I think one of the words ``sentence'' below should be something else, btu
-%not sure what you meant. Ah.. I think ``salience''?
-%We adopt Gaussian process regression in order to predict sentence sentence
-We adopt Gaussian process regression in order to predict sentence salience
-\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
+
+Nuggets represent ground truth data which would not be available at test time;
+we need to find some function $f : \mathbb{R}^d \rightarrow \mathbb{R}$ 
+such that 
+\[ \operatorname{salience}(s) = f(x_s) \]
+where $x_s$ is a $d$-dimensional feature vector for sentence $s$. 
+Given that we have many thousands of sentences of varying relevance to each
+event in our stream and salience scores are derived in an unsupervised fashion
+(assuming event nuggets are known), we can generate large amounts of training
+data for our model by simply sampling sentences 
+from the stream and computing the maximum
+semantic similarity to the nugget set.
+
+To learn $f$ and predict salience,
+we adopt Gaussian process (GP) regression 
+\cite{rasmussen:gaussian-process-book}.  GP regressors are a
 class of data-driven, non-parametric model generalizing the multi-variate
-Gaussian to the infinite dimensional setting.  Gaussian processes are general
-and are state of the art for many regression tasks.  A full treatment of
-Gaussian process regression is beyond the scope of this article and can be
-found in standard textbooks.
+Gaussian to the infinite dimensional setting.  GP's are general
+and are state of the art for many regression tasks.  
+%A full treatment of
+%Gaussian process regression is beyond the scope of this article and can be
+%found in standard textbooks.
 
-As a non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.
-In our experiments, we used a radial basis function (RBF) kernel.
-Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
+GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
+case sentences.  In our experiments, we used a radial basis 
+function (RBF) kernel.  Given two featurized sentences, 
+$\features$ and $\features^\prime$, the RBF kernel is defined as,
 \begin{align*}
@@
 The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
+Kernel matrices are also closed under addition, i.e. the sum of kernels
+is also a kernel; we exploit this fact by organizing our features into 
+meaningful groups with their own RBF kernel using the sum of their kernels 
+to parameterize the model.
 
-\subsubsection{Training Data}
-\ckcomment{Explain how we train the model here...}
-%In order to train the salience model, we need to quantify the notion of salience. We assume for each event we have a set of atomic facts, or nuggets.
-
-%
 \begin{figure}[t!]
@@
 Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
-These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
+%These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
 
@@
 These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
-Since training is one on grammatical English, some of these features help
+Since training is done on grammatical English, some of these features help
 to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
@@
 
-\subsection{Affinity Propagation Clustering}
-\ckcomment{Explain AP and how salience gets plugged in here}
+\subsection{Exemplar Selection}
+\label{sec:exsel}
+
+We combine the output of our salience prediction model with the affinity
+propagation algorithm to identify a set of exemplar sentences 
+for each input batch. 
+Affinity propagation (AP) is a clustering algorithm
+that identifies a subset of datapoints as exemplars and forms clusters
+by assigning the remaining points to one of the exemplars. AP can be seen
+as a generalization of other $k$-centers algorithms that attempts to 
+maximize the net similarity objective 
+\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
++ \sum_{i : i = e_i}^n \operatorname{pref}(e_i)  \]
+where $e_i$ is the exemplar of the $i$-th data point, and functions
+$\operatorname{sim}$ and $\operatorname{pref}$ express the pairwise similarity
+of data points and the apriori preference of a data point to be an exemplar
+respectively. There are very 
+few restrictions on the nature of the similarities and preferences
+other than that they be real-valued.
+
+AP has several useful properties and interpretations. Chiefly, the number
+of clusters $k$ is not a model hyper-parameter. Given that our task requires
+clustering many batches of streaming data, searching for an optimal $k$ 
+would be computationally prohibitive. With AP, $k$ is determined by the
+similarities and preferences of the data. Generally lower preferences will
+result in fewer clusters.
+
+In our summarization system we use the output of our salience model as 
+the preferences, i.e.
+$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
+function we use semantic similarity. Recall that $\operatorname{salience}(s)$
+is a prediction of the semantic similarity of $s$ to information about the 
+event be summarized, i.e. the set of event nuggets.
+Intuitively, when maximizing objective function $\mathcal{S}$, AP must balance
+between best representing the input data and representing the most salient
+input. Additionally, when the level of input is high but the salience
+predictions are low, the preference term will guide AP toward a solution
+with fewer clusters; vice-versa when input is very salient on average but
+the volume of input is low. The adaptive nature of our model differentiates
+our method from most out update summarization systems.
+
 \subsection{Update Selection}
-\ckcomment{Explain thresholds}
+\label{sec:upsel}
+
+After exemplar sentences have been identified, two more filters are applied 
+before updating the summary. The exemplars are examined
+sequencially in order of highest salience. First, a salience threshold 
+is applied, and any exemplars whose salience is less than
+$\lambda_{sal}$ are discarded. Second, a similarity threshold 
+is applied, where the exemplar is discarded if its
+maximum semantic similarity to any previous updates in the summary is
+greater than $\lambda_{sim}$. If the exemplar passes these two filters,
+it is selected as an update and added to the summary.

### 2015-02-26 relatedwork.tex
latest revision changes

@@
 official press releases related to Hurricane Wilma  in 2005 using Rouge score
-between the system summary and a manually generated target summary.
+between the system summary and a manually generated target summary. In addition
+to natural disasters, we evaluate our system on a variety of event types 
+ such as terrorism and mass shootings (e.g., the 2012
+shooting in Aurora, Col.), accidents
+(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
+Arab spring). The size and scope of the events varies considerably. Hurricane
+Sandy, for example, affected multiple countries over weeks while the
+2012 Aurora shooting was contained to a single location spanning several hours.
 
-\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}
+%\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

### 2015-02-26 results.tex
latest revision changes

@@
 \subsection{Nugget Evaluation}
+
+Figure~\ref{fig:nuggets} shows the average gain across a range of 
+similarity thresholds, where thresholds closer to 1 are more conservative
+estimates. The ranking of the systems remains relatively constant across the 
+sweep with 
+AP+Salience and AP+Salience TR attaining comparable performance, as well
+as beating all baseline systems.
+Predicting salience in general is helpful for keeping a summary on topic as
+the Rank by salience approach out performs the clustering only approaches
+on average gain.
+
+When looking at the comprehensiveness of the summaries AP outperforms
+both AP+Salience variants. The compromise encoded in the AP+Salience objective
+function, between being representitive and being salient, is seen clearly here
+where the performance of the AP+Salience methods is lower bounded by 
+the salience focused Rank
+system and upper bounded by the clustering only AP system.
 
@@
 
-\begin{figure*}
-  \includegraphics[width=\textwidth]{rouge-time.png}
+\begin{figure}
+  \includegraphics[width=3.2in,height=5in]{rouge-time.png}
 \caption{System ROUGE-1 performance over time.}
 \label{fig:trouge}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=3.2in,height=5in]{nugget-metrics.png}
+\caption{Nugget performance.}
+\label{fig:nperf}
+\end{figure}

### 2015-02-26 experiments.tex
edits; nuggets

@@
 This data includes 25 events and event metadata (e.g., a user
-search query for the event, the event type, and event time frame).
-Additionally, TREC assessors constructed a set of ground truth facts, or 
-nuggets, for each event. The events took place from 20??-20?? and had 
-an associated Wikipedia (WP) page. The WP pages were edited and updated 
-concurrently with the events, so the diffs between revisions produce a similar
-ouput to the ideal update summarization system. Assessors analyzed the diffs
-and selected phrases that were important enough to include in an update 
-summary.  
+search query for the event, the event type, and event timeframe).  All events
+occurred during the span of the corpus.
+
+NIST assessors constructed a set of ground truth nuggets for each event. In order
+to accomplish this, for each event, assessors were provided with the
+revision history of the Wikipedia page associated with the.  For example, 
+the revision history for the Wikipedia page for `Hurricane Sandy' will 
+contain text additions including those related to individual nuggets.  The assessment
+task involves reviewing the Wikipedia revisions in the evaluation timeframe 
+and marking the text additions capturing a new, unique nugget.  More detail
+on this process can be found in the track description \cite{?}.
+
 For the document stream, we use the news portion of the

### 2015-02-26 introduction.tex
edits; nuggets

@@
 While we ground these features in the domain of disaster, their templates
-are generally applicable to many update summarization domains.
+are generally applicable to any update summarization task where queries can be classified into types.
 % to represent its likelihood of referring to the input disaster. 

### 2015-02-26 methods.tex
edits; nuggets

@@
 we assume the event can be represented as a set of atomic facts, or nuggets.
-Nuggets are brief and important text snippets that should be conveyed
+Nuggets are brief and important text snippets that represent sub-event that should be conveyed
 by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
@@
 \operatorname{sim}(u, n) \]
-where $\operatorname{sim}(\cdot)$ is the semantic similarity \cite{?} (i.e.
+where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
 the cosine similarity of latent vectors associated with the update and 
-nugget text). Since we will not have nuggets for new events, we train
-a regression model to predict salience using features derived from 
-the updates themselves
-and the document stream.
+nugget text \cite{?}. % We train
+% a regression model to predict salience using features derived from
+% the updates themselves
+% and the document stream.
 
@@
 
-Our system archictecture follows a simple pipeline design where each
+Our system architecture follows a simple pipeline design where each
 stage provides an additional level of processing or filtering of the input
@@
 Step 2 uses the affinity propagation algorithm to incorporate the salience
-predictions from step 1. For this reason we dub our summarization 
-system the \textbf{AP+Salience} model.
+predictions from step 1. We refer to this summarization 
+system as the \textbf{AP+Salience} model.
 
@@
 before updating the summary. The exemplars are examined
-sequencially in order of highest salience. First, a salience threshold 
+sequentially in order of highest salience. First, a salience threshold 
 is applied, and any exemplars whose salience is less than

### 2015-02-26 dataset.tex
Latest revision ck

For the document stream, we use the news portion of the
 2014 Trec KBA Stream Corpus.
The documents from this corpus come from hourly crawls of the web covering 
 ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}

Our experiments also make use of the Trec Temporal Summarization (TS) Track
 data from 2013 and 2014. 
This data includes 25 events and event metadata (e.g., a user
search query for the event, the event type, and event evaluation timeframe).  
All events occurred during the time span of the Trec KBA Stream Corpus.

Along with the metadata, NIST assessors constructed a set of ground truth nuggets for each event. 
Nuggets are brief and important text snippets that represent sub-events that should be conveyed
by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
In order
to accomplish this, for each event, assessors were provided with the
revision history of the Wikipedia page associated with the event.  
For example, 
the revision history for the Wikipedia page for `Hurricane Sandy' will 
contain text additions including those related to individual nuggets.  The assessment
task involves reviewing the Wikipedia revisions in the evaluation timeframe 
and marking the text additions capturing a new, unique nugget.  More detail
on this process can be found in the track description \cite{?}.

### 2015-02-26 experiments.tex
Latest revision ck

--- 2015-02-26 experiments.tex
@@
-We evaluate our system on two metrics: ROUGE \cite{?}, the most common automatic
-summarization method and a modified method adapted 
-from the Text Retrieval Conference Temporal Summarization (TREC-TS) track \cite{?}.
-
-\subsection{Dataset}
-
-Both evaluations make use of the TREC TS data from 2013 and 2014. 
-This data includes 25 events and event metadata (e.g., a user
-search query for the event, the event type, and event time frame).
-Additionally, TREC assessors constructed a set of ground truth facts, or 
-nuggets, for each event. The events took place from 20??-20?? and had 
-an associated Wikipedia (WP) page. The WP pages were edited and updated 
-concurrently with the events, so the diffs between revisions produce a similar
-ouput to the ideal update summarization system. Assessors analyzed the diffs
-and selected phrases that were important enough to include in an update 
-summary.  
-For the document stream, we use the news portion of the
- 2014 TREC Stream Corpus.
-The documents from this corpus come from hourly crawls of the web covering 
- ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}
+We evaluate our system on two metrics: ROUGE \cite{?}, a common automatic
+summarization method and an evaluation of system expected gain 
+and comprehensiveness---metrics adapted 
+from the Trec TS track \cite{?}.
 
@@
 
-Of the 25 events in the TREC TS data, 24 are covered by the Stream Corpus.
+Of the 25 events in the TREC TS data, 24 are covered by the news portion 
+of the Trec KBA Stream Corpus.
 From these 24, we set aside three events to use as a development set.
@@
 the model's log likelihood using ???.
+
+We perform a leave-one-out evaluation of each event.
 At test time, we predict a sentence's salience using the average predictions
-of the 23 other models---essentially a leave-one-out approach.
+of the 23 other models.
 
@@
 systems at 12 hour intervals using the same regime as above. 
-The model summaries in this case are retrospective, and so we can see how quickly
-each system takes to achieve its maximum performance.
+The model summaries in this case are retrospective, 
+and so we can see how quickly
+each system is able to achieve its maximum performance.
 
-\subsection{Average Gain and Comprehensiveness}
+\subsection{Expected Gain and Comprehensiveness}
 
-We treat the event's nuggets as unique units of information.
+NIST developed a series metrics for evaluating update summarization systems
+as part of the Trec TS track.
+We present results on two of these metrics, the expected gain and 
+comprehensiveness.
+
+\textbf{Expected Gain } We treat the event's nuggets as unique units of 
+information.
 When a system adds an update to its summary, it is potentially adding some
@@
 
-\[ \mathrm{Avg. Gain} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
+\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
 \] 
@@
 Because an update can map to more than one nugget, it is possible to receive an
-average gain greater than 1. An average gain of 1 would indicate that every
+expected gain greater than 1; values greater than one can be thought of as a 
+compression ration, i.e. an expected gain of 10 indicates the average update
+contains 10 unique nuggets on average and that the summary is a 10~:~1 
+compression of the nugget information.
+An expected gain of 1 would indicate that every
 sentence was both relevant and contained a unique piece of information.
 
-Additionally, we can use the nuggets to measure the completeness of an update 
+\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure the completeness of an update 
 summary. We define
@@
 
+Update-nugget matches are computed automatically, where a
+match exists if the semantic similarity of the update/nugget pair is above 
+a threshold.
 Determining an optimal threshold to count matches is difficult so 
@@
 performance.
-A qualitative assessment of matches 
+A manual inspection of matches 
 suggests that semantic similarity values around .7 produce reasonable matches.
@@
 The average semantic similarity of manual matches performed 
-by TREC assessors was considerably lower at approximately .25, increasing
-our confidence in the automatic matches in the .5-1 range.
-
-An assessment of manual matchings done by TREC 
-assessors revealed that the average cosine similarity
+by NIST assessors was considerably lower at approximately .25, increasing
+our confidence in the automatic matches in the .5--1 range.
 
@@
 
-\textbf{AP+Salience Time Ranked} : This model is everywhere the same as AP+Salience except at the update selection stage, 
-where the algorithm selects exemplars in temporal order rather than salience order.
+\textbf{Affinity Propagation only (AP) } The purpose of this model
+is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
+clustering component, but without the salience information.
+In this model, input sentences are apriori equally likely to be exemplars;
+%clustering algorithm without the salience predictions.
+the salience values are uniformly set as the median value of the 
+input similarity scores, as is commonly used in the AP literature \cite{?}.
+After clustering a sentence batch, the exemplars are examined in order
+of increasing time since event start and selected as updates if their
+maximum similarity to the previous updates is less than $\lambda_{sim}$,
+as in the novelty filtering stage of AP+Salience.
 
-\textbf{AP} : In this model we use the AP clustering algorithm without the salience predictions.
-The preferences are chosen as the median value of the ? matrix, as is commonly used with AP \cite{?}.
-The update selection procedes in time order. As in the AP+Salience model we tune a similarity threshold,
-and ignore exemplar sentences whose maximum similarity to a previous update is above the threshold.
-
-\textbf{Rank by Salience} : In this model we rank the sentences by salience and select the updates at each hour that are
-above a salience threshold and are below a similarity threshold (similar to the AP+Salience model).
-
-\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We also provide another clustering baseline, HAC using ?-linkage. 
-This method was chosen over other clustering approaches because the number of clusters 
-is not an explicit hyper parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
+\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
+We include this baseline to show that AP+Salience is not just an improvement
+over AP but centrality driven methods in general.
+HAC was chosen over other clustering approaches because the number of clusters 
+is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
 using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
 Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
-Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold
-(like the previous models).
+Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.
+
+\textbf{Rank by Salience (RS)} 
+We also isolate the impact of our salience model in order to demonstrate 
+the that the fusion of clustering and salience prediction improves over
+predicting salience alone.
+In this model we predict the salience of sentences as in step 1 for 
+AP+Salience. %However, rather than clustering, we simply rank the sentences
+%by salience.
+We omit the clustering phase (step 2).
+Updates are selected identically to step 3 of AP+Salience, proceeding 
+in order of decreasing salience, selecting updates that are above a salience 
+threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
+with respect to the previously selected updates.
+
+\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
+to the
+AP+Salience model except at the novelty filter stage.
+Within the novelty filtering stage, 
+AP+Salience inspects exemplars in order of decreasing salience when computing
+similarity to previous updates. Because the order in which we add updates
+affects future similarity calculations, we evaluate an alternate ordering 
+scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
+time since event start. The purpose of this model is to give preference to 
+more recent updates, since timelyness is one of the desirable traits of our update
+summaries.

### 2015-02-26 feature_table.tex
Latest revision ck

--- 2015-02-23 methods.tex
@@
-In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
-We provide a brief overview of the problem.
-Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.
-
-An update summarization system takes as input 
-\begin{enumerate*}[label=\itshape\alph*\upshape)]
-  \item a short query $\query$ defining the event to be tracked (e.g. `Hurricane Sandy'), 
-  \item an event category $\eventcategory$ defining the type of event to be tracked (e.g. `hurricane'), 
-  \item a stream of time-stamped documents, $(\doc_0, \doc_1,\ldots,\doc_T)$,
-  presented in temporal order, and \item an evaluation time period of interest,
-      $(\stime,\etime)$.  \end{enumerate*} While processing documents
-      throughout the time period of interest, the system must output sentences,
-      known as updates, which are \emph{relevant} to the query,
-      \emph{comprehensive} with respect to the event, \emph{novel} to the user,
-      and \emph{timely} with respect to when the update occurred.  Precisely
-      how to measure these properties will be discussed in Section
-      \ref{sec:methods}.  In our version of this problem, we assume that that
-      the system receives one batch of new sentence-segmented documents every
-      hour throughout the period of interest.
-
-We address disaster types such as terrorism and mass shootings (e.g., the 2012
-shooting in Aurora, Col.), natural disasters (e.g., Hurricane Sandy), accidents
-(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
-Arab spring). The size and scope of the events varies considerably. Hurricane
-Sandy, for example, effected multiple countries over weeks while the
-2012 Aurora shooting was contained to a single location spanning several hours.
-
-%\kmcomment{This was as best as I could do on this. These are the example
-%    categories from the wikipedia page. However you refer in the language model
-%    description to the event type earthquake which is more specific. Also, in
-%    the TREC data you refer to event types that are specified in the metadata.
-%So I'm wondering how many different categories you have and where they come
-%from.}
-
-%\fdcomment{include description of nuggets here.} 
-%%\kmcomment{We should jointly
-%    discuss what goes in data and what goes in problem definition. I have moved
-%    event types here because I think knowing how many types and examples of
-%    them would be helpful upfront. I think where nuggets and wikipedia pages go
-%is questionable.}
-
-Figure \ref{alg:temporal-summarization} outlines our general update
-summarization algorithm.  At each hour, the system processes each input 
-sentence batch $S_t$.
-We predict the salience $P$ for all input sentences $s\in S_t$.
-Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
-obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
-to be updates and add those to our set of summary updates $U$.
-
-\begin{algorithm}%[H]
- \KwData{ 
-     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
-     batches\\
- \KwResult{U --- a list of updates, i.e. the update summary} 
-}
- ~\\
- Initialize empty list $U$ of updates\\
-    \For{$t \gets \stime,\ldots,\etime $}{
-        $P \gets \operatorname{PredictSalience}(S_t)$\\
-        $E \gets \operatorname{APCluster}(P, S_t)$ \\
-        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
-        $U \gets U \cup U_t$
-}
- \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
-\end{algorithm}
-
-\subsection{Salience Prediction}
-
-\subsubsection{Model}
-%KM - I think one of the words ``sentence'' below should be something else, btu
-%not sure what you meant. Ah.. I think ``salience''?
-%We adopt Gaussian process regression in order to predict sentence sentence
-We adopt Gaussian process regression in order to predict sentence salience
-\cite{rasmussen:gaussian-process-book}.  Gaussian process regressors are a
-class of data-driven, non-parametric model generalizing the multi-variate
-Gaussian to the infinite dimensional setting.  Gaussian processes are general
-and are state of the art for many regression tasks.  A full treatment of
-Gaussian process regression is beyond the scope of this article and can be
-found in standard textbooks.
-
-As a non-parametric, the Gaussian process relies on a  covariance matrix $\kernelMatrix$, measuring the affinity between pairs of instances, in our case candidate sentences.
-In our experiments, we used a radial basis function (RBF) kernel.
-Given two featurized sentences, $\features$ and $\features^\prime$, the RBF kernel is defined as,
-\begin{align*}
-        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
-\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
-\end{align*}
-where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
-
-\subsubsection{Training Data}
-\ckcomment{Explain how we train the model here...}
-%In order to train the salience model, we need to quantify the notion of salience. We assume for each event we have a set of atomic facts, or nuggets.
-
-%
 \begin{figure}[t!]
@@
 \end{figure}
-
-\subsubsection{Features}
-We want our model to be predictive across different kinds of events so we avoid lexical features.
-Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
-These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
-
-\paragraph{Basic Features}
-%KM - Would be good to have quick justification of these features. I added a
-%sentence. Feel free to edit or remove.
-
-We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
-Since training is one on grammatical English, some of these features help
-to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
-Others help to more heavily weight important sentences (e.g., that appear in
-prominent positions such as paragraph initial or article initial).
-
-\paragraph{Query Features}
-
-Query features measure the relationship between the sentence and the event query and type.
-These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
-For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
-\paragraph{Language Model Features}\label{subsubsec:lm}
-Language models allow us to measure the likelihood of a sentence having been produced from a particular source.
-We consider two types of language model features.
-The first model is estimated from a corpus of generic news articles.
-This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).
-The second model is estimated from text specific to our event types.
-For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.
-These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g.
-most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
-%KM - Note: Someplace the exact list of event types should appear. Probably not
-%here.
-%KM - I note you have it in a later section but it is labeled as data you use
-%to train language models and semantic similarity. I think it would be good to
-%have up front in definition of task.
-
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
-
-\paragraph{Geographic Relevance Features}
-
-Locations are identified using a named entity tagger.
-For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.
-We then compute its distance to that of the event location.
-It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-%KM - Probably should say how you determine event location. Some events move.
-%KM - In your figure you only include one distance.
-
-\paragraph{Temporal Relevance Features}
-
-Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
-the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.
-
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
-We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
-%\begin{align*}
-%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-%where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-%\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-%\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-%\end{align*}
-
-[\textrm{Simpler explanation goes here}]
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.
-
-% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-% of an input space $\mathcal{X}$ to the reals,
-%
-% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
-% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
-%  an $n$-dimensional vector whose elements are evaluations of the function $f$
-% at points $x_i \in \mathcal{X}$.
-% We say $p(f)$ is a Gaussian process if for any finite subset
-% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
-% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
-% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-% set to 0, leaving the distribution to be completely characterized by the
-% kernel function on the data.
-%
-% In the regression setting, we typically have a response variable $y$ that
-% is the sum of our model prediction  and
-% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
-% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-% two distributions
-% of principal interest are the marginal likelihood
-% $p(\mathbf{y}|\mathbf{X}) =
-% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-% distribution,
-%
-% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-%
-% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-% response, and
-% \begin{align*}
-% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-% \boldsymbol{\sigma}^2_* &
-% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-% \mathbf{K}_*^T + \sigma^2\\
-% \end{align*}.
-%
-% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
-% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-%
-%
-
-\subsection{Affinity Propagation Clustering}
-\ckcomment{Explain AP and how salience gets plugged in here}
-\subsection{Update Selection}
-\ckcomment{Explain thresholds}

### 2015-02-26 methods.tex
Latest revision ck

@@
 
-An update summarization system takes as input 
+Our update summarization system takes as input 
 \begin{enumerate*}[label=\itshape\alph*\upshape)]
@@
       \end{enumerate*} While processing documents
-      throughout the time period of interest, the system must output sentences,
-      known as updates.
+      throughout the time period of interest, the system outputs sentences,
+      which we refer to as updates.
       Updates should not only be relevant to the event but also salient,
@@
 
-In order to quantify the salience of a sentence with respect to an event,
-we assume the event can be represented as a set of atomic facts, or nuggets.
-Nuggets are brief and important text snippets that represent sub-event that should be conveyed
-by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
+\begin{figure}
 \setlength{\fboxsep}{10pt}
-\begin{figure}
     \fbox{\parbox{6.9cm}{
-        $\cdot$ hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay\\
+        -- hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay
 
-$\cdot$ Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief\\
+        -- Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief
 
-        $\cdot$ over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
+        -- over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
  }}
@@
 
-Given an update $u$ and a set of nuggets $N$ we define the salience of 
+In order to quantify the salience of an update with respect to an event,
+we match the update to its most similar event nugget.
+%we assume the event can be represented as a set of atomic facts, or nuggets.
+Formally, given an update $u$ and a set of nuggets $N$ we define the salience of 
 $u$ as,
@@
 sentences.
-At the highest level, we begin with an empty update summary $U$.
+We begin with an empty update summary $U$.
 At each hour we receive a new batch of sentences $b_t$ from the stream
@@
 
-Nuggets represent ground truth data which would not be available at test time;
-we need to find some function $f : \mathbb{R}^d \rightarrow \mathbb{R}$ 
-such that 
-\[ \operatorname{salience}(s) = f(x_s) \]
-where $x_s$ is a $d$-dimensional feature vector for sentence $s$. 
-Given that we have many thousands of sentences of varying relevance to each
-event in our stream and salience scores are derived in an unsupervised fashion
-(assuming event nuggets are known), we can generate large amounts of training
-data for our model by simply sampling sentences 
-from the stream and computing the maximum
-semantic similarity to the nugget set.
+When evaluating our summarization system
+we do not have access to event nuggets. However, we can use the nuggets
+in our training data to learn a model to predict salience for sentences
+from new events.
+For each event in our training data, we sample a set of sentences and  each 
+sentence's salience is computed, i.e. we compute the maximum semantic 
+similarity to the event's nuggets.
+We now have a training set of set sentences and the target values 
+(their salience) to predict.
 
-To learn $f$ and predict salience,
-we adopt Gaussian process (GP) regression 
+In order to fit a regression model, the sentences are represented as feature
+vectors (described in the next section).
+We fit a Gaussian process (GP) regression model to this data.
 \cite{rasmussen:gaussian-process-book}.  GP regressors are a
 class of data-driven, non-parametric model generalizing the multi-variate
-Gaussian to the infinite dimensional setting.  GP's are general
-and are state of the art for many regression tasks.  
-%A full treatment of
-%Gaussian process regression is beyond the scope of this article and can be
-%found in standard textbooks.
-
+Gaussian to the infinite dimensional setting.  
+%GP's are general
+%and are state of the art for many regression tasks.  
 GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
 case sentences.  In our experiments, we used a radial basis 
-function (RBF) kernel.  Given two featurized sentences, 
-$\features$ and $\features^\prime$, the RBF kernel is defined as,
-\begin{align*}
-        \kernelMatrixij{\features}{\features^\prime}&= \sigma^2 \exp\left(- \frac{1}{2} 
-\sum_{i=1}^{\numfeatures} \frac{ (\featuresi{i}-\featuresi{i}^\prime)^2}{\ell_i^2} \right)
-\end{align*}
-where $\sigma$ and the $\ell_i$ are parameters we fit to our observed training data.
-The $\ell_i$ are feature dependent scaling parameters; once learned, they not only improve the accuracy of the model, but give us some introspection  into which features are more important.
-Kernel matrices are also closed under addition, i.e. the sum of kernels
-is also a kernel; we exploit this fact by organizing our features into 
-meaningful groups with their own RBF kernel using the sum of their kernels 
-to parameterize the model.
+function (RBF) kernel.  
+Our features fall naturally into five groups and we use a separate RBF kernel
+for each, using the sum of each feature group kernel as the final input
+to the GP model.
 
-\begin{figure}[t!]
-\begin{tabular}{| l |} 
-\hline
-\textbf{Basic Features}\\
-$\cdot$ Sent. position (normalized by doc. length) \\
-$\cdot$ Sent. length \\
-$\cdot$ Ratio of punc. to non-punc. chars. \\
-$\cdot$ Ratio of caps. to non-caps. chars. \\
-$\cdot$ Ratio of lowercase to other chars. \\
-$\cdot$ Ratio of uppercase to other chars. \\
-$\cdot$ \# of caps. words (normalized by \# of words)\\
-$\cdot$ \# of Person, Location, Org. Date, Number,\\
-$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
-$\;\;$ tags (normalized by \# of words)\\
-\hline
-\textbf{Query Features}\\
-$\cdot$ \% of query words covered by sent.\\
-$\cdot$ Total query matches.\\
-$\cdot$ Total event-type synonyms/hypernyms/ \\
-$\;\;$ hyponyms coverage.\\
-$\cdot$ Total event-type synonyms/hypernyms/ \\
-$\;\;$ hyponyms matches.\\
-\hline
-\textbf{Language Model Features}\\
-$\cdot$ Avg. token log probability (domain lang. \\
-$\;\;$ model)\\
-$\cdot$ Avg. token log probability (background \\
-$\;\;$ lang. model)\\
-\hline
-\textbf{Geo-tag Features}\\
-$\cdot$ Median document distance to nearest  \\
-$\;\;$ location cluster (current hour).\\
-$\cdot$ Median document distance to nearest  \\
-$\;\;$ location clusters (previous hour).\\
-$\cdot$ Distance of first location in doc. to nearest \\
-$\;\;$ location cluster (current hour).\\
-$\cdot$ Distance of first location in doc. to nearest \\
-$\;\;$ location cluster (previous hour).\\
-\hline
-\textbf{Temporal Features}\\
-$\cdot$ Avg. tf-idf at current time.\\
-$\cdot$ Change in avg. tf-idf since previous hour \\
-$\;\;$  (up to 24 hours).\\
-$\cdot$ Time since query/event start.\\
-\hline
-\end{tabular}
-\caption{Salience model features.}
-\end{figure}
+%include feature table
+\input{feature_table.tex}
 
 \subsubsection{Features}
-We want our model to be predictive across different kinds of events so we avoid lexical features.
+We want our model to be predictive across different event instances so we avoid lexical features.
 Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
-%These features are used to fit a Gaussian process regression model that can predict the similarity of a sentence to a gold summary \cite{preotiuc2013temporal}.
 
@@
 These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
-Since training is done on grammatical English, some of these features help
-to downweight sentences that are ungrammatical (e.g., have too many capitalized words or are too short).
-Others help to more heavily weight important sentences (e.g., that appear in
+The data stream comprises text extracted from raw html documents;
+these features help to downweight sentences that are not actually article 
+content (e.g. web page titles, links to other content) or
+more heavily weight important sentences (e.g., that appear in
 prominent positions such as paragraph initial or article initial).
@@
 These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
-For example, for event type \emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'', and ``aftershock''.
+For example, for event type \emph{earthquake},  we match sentence terms 
+``quake'', ``temblor'', ``seism'', and ``aftershock''.
+
 \paragraph{Language Model Features}\label{subsubsec:lm}
-Language models allow us to measure the likelihood of a sentence having been produced from a particular source.
-We consider two types of language model features.
-The first model is estimated from a corpus of generic news articles.
-This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps us to filter out text snippets which are not sentences (e.g., web page titles).
+Language models allow us to measure the likelihood of a sentence having been 
+produced from a particular source.  We consider two types of language model 
+features.  The first model is estimated from a corpus of generic news 
+articles \ckcomment{(we used the 199?-200? Associated Press and New York Times sections of the Gigaword corpus)}.  
+This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
+written in the newswire style.
+
 The second model is estimated from text specific to our event types.  
-For example, the language model for event type `earthquake' is estimated from Wikipedia pages under the category \emph{Category:Earthquakes}.
-These models are intended to detect sentences similar to those appearing in summaries of other events in the same category (e.g.
-most earthquake summaries are likely to include higher probability for ngrams including the token `magnitude').
+For each event type we create a corpus of related documents using pages
+and subcategories listed under a related Wikipedia category.
+For example, the language model for event type `earthquake' is estimated 
+from Wikipedia pages under the category \emph{Category:Earthquakes}.
+
+These models are intended to detect sentences similar to those appearing in 
+summaries of other events in the same category 
+(e.g. most earthquake summaries are likely to include higher probability for 
+ngrams including the token `magnitude').
+
+We use the SRILM toolkit to implement a 5-gram Kneser-Ney model for both
+the background language and model and the event specific language models.
+For each sentence we use the average token log probability under each model
+as a feature.
+
 %KM - Note: Someplace the exact list of event types should appear. Probably not
@@
 
-Locations are identified using a named entity tagger.
-For each location in a sentence, we obtain its latitude and longitude using the a publicly available geolocation service.
-We then compute its distance to that of the event location.
-It is possible for a sentence and an event to have multiple locations so we take as features the minimum, maximum, and average distance of all sentence-event location pairs.
-Distances are calculated using the Vincenty distance.
-%KM - Probably should say how you determine event location. Some events move.
-%KM - In your figure you only include one distance.
+There are two challenges to using geographic features. First we do not 
+know where the event is and second most sentences do not contain references
+to a location.
+We address the first issue by extracting all locations from sentences at the
+current hour and looking up their latitude and 
+longitude using a publicly available geolocation service. Since the document
+stream contains documents that are at least somewhat relevant to the event,
+we assume in aggregate the locations should give us a rough area of interest.
+The locations are clustered (using affinity propagation and uniform salience)
+and we treat the resulting cluster centers
+as the event locations for the current time.
+
+To address the second issue, 
+we compute geographic relevance features for the document as a whole and all
+sentences from that document receive the same feature value.
+Using the locations in each document, we compute the median distance to the 
+nearest event location. Because document position is a good indicator 
+of importance we also compute the distance of the first mentioned
+location to the nearest event location. Because some events can move, we also
+compute these distances to event locations from the previous hour.
 
@@
 
-Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
+We compute the IDF for each hour in our data stream. 
+For each sentence, the average TFIDF for the current hour $t$ is taken as a 
+feature. Additionally, we use the difference in average TFIDF from time $t$
+to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
+sentence have changed over the last 24 hours.
+
+The final temporal feature is the hours since the event started.
+%difference 
+%Using these IDF 
+%calculations, the avg TFIDF is calculated for
+%Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
 We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
@@
 
-[\textrm{Simpler explanation goes here}]
+%[\textrm{Simpler explanation goes here}]
 % 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
@@
 
-We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and hours since the event in question started as features.
+%We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and
 
@@
 that identifies a subset of datapoints as exemplars and forms clusters
-by assigning the remaining points to one of the exemplars. AP can be seen
-as a generalization of other $k$-centers algorithms that attempts to 
+by assigning the remaining points to one of the exemplars. AP attempts to 
 maximize the net similarity objective 
@@
 where $e_i$ is the exemplar of the $i$-th data point, and functions
-$\operatorname{sim}$ and $\operatorname{pref}$ express the pairwise similarity
+$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
+similarity
 of data points and the apriori preference of a data point to be an exemplar
-respectively. There are very 
-few restrictions on the nature of the similarities and preferences
-other than that they be real-valued.
+respectively. 
+AP differs from other $k$-centers algorithms in that it simultaneously 
+considers all data points as exemplars, making it less prone to finding
+local optima as a result of poor initialization. Furthermore, the 
+second term in $\mathcal{S}$ incorporates the individual importance of 
+data points as candidate exemplars; most other clustering algorithms only make
+use of the first term, i.e. the pairwise similarities between data points.
+
+%There are very 
+%few restrictions on the nature of the similarities and preferences
+%other than that they be real-valued.
 
@@
 
-In our summarization system we use the output of our salience model as 
-the preferences, i.e.
-$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
-function we use semantic similarity. Recall that $\operatorname{salience}(s)$
+%In our summarization system we use the output of our salience model as 
+%the preferences, i.e.
+%$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
+%function we use semantic similarity. 
+Recall that $\operatorname{salience}(s)$
 is a prediction of the semantic similarity of $s$ to information about the 
@@
 
-After exemplar sentences have been identified, two more filters are applied 
-before updating the summary. The exemplars are examined
-sequentially in order of highest salience. First, a salience threshold 
-is applied, and any exemplars whose salience is less than
-$\lambda_{sal}$ are discarded. Second, a similarity threshold 
-is applied, where the exemplar is discarded if its
+\textbf{Salience Filter } To 
+ensure that only the most salient updates are selected we apply a minimum
+salience threshold;
+after exemplar sentences have been identified, any exemplars whose salience is 
+less than $\lambda_{sal}$ are removed from consideration.
+
+\textbf{Novelty Filter } Next,
+to prevent adding updates that are redundant, we filter out exemplars
+that are too similar to previous updates.
+The exemplars are examined
+sequencially in order of decreasing salience and  a similarity threshold 
+is applied, where the exemplar is ignored if its
 maximum semantic similarity to any previous updates in the summary is
-greater than $\lambda_{sim}$. If the exemplar passes these two filters,
-it is selected as an update and added to the summary.
+greater than $\lambda_{sim}$.
+
+Exemplars that pass these thresholds are added selected as updates and added
+to the summary.

### 2015-02-27 dataset.tex
edits; nuggets

@@
 For the document stream, we use the news portion of the
- 2014 Trec KBA Stream Corpus.
+ 2014 TREC KBA Stream Corpus.
 The documents from this corpus come from hourly crawls of the web covering 
@@
 
-Our experiments also make use of the Trec Temporal Summarization (TS) Track
+Our experiments also make use of the TREC Temporal Summarization (TS) Track
  data from 2013 and 2014. 
@@
 search query for the event, the event type, and event evaluation timeframe).  
-All events occurred during the time span of the Trec KBA Stream Corpus.
+All events occurred during the time span of the TREC KBA Stream Corpus.
 

### 2015-02-27 experiments.tex
edits; nuggets

@@
 NIST developed a series metrics for evaluating update summarization systems
-as part of the Trec TS track.
+as part of the TREC TS track.
 We present results on two of these metrics, the expected gain and 
@@
 \subsection{Model Comparisons}
-
-Our proposed model using AP+Salience is compared against several variants and baselines. All thresholds
-are tuned on the development set.
+We refer to our complete model as \textbf{AP+Salience}.  We compare this model again several variants and baselines intended to measure the contribution of different components. All thresholds
+for all runs are tuned on the development set.
 
@@
 
+\fdcomment{consider removing AP+SalienceTR}
 \textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 

### 2015-02-27 introduction.tex
edits; nuggets

--- 2015-02-26 introduction.tex
@@
 a clustering based multi-document summarization system. 
-Not only does incorporating salience produce more relevant summaries,
-our system is better able to adapt to
+We demonstrate that combining salience with clustering produces more relevant summaries
+compared to baselines using clustering alone.  We believe this is
+because our system is better able to adapt to
 dynamic changes in input volume which adversely 
@@
 the event, and the change in word frequencies over the time of the event.
-While we ground these features in the domain of disaster, their templates
-are generally applicable to many update summarization domains.
+While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
+expected similarity of their ideal summaries.
 % to represent its likelihood of referring to the input disaster. 

### 2015-02-27 methods.tex
edits; nuggets

@@
     %  $(\stime,\etime)$.  
-      \end{enumerate*} While processing documents
-      throughout the time period of interest, the system outputs sentences,
-      which we refer to as updates.
-      Updates should not only be relevant to the event but also salient,
-      i.e. worthy of
-      reporting.
-      For most domains, updates also need to be 
-      timely and novel, i.e. users need to be informed of changes in the 
-      event as quickly as possible.
-      Finally, in aggregate the updates must comprehensively summarize the 
-      event.
+\end{enumerate*} 
+While processing documents
+throughout the time period of interest, the system outputs sentences
+from these documents likely to be useful to query issuer.  We refer
+to these selected sentences as \emph{updates}.
+
+In order to measure the usefulness of a system's updates, we consider
+the degree to which the system output reflects the different
+aspects of an event.  Events are often composed of a variety of sub-events.  
+For example, the Hurricane Sandy
+event includes sub-events related to the storm making landfall,
+the ensuing flooding, the many transportation issues, amongst many
+others.  An ideal system would update the user about each of these
+sub-events as they occur, with low latency.  In order to use
+language consistent with previous literature, we refer to these
+sub-events as the \emph{nuggets} associated with an event.  Nuggets are
+defined as a fine-grained atomic sub-event associated with an event.  
+We present several example nuggets associated with the Hurricane
+Sandy event in Figure \ref{fig:nuggets}.  We describe how these 
+nuggets are found in Section \ref{sec:data}.
+
+% Updates should not only be relevant to the event but also salient,
+% i.e. worthy of
+% reporting.
+% For most domains, updates also need to be
+% timely and novel, i.e. users need to be informed of changes in the
+% event as quickly as possible.
+% Finally, in aggregate the updates must comprehensively summarize the
+% event.
 
@@
 
-In order to quantify the salience of an update with respect to an event,
-we match the update to its most similar event nugget.
-%we assume the event can be represented as a set of atomic facts, or nuggets.
-Formally, given an update $u$ and a set of nuggets $N$ we define the salience of 
-$u$ as,
-\[\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
-\operatorname{sim}(u, n) \]
+Throughout our treatment of our algorithm, the \emph{salience} 
+of an update captures the degree to which it reflects an 
+event's nuggets.  Assuming that we have a text representation 
+of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
+defined as,
+\begin{align}
+\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
+\operatorname{sim}(u, n) \ref{eq:salience}
+\end{align}
 where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
@@
     \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
-  \item select a set of exemplar sentences (by combining clustering with 
-      salience predictions) (Section~\ref{sec:exsel}),
+  \item select a set of exemplar sentences  in $b_t$ by combining clustering with 
+      salience predictions (Section~\ref{sec:exsel}),
   \item add the most novel and salient exemplars to $U$ (Section~\ref{sec:upsel}).
@@
 The resultant list of updates $U$ is our summary of the event.
-Step 2 uses the affinity propagation algorithm to incorporate the salience
-predictions from step 1. We refer to this summarization 
-system as the \textbf{AP+Salience} model.
+% Step 2 uses the affinity propagation algorithm to incorporate the salience
+% predictions from step 1. We refer to this summarization
+% system as the \textbf{AP+Salience} model.
 
@@
 in our training data to learn a model to predict salience for sentences
-from new events.
+in new events.
 For each event in our training data, we sample a set of sentences and  each 
-sentence's salience is computed, i.e. we compute the maximum semantic 
-similarity to the event's nuggets.
-We now have a training set of set sentences and the target values 
+sentence's salience is computed according to Equation \ref{eq:salience}.
+This results in training set of set sentences and the target values 
 (their salience) to predict.
@@
 
+\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
+expected summary is similar to some previously seen target or some side information.}
+
 These models are intended to detect sentences similar to those appearing in 
@@
 \paragraph{Geographic Relevance Features}
-
+\fdcomment{need introduction here.  what are geographic features?}
 There are two challenges to using geographic features. First we do not 
@@
 current hour and looking up their latitude and 
-longitude using a publicly available geolocation service. Since the document
+longitude using a publicly available geolocation service. 
+\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
+Since the document
 stream contains documents that are at least somewhat relevant to the event,
@@
 The locations are clustered (using affinity propagation and uniform salience)
+\fdcomment{affinity propagation undefined}
 and we treat the resulting cluster centers
@@
 \paragraph{Temporal Relevance Features}
-
+\fdcomment{motivate burstiness; explain what it is before using the expression.}
 Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
@@
 to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
-sentence have changed over the last 24 hours.
+sentence have changed over the last 24 hours. \fdcomment{unclear.  is it just changing the IDF part of the 
+items in the average?}
 

### 2015-02-27 conclusion.tex
Latest rev ck

In this paper, we have presented an update summarization system for the disaster domain,
and demonstrated improved system performance by integrating sentence salience with clustering.
Addtionally we have presented several interesting feature groups for use in disaster summarization.
In the future we would like to explore the applicability of this model and features
to more general domains.

### 2015-02-27 experiments.tex
Latest rev ck

@@
 Because an update can map to more than one nugget, it is possible to receive an
-expected gain greater than 1; values greater than one can be thought of as a 
-compression ration, i.e. an expected gain of 10 indicates the average update
-contains 10 unique nuggets on average and that the summary is a 10~:~1 
-compression of the nugget information.
+expected gain greater than 1.
+%values greater than one can be thought of as a 
+%compression ration, i.e. an expected gain of 10 indicates the average update
+%contains 10 unique nuggets on average and that the summary is a 10~:~1 
+%compression of the nugget information.
 An expected gain of 1 would indicate that every
@@
 
-\fdcomment{consider removing AP+SalienceTR}
-\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
-to the
-AP+Salience model except at the novelty filter stage.
-Within the novelty filtering stage, 
-AP+Salience inspects exemplars in order of decreasing salience when computing
-similarity to previous updates. Because the order in which we add updates
-affects future similarity calculations, we evaluate an alternate ordering 
-scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
-time since event start. The purpose of this model is to give preference to 
-more recent updates, since timelyness is one of the desirable traits of our update
-summaries.
+%\fdcomment{consider removing AP+SalienceTR}
+%\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
+%to the
+%AP+Salience model except at the novelty filter stage.
+%Within the novelty filtering stage, 
+%AP+Salience inspects exemplars in order of decreasing salience when computing
+%similarity to previous updates. Because the order in which we add updates
+%affects future similarity calculations, we evaluate an alternate ordering 
+%scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
+%time since event start. The purpose of this model is to give preference to 
+%more recent updates, since timelyness is one of the desirable traits of our update
+%summaries.
+%

### 2015-02-27 feature_table.tex
Latest rev ck

@@
 $\cdot$ Sent. length \\
-$\cdot$ Ratio of punc. to non-punc. chars. \\
-$\cdot$ Ratio of caps. to non-caps. chars. \\
-$\cdot$ Ratio of lowercase to other chars. \\
-$\cdot$ Ratio of uppercase to other chars. \\
+$\cdot$ Ratio of punc./caps./lower/upper to \\
+$\;\;$ other chars. \\
+%$\cdot$ Ratio of caps. to non-caps. chars. \\
+%$\cdot$ Ratio of lowercase to other chars. \\
+%$\cdot$ Ratio of uppercase to other chars. \\
 $\cdot$ \# of caps. words (normalized by \# of words)\\
-$\cdot$ \# of Person, Location, Org. Date, Number,\\
-$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
-$\;\;$ tags (normalized by \# of words)\\
+$\cdot$ \# of Person, Location, Org., etc.\\
+$\;\;$ (normalized by \# of words)\\
+%$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
 \hline
 \textbf{Query Features}\\
-$\cdot$ \% of query words covered by sent.\\
-$\cdot$ Total query matches.\\
+$\cdot$ \% of query words coverage/matches\\
 $\cdot$ Total event-type synonyms/hypernyms/ \\
-$\;\;$ hyponyms coverage.\\
-$\cdot$ Total event-type synonyms/hypernyms/ \\
-$\;\;$ hyponyms matches.\\
+$\;\;$ hyponyms coverage/matches.\\
+%$\cdot$ Total event-type synonyms/hypernyms/ \\
+%$\;\;$ hyponyms matches.\\
 \hline
@@
 $\cdot$ Median document distance to nearest  \\
-$\;\;$ location cluster (current hour).\\
-$\cdot$ Median document distance to nearest  \\
-$\;\;$ location clusters (previous hour).\\
+$\;\;$ location cluster (current hour/prev. hour).\\
+%$\cdot$ Median document distance to nearest  \\
+%$\;\;$ location clusters (previous hour).\\
 $\cdot$ Distance of first location in doc. to nearest \\
-$\;\;$ location cluster (current hour).\\
-$\cdot$ Distance of first location in doc. to nearest \\
-$\;\;$ location cluster (previous hour).\\
+$\;\;$ location cluster (current hour/prev. hour).\\
+%$\cdot$ Distance of first location in doc. to nearest \\
+%$\;\;$ location cluster (previous hour).\\
 \hline

### 2015-02-27 methods.tex
Latest rev ck

@@
 
-% Updates should not only be relevant to the event but also salient,
-% i.e. worthy of
-% reporting.
-% For most domains, updates also need to be
-% timely and novel, i.e. users need to be informed of changes in the
-% event as quickly as possible.
-% Finally, in aggregate the updates must comprehensively summarize the
-% event.
-
-\begin{figure}
-\setlength{\fboxsep}{10pt}
-    \fbox{\parbox{6.9cm}{
-        -- hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay
-
-        -- Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief
-
-        -- over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
- }}
- \caption{Example nuggets from Hurricane Sandy.\label{fig:nuggets}}
-\end{figure}
+\input{nugget_examples.tex}
 
@@
 \operatorname{salience}(u) = \operatorname{max}_{n \in N} 
-\operatorname{sim}(u, n) \ref{eq:salience}
+\operatorname{sim}(u, n) \label{eq:salience}
 \end{align}
@@
 the cosine similarity of latent vectors associated with the update and 
-nugget text \cite{?}. % We train
-% a regression model to predict salience using features derived from
-% the updates themselves
-% and the document stream.
+nugget text \cite{?}.
 
-     % \ref{sec:methods}.  In our version of this problem, we assume that that
-     % the system receives one batch of new sentence-segmented documents every
-     % hour throughout the period of interest.
 \subsection{Update Summarization}
@@
   \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
-  \item select a set of exemplar sentences  in $b_t$ by combining clustering with 
+  \item select a set of exemplar sentences in $b_t$ by combining 
+      clustering with 
       salience predictions (Section~\ref{sec:exsel}),
-  \item add the most novel and salient exemplars to $U$ (Section~\ref{sec:upsel}).
+  \item add the most novel and salient exemplars 
+      to $U$ (Section~\ref{sec:upsel}).
 \end{enumerate}
-
 The resultant list of updates $U$ is our summary of the event.
-% Step 2 uses the affinity propagation algorithm to incorporate the salience
-% predictions from step 1. We refer to this summarization
-% system as the \textbf{AP+Salience} model.
-
-%\kmcomment{This was as best as I could do on this. These are the example
-%    categories from the wikipedia page. However you refer in the language model
-%    description to the event type earthquake which is more specific. Also, in
-%    the TREC data you refer to event types that are specified in the metadata.
-%So I'm wondering how many different categories you have and where they come
-%from.}
-
-%\fdcomment{include description of nuggets here.} 
-%%\kmcomment{We should jointly
-%    discuss what goes in data and what goes in problem definition. I have moved
-%    event types here because I think knowing how many types and examples of
-%    them would be helpful upfront. I think where nuggets and wikipedia pages go
-%is questionable.}
-
-%Figure \ref{alg:temporal-summarization} outlines our general update
-%summarization algorithm.  At each hour, the system processes each input 
-%sentence batch $S_t$.
-%We predict the salience $P$ for all input sentences $s\in S_t$.
-%Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
-%obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
-%to be updates and add those to our set of summary updates $U$.
-%
-%\begin{algorithm}%[H]
-% \KwData{ 
-%     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
-%     batches\\
-% \KwResult{U --- a list of updates, i.e. the update summary} 
-%}
-% ~\\
-% Initialize empty list $U$ of updates\\
-%    \For{$t \gets \stime,\ldots,\etime $}{
-%        $P \gets \operatorname{PredictSalience}(S_t)$\\
-%        $E \gets \operatorname{APCluster}(P, S_t)$ \\
-%        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
-%        $U \gets U \cup U_t$
-%}
-% \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
-%\end{algorithm}
 
@@
 In order to fit a regression model, the sentences are represented as feature
-vectors (described in the next section).
+vectors (described in Section~\ref{sec:features}).
 We fit a Gaussian process (GP) regression model to this data.
@@
 \subsubsection{Features}
+\label{sec:features}
 We want our model to be predictive across different event instances so we avoid lexical features.
@@
 \paragraph{Basic Features}
-%KM - Would be good to have quick justification of these features. I added a
-%sentence. Feel free to edit or remove.
-
 We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
@@
 features.  The first model is estimated from a corpus of generic news 
-articles \ckcomment{(we used the 199?-200? Associated Press and New York Times sections of the Gigaword corpus)}.  
+articles (we used the 1995-2010 Associated Press section of the 
+Gigaword corpus).  
 This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
@@
 
-\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
-expected summary is similar to some previously seen target or some side information.}
+%\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
+%expected summary is similar to some previously seen target or some side information.}
 
@@
 (e.g. most earthquake summaries are likely to include higher probability for 
-ngrams including the token `magnitude').
+ngrams including the token `magnitude'). While we focus our system on the 
+language of news and disaster, we emphasize that the use of language modeling 
+can be an effective feature for multi-document summarization for other 
+domains that have related text corpora.
 
@@
 
-%For both models, we Finally, we extract the percentage of capitalized words,
-%and sentence length as features. These last two features also help to
-%identify sentences that are less likely to contain relevant content-- overly
-%long and heavily capitalized sentences in our corpus were likely to be long
-%strings of web-page headlines, section headers, and other irrelevant page
-%structure.
+\paragraph{Geographic Relevance Features}
+%\fdcomment{need introduction here.  what are geographic features?}
+The disasters in our corpus are all phenomena that
+affect some part of the word. 
+Where possible, we would like to capture a sentence's proximity to the event,
+i.e. when a sentence references a location, it should be close to the 
+area of the disaster.
 
-\paragraph{Geographic Relevance Features}
-\fdcomment{need introduction here.  what are geographic features?}
 There are two challenges to using geographic features. First we do not 
@@
 to a location.
-We address the first issue by extracting all locations from sentences at the
+We address the first issue by extracting all locations from 
+documents relevant to the event at the
 current hour and looking up their latitude and 
 longitude using a publicly available geolocation service. 
-\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
-Since the document
-stream contains documents that are at least somewhat relevant to the event,
+%\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
+Since the documents that are at least somewhat relevant to the event,
 we assume in aggregate the locations should give us a rough area of interest.
-The locations are clustered (using affinity propagation and uniform salience)
-\fdcomment{affinity propagation undefined}
+The locations are clustered %(using affinity propagation and uniform salience)
+%\fdcomment{affinity propagation undefined}
 and we treat the resulting cluster centers
@@
 To address the second issue, 
-we compute geographic relevance features for the document as a whole and all
+we compute features for the document as a whole and all
 sentences from that document receive the same feature value.
+
 Using the locations in each document, we compute the median distance to the 
@@
 of importance we also compute the distance of the first mentioned
-location to the nearest event location. Because some events can move, we also
+location to the nearest event location. All sentences in the document take
+as features these two distance calculations.
+Because some events can move, we also
 compute these distances to event locations from the previous hour.
@@
 \paragraph{Temporal Relevance Features}
-\fdcomment{motivate burstiness; explain what it is before using the expression.}
-Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.
-the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.
+%\fdcomment{motivate burstiness; explain what it is before using the expression.}
+As we track events over time, it is likely that the coverage of the event 
+may die down, only to spike back up when there is a breaking development.
+Alternatively, words that were salient initially may become background context,
+and over time lose their importance.
+Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
+can help to locate novel sentences that are part of the most recent reportage
+and have yet to fall into the background.
 
@@
 to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
-sentence have changed over the last 24 hours. \fdcomment{unclear.  is it just changing the IDF part of the 
-items in the average?}
-
-The final temporal feature is the hours since the event started.
+sentence have changed over the last 24 hours, i.e. we keep the sentence
+term frequencies fixed and compute the difference in IDF. Large changes
+in the is value indicate the sentence contains bursty terms.
+%\fdcomment{unclear.  is it just changing the IDF part of the 
+%items in the average?}
+We also use the time (in hours) since the event started as a feature.
 %difference 
@@
 \[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
-+ \sum_{i : i = e_i}^n \operatorname{pref}(e_i)  \]
++ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
 where $e_i$ is the exemplar of the $i$-th data point, and functions

### 2015-02-27 nugget_examples.tex
Latest rev ck

\begin{figure}
\setlength{\fboxsep}{10pt}
    \fbox{\parbox{6.9cm}{
        -- hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay

%        -- Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief

        -- over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
 }}
 \caption{Example nuggets from Hurricane Sandy.\label{fig:nuggets}}
\end{figure}

### 2015-02-27 results.tex
Latest revision ck
Latest rev ck

--- 2015-02-23 results.tex
@@
 \subsection{ROUGE}
+\input{rouge_cum.tex}
 
 Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
-Both versions of AP+Salience show improvement on ngram precision and recall over the 
-baselines. This improvement is statistically significant at the $\alpha = .01$ level
-using the Wilcoxon signed-rank test. Additionally, AP+Salience's improvement over
-AP+Salience TR in recall and F-measure is also statistically significant at the same level.
+AP+Salience shows improvement on ngram precision and recall over the 
+baselines. This improvement is statistically significant for all ngram and longest common substring
+precision, recall, and F-measure
+at the $\alpha = .01$ level
+using the Wilcoxon signed-rank test.
 
-AP+Salience maintains its performance above the baselines over time as well. Table~\ref{fig:trouge}
-shows the ROUGE-1 performance over time. Within the initial days of the event, recall is 
-competitive amongst system. However, both AP+Salience variants are able to achieve higher
-precision early on.
+AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
+shows the ROUGE-1 scores over time. 
+We emphasize the difference in unigram precision (bigram precision is not shown but it follows
+similar curve).
+Within the initial days of the event, AP+Salience is able to take the lead over the over 
+systems in ngram precision. 
+The AP+Salience model is better able to find salient updates
+earlier on; for the disaster domain, this is an especially important quality of the model.
 
-\subsection{Nugget Evaluation}
+Moreover, the AP+Salience's recall is not diminished by the high precision and remains competitive with AP.
+Over time AP+Salience's recall also begins to pull away, while the other models start to suffer
+from topic drift.
+
+\subsection{Expected Gain and Comprehensiveness}
+
+Figure~\ref{fig:nperf} shows the expected gain across a range of 
+similarity thresholds, where thresholds closer to 1 are more conservative
+estimates. The ranking of the systems remains constant across the 
+sweep with 
+AP+Salience beating all baseline systems.
+Predicting salience in general is helpful for keeping a summary on topic as
+the RS approach out performs the clustering only approaches
+on expected gain.
+
+When looking at the comprehensiveness of the summaries AP outperforms
+ AP+Salience. The compromise encoded in the AP+Salience objective
+function, between being representitive and being salient, is seen clearly here
+where the performance of the AP+Salience methods is lower bounded by 
+the salience focused RS
+system and upper bounded by the clustering only AP system.
+Overall, AP+Salience achieves the best balance of these two metrics.
 
 \subsection{Feature Ablation}
+\input{rouge_fa.tex}
 
-\begin{table}[h]
-\centering
-% centering table
-\begin{tabular}{l c c c}
-% creating 10 columns
-\multicolumn{4}{c}{ROUGE-1}\\
-\hline
-\hline
-% inserting double-line
-$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\
-[0.5ex]
-\hline
-AP+Salience &          0.289   &    0.344  &  0.312  \\
-AP+Salience TR &       0.278   &    0.342  &  0.303 \\
-AP    &         0.246   &    0.285  &  0.263 \\
-HAC     &        0.238   &    0.276  &  0.255 \\
-Rank    &        0.229   &    0.271  &  0.247 \\
-% inserts single-line
-% Entering 1st row
-%& &soft &1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
-%[-1ex]
-%\raisebox{1.5ex}{Police} &\raisebox{1.5ex}{5}&hard& 2 & $-4$ & 4 & 4 & $-2$ & $-4$ & 4\\
-%[1ex]
-% Entering 2nd row
-%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
-%[-1ex]
-%\raisebox{1.5ex}{Beethoven}&\raisebox{1.5ex}{5}& hard&8 & $-8$ & 2 & 8 & $-8$ & $-8$ & 6
-%\\
-%[1ex]
-% Entering 3rd row
-%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1
-%\\
-%[-1ex]
-%\raisebox{1.5ex}{Metallica} &\raisebox{1.5ex}{5}& hard&4 & $-8$ & 8 & 4 & $-8$ & $-8$ & 8\\
-%[1ex]
-% [1ex]adds vertical space
-\hline % inserts single-line
-\end{tabular}
-~\\[1ex]
-~\\
-\begin{tabular}{l c c c}
-% creating 10 columns
-\multicolumn{4}{c}{ROUGE-2}\\
-\hline
-\hline
-% inserting double-line
-$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\[0.5ex]
-\hline
-AP+Salience & 0.047 &     0.056 &   0.051 \\
-AP+Salience TR & 0.044  &    0.054 &  0.047 \\
-AP & 0.033  &    0.038 &  0.035 \\
-HAC & 0.031  &    0.035 &  0.034 \\
-Rank & 0.031  &    0.037 &  0.034 \\
-\hline % inserts single-line
-\end{tabular}
-\caption{System ROUGE performance.} % title name of the table
-\label{tab:rouge}
-\end{table}
+Table~\ref{tab:farouge} shows the results of our feature ablation
+tests. Removing the language models yields a statistically 
+significant drop in both ngram recall and F-measure. 
+Interestingly, removing the basic features leads to an
+increase in both unigram and bigram precision; in the bigram
+case this is enough to cause a statistically significant increase
+in F-measure over the full model. In other words, the generic features
+actually lead to an inferior model when we can incorporate more appropriate
+domain specific features. \ckcomment{Kathy: maybe we should mention Spark Jones here about the idea that there is no such thing as generic summarization?}
 
-\begin{figure*}
-  \includegraphics[width=\textwidth]{rouge-time.png}
+\begin{figure}
+    \includegraphics[]{rouge-time.eps}
 \caption{System ROUGE-1 performance over time.}
 \label{fig:trouge}
-\end{figure*}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[]{nuggets-metrics.eps}
+\caption{Avg. Gain and Comprehensiveness performance.}
+\label{fig:nperf}
+\end{figure}

### 2015-02-28 conclusion.tex
Latest rev ck

@@
 and demonstrated improved system performance by integrating sentence salience with clustering.
-Addtionally we have presented several interesting feature groups for use in disaster summarization.
+
+Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
+better summaries; specifically features that capture the language and location
+of the domain.
+
 In the future we would like to explore the applicability of this model and features
-to more general domains.
+%KM3 - I think not just for more general domains (as it says something wrong about this one) but to a wider class of events.
+to a wider class of events. Additionally, we want to further analyze the failure 
+of the temporal relevance features; intuitively we believe time to be important
+for the update summarization task.
+
+%Addtionally we have presented several interesting feature groups for use in disaster summarization.
+%In the future we would like to explore the applicability of this model and features
+%to more general domains.

### 2015-02-28 dataset.tex
Latest rev ck

@@
 The documents from this corpus come from hourly crawls of the web covering 
- ?, 20?? to ?, 20??. \ckcomment{Need to mention we further filtered the corpus}
+ October 2011 through February 2013.
 
 Our experiments also make use of the TREC Temporal Summarization (TS) Track
- data from 2013 and 2014. 
+ data from 2013 and 2014 \cite{aslam2013trec}. 
 This data includes 25 events and event metadata (e.g., a user
@@
 All events occurred during the time span of the TREC KBA Stream Corpus.
+For each event we create a stream of relevant documents by 
+selecting only documents with article content that contains the complete set 
+of query words. 
+%identifying the 
+%article start/end within each web page in the corpus using a classifier
+%tra
+%those documents
 
@@
 Nuggets are brief and important text snippets that represent sub-events that should be conveyed
-by an ideal update summary (see Figure~\ref{fig:nuggets} for examples).
+by an ideal update summary. % (see Figure~\ref{fig:nuggets} for examples).
 In order
@@
 and marking the text additions capturing a new, unique nugget.  More detail
-on this process can be found in the track description \cite{?}.
+on this process can be found in the track description \cite{aslam2013trec}.

### 2015-02-28 experiments.tex
Latest rev ck

@@
-We evaluate our system on two metrics: ROUGE \cite{?}, a common automatic
+We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, a common automatic
 summarization method and an evaluation of system expected gain 
 and comprehensiveness---metrics adapted 
-from the Trec TS track \cite{?}.
+from the TREC TS track \cite{aslam2013trec}.
 
@@
 Of the 25 events in the TREC TS data, 24 are covered by the news portion 
-of the Trec KBA Stream Corpus.
+of the TREC KBA Stream Corpus.
 From these 24, we set aside three events to use as a development set.
 All system salience and similarity threshold parameters are tuned on the
-development set.
+development set to maximize ROUGE-2 F1 scores.
 
 We train a salience model for each event using 1000 sentences randomly sampled
-from the event's document stream. Kernel length scale parameters  for each
-feature group are fit to maximize
-the model's log likelihood using ???.
+from the event's document stream. 
+%Kernel length scale parameters  for each
+%feature group are fit to maximize
+%the model's log likelihood using ???.
 
@@
 nuggets. 
-See figure~\ref{fig:test} for an example model summary excerpt.
+%See figure~\ref{fig:test} for an example model summary excerpt.
 Generally, ROUGE evaluation assumes both model and system summaries
@@
 The model summaries in this case are retrospective, 
-and so we can see how quickly
-each system is able to achieve its maximum performance.
+%and so we can see how quickly
+%each system is able to achieve its maximum performance.
+and this evaluation reveals how quickly each system is able to cover  information in the summary.
 
@@
 
-NIST developed a series metrics for evaluating update summarization systems
+%NIST developed a series metrics for evaluating update summarization systems
+NIST developed a series of metrics for evaluating update summarization systems
 as part of the TREC TS track.
@@
 suggests that semantic similarity values around .7 produce reasonable matches.
-For example, the nugget 
-\begin{blockquote}
-watches and warnings in effect for much of the Mid-Atlantic states and 
-southern New England 
-\end{blockquote}
-\noindent
-matches with the update 
-\begin{blockquote}
-A tropical storm warning remains in effect for much of the coasts 
-of North and South Carolina\end{blockquote}
-\noindent
-with a similarity of .7. 
+%For example, the nugget 
+%\begin{blockquote}
+%watches and warnings in effect for much of the Mid-Atlantic states and 
+%southern New England 
+%\end{blockquote}
+%\noindent
+%matches with the update 
+%\begin{blockquote}
+%A tropical storm warning remains in effect for much of the coasts 
+%of North and South Carolina\end{blockquote}
+%\noindent
+%with a similarity of .7. 
 The average semantic similarity of manual matches performed 
@@
 the salience values are uniformly set as the median value of the 
-input similarity scores, as is commonly used in the AP literature \cite{?}.
+input similarity scores, as is commonly used in the AP literature \cite{frey2007clustering}.
 After clustering a sentence batch, the exemplars are examined in order

### 2015-02-28 introduction.tex
Latest rev ck

@@
 overwhelm those trying to monitor the situation. 
-Crisis informatics \cite{?} is dedicated to finding methods for sharing the
+Crisis informatics \cite{palen2010vision} is dedicated to finding methods for sharing the
 right information in a timely fashion during such an event.
@@
 human-in-the-loop approaches ranging from on the ground information 
-gathering to crowdsourced reporting and disaster management.
+gathering to crowdsourced reporting and disaster management \cite{starbird2013working}.
 
@@
 We demonstrate that combining salience with clustering produces more relevant summaries
-compared to baselines using clustering alone.  We believe this is
+compared to baselines using clustering alone.  
+We suggest that this
 because our system is better able to adapt to
@@
 expected similarity of their ideal summaries.
+%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
+%expected similarity of their ideal summaries.
 % to represent its likelihood of referring to the input disaster. 
@@
 in the information retrieval and multi-document
-summarization literature. (Rest of outline goes here...)
+summarization literature. Section~\ref{sec:methods} outlines the details
+of our salience and summarization models. Next we describe our
+data (Section~\ref{sec:data}) and experiments (Section~\ref{sec:exper}). Finally,
+we discuss our results (Section~\ref{sec:results}) and conclude the paper.
 

### 2015-02-28 methods.tex
Latest rev ck

@@
+\label{sec:methods}
 %In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
@@
 others.  An ideal system would update the user about each of these
-sub-events as they occur, with low latency.  In order to use
-language consistent with previous literature, we refer to these
+sub-events as they occur, with low latency.  
+%In order to use
+%language consistent with previous literature, 
+We refer to these
 sub-events as the \emph{nuggets} associated with an event.  Nuggets are
 defined as a fine-grained atomic sub-event associated with an event.  
+%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
 We present several example nuggets associated with the Hurricane
-Sandy event in Figure \ref{fig:nuggets}.  We describe how these 
+Sandy event 
+in Figure \ref{fig:nuggets}.  
+We describe how these 
 nuggets are found in Section \ref{sec:data}.
 
-\input{nugget_examples.tex}
+%\input{nugget_examples.tex}
 
@@
 the cosine similarity of latent vectors associated with the update and 
-nugget text \cite{?}.
+nugget text \cite{guo2012simple}.
 
@@
 
-\subsubsection{Model}
-
-When evaluating our summarization system
-we do not have access to event nuggets. However, we can use the nuggets
-in our training data to learn a model to predict salience for sentences
-in new events.
-For each event in our training data, we sample a set of sentences and  each 
-sentence's salience is computed according to Equation \ref{eq:salience}.
-This results in training set of set sentences and the target values 
-(their salience) to predict.
-
-In order to fit a regression model, the sentences are represented as feature
-vectors (described in Section~\ref{sec:features}).
-We fit a Gaussian process (GP) regression model to this data.
-\cite{rasmussen:gaussian-process-book}.  GP regressors are a
-class of data-driven, non-parametric model generalizing the multi-variate
-Gaussian to the infinite dimensional setting.  
-%GP's are general
-%and are state of the art for many regression tasks.  
-GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
-case sentences.  In our experiments, we used a radial basis 
-function (RBF) kernel.  
-Our features fall naturally into five groups and we use a separate RBF kernel
-for each, using the sum of each feature group kernel as the final input
-to the GP model.
-
 %include feature table
-\input{feature_table.tex}
+%\input{feature_table.tex}
 
@@
 \label{sec:features}
-We want our model to be predictive across different event instances so we avoid lexical features.
+We want our model to be predictive of sentence salience across different event instances so we avoid lexical features.
 Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
@@
 articles (we used the 1995-2010 Associated Press section of the 
-Gigaword corpus).  
+Gigaword corpus \cite{graff2003english}).  
 This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
@@
 
-We use the SRILM toolkit to implement a 5-gram Kneser-Ney model for both
+We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram Kneser-Ney model for both
 the background language and model and the event specific language models.
@@
 term frequencies fixed and compute the difference in IDF. Large changes
-in the is value indicate the sentence contains bursty terms.
+in IDF value indicate the sentence contains bursty terms.
 %\fdcomment{unclear.  is it just changing the IDF part of the 
@@
 
+\subsubsection{Model}
+
+At run-time
+our summarization system
+does not have access to event nuggets to compute sentence salience. 
+However, we can use the nuggets
+in our training data to train a model to predict salience for sentences
+in new events.
+For each event in our training data, we sample a set of sentences and  each 
+sentence's salience is computed according to Equation \ref{eq:salience}.
+This results in a training set of sentences and the target values 
+(their salience) to predict.
+
+Now that we have a feature representation for the sentences, we can use
+any off the shelf regression technique to predict salience.
+We opt to use a Gaussian process (GP) regression model
+\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
+kernel.  
+Our features fall naturally into five groups and we use a separate RBF kernel
+for each, using the sum of each feature group RBF kernel as the final input
+to the GP model.
+%GP regressors are a
+%class of data-driven, non-parametric model generalizing the multi-variate
+%Gaussian to the infinite dimensional setting.
+
+%When evaluating
+%In order to fit a regression model, the sentences are represented as feature
+%vectors (described in Section~\ref{sec:features}).
+%GP's are general
+%and are state of the art for many regression tasks.  
+%GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
+%case sentences.  In our experiments, we used a radial basis 
+%function (RBF) kernel.
+
 \subsection{Exemplar Selection}
@@
 
-We combine the output of our salience prediction model with the affinity
-propagation algorithm to identify a set of exemplar sentences 
-for each input batch. 
+Once we have predicted the salience for a batch of sentences, we must
+now select a set of update candidates, i.e. sentences that are both salient
+and representative of the current batch.
+To accomplish this, we combine the output of our 
+salience prediction model with the affinity
+propagation algorithm 
+%to identify a set of exemplar sentences 
+%for each input batch. 
 Affinity propagation (AP) is a clustering algorithm
 that identifies a subset of datapoints as exemplars and forms clusters
-by assigning the remaining points to one of the exemplars. AP attempts to 
+by assigning the remaining points to one of the exemplars 
+\cite{frey2007clustering}. AP attempts to 
 maximize the net similarity objective 
@@
 the volume of input is low. The adaptive nature of our model differentiates
-our method from most out update summarization systems.
+our method from most other update summarization systems.
 
@@
 
-\textbf{Salience Filter } To 
+The exemplar sentences from the exemplar selection stage are the most 
+salient and representative of the input for the current hour. However,
+we need to reconcile these sentences with updates from the previous hour
+to ensure that the most salient and least redundant  updates are selected.
+To 
 ensure that only the most salient updates are selected we apply a minimum
@@
 
-\textbf{Novelty Filter } Next,
+ Next,
 to prevent adding updates that are redundant, we filter out exemplars

### 2015-02-28 relatedwork.tex
Latest rev ck

@@
 
-First, centrality-focused approaches (including graph,
-cluster, and centroid methods) are very natural for retrospective analysis in
+First, centrality-focused approaches (including graph \cite{erkan2004lexrank},
+cluster \cite{hatzivassiloglou2001simfinder}, and centroid \cite{radev2004centroid} methods) are very natural for retrospective analysis in
 the sense that they let the data "speak for itself." 
@@
 
-The last category includes probabilistic, information theoritic, and set cover
+The last category includes probabilistic \cite{haghighi2009exploring}, 
+information theoritic, and set cover \cite{lin2011class}
 approaches. While these methods are generally focused on producing diverse
@@
 related to 197 natural and human disasters from 2009 to 2011 using variants of
-Rouge modified to capture novelty, relevance, and comprehensiveness
+ROUGE modified to capture novelty, relevance, and comprehensiveness
 \cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
@@
 narrative structure than Guo \textit{et al.}.  The method is evaluated on
-official press releases related to Hurricane Wilma  in 2005 using Rouge score
+official press releases related to Hurricane Wilma  in 2005 using ROUGE score
 between the system summary and a manually generated target summary. In addition

### 2015-02-28 results.tex
Latest rev ck

@@
 
-Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
-AP+Salience shows improvement on ngram precision and recall over the 
-baselines. This improvement is statistically significant for all ngram and longest common substring
+Table~\ref{tab:rouge} shows our results for system output samples against the full summary of nuggets using ROUGE. 
+%shows our results for the evaluation using ROUGE. 
+%AP+Salience shows improvement on ngram precision and recall over the 
+%baselines. 
+This improvement is statistically significant for all ngram and longest common substring
 precision, recall, and F-measure
@@
 
+\begin{figure}
+    \includegraphics[]{rouge-time.eps}
+\caption{System ROUGE-1 performance over time.}
+\label{fig:trouge}
+\end{figure}
+
 AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
 shows the ROUGE-1 scores over time. 
-We emphasize the difference in unigram precision (bigram precision is not shown but it follows
+We show the difference in unigram precision (bigram precision is not shown but it follows
 similar curve).
@@
 \subsection{Expected Gain and Comprehensiveness}
+\begin{figure}
+  \includegraphics[]{nuggets-metrics.eps}
+\caption{Expected Gain and Comprehensiveness performance.}
+\label{fig:nperf}
+\end{figure}
 
@@
 actually lead to an inferior model when we can incorporate more appropriate
-domain specific features. \ckcomment{Kathy: maybe we should mention Spark Jones here about the idea that there is no such thing as generic summarization?}
+domain specific features.
+This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
 
-\begin{figure}
-    \includegraphics[]{rouge-time.eps}
-\caption{System ROUGE-1 performance over time.}
-\label{fig:trouge}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[]{nuggets-metrics.eps}
-\caption{Avg. Gain and Comprehensiveness performance.}
-\label{fig:nperf}
-\end{figure}
+Removing the language model and geographic relevance features leads to a
+statistically significant drop in ROUGE-1 F1 scores. Unfortunately,
+this is not the case for the temporal relevance features. We surmise that
+these features are too strongly correlated with each other, 
+i.e. the differences in TFIDF between hours are definitely not IID variables.

### 2015-02-28 conclusion.tex
ck rev

@@
 
-In the future we would like to explore the applicability of this model and features
+In the future we would like to explore the application of the AP+Salience
+ model and features
 %KM3 - I think not just for more general domains (as it says something wrong about this one) but to a wider class of events.
-to a wider class of events. Additionally, we want to further analyze the failure 
+to a wider class of events.
+
+We would also like to further analyze the failure 
 of the temporal relevance features; intuitively we believe time to be important

### 2015-02-28 introduction.tex
ck rev

@@
 %KM - I think we should consider adding other event summarizers. 
-(e.g.,
-\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}), 
+\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
 these experiments were not developed to handle streaming data from a
@@
 compared to baselines using clustering alone.  
-We suggest that this
+Our experiments suggest that this
 because our system is better able to adapt to
@@
 event
-type (i.e., industrial accident, hurricane, riot) and whether it specifies 
+type (e.g., industrial accident, hurricane, riot) and whether it specifies 
 information
@@
 the event, and the change in word frequencies over the time of the event.
-While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
+While we evaluate these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
 expected similarity of their ideal summaries.

### 2015-02-28 methods.tex
ck rev

@@
 We refer to these
-sub-events as the \emph{nuggets} associated with an event.  Nuggets are
+sub-events as the \emph{nuggets} associated with an event.  A nugget is
 defined as a fine-grained atomic sub-event associated with an event.  
 %An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
-We present several example nuggets associated with the Hurricane
+We present 2 example nuggets associated with the Hurricane
 Sandy event 
-in Figure \ref{fig:nuggets}.  
+in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
+several hunder nuggets in total.
 We describe how these 
@@
 
-%\input{nugget_examples.tex}
+\input{nugget_examples.tex}
 
@@
 event's nuggets.  Assuming that we have a text representation 
-of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
+for each of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
 defined as,
@@
 At each hour we receive a new batch of sentences $b_t$ from the stream
+of event relevant documents
 and perform the following actions:
-\begin{enumerate}
+\begin{enumerate}[nosep]
   \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
@@
 from Wikipedia pages under the category \emph{Category:Earthquakes}.  
-
 %\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
 %expected summary is similar to some previously seen target or some side information.}
-
 These models are intended to detect sentences similar to those appearing in 
@@
 However, we can use the nuggets
-in our training data to train a model to predict salience for sentences
-in new events.
+in our training data to model salience as a function of our features.
+Such a model generalizes to new events.
 For each event in our training data, we sample a set of sentences and  each 
@@
 
-Now that we have a feature representation for the sentences, we can use
-any off the shelf regression technique to predict salience.
+Given a feature representation for the sentences, we can use
+any regression technique to predict salience.
 We opt to use a Gaussian process (GP) regression model
@@
 salience prediction model with the affinity
-propagation algorithm 
+propagation algorithm.
 %to identify a set of exemplar sentences 
@@
 similarity
-of data points and the apriori preference of a data point to be an exemplar
+of data points and our predicted apriori preference of a data point to be an exemplar
 respectively. 
@@
  Next,
-to prevent adding updates that are redundant, we filter out exemplars
+to prevent adding updates that are redundant with previous output updates, 
+we filter out exemplars
 that are too similar to previous updates.

### 2015-02-28 relatedwork.tex
ck rev

--- 2015-02-22 relatedwork.tex
@@
 
-First, centrality-focused approaches (including graph,
-cluster, and centroid methods) are very natural for retrospective analysis in
-the sense that they let the data "speak for itself." 
+First, centrality-focused approaches (including graph \cite{erkan2004lexrank},
+cluster \cite{hatzivassiloglou2001simfinder}, and centroid \cite{radev2004centroid} methods) are very natural for retrospective analysis in
+the sense that they let the data ``speak for itself.'' 
 These methods equate salience with centrality, either to the input or some other
@@
 Several researchers have recognized the importance of summarization during
-natural disasters.  Guo \textit{et al.} developed a system for detecting
+natural disasters.  \cite{qi:temporal-summarization} developed a system for detecting
 novel, relevant, and comprehensive sentences immediately after a natural
-disaster \cite{qi:temporal-summarization}.  The method uses a model of
-sentence relevance and novelty in order to select appropriate updates.
-Training data for regression targets is automatically generated from
-retrospective Wikipedia data.  The system is evaluated on news documents
-related to 197 natural and human disasters from 2009 to 2011 using variants of
-Rouge modified to capture novelty, relevance, and comprehensiveness
-\cite{lin2004rouge}.  Wang and Li present a clustering-based approach to
-efficiency detect important updates during natural disasters
-\cite{wang:update-summarization}.  The algorithm works by hierarchically
+disaster.  
+%The method uses a model of
+%sentence relevance and novelty in order to select appropriate updates.
+%Training data for regression targets is automatically generated from
+%retrospective Wikipedia data.  The system is evaluated on news documents
+%related to 197 natural and human disasters from 2009 to 2011 using variants of
+%ROUGE modified to capture novelty, relevance, and comprehensiveness
+%\cite{lin2004rouge}.  
+\cite{wang:update-summarization} present a clustering-based approach to
+efficiency detect important updates during natural disasters.  
+The algorithm works by hierarchically
 clustering sentences online, allowing the system to output a more expressive
-narrative structure than Guo \textit{et al.}.  The method is evaluated on
-official press releases related to Hurricane Wilma  in 2005 using Rouge score
-between the system summary and a manually generated target summary.
+narrative structure than \cite{qi:temporal-summarization}.  
+Our system attempts to unify these system's approaches 
+(predictive ranking and clustering respectively). 
+%The method is evaluated on
+%official press releases related to Hurricane Wilma  in 2005 using ROUGE score
+%between the system summary and a manually generated target summary. 
+%In addition
+%to natural disasters, we evaluate our system on a variety of event types 
+% such as terrorism and mass shootings (e.g., the 2012
+%shooting in Aurora, Col.), accidents
+%(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
+%Arab spring). The size and scope of the events varies considerably. Hurricane
+%Sandy, for example, affected multiple countries over weeks while the
+%2012 Aurora shooting was contained to a single location spanning several hours.
 
-\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}
+%\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}

### 2015-02-28 results.tex
ck rev

@@
 %baselines. 
-This improvement is statistically significant for all ngram and longest common substring
-precision, recall, and F-measure
+This improvement is statistically significant for all ngram
+precision, recall, and F-measures
 at the $\alpha = .01$ level
@@
 \label{fig:trouge}
+\vspace{-14pt}
 \end{figure}
@@
 \label{fig:nperf}
+\vspace{-12pt}
 \end{figure}
@@
 domain specific features.
-This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
+%This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
 

### 2015-02-28 conclusion.tex
almost final draft

@@
 
-Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
-better summaries; specifically features that capture the language and location
-of the domain.
+%Addtionally, we have shown that feature groups specifically targeted for the domain of disaster yield
+%better summaries; specifically features that capture the language and location
+%of the domain.
+
+We also have shown that features specifically targeted to the domain of disaster yield better summaries.
+We developed novel features that capture the language typical of different event types and that identify sentences specific to the particular disaster based on location.
 
@@
 
-We would also like to further analyze the failure 
-of the temporal relevance features; intuitively we believe time to be important
-for the update summarization task.
+%We would also like to further analyze the failure 
+%of the temporal relevance features; intuitively we believe time to be important
+%for the update summarization task.
 

### 2015-02-28 dataset.tex
ck rev
almost final draft

@@
 For the document stream, we use the news portion of the
- 2014 TREC KBA Stream Corpus.
+ 2014 TREC KBA Stream Corpus \cite{frank2012building}.
 The documents from this corpus come from hourly crawls of the web covering 
@@
 For each event we create a stream of relevant documents by 
-selecting only documents with article content that contains the complete set 
+selecting only documents that contain the complete set 
 of query words. 

### 2015-02-28 experiments.tex
almost final draft

@@
-We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, a common automatic
+We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, an automatic
 summarization method and an evaluation of system expected gain 
@@
 
-ROUGE measures the ngram overlap between a model (i.e. human) summary 
+ROUGE measures the ngram overlap between a model summary 
 and an automatically generated system summary. 
@@
 1000 random summaries of length less than or equal to the model summary 
-(truncating the last sentence when neccessary). The final ROUGE score for the 
+(truncating the last sentence when neccessary). The final ROUGE scores for the 
 system are the average scores from these 1000 samples.
@@
 %each system is able to achieve its maximum performance.
-and this evaluation reveals how quickly each system is able to cover  information in the summary.
+and this evaluation reveals how quickly systems can cover  information in the model.
 
@@
 %NIST developed a series metrics for evaluating update summarization systems
-NIST developed a series of metrics for evaluating update summarization systems
+NIST developed metrics for evaluating update summarization systems
 as part of the TREC TS track.
@@
 where $N$ is the set of event nuggets.
-A comprehensiveness of 1 indicates that the summary has cover 
+A comprehensiveness of 1 indicates that the summary has covered
 all nugget information for the event; the maximum
@@
 
-Update-nugget matches are computed automatically, where a
+Update-nugget matches are computed automatically; a
 match exists if the semantic similarity of the update/nugget pair is above 
@@
 The average semantic similarity of manual matches performed 
-by NIST assessors was considerably lower at approximately .25, increasing
+by NIST assessors was much lower at approximately .25, increasing
 our confidence in the automatic matches in the .5--1 range.
@@
 \subsection{Model Comparisons}
-We refer to our complete model as \textbf{AP+Salience}.  We compare this model again several variants and baselines intended to measure the contribution of different components. All thresholds
+We refer to our complete model as \textbf{AP+Salience}.  We compare this model against several variants and baselines intended to measure the contribution of different components. All thresholds
 for all runs are tuned on the development set.

### 2015-02-28 introduction.tex
almost final draft

@@
 We demonstrate that combining salience with clustering produces more relevant summaries
-compared to baselines using clustering alone.  
+compared to baselines using clustering or relevance alone.  
 Our experiments suggest that this
-because our system is better able to adapt to
+is because our system is better able to adapt to
 dynamic changes in input volume which adversely 
-effect methods that use redundancy as a proxy for salience. 
+affect methods that use redundancy as a proxy for salience. 
 %or predict salience but
@@
 prediction, our approach also exploits knowledge about the event to determine
-salience. Thus, salience does not just represent importance within a set of
-documents; it also represents both how typical a sentence is of the  
+salience. Thus, salience represents both how typical a sentence is of the  
 event
@@
 the event, and the change in word frequencies over the time of the event.
-While we evaluate these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
-expected similarity of their ideal summaries.
+While we evaluate these features in the domain of disaster, this approach is generally applicable to many update summarization tasks.
+% where queries can be categorized by the 
+%expected similarity of their ideal summaries.
 %While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 

### 2015-02-28 methods.tex
almost final draft

@@
 throughout the time period of interest, the system outputs sentences
-from these documents likely to be useful to query issuer.  We refer
+from these documents likely to be useful to the query issuer.  We refer
 to these selected sentences as \emph{updates}.
@@
 in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
-several hunder nuggets in total.
+several hundred nuggets in total.
 We describe how these 
@@
 We want our model to be predictive of sentence salience across different event instances so we avoid lexical features.
-Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.
+Instead, we extract features such as language model scores, geographic relevance, and temporal relevance from each sentence.
 
-\paragraph{Basic Features}
+\textbf{Basic Features}
 We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
@@
 The data stream comprises text extracted from raw html documents;
-these features help to downweight sentences that are not actually article 
+these features help to downweight sentences that are not 
 content (e.g. web page titles, links to other content) or
@@
 
-\paragraph{Query Features}
-
+\textbf{Query Features}
 Query features measure the relationship between the sentence and the event query and type.
@@
 
-\paragraph{Language Model Features}\label{subsubsec:lm}
-Language models allow us to measure the likelihood of a sentence having been 
-produced from a particular source.  We consider two types of language model 
+\textbf{Language Model Features}\label{subsubsec:lm}
+Language models allow us to measure the likelihood of producing a sentence 
+from a particular source.  We consider two types of language model 
 features.  The first model is estimated from a corpus of generic news 
@@
 We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram Kneser-Ney model for both
-the background language and model and the event specific language models.
+the background language model and the event specific language models.
 For each sentence we use the average token log probability under each model
@@
 
-\paragraph{Geographic Relevance Features}
+\textbf{Geographic Relevance Features}
 %\fdcomment{need introduction here.  what are geographic features?}
@@
 
-\paragraph{Temporal Relevance Features}
+\textbf{Temporal Relevance Features}
 %\fdcomment{motivate burstiness; explain what it is before using the expression.}
@@
 may die down, only to spike back up when there is a breaking development.
-Alternatively, words that were salient initially may become background context,
-and over time lose their importance.
+%Alternatively, words that were salient initially may become background context,
+%and over time lose their importance.
 Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
@@
 
-Exemplars that pass these thresholds are added selected as updates and added
+Exemplars that pass these thresholds are selected as updates and added
 to the summary.

### 2015-02-28 relatedwork.tex
almost final draft

@@
 information theoritic, and set cover \cite{lin2011class}
-approaches. While these methods are generally focused on producing diverse
+approaches. While these methods are  focused on producing diverse
 summaries, they are difficult to adapt to the streaming setting, where 
@@
 \cite{wang:update-summarization} present a clustering-based approach to
-efficiency detect important updates during natural disasters.  
+efficiently detect important updates during natural disasters.  
 The algorithm works by hierarchically

### 2015-05-26 experiments.tex
Addressed comments. Fixed margins/sizings. Added authors.

@@
 clustering component, but without the salience information.
-In this model, input sentences are apriori equally likely to be exemplars;
+In this model, input sentences are \textit{apriori} equally likely to be exemplars;
 %clustering algorithm without the salience predictions.
@@
 
-\textbf{Hierarchichal Agglomerative Clustering (HAC)} : We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
+\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
 We include this baseline to show that AP+Salience is not just an improvement
@@
 We also isolate the impact of our salience model in order to demonstrate 
-the that the fusion of clustering and salience prediction improves over
+that the fusion of clustering and salience prediction improves over
 predicting salience alone.

### 2015-05-26 introduction.tex
Addressed comments. Fixed margins/sizings. Added authors.

@@
 is because our system is better able to adapt to
-dynamic changes in input volume which adversely 
+dynamic changes in input volume that adversely 
 affect methods that use redundancy as a proxy for salience. 

### 2015-05-26 methods.tex
Addressed comments. Fixed margins/sizings. Added authors.

@@
 event includes sub-events related to the storm making landfall,
-the ensuing flooding, the many transportation issues, amongst many
+the ensuing flooding, the many transportation issues, among many
 others.  An ideal system would update the user about each of these
-sub-events as they occur, with low latency.  
+sub-events as they occur. %, with low latency.  
 %In order to use
@@
 Sandy event 
-in Figure \ref{fig:nuggets}. Note that each event has anywhere from 50 to
-several hundred nuggets in total.
+in Figure \ref{fig:nuggets}. 
+Note that each event has anywhere from 50 to
+several hundred nuggets in total in our gold dataset.
+%Note that each event has anywhere from 50 to
+%several hundred nuggets in total.
 We describe how these 
@@
 We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.
+These include sentence length, the number of capitalized words normalized by sentence length, document position, and number of named entities.
 The data stream comprises text extracted from raw html documents;
@@
 \textbf{Query Features}
-Query features measure the relationship between the sentence and the event query and type.
-These include the number of query words present in the sentence in addition to the number of event type synonyms, hypernyms, and hyponyms as found in WordNet \cite{miller1995wordnet}.
+Query features measure the relationship between the sentence and the event 
+query and type.  These include the number of query words present in the 
+sentence in addition to the number of event type synonyms, hypernyms, and 
+hyponyms using WordNet \cite{miller1995wordnet}.  
 For example, for event type \emph{earthquake},  we match sentence terms 
@@
 The disasters in our corpus are all phenomena that
-affect some part of the word. 
+affect some part of the world. 
 Where possible, we would like to capture a sentence's proximity to the event,
@@
 There are two challenges to using geographic features. First we do not 
-know where the event is and second most sentences do not contain references
+know where the event is, and second, most sentences do not contain references
 to a location.
@@
 
-To address the second issue, 
-we compute features for the document as a whole and all
-sentences from that document receive the same feature value.
-
+%\ckcomment{(Paragraphs in blue should replace last two paragraphs in geo relevance section)\\
+The second issue arises from the fact that the majority of sentences in our 
+data do not contain explict references to locations, i.e. a sequence of tokens
+tagged as location named entities.   
+Our intuition is that geographic relevance is important in the disaster
+domain, and we would like to take advantage of the sentences that do have 
+location information present. To make up for this imbalance, we  
+%Extracting features for just these sentences
+%would leave 
+% Last sentence is unclear
+%Assigning a null or default value to sentences without explicit location is 
+%not a viable option as 
+%the imbalance of null/default values would make it difficult for the model
+%to learn meaningful weights for these features.
+% also unclear
+instead compute an overall location for the document and derive 
+geographic features based on the document's proximity to the event in 
+question.
+These features are assigned to all sentences in the document.\\\\
+Our method of computing document-level geographic relevance features is as 
+follows.
 Using the locations in each document, we compute the median distance to the 
@@
 compute these distances to event locations from the previous hour.
+%}
+
+%To address the second issue, 
+%we compute features for the document as a whole and all
+%sentences from that document receive the same feature value.
+
+%Using the locations in each document, we compute the median distance to the 
+%nearest event location. Because document position is a good indicator 
+%of importance we also compute the distance of the first mentioned
+%location to the nearest event location. All sentences in the document take
+%as features these two distance calculations.
+%Because some events can move, we also
+%compute these distances to event locations from the previous hour.
 
@@
 \subsubsection{Model}
-
-At run-time
-our summarization system
-does not have access to event nuggets to compute sentence salience. 
-However, we can use the nuggets
-in our training data to model salience as a function of our features.
-Such a model generalizes to new events.
+%\ckcomment{(replace section with this)\\
+Given our feature representation of the input sentences, we need only target 
+salience values for model learning. 
 For each event in our training data, we sample a set of sentences and  each 
 sentence's salience is computed according to Equation \ref{eq:salience}.
-This results in a training set of sentences and the target values 
-(their salience) to predict.
+This results in a training set of sentences, their feature representations,
+and their target salience values to predict.
 
-Given a feature representation for the sentences, we can use
-any regression technique to predict salience.
+%Given the feature representation/salience pairs, we can use
+%any regression technique to predict salience.
 We opt to use a Gaussian process (GP) regression model
 \cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
-kernel.  
+kernel for the salience prediction task.  
 Our features fall naturally into five groups and we use a separate RBF kernel
@@
 to the GP model.
+%}
+
+%At run-time
+%our summarization system
+%does not have access to event nuggets to compute sentence salience. 
+%However, we can use the nuggets
+%in our training data to model salience as a function of our features.
+%Such a model generalizes to new events.
+%For each event in our training data, we sample a set of sentences and  each 
+%sentence's salience is computed according to Equation \ref{eq:salience}.
+%This results in a training set of sentences and the target values 
+%(their salience) to predict.
+%
+%Given a feature representation for the sentences, we can use
+%any regression technique to predict salience.
+%We opt to use a Gaussian process (GP) regression model
+%\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
+%kernel.  
+%Our features fall naturally into five groups and we use a separate RBF kernel
+%for each, using the sum of each feature group RBF kernel as the final input
+%to the GP model.
+
 %GP regressors are a
@@
 similarity
-of data points and our predicted apriori preference of a data point to be an exemplar
+of data points and our predicted \textit{apriori} preference of a data point to be an exemplar
 respectively. 
@@
 predictions are low, the preference term will guide AP toward a solution
-with fewer clusters; vice-versa when input is very salient on average but
+with fewer clusters; \textit{vice-versa} when input is very salient on average but
 the volume of input is low. The adaptive nature of our model differentiates

### 2015-05-26 results.tex
almost final draft
Addressed comments. Fixed margins/sizings. Added authors.

@@
 \label{fig:trouge}
-\vspace{-14pt}
+%\vspace{-14pt}
 \end{figure}
@@
 \subsection{Expected Gain and Comprehensiveness}
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[]{nuggets-metrics.eps}
@@
 \label{fig:nperf}
-\vspace{-12pt}
+%\vspace{-12pt}
 \end{figure}
@@
  AP+Salience. The compromise encoded in the AP+Salience objective
-function, between being representitive and being salient, is seen clearly here
+function, between being representative and being salient, is seen clearly here
 where the performance of the AP+Salience methods is lower bounded by 
@@
 domain specific features.
+The result mirrors Sparck Jones' claim that generic approaches to summarization cannot produce a useful summary \cite{ksj98}.
 %This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.

### 2015-05-29 dataset.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
-For the document stream, we use the news portion of the
- 2014 TREC KBA Stream Corpus \cite{frank2012building}.
-The documents from this corpus come from hourly crawls of the web covering 
- October 2011 through February 2013.
+For the document stream, we use the news portion of the 2014 TREC KBA Stream 
+Corpus \cite{frank2012building}. The documents from this corpus come from 
+hourly crawls of the web covering October 2011 through February 2013.
 
 Our experiments also make use of the TREC Temporal Summarization (TS) Track
- data from 2013 and 2014 \cite{aslam2013trec}. 
-This data includes 25 events and event metadata (e.g., a user
-search query for the event, the event type, and event evaluation timeframe).  
-All events occurred during the time span of the TREC KBA Stream Corpus.
-For each event we create a stream of relevant documents by 
-selecting only documents that contain the complete set 
-of query words. 
-%identifying the 
-%article start/end within each web page in the corpus using a classifier
-%tra
-%those documents
+ data from 2013 and 2014 \cite{aslam2013trec}.  This data includes 25 events 
+and event metadata (e.g., a user search query for the event, the event type, 
+and event evaluation time frame). All events occurred during the time span of 
+the TREC KBA Stream Corpus. For each event we create a stream of relevant 
+documents by selecting only documents that contain the complete set of query 
+words.
 
-Along with the metadata, NIST assessors constructed a set of ground truth nuggets for each event. 
-Nuggets are brief and important text snippets that represent sub-events that should be conveyed
-by an ideal update summary. % (see Figure~\ref{fig:nuggets} for examples).
-In order
-to accomplish this, for each event, assessors were provided with the
-revision history of the Wikipedia page associated with the event.  
-For example, 
-the revision history for the Wikipedia page for `Hurricane Sandy' will 
-contain text additions including those related to individual nuggets.  The assessment
-task involves reviewing the Wikipedia revisions in the evaluation timeframe 
+Along with the metadata, NIST assessors constructed a set of ground truth 
+nuggets for each event. Nuggets are brief and important text snippets that 
+represent sub-events that should be conveyed by an ideal update summary. 
+In order to accomplish this, for each event, assessors were provided with the
+revision history of the Wikipedia page associated with the event. For example,
+the revision history for the Wikipedia page for `Hurricane Sandy' will contain
+text additions including those related to individual nuggets.  The assessment
+task involves reviewing the Wikipedia revisions in the evaluation time frame 
 and marking the text additions capturing a new, unique nugget.  More detail

### 2015-05-29 experiments.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
 We evaluate our system on two metrics: ROUGE \cite{lin2004rouge}, an automatic
-summarization method and an evaluation of system expected gain 
-and comprehensiveness---metrics adapted 
-from the TREC TS track \cite{aslam2013trec}.
+summarization method and an evaluation of system expected gain and 
+comprehensiveness---metrics adapted from the TREC TS track 
+\cite{aslam2013trec}.
 
@@
 
-Of the 25 events in the TREC TS data, 24 are covered by the news portion 
-of the TREC KBA Stream Corpus.
-From these 24, we set aside three events to use as a development set.
-All system salience and similarity threshold parameters are tuned on the
-development set to maximize ROUGE-2 F1 scores.
+Of the 25 events in the TREC TS data, 24 are covered by the news portion of 
+the TREC KBA Stream Corpus. From these 24, we set aside three events to use as
+a development set. All system salience and similarity threshold parameters are
+tuned on the development set to maximize ROUGE-2 F1 scores.
 
@@
 from the event's document stream.
-%Kernel length scale parameters  for each
-%feature group are fit to maximize
-%the model's log likelihood using ???.
 
-We perform a leave-one-out evaluation of each event.
-At test time, we predict a sentence's salience using the average predictions
-of the 23 other models.
+We perform a leave-one-out evaluation of each event. At test time, we predict 
+a sentence's salience using the average predictions of the 23 other models.
 
@@
 
-ROUGE measures the ngram overlap between a model summary 
-and an automatically generated system summary. 
-Model summaries for each event were constructed by concatenating the event's 
-nuggets. 
-%See figure~\ref{fig:test} for an example model summary excerpt.
-Generally, ROUGE evaluation assumes both model and system summaries
-are of a bounded length. Since our systems are summarizing events over a span
-of two weeks time, the total length of our system output is much longer than
-the model. To address this, for each system/event pair, we sample with replacement
-1000 random summaries of length less than or equal to the model summary 
-(truncating the last sentence when neccessary). The final ROUGE scores for the 
-system are the average scores from these 1000 samples.
+ROUGE measures the ngram overlap between a model summary and an automatically 
+generated system summary. Model summaries for each event were constructed by 
+concatenating the event's nuggets. Generally, ROUGE evaluation assumes both 
+model and system summaries are of a bounded length. Since our systems are 
+summarizing events over a span of two weeks time, the total length of our 
+system output is much longer than the model. To address this, for each 
+system/event pair, we sample with replacement 1000 random summaries of length 
+less than or equal to the model summary (truncating the last sentence when 
+neccessary). The final ROUGE scores for the system are the average scores 
+from these 1000 samples.
 
 Because we are interested in system performance over time, we also evaluate 
-systems at 12 hour intervals using the same regime as above. 
-The model summaries in this case are retrospective, 
-%and so we can see how quickly
-%each system is able to achieve its maximum performance.
-and this evaluation reveals how quickly systems can cover  information in the model.
+systems at 12 hour intervals using the same regime as above. The model 
+summaries in this case are retrospective, and this evaluation reveals how 
+quickly systems can cover  information in the model.
 
@@
 
-%NIST developed a series metrics for evaluating update summarization systems
-NIST developed metrics for evaluating update summarization systems
-as part of the TREC TS track.
-We present results on two of these metrics, the expected gain and 
-comprehensiveness.
+NIST developed metrics for evaluating update summarization systems as part of 
+the TREC TS track. We present results on two of these metrics, the expected 
+gain and comprehensiveness.
 
 \textbf{Expected Gain } We treat the event's nuggets as unique units of 
-information.
-When a system adds an update to its summary, it is potentially adding some
-of this nugget information. It would be instructive to know how much unique
-and novel information each update is adding on average to the summary.
-To that end, we define
+information. When a system adds an update to its summary, it is potentially 
+adding some of this nugget information. It would be instructive to know how 
+much unique and novel information each update is adding on average to the 
+summary. To that end, we define
+\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}
+\] 
+where $S$ is the set of system updates, $S_n$ is the set of nuggets contained 
+in $S$, and $|\cdot|$ is the number of elements in the set. To compute the 
+set $S_n$ we match each system update to 0 or more nuggets, where an update 
+matches a nugget if their semantic similarity is above a threshold. $S_n$ 
+results from the unique set of nuggets matched. Because an update can map to 
+more than one nugget, it is possible to receive an expected gain greater 
+than 1. An expected gain of 1 would indicate that every sentence was both 
+relevant and contained a unique piece of information.
 
-\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
-\] 
-where $S$ is the set of system updates, 
-$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
-elements in the set.
-To compute the set $S_n$ we match each system update to 0 or more nuggets, 
-where an update matches a nugget if their semantic similarity is above 
-a threshold. $S_n$ results from the unique set of nuggets matched.
-Because an update can map to more than one nugget, it is possible to receive an
-expected gain greater than 1.
-%values greater than one can be thought of as a 
-%compression ration, i.e. an expected gain of 10 indicates the average update
-%contains 10 unique nuggets on average and that the summary is a 10~:~1 
-%compression of the nugget information.
-An expected gain of 1 would indicate that every
-sentence was both relevant and contained a unique piece of information.
-
-\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure the completeness of an update 
-summary. We define
-
+\textbf{Comprehensiveness } Additionally, we can use the nuggets to measure 
+the completeness of an update summary. We define
 \[ \mathrm{Comprehensiveness} = \frac{|S_n|}{|N|}\]
-
-where $N$ is the set of event nuggets.
-A comprehensiveness of 1 indicates that the summary has covered
-all nugget information for the event; the maximum
+where $N$ is the set of event nuggets. A comprehensiveness of 1 indicates that
+the summary has covered all nugget information for the event; the maximum
 attainable comprehensiveness is 1.
 
-Update-nugget matches are computed automatically; a
-match exists if the semantic similarity of the update/nugget pair is above 
-a threshold.
-Determining an optimal threshold to count matches is difficult so 
-we evaluate at 
-threshold values ranging
-from .5 to 1, where values closer to 1 are more conservative estimates of
-performance.
-A manual inspection of matches 
-suggests that semantic similarity values around .7 produce reasonable matches.
-%For example, the nugget 
-%\begin{blockquote}
-%watches and warnings in effect for much of the Mid-Atlantic states and 
-%southern New England 
-%\end{blockquote}
-%\noindent
-%matches with the update 
-%\begin{blockquote}
-%A tropical storm warning remains in effect for much of the coasts 
-%of North and South Carolina\end{blockquote}
-%\noindent
-%with a similarity of .7. 
-The average semantic similarity of manual matches performed 
-by NIST assessors was much lower at approximately .25, increasing
-our confidence in the automatic matches in the .5--1 range.
+Update-nugget matches are computed automatically; a match exists if the 
+semantic similarity of the update/nugget pair is above a threshold. 
+Determining an optimal threshold to count matches is difficult so we evaluate 
+at threshold values ranging from .5 to 1, where values closer to 1 are more 
+conservative estimates of performance. A manual inspection of matches suggests
+that semantic similarity values around .7 produce reasonable matches.
+The average semantic similarity of manual matches performed by NIST assessors 
+was much lower at approximately .25, increasing our confidence in the 
+automatic matches in the .5--1 range.
 
 \subsection{Model Comparisons}
-We refer to our complete model as \textbf{AP+Salience}.  We compare this model against several variants and baselines intended to measure the contribution of different components. All thresholds
-for all runs are tuned on the development set.
 
-\textbf{Affinity Propagation only (AP) } The purpose of this model
-is to directly measure the effect of integrating salience and clustering by providing a baseline that uses the identical 
-clustering component, but without the salience information.
-In this model, input sentences are \textit{apriori} equally likely to be exemplars;
-%clustering algorithm without the salience predictions.
-the salience values are uniformly set as the median value of the 
-input similarity scores, as is commonly used in the AP literature \cite{frey2007clustering}.
-After clustering a sentence batch, the exemplars are examined in order
-of increasing time since event start and selected as updates if their
-maximum similarity to the previous updates is less than $\lambda_{sim}$,
-as in the novelty filtering stage of AP+Salience.
+We refer to our complete model as \textsc{AP+Salience}.  We compare this model
+against several variants and baselines intended to measure the contribution of
+different components. All thresholds for all runs are tuned on the development
+set.
 
-\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another clustering baseline, single-linkage hierarchichal agglomerative clustering.
-We include this baseline to show that AP+Salience is not just an improvement
-over AP but centrality driven methods in general.
-HAC was chosen over other clustering approaches because the number of clusters 
-is not an explicit hyper-parameter. To produce flat clusters from the hierarchical clustering, we flatten the HAC dendrogram
-using the cophenetic distance criteria, i.e. observations in each flat cluster have no greater a cophenetic distance than a threshold.
-Cluster centers are determined to be the sentence with highest cosine similariy to the flat cluster mean.
-Cluster centers are examined in time order and are added to the summary if their similarity to previous updates is below a similarity threshold $\lambda_{sim}$ as is done in the AP model.
+\textbf{Affinity Propagation only (\textsc{AP}) } The purpose of this model 
+is to 
+directly measure the effect of integrating salience and clustering by 
+providing a baseline that uses the identical clustering component, but without
+the salience information. In this model, input sentences are \textit{apriori}
+equally likely to be exemplars; the salience values are uniformly set as the 
+median value of the  input similarity scores, as is commonly used in the 
+\textsc{AP} 
+literature \cite{frey2007clustering}. After clustering a sentence batch, the 
+exemplars are examined in order of increasing time since event start and 
+selected as updates if their maximum similarity to the previous updates is 
+less than $\lambda_{sim}$, as in the novelty filtering stage of 
+\textsc{AP+Salience}.
+
+\textbf{Hierarchichal Agglomerative Clustering (HAC)} We provide another 
+clustering baseline, single-linkage hierarchichal agglomerative clustering.
+We include this baseline to show that \textsc{AP+Salience} is not just an 
+improvement
+over \textsc{AP} but centrality driven methods in general. \textsc{HAC} was 
+chosen over other 
+clustering approaches because the number of clusters is not an explicit 
+hyper-parameter. To produce flat clusters from the hierarchical clustering, we
+flatten the \textsc{HAC} dendrogram using the cophenetic distance criteria, 
+i.e. 
+observations in each flat cluster have no greater a cophenetic distance than a
+threshold. Cluster centers are determined to be the sentence with highest 
+cosine similariy to the flat cluster mean. Cluster centers are examined in 
+time order and are added to the summary if their similarity to previous 
+updates is below a similarity threshold $\lambda_{sim}$ as is done in the 
+\textsc{AP} model.
 
@@
 that the fusion of clustering and salience prediction improves over
-predicting salience alone.
-In this model we predict the salience of sentences as in step 1 for 
-AP+Salience. %However, rather than clustering, we simply rank the sentences
-%by salience.
-We omit the clustering phase (step 2).
-Updates are selected identically to step 3 of AP+Salience, proceeding 
-in order of decreasing salience, selecting updates that are above a salience 
+predicting salience alone. In this model we predict the salience of sentences 
+as in step 1 for \textsc{AP+Salience}. We omit the clustering phase (step 2).
+Updates are selected identically to step 3 of \textsc{AP+Salience}, 
+proceeding in order
+of decreasing salience, selecting updates that are above a salience 
 threshold $\lambda_{sal}$ and below a similarity threshold $\lambda_{sim}$
 with respect to the previously selected updates.
-
-%\fdcomment{consider removing AP+SalienceTR}
-%\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
-%to the
-%AP+Salience model except at the novelty filter stage.
-%Within the novelty filtering stage, 
-%AP+Salience inspects exemplars in order of decreasing salience when computing
-%similarity to previous updates. Because the order in which we add updates
-%affects future similarity calculations, we evaluate an alternate ordering 
-%scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
-%time since event start. The purpose of this model is to give preference to 
-%more recent updates, since timelyness is one of the desirable traits of our update
-%summaries.
-%

### 2015-05-29 introduction.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
 \label{sec:introduction}
-During crises, information is critical for first responders,
- crisis management organizations, and those caught in the event.  
-When the event is significant, as in the case of Hurricane
-Sandy, the amount of content produced by traditional news outlets,
-government agencies, relief organizations, and social media can vastly
-overwhelm those trying to monitor the situation. 
-Crisis informatics \cite{palen2010vision} is dedicated to finding methods for sharing the
-right information in a timely fashion during such an event.
-Research in this field has focused on
-human-in-the-loop approaches ranging from on the ground information 
-gathering to crowdsourced reporting and disaster management \cite{starbird2013working}.
+
+During crises, information is critical for first responders, crisis management
+organizations, and those caught in the event. When the event is significant, 
+as in the case of Hurricane Sandy, the amount of content produced by 
+traditional news outlets, government agencies, relief organizations, and 
+social media can vastly overwhelm those trying to monitor the situation. 
+Crisis informatics \cite{palen2010vision} is dedicated to finding methods for 
+sharing the right information in a timely fashion during such an event.
+Research in this field has focused on human-in-the-loop approaches ranging 
+from on the ground information gathering to crowdsourced reporting and 
+disaster management \cite{starbird2013working}.
 
 Multi-document summarization has the potential to assist the crisis 
-informatics community. 
-Automatic summarization could deliver
-relevant and salient information at regular intervals, 
-even when human volunteers are unable to. 
-Perhaps more importantly it could help filter out
-unnecessary and irrelevant detail when the volume of incoming information
-is large. 
-While methods for identifying,
-tracking, and summarizing events from text based input have been explored
-extensively
-%KM - I think we should consider adding other event summarizers. 
+informatics community. Automatic summarization could deliver relevant and 
+salient information at regular intervals, even when human volunteers are 
+unable to. Perhaps more importantly it could help filter out unnecessary and 
+irrelevant detail when the volume of incoming information is large. While 
+methods for identifying, tracking, and summarizing events from text based 
+input have been explored extensively
 \cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}, 
 these experiments were not developed to handle streaming data from a
- heterogeneous environment at web scale. 
-These methods also rely heavily on redundancy which is suboptimal for time
-sensitive domains where there is a high cost in delaying information.
+heterogeneous environment at web scale. These methods also rely heavily on 
+redundancy which is suboptimal for time sensitive domains where there is a 
+high cost in delaying information.
 
@@
 large-scale event, such as a disaster, and integrates these predictions into
-a clustering based multi-document summarization system. 
-We demonstrate that combining salience with clustering produces more relevant summaries
-compared to baselines using clustering or relevance alone.  
-Our experiments suggest that this
-is because our system is better able to adapt to
-dynamic changes in input volume that adversely 
-affect methods that use redundancy as a proxy for salience. 
-%or predict salience but
-%ignore redundancy in the input alone.
+a clustering based multi-document summarization system. We demonstrate that 
+combining salience with clustering produces more relevant summaries compared 
+to baselines using clustering or relevance alone.  Our experiments suggest 
+that this is because our system is better able to adapt to dynamic changes in 
+input volume that adversely affect methods that use redundancy as a proxy for 
+salience.
 
@@
 the event, and the change in word frequencies over the time of the event.
-While we evaluate these features in the domain of disaster, this approach is generally applicable to many update summarization tasks.
-% where queries can be categorized by the 
-%expected similarity of their ideal summaries.
-%While we ground these features in the domain of disaster, this approach is generally applicable to any update summarization task where queries can be categorized by the 
-%expected similarity of their ideal summaries.
-% to represent its likelihood of referring to the input disaster. 
-%We also utilize the change in word frequencies as we process the stream 
-%to measure the changing importance of updates over time.
+While we evaluate these features in the domain of disasters, this approach is 
+generally applicable to many update summarization tasks.
 
-Our approach achieves a statistically significant improvement in 
- ROUGE scores compared to multiple baselines.
-Additionally, we introduce novel methods for estimating the average information
-gain each update provides and how completely the update summary covers 
-the event it is tracking; our system's updates contain more relevant
-information on average than the competeing baselines.
+Our approach achieves a statistically significant improvement in ROUGE scores 
+compared to multiple baselines. Additionally, we introduce novel methods for 
+estimating the average information gain each update provides and how 
+completely the update summary covers the event it is tracking; our system's 
+updates contain more relevant information on average than the competing 
+baselines.
 
-The remainder of the paper is organized as follows.
-We begin with a review of related work
-in the information retrieval and multi-document
-summarization literature. Section~\ref{sec:methods} outlines the details
-of our salience and summarization models. Next we describe our
-data (Section~\ref{sec:data}) and experiments (Section~\ref{sec:exper}). Finally,
-we discuss our results (Section~\ref{sec:results}) and conclude the paper.
-
-%Our main contributions are three fold. First, we present a straightforward 
-%framework for 
-%combining prior beliefs with a clustering algorithm. Second, we demonstrate
-%the empirical effectiveness of this approach on the update summarization
-%task. Finally, we present a novel feature represention for sentence salience
-%in the crisis/disaster domain.
+The remainder of the paper is organized as follows. We begin with a review of 
+related work in the information retrieval and multi-document summarization 
+literature. Section~\ref{sec:methods} outlines the details of our salience 
+and summarization models. Next we describe our data (Section~\ref{sec:data}) 
+and experiments (Section~\ref{sec:exper}). Finally, we discuss our results 
+(Section~\ref{sec:results}) and conclude the paper.

### 2015-05-29 methods.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
 \label{sec:methods}
-%In order to evaluate an update summarization system, we adopt the simulation-based approach used in the TREC Temporal Summarization (TS) Track.
-We provide a brief overview of the problem.
-Details on the formulation can be found in the track overview paper \cite{aslam2013trec}.
 
@@
 \begin{enumerate*}[label=\itshape\alph*\upshape)]
-  \item a short query defining the event to be tracked (e.g. `Hurricane Sandy'), 
-  \item an event category defining the type of event to be tracked (e.g. `hurricane'), 
-  \item a stream of time-stamped documents %, $(\doc_0, \doc_1,\ldots,\doc_T)$,
-  presented in temporal order, and \item an evaluation time period of interest.
-    %  $(\stime,\etime)$.  
+  \item a short query defining the event to be tracked 
+  (e.g. `Hurricane Sandy'), 
+  \item an event category defining the type of event to be tracked 
+  (e.g. `hurricane'), 
+  \item a stream of time-stamped documents presented in temporal order, 
+  and \item an evaluation time period of interest.
 \end{enumerate*} 
-While processing documents
-throughout the time period of interest, the system outputs sentences
-from these documents likely to be useful to the query issuer.  We refer
-to these selected sentences as \emph{updates}.
+While processing documents throughout the time period of interest, the system 
+outputs sentences from these documents likely to be useful to the query 
+issuer.  We refer to these selected sentences as \emph{updates}.
 
-In order to measure the usefulness of a system's updates, we consider
-the degree to which the system output reflects the different
-aspects of an event.  Events are often composed of a variety of sub-events.  
-For example, the Hurricane Sandy
-event includes sub-events related to the storm making landfall,
-the ensuing flooding, the many transportation issues, among many
+In order to measure the usefulness of a system's updates, we consider the 
+degree to which the system output reflects the different aspects of an event.
+Events are often composed of a variety of sub-events.  For example, the 
+Hurricane Sandy event includes sub-events related to the storm making 
+landfall, the ensuing flooding, the many transportation issues, among many
 others.  An ideal system would update the user about each of these
-sub-events as they occur. %, with low latency.  
-%In order to use
-%language consistent with previous literature, 
-We refer to these
-sub-events as the \emph{nuggets} associated with an event.  A nugget is
-defined as a fine-grained atomic sub-event associated with an event.  
-%An example nugget for Hurricane Sandy might be ``Sandy on course for the Jersey Shore.''
-We present 2 example nuggets associated with the Hurricane
-Sandy event 
-in Figure \ref{fig:nuggets}. 
-Note that each event has anywhere from 50 to
-several hundred nuggets in total in our gold dataset.
-%Note that each event has anywhere from 50 to
-%several hundred nuggets in total.
-We describe how these 
-nuggets are found in Section \ref{sec:data}.
+sub-events as they occur. We refer to these sub-events as the \emph{nuggets} 
+associated with an event.  A nugget is defined as a fine-grained atomic 
+sub-event associated with an event.  We present 2 example nuggets associated 
+with the Hurricane Sandy event in Figure \ref{fig:nuggets}. Each event has 
+anywhere from 50 to several hundred nuggets in total in our gold dataset.
+We describe how these nuggets are found in Section \ref{sec:data}.
 
@@
 
-Throughout our treatment of our algorithm, the \emph{salience} 
-of an update captures the degree to which it reflects an 
-event's nuggets.  Assuming that we have a text representation 
-for each of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
-defined as,
+Throughout our treatment of our algorithm, the \emph{salience} of an update 
+captures the degree to which it reflects an event's unobserved nuggets.  
+Assuming that we have a text representation for each of our nuggets, the 
+salience of an update $u$  with respect to a set of nuggets $N$ is defined as,
 \begin{align}
@@
 \end{align}
-where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
-the cosine similarity of latent vectors associated with the update and 
-nugget text \cite{guo2012simple}.
+where $\operatorname{sim}(\cdot)$ is the semantic similarity such as the 
+cosine similarity of latent vectors associated with the update and nugget 
+text \cite{guo2012simple}.
 
@@
 
-Our system architecture follows a simple pipeline design where each
-stage provides an additional level of processing or filtering of the input
-sentences.
-We begin with an empty update summary $U$.
-At each hour we receive a new batch of sentences $b_t$ from the stream
-of event relevant documents
+Our system architecture follows a simple pipeline design where each stage 
+provides an additional level of processing or filtering of the input 
+sentences. We begin with an empty update summary $U$. At each hour we receive 
+a new batch of sentences $b_t$ from the stream of event relevant documents
 and perform the following actions:
-\begin{enumerate}[nosep]
+\begin{enumerate}
   \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
@@
 
-%include feature table
-%\input{feature_table.tex}
-
 \subsubsection{Features}
 \label{sec:features}
-We want our model to be predictive of sentence salience across different event instances so we avoid lexical features.
+
+We want our model to be predictive of sentence salience across different event instances so we avoid event-specific lexical features.
 Instead, we extract features such as language model scores, geographic relevance, and temporal relevance from each sentence.
@@
 \textbf{Basic Features}
-We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}.
-These include sentence length, the number of capitalized words normalized by sentence length, document position, and number of named entities.
-The data stream comprises text extracted from raw html documents;
-these features help to downweight sentences that are not 
-content (e.g. web page titles, links to other content) or
-more heavily weight important sentences (e.g., that appear in
-prominent positions such as paragraph initial or article initial).
+We employ several basic features that have been used previously in supervised 
+models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. 
+These include sentence length, the number of capitalized words normalized by 
+sentence length, document position, and number of named entities.  The data 
+stream comprises text extracted from raw html documents; these features help 
+to downweight sentences that are not content (e.g. web page titles, links to 
+other content) or more heavily weight important sentences (e.g., that appear 
+in prominent positions such as paragraph initial or article initial).
 
@@
 sentence in addition to the number of event type synonyms, hypernyms, and 
-hyponyms using WordNet \cite{miller1995wordnet}.  
-For example, for event type \emph{earthquake},  we match sentence terms 
-``quake'', ``temblor'', ``seism'', and ``aftershock''.
+hyponyms using WordNet \cite{miller1995wordnet}.  For example, for event type 
+\emph{earthquake},  we match sentence terms ``quake'', ``temblor'', ``seism'',
+and ``aftershock''.
 
@@
 Language models allow us to measure the likelihood of producing a sentence 
-from a particular source.  We consider two types of language model 
-features.  The first model is estimated from a corpus of generic news 
-articles (we used the 1995-2010 Associated Press section of the 
-Gigaword corpus \cite{graff2003english}).  
-This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
+from a particular source.  We consider two types of language model features.  
+The first model is estimated from a corpus of generic news articles (we used 
+the 1995-2010 Associated Press section of the Gigaword corpus 
+\cite{graff2003english}).  
+This model is intended to assess the general writing quality (grammaticality, 
+word usage) of an input sentence and helps our model to select sentences
 written in the newswire style.
 
-The second model is estimated from text specific to our event types.  
-For each event type we create a corpus of related documents using pages
-and subcategories listed under a related Wikipedia category.
-For example, the language model for event type `earthquake' is estimated 
-from Wikipedia pages under the category \emph{Category:Earthquakes}.  
-%\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
-%expected summary is similar to some previously seen target or some side information.}
-These models are intended to detect sentences similar to those appearing in 
-summaries of other events in the same category 
-(e.g. most earthquake summaries are likely to include higher probability for 
-ngrams including the token `magnitude'). While we focus our system on the 
-language of news and disaster, we emphasize that the use of language modeling 
-can be an effective feature for multi-document summarization for other 
-domains that have related text corpora.
+The second model is estimated from text specific to our event types.  For each
+event type we create a corpus of related documents using pages and 
+subcategories listed under a related Wikipedia category. For example, the 
+language model for event type `earthquake' is estimated from Wikipedia pages 
+under the category \emph{Category:Earthquakes}.  These models are intended to 
+detect sentences similar to those appearing in summaries of other events in 
+the same category (e.g. most earthquake summaries are likely to include 
+higher probability for ngrams including the token `magnitude'). While we focus
+our system on the language of news and disaster, we emphasize that the use of 
+language modeling can be an effective feature for multi-document summarization
+for other domains that have related text corpora.
 
-We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram Kneser-Ney model for both
-the background language model and the event specific language models.
-For each sentence we use the average token log probability under each model
-as a feature.
-
-%KM - Note: Someplace the exact list of event types should appear. Probably not
-%here.
-%KM - I note you have it in a later section but it is labeled as data you use
-%to train language models and semantic similarity. I think it would be good to
-%have up front in definition of task.
+We use the SRILM toolkit \cite{stolcke2002srilm} to implement a 5-gram 
+Kneser-Ney model for both the background language model and the event specific
+language models. For each sentence we use the average token log probability 
+under each model as a feature.
 
 \textbf{Geographic Relevance Features}
-%\fdcomment{need introduction here.  what are geographic features?}
-The disasters in our corpus are all phenomena that
-affect some part of the world. 
-Where possible, we would like to capture a sentence's proximity to the event,
-i.e. when a sentence references a location, it should be close to the 
+The disasters in our corpus are all phenomena that affect some part of the 
+world. Where possible, we would like to capture a sentence's proximity to the 
+event, i.e. when a sentence references a location, it should be close to the 
 area of the disaster.
 
-There are two challenges to using geographic features. First we do not 
-know where the event is, and second, most sentences do not contain references
-to a location.
-We address the first issue by extracting all locations from 
-documents relevant to the event at the
-current hour and looking up their latitude and 
-longitude using a publicly available geolocation service. 
-%\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
-Since the documents that are at least somewhat relevant to the event,
-we assume in aggregate the locations should give us a rough area of interest.
-The locations are clustered %(using affinity propagation and uniform salience)
-%\fdcomment{affinity propagation undefined}
-and we treat the resulting cluster centers
-as the event locations for the current time.
+There are two challenges to using geographic features. First, we do not know 
+where the event is, and second, most sentences do not contain references to a 
+location. We address the first issue by extracting all locations from 
+documents relevant to the event at the current hour and looking up their 
+latitude and longitude using a publicly available geo-location service. 
+Since the documents that are at least somewhat relevant to the event, we 
+assume in aggregate the locations should give us a rough area of interest.
+The locations are clustered and we treat the resulting cluster centers as the 
+event locations for the current time.
 
-%\ckcomment{(Paragraphs in blue should replace last two paragraphs in geo relevance section)\\
 The second issue arises from the fact that the majority of sentences in our 
-data do not contain explict references to locations, i.e. a sequence of tokens
-tagged as location named entities.   
-Our intuition is that geographic relevance is important in the disaster
-domain, and we would like to take advantage of the sentences that do have 
-location information present. To make up for this imbalance, we  
-%Extracting features for just these sentences
-%would leave 
-% Last sentence is unclear
-%Assigning a null or default value to sentences without explicit location is 
-%not a viable option as 
-%the imbalance of null/default values would make it difficult for the model
-%to learn meaningful weights for these features.
-% also unclear
-instead compute an overall location for the document and derive 
-geographic features based on the document's proximity to the event in 
-question.
-These features are assigned to all sentences in the document.\\\\
+data do not contain explicit references to locations, i.e. a sequence of 
+tokens tagged as location named entities. Our intuition is that geographic 
+relevance is important in the disaster domain, and we would like to take 
+advantage of the sentences that do have location information present. To make 
+up for this imbalance, we instead compute an overall location for the document
+and derive geographic features based on the document's proximity to the event
+in question. These features are assigned to all sentences in the document.
+
 Our method of computing document-level geographic relevance features is as 
-follows.
-Using the locations in each document, we compute the median distance to the 
-nearest event location. Because document position is a good indicator 
-of importance we also compute the distance of the first mentioned
-location to the nearest event location. All sentences in the document take
-as features these two distance calculations.
-Because some events can move, we also
-compute these distances to event locations from the previous hour.
-%}
-
-%To address the second issue, 
-%we compute features for the document as a whole and all
-%sentences from that document receive the same feature value.
-
-%Using the locations in each document, we compute the median distance to the 
-%nearest event location. Because document position is a good indicator 
-%of importance we also compute the distance of the first mentioned
-%location to the nearest event location. All sentences in the document take
-%as features these two distance calculations.
-%Because some events can move, we also
-%compute these distances to event locations from the previous hour.
+follows. Using the locations in each document, we compute the median distance 
+to the nearest event location. Because document position is a good indicator 
+of importance we also compute the distance of the first mentioned location to 
+the nearest event location. All sentences in the document take as features 
+these two distance calculations. Because some events can move, we also compute
+these distances to event locations from the previous hour.
 
 \textbf{Temporal Relevance Features}
-%\fdcomment{motivate burstiness; explain what it is before using the expression.}
 As we track events over time, it is likely that the coverage of the event 
 may die down, only to spike back up when there is a breaking development.
-%Alternatively, words that were salient initially may become background context,
-%and over time lose their importance.
 Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
@@
 
-We compute the IDF for each hour in our data stream. 
-For each sentence, the average TFIDF for the current hour $t$ is taken as a 
-feature. Additionally, we use the difference in average TFIDF from time $t$
-to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
-sentence have changed over the last 24 hours, i.e. we keep the sentence
-term frequencies fixed and compute the difference in IDF. Large changes
-in IDF value indicate the sentence contains bursty terms.
-%\fdcomment{unclear.  is it just changing the IDF part of the 
-%items in the average?}
+We compute the IDF for each hour in our data stream. For each sentence, the 
+average TF*IDF for the current hour $t$ is taken as a feature. Additionally, 
+we use the difference in average TF*IDF from time $t$ to $t-i$ for 
+$i = \{1, \ldots, 24\}$ to measure how the TF*IDF scores for the sentence have
+changed over the last 24 hours, i.e. we keep the sentence term frequencies 
+fixed and compute the difference in IDF. Large changes in IDF value indicate 
+the sentence contains bursty terms.
 We also use the time (in hours) since the event started as a feature.
-%difference 
-%Using these IDF 
-%calculations, the avg TFIDF is calculated for
-%Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.
-We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as
-%\begin{align*}
-%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
-%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
-%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
-%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}
-
-%where \begin{align*} \operatorname{tf-idf}_t(w,d) =&
-%\log\left(1+\sum_{w^\prime \in d}1\{w=w^\prime\}  \right)\\ & \times
-%\log\left(\frac{|D_t|}{1 + \sum_{d^\prime \in D_t}1\{w \in d^\prime\}}\right).
-%\end{align*}
-
-%[\textrm{Simpler explanation goes here}]
-% 1\{w = w^\prime} %- \operatorname{avg-tf-idf}_{t_{i-1}}(w).
-%\end{align*}
-
-%We similarly find the sentence's 5-hour burstiness.  In addition to burstiness, we also include the sentence's average tf-idf and
-
-% Formally, let $p(f)$ be a distribution over functions where $f$ is any mapping
-% of an input space $\mathcal{X}$ to the reals,
-%
-% $$f: \mathcal{X} \rightarrow \mathcal{R}.$$
-% Let the random variable $\mathbf{f} = (f(x_1),\ldots,f(x_n) )$ be
-%  an $n$-dimensional vector whose elements are evaluations of the function $f$
-% at points $x_i \in \mathcal{X}$.
-% We say $p(f)$ is a Gaussian process if for any finite subset
-% $\{x_1,\ldots,x_n\} \subset \mathcal{X}$, the marginal distribution over
-% that finite subset $p(\mathbf{f})$ has a multivariate Gaussian distribution.
-% A GP is parameterized by a mean function $\mu(\mathbf{x})$ and a
-% covariance function $K(x,x^\prime)$. Generally, the mean function is simply
-% set to 0, leaving the distribution to be completely characterized by the
-% kernel function on the data.
-%
-% In the regression setting, we typically have a response variable $y$ that
-% is the sum of our model prediction  and
-% some Gaussian noise, i.e. $y = f(x) + \epsilon$ with
-% $\epsilon \sim \mathcal{N}(0, \sigma^2)$. When
-% $f \sim \operatorname{GP}(\mathbf{0}, \mathbf{K})$, the
-% two distributions
-% of principal interest are the marginal likelihood
-% $p(\mathbf{y}|\mathbf{X}) =
-% \mathcal{N}(\mathbf{0},\mathbf{K} + \sigma^2\mathbf{I})$ and the predictive
-% distribution,
-%
-% $$p(\mathbf{y_*}|\mathbf{x_*},\mathbf{X},\mathbf{y}) =
-% \mathcal{N}(\boldsymbol{\mu}_*, \boldsymbol{\sigma}^2_*) $$
-%
-% where $\mathbf{x_*}$ is a new or unseen input, $\mathbf{y_*}$ our predicted
-% response, and
-% \begin{align*}
-% \boldsymbol{\mu}_* & = \mathbf{K_*}(\mathbf{K} + \sigma^2\mathbf{I})^{-1}\mathbf{y} \\
-% \boldsymbol{\sigma}^2_* &
-% = \mathbf{K}_{**} - \mathbf{K}_*(\mathbf{K} + \sigma^2\mathbf{I})^{-1}
-% \mathbf{K}_*^T + \sigma^2\\
-% \end{align*}.
-%
-% Here $\mathbf{K}_* = K(\mathbf{x}_*, \mathbf{X})$, and
-% $\mathbf{K}_{**} = K(\mathbf{x}_*, \mathbf{x}_*)$.
-%
-%
 
 \subsubsection{Model}
-%\ckcomment{(replace section with this)\\
+
 Given our feature representation of the input sentences, we need only target 
-salience values for model learning. 
-For each event in our training data, we sample a set of sentences and  each 
-sentence's salience is computed according to Equation \ref{eq:salience}.
-This results in a training set of sentences, their feature representations,
-and their target salience values to predict.
+salience values for model learning. For each event in our training data, we 
+sample a set of sentences and  each sentence's salience is computed according 
+to Equation \ref{eq:salience}. This results in a training set of sentences, 
+their feature representations, and their target salience values to predict.
 
-%Given the feature representation/salience pairs, we can use
-%any regression technique to predict salience.
 We opt to use a Gaussian process (GP) regression model
 \cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
-kernel for the salience prediction task.  
-Our features fall naturally into five groups and we use a separate RBF kernel
-for each, using the sum of each feature group RBF kernel as the final input
-to the GP model.
-%}
-
-%At run-time
-%our summarization system
-%does not have access to event nuggets to compute sentence salience. 
-%However, we can use the nuggets
-%in our training data to model salience as a function of our features.
-%Such a model generalizes to new events.
-%For each event in our training data, we sample a set of sentences and  each 
-%sentence's salience is computed according to Equation \ref{eq:salience}.
-%This results in a training set of sentences and the target values 
-%(their salience) to predict.
-%
-%Given a feature representation for the sentences, we can use
-%any regression technique to predict salience.
-%We opt to use a Gaussian process (GP) regression model
-%\cite{rasmussen:gaussian-process-book} with a Radial Basis Function (RBF) 
-%kernel.  
-%Our features fall naturally into five groups and we use a separate RBF kernel
-%for each, using the sum of each feature group RBF kernel as the final input
-%to the GP model.
-
-%GP regressors are a
-%class of data-driven, non-parametric model generalizing the multi-variate
-%Gaussian to the infinite dimensional setting.
-
-%When evaluating
-%In order to fit a regression model, the sentences are represented as feature
-%vectors (described in Section~\ref{sec:features}).
-%GP's are general
-%and are state of the art for many regression tasks.  
-%GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
-%case sentences.  In our experiments, we used a radial basis 
-%function (RBF) kernel.
+kernel for the salience prediction task. Our features fall naturally into 
+five groups and we use a separate RBF kernel for each, using the sum of each 
+feature group RBF kernel as the final input to the GP model.
 
@@
 now select a set of update candidates, i.e. sentences that are both salient
-and representative of the current batch.
-To accomplish this, we combine the output of our 
-salience prediction model with the affinity
-propagation algorithm.
-%to identify a set of exemplar sentences 
-%for each input batch. 
+and representative of the current batch. To accomplish this, we combine the 
+output of our salience prediction model with the affinity propagation 
+algorithm.
 Affinity propagation (AP) is a clustering algorithm
-that identifies a subset of datapoints as exemplars and forms clusters
+that identifies a subset of data points as exemplars and forms clusters
 by assigning the remaining points to one of the exemplars 
-\cite{frey2007clustering}. AP attempts to 
-maximize the net similarity objective 
+\cite{frey2007clustering}. AP attempts to maximize the net similarity 
+objective 
 \[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
@@
 $\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
-similarity
-of data points and our predicted \textit{apriori} preference of a data point to be an exemplar
-respectively. 
+similarity of data points and our predicted \textit{apriori} preference of a 
+data point to be an exemplar respectively. 
 AP differs from other $k$-centers algorithms in that it simultaneously 
-considers all data points as exemplars, making it less prone to finding
-local optima as a result of poor initialization. Furthermore, the 
-second term in $\mathcal{S}$ incorporates the individual importance of 
-data points as candidate exemplars; most other clustering algorithms only make
-use of the first term, i.e. the pairwise similarities between data points.
-
-%There are very 
-%few restrictions on the nature of the similarities and preferences
-%other than that they be real-valued.
+considers all data points as exemplars, making it less prone to finding local 
+optima as a result of poor initialization. Furthermore, the second term in 
+$\mathcal{S}$ incorporates the individual importance of data points as 
+candidate exemplars; most other clustering algorithms only make use of the 
+first term, i.e. the pairwise similarities between data points.
 
@@
 
-%In our summarization system we use the output of our salience model as 
-%the preferences, i.e.
-%$\operatorname{salience}(s) = \operatorname{pref}(s)$; for the similarity 
-%function we use semantic similarity. 
-Recall that $\operatorname{salience}(s)$
-is a prediction of the semantic similarity of $s$ to information about the 
-event be summarized, i.e. the set of event nuggets.
-Intuitively, when maximizing objective function $\mathcal{S}$, AP must balance
-between best representing the input data and representing the most salient
-input. Additionally, when the level of input is high but the salience
-predictions are low, the preference term will guide AP toward a solution
-with fewer clusters; \textit{vice-versa} when input is very salient on average but
-the volume of input is low. The adaptive nature of our model differentiates
-our method from most other update summarization systems.
+Recall that $\operatorname{salience}(s)$ is a prediction of the semantic 
+similarity of $s$ to information about the event be summarized, i.e. the set 
+of event nuggets. Intuitively, when maximizing objective function 
+$\mathcal{S}$, AP must balance between best representing the input data and 
+representing the most salient input. Additionally, when the level of input is 
+high but the salience predictions are low, the preference term will guide AP 
+toward a solution with fewer clusters; \textit{vice-versa} when input is very 
+salient on average but the volume of input is low. The adaptive nature of our 
+model differentiates our method from most other update summarization systems.
 
@@
 
-The exemplar sentences from the exemplar selection stage are the most 
-salient and representative of the input for the current hour. However,
-we need to reconcile these sentences with updates from the previous hour
-to ensure that the most salient and least redundant  updates are selected.
-To 
-ensure that only the most salient updates are selected we apply a minimum
-salience threshold;
-after exemplar sentences have been identified, any exemplars whose salience is 
-less than $\lambda_{sal}$ are removed from consideration.
+The exemplar sentences from the exemplar selection stage are the most salient 
+and representative of the input for the current hour. However, we need to 
+reconcile these sentences with updates from the previous hour to ensure that 
+the most salient and least redundant  updates are selected. To ensure that 
+only the most salient updates are selected we apply a minimum salience 
+threshold; after exemplar sentences have been identified, any exemplars whose 
+salience is less than $\lambda_{sal}$ are removed from consideration.
 
- Next,
-to prevent adding updates that are redundant with previous output updates, 
-we filter out exemplars
-that are too similar to previous updates.
-The exemplars are examined
-sequencially in order of decreasing salience and  a similarity threshold 
-is applied, where the exemplar is ignored if its
-maximum semantic similarity to any previous updates in the summary is
-greater than $\lambda_{sim}$.
+Next, to prevent adding updates that are redundant with previous output 
+updates, we filter out exemplars that are too similar to previous updates.
+The exemplars are examined sequentially in order of decreasing salience and  a similarity threshold is applied, where the exemplar is ignored if its maximum 
+semantic similarity to any previous updates in the summary is greater than 
+$\lambda_{sim}$.
 

### 2015-05-29 relatedwork.tex
Addressed comments. Fixed margins/sizings. Added authors.
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
 \cite{nenkova2012survey}. 
-%We conceptualize 
-Existing approaches generally fall into  %falling into
-one of three categories, each with specific tradeoffs with respect to update 
-summarization.
+Existing approaches generally fall into one of three categories, each with 
+specific trade-offs with respect to update summarization.
 
 First, centrality-focused approaches (including graph \cite{erkan2004lexrank},
-cluster \cite{hatzivassiloglou2001simfinder}, and centroid \cite{radev2004centroid} methods) are very natural for retrospective analysis in
-the sense that they let the data ``speak for itself.'' 
-These methods equate salience with centrality, either to the input or some other
-aggregate object (i.e. a cluster center or input centroid).
-However, they rely
-chiefly on
-redundancy. When applied to an unfolding event, there may not exist enough
-redundant content at the event onset for these methods to exploit.
+cluster \cite{hatzivassiloglou2001simfinder}, and centroid 
+\cite{radev2004centroid} methods) are very natural for retrospective analysis 
+in the sense that they let the data ``speak for itself.'' These methods 
+equate salience with centrality, either to the input or some other aggregate 
+object (i.e. a cluster center or input centroid). However, they rely
+chiefly on redundancy. When applied to an unfolding event, there may not exist
+enough redundant content at the event onset for these methods to exploit.
 Once the event onset has passed, however, the redundancy reduction of these 
@@
 
-The second category, predictive approaches,
-includes ranking and classification based methods.
-Sentences have been ranked by the average word probability, average tf-idf
-score, and the number of topically related words (topic-signatures in the
-summarization literature)
+The second category, predictive approaches, includes ranking and 
+classification based methods. Sentences have been ranked by the average word 
+probability, average TF*IDF score, and the number of topically related words 
+(topic-signatures in the summarization literature)
 \cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
 statistics are easily computable from the input sentences, while the third
-only requires an additional, generic background corpus.  
-In classification based methods, model features are
-usually derived from human generated summaries, and are non-lexical in nature
-(e.g., sentence starting position, number of topic-signatures, number of
-unique words). Seminal work in this area has employed naive
-Bayes and logistic regression classifiers to identify sentences for summary
-inclusion \cite{kupiec1995trainable,conroy2001using}. 
+only requires an additional, generic background corpus. In classification 
+based methods, model features are usually derived from human generated 
+summaries, and are non-lexical in nature (e.g., sentence starting position, 
+number of topic-signatures, number of unique words). Seminal work in this 
+area has employed na\"ive Bayes and logistic regression classifiers to 
+identify sentences for summary inclusion 
+\cite{kupiec1995trainable,conroy2001using}. 
 While these methods are less dependent on redundancy, the expressiveness of
@@
 The last category includes probabilistic \cite{haghighi2009exploring}, 
-information theoritic, and set cover \cite{lin2011class}
+information theoretic, and set cover \cite{lin2011class}
 approaches. While these methods are  focused on producing diverse
@@
 we do not necessarily have a fixed summary length and the corpus to be
-summarized contains many irrelevent sentences, i.e. there are large
-portions of the corpora that we specificly want to avoid.
+summarized contains many irrelevant sentences, i.e. there are large
+portions of the corpora that we specifically want to avoid.
 
-% This has often been approached as a ranking
-%problem.
-%We broadly conceptualize this decision as either an intrinsic or extrinsic
-%sentence evaluation process. Intrinsic approaches evaluate sentences
-%individually, possibly by predicting the impact on summary quality using
-%sentece level features. 
-%Sentences have been ranked by the average word probability, average tf-idf
-%score, and the number of topically related words (topic-signatures in the
-%summarization literature)
-%\cite{nenkova2005impact,hovy1998automated,lin2000automated}. The first two
-%statistics are easily computable from the input sentences, while the third
-%only requires an additional, generic background corpus.  Another ranking
-%approach, centroid summarization, involves creating an average bag of words
-%(BOW) vector, the centroid, from the input sentences and ranking sentences by
-%their similarity to the centroid \cite{radev2004centroid}.  Graph
-%\cite{erkan2004lexrank} and clustering
-%\cite{hatzivassiloglou2001simfinder,mckeown1999towards,siddharthan2004syntactic}
-%based approaches, on the other hand, make use of pair-wise similarity
-%comparisons amongst input sentences.  In these models, salient sentences are
-%more central to the input or cluster, respectively.
-
-%identify salient regions of the input space while simultaneously coping with
-%redundancy.  Graph-based algorithms have been used to rank sentences
-%Clustering algorithms, e.g., are commonly used to exploit redundancy in
-%input. Input sentences are clustered and summaries are generated by selecting
-%the most representative sentence from each cluster.  Graph-based models have
-%also been used for summarization.  E.g., the LexRank algorithm treats
-%sentences as nodes in a graph, where edges are constructed by way of cosine
-%similarity between sentence nodes; edges are either continuosly weighted by
-%similarity or discrete, existing only when the similarity is above a
-%threshold.  The PageRank algorithm is used on the graph to find the most
-%important sentence nodes. In both clustering and graph-based approaches,
-%sentence salience is largely determined by the pairwise relations between
-%sentences.
-
-%\fdadd{
 Several researchers have recognized the importance of summarization during
-natural disasters.  \cite{qi:temporal-summarization} developed a system for detecting
-novel, relevant, and comprehensive sentences immediately after a natural
-disaster.  
-%The method uses a model of
-%sentence relevance and novelty in order to select appropriate updates.
-%Training data for regression targets is automatically generated from
-%retrospective Wikipedia data.  The system is evaluated on news documents
-%related to 197 natural and human disasters from 2009 to 2011 using variants of
-%ROUGE modified to capture novelty, relevance, and comprehensiveness
-%\cite{lin2004rouge}.  
-\cite{wang:update-summarization} present a clustering-based approach to
-efficiently detect important updates during natural disasters.  
-The algorithm works by hierarchically
-clustering sentences online, allowing the system to output a more expressive
-narrative structure than \cite{qi:temporal-summarization}.  
-Our system attempts to unify these system's approaches 
-(predictive ranking and clustering respectively). 
-%The method is evaluated on
-%official press releases related to Hurricane Wilma  in 2005 using ROUGE score
-%between the system summary and a manually generated target summary. 
-%In addition
-%to natural disasters, we evaluate our system on a variety of event types 
-% such as terrorism and mass shootings (e.g., the 2012
-%shooting in Aurora, Col.), accidents
-%(e.g., the 2012 Pakistan garment factory fire), and social activism (e.g., the
-%Arab spring). The size and scope of the events varies considerably. Hurricane
-%Sandy, for example, affected multiple countries over weeks while the
-%2012 Aurora shooting was contained to a single location spanning several hours.
-
-%\fdcomment{Glasgow temporal summarization system \cite{mccreadie:temporal-summarization}.}
+natural disasters.  \cite{qi:temporal-summarization} developed a system for 
+detecting novel, relevant, and comprehensive sentences immediately after a 
+natural disaster.  \cite{wang:update-summarization} present a clustering-based
+approach to efficiently detect important updates during natural disasters.  
+The algorithm works by hierarchically clustering sentences online, allowing 
+the system to output a more expressive narrative structure than 
+\cite{qi:temporal-summarization}.  Our system attempts to unify these system's
+approaches (predictive ranking and clustering respectively).

### 2015-05-29 results.tex
spellinggit add acl2015.tex conclusion.tex dataset.tex experiments.tex introduction.tex methods.tex relatedwork.tex results.tex rouge_cum.tex nuggets-metrics-eps-converted-to.pdf rouge-time-eps-converted-to.pdf !

@@
 
-Table~\ref{tab:rouge} shows our results for system output samples against the full summary of nuggets using ROUGE. 
-%shows our results for the evaluation using ROUGE. 
-%AP+Salience shows improvement on ngram precision and recall over the 
-%baselines. 
-This improvement is statistically significant for all ngram
-precision, recall, and F-measures
-at the $\alpha = .01$ level
-using the Wilcoxon signed-rank test.
+Table~\ref{tab:rouge} shows our results for system output samples against the 
+full summary of nuggets using ROUGE. This improvement is statistically 
+significant for all ngram precision, recall, and F-measures at the 
+$\alpha = .01$ level using the Wilcoxon signed-rank test.
 
@@
 
-AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
-shows the ROUGE-1 scores over time. 
-We show the difference in unigram precision (bigram precision is not shown but it follows
-similar curve).
-Within the initial days of the event, AP+Salience is able to take the lead over the over 
-systems in ngram precision. 
-The AP+Salience model is better able to find salient updates
-earlier on; for the disaster domain, this is an especially important quality of the model.
+\textsc{AP+Salience} maintains its performance above the baselines over time 
+as well. Figure~\ref{fig:trouge} shows the ROUGE-1 scores over time. We show 
+the difference in unigram precision (bigram precision is not shown but it 
+follows similar curve). Within the initial days of the event, 
+\textsc{AP+Salience} is able to take the lead over the over systems in ngram 
+precision. The \textsc{AP+Salience} model is better able to find salient 
+updates earlier on; for the disaster domain, this is an especially important 
+quality of the model.
 
-Moreover, the AP+Salience's recall is not diminished by the high precision and remains competitive with AP.
-Over time AP+Salience's recall also begins to pull away, while the other models start to suffer
-from topic drift.
+Moreover, the \textsc{AP+Salience}'s recall is not diminished by the high 
+precision and remains competitive with \textsc{AP}. Over time 
+\textsc{AP+Salience}'s recall also begins to pull away, while the other models
+start to suffer from topic drift.
 
@@
 
-Figure~\ref{fig:nperf} shows the expected gain across a range of 
-similarity thresholds, where thresholds closer to 1 are more conservative
-estimates. The ranking of the systems remains constant across the 
-sweep with 
-AP+Salience beating all baseline systems.
-Predicting salience in general is helpful for keeping a summary on topic as
-the RS approach out performs the clustering only approaches
-on expected gain.
+Figure~\ref{fig:nperf} shows the expected gain across a range of similarity 
+thresholds, where thresholds closer to 1 are more conservative estimates. 
+The ranking of the systems remains constant across the sweep with 
+\textsc{AP+Salience} beating all baseline systems. Predicting salience in 
+general is helpful for keeping a summary on topic as the \textsc{RS} approach 
+out performs the clustering only approaches on expected gain.
 
-When looking at the comprehensiveness of the summaries AP outperforms
- AP+Salience. The compromise encoded in the AP+Salience objective
-function, between being representative and being salient, is seen clearly here
-where the performance of the AP+Salience methods is lower bounded by 
-the salience focused RS
-system and upper bounded by the clustering only AP system.
-Overall, AP+Salience achieves the best balance of these two metrics.
+When looking at the comprehensiveness of the summaries \textsc{AP} outperforms
+\textsc{AP+Salience}. The compromise encoded in the \textsc{AP+Salience} 
+objective function, between being representative and being salient, is seen 
+clearly here where the performance of the \textsc{AP+Salience} methods is 
+lower bounded by the salience focused \textsc{RS} system and upper bounded by 
+the clustering only \textsc{AP} system. Overall, \textsc{AP+Salience} achieves
+the best balance of these two metrics.
 
@@
 
-Table~\ref{tab:farouge} shows the results of our feature ablation
-tests. Removing the language models yields a statistically 
-significant drop in both ngram recall and F-measure. 
-Interestingly, removing the basic features leads to an
-increase in both unigram and bigram precision; in the bigram
-case this is enough to cause a statistically significant increase
-in F-measure over the full model. In other words, the generic features
-actually lead to an inferior model when we can incorporate more appropriate
-domain specific features.
-The result mirrors Sparck Jones' claim that generic approaches to summarization cannot produce a useful summary \cite{ksj98}.
-%This result mirrors Sparck Jones' claim that summarization should be in service of a purpose~\cite{?}.
+Table~\ref{tab:farouge} shows the results of our feature ablation tests. 
+Removing the language models yields a statistically significant drop in both 
+ngram recall and F-measure. Interestingly, removing the basic features leads 
+to an increase in both unigram and bigram precision; in the bigram case this 
+is enough to cause a statistically significant increase in F-measure over the 
+full model. In other words, the generic features actually lead to an inferior 
+model when we can incorporate more appropriate domain specific features.
+The result mirrors Sparck Jones' claim that generic approaches to 
+summarization cannot produce a useful summary \cite{ksj98}.
 
@@
 these features are too strongly correlated with each other, 
-i.e. the differences in TFIDF between hours are definitely not IID variables.
+i.e. the differences in TF*IDF between hours are definitely not i.i.d. 
+variables.

### 2015-05-29 outputs.tex
Added example outputs.

\begin{figure*}
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{Hurricane Sandy}}\\
\hline
\hline
%\small
%\tabitem The forecast map shows Sandy reaching eastern Cuba by early Thursday 
%    before heading to the Bahamas.\\
%\small
%\tabitem Jamaica's government issued a hurricane warning on Tuesday morning and 
%    announced schools would be closed on Wednesday.\\
\small
\tabitem  & \small Dangerous flash floods and mudslides set off by Sandy were a threat for the
   island of roughly 2.7 million inhabitants, Jamaica's meteorological service
   said.\\
\small
\tabitem  & \small A few reliable forecast models bring Sandy close enough to the coast to 
   produce heavy rains, strong winds and beach erosion.\\
\small
\tabitem  & \small Max winds are 65 mph with strengthening to a hurricane expected in the next
   12 to 18 hours.\\
\small
\tabitem & \small The two international airports, as well as schools and businesses, 
         will remain closed today until the hurricane warning, which is now in
         effect for the island, is lifted. \\
\hline
%at 11 a.m., sandy was in the caribbean about 65 miles south of kingston , jamaica , moving north at 13 mph with sustained winds of 80 mph.
%"""""""we heard on the radio that the hurricane was coming this way,"""""""" he said in the poor kingston community of standpipe, situated next to one of the debris-clogged gullies that crisscross the capital."""""""
\end{tabular}

~\\
~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Pakistan Garment Factory Fires}}\\
\hline
\hline
%\small
%\tabitem Medical officials said that some had died of suffocation while others
%         were burnt alive as the fire took hold. \\
\small
\tabitem & \small The fire broke out when people in the building were trying to start 
         their generator after the electricity went out. \\
\small
\tabitem & \small Pakistani television showed pictures of what appeared to be a 
         three-story building with flames leaping from the top-floor windows 
         and smoke billowing into the night sky. \\
\small
\tabitem & \small The people went to the back side of the building but there was no 
         access, so we had to made forceful entries and rescue the people, 
         said Numan Noor, a firefighter on the scene. \\
\small
\tabitem & \small ``We have recovered 63 bodies, including three found when we reached 
         the basement of the building,'' Karachi fire chief Ehtesham Salim told
         AFP on Wednesday. \\
%\small
%\tabitem Salim added that the blaze was Karachi’s ``biggest fire in terms of 
%         deaths in decades.'' \\
%\small
%\tabitem The garment trade as a whole is vital to Pakistan’s shaky economy. \\
\hline
\end{tabular}

~\\
~\\
\begin{tabular}{|l m{15cm}|}
\multicolumn{2}{c}{\textbf{2012 Romanian Protests}}\\
\hline
\hline
\small
\tabitem & \small Clashes between riot police and demonstrators have also erupted in 
         the Romanian capital Bucharest for a third day in a row. \\
\small
\tabitem & \small BOC urged Romanians to understand that tough austerity measures are 
         needed to avoid a default. \\
\small
\tabitem & \small More than 1,000 protesters rallied in Bucharest's main university 
         square, blocking traffic. \\
\small
\tabitem & \small Bucharest : a Romanian medical official says 59 people suffered 
         injuries as days of protests against the government and austerity 
         measures turned violent. \\
%\small
%\tabitem Most Romanians agree that only fundamental reform can save the 
%         country's ailing and corrupt system. \\
%\small
%\tabitem Many Romanians feared this would only increase corruption and create 
%         a further divide between the classes, leading to a two-tier system in
%         which only the wealthy would be able to afford care. \\
\hline
\end{tabular}

\caption{\textsc{AP+Salience} summary excerpts.}
\label{fig:summaries}
\end{figure*}

### 2015-05-29 outputs.tex
Added event type examples.

@@
 \begin{figure*}
-\begin{tabular}{|l m{15cm}|}
-\multicolumn{2}{c}{\textbf{Hurricane Sandy}}\\
-\hline
-\hline
+%\begin{tabular}{|l m{15cm}|}
+%\multicolumn{2}{c}{\textbf{Hurricane Sandy}}\\
+%\hline
+%\hline
+%%\small
+%%\tabitem The forecast map shows Sandy reaching eastern Cuba by early Thursday 
+%%    before heading to the Bahamas.\\
+%%\small
+%%\tabitem Jamaica's government issued a hurricane warning on Tuesday morning and 
+%%    announced schools would be closed on Wednesday.\\
 %\small
-%\tabitem The forecast map shows Sandy reaching eastern Cuba by early Thursday 
-%    before heading to the Bahamas.\\
+%\tabitem  & \small Dangerous flash floods and mudslides set off by Sandy were a threat for the
+%   island of roughly 2.7 million inhabitants, Jamaica's meteorological service
+%   said.\\
 %\small
-%\tabitem Jamaica's government issued a hurricane warning on Tuesday morning and 
-%    announced schools would be closed on Wednesday.\\
-\small
-\tabitem  & \small Dangerous flash floods and mudslides set off by Sandy were a threat for the
-   island of roughly 2.7 million inhabitants, Jamaica's meteorological service
-   said.\\
-\small
-\tabitem  & \small A few reliable forecast models bring Sandy close enough to the coast to 
-   produce heavy rains, strong winds and beach erosion.\\
-\small
-\tabitem  & \small Max winds are 65 mph with strengthening to a hurricane expected in the next
-   12 to 18 hours.\\
-\small
-\tabitem & \small The two international airports, as well as schools and businesses, 
-         will remain closed today until the hurricane warning, which is now in
-         effect for the island, is lifted. \\
-\hline
-%at 11 a.m., sandy was in the caribbean about 65 miles south of kingston , jamaica , moving north at 13 mph with sustained winds of 80 mph.
-%"""""""we heard on the radio that the hurricane was coming this way,"""""""" he said in the poor kingston community of standpipe, situated next to one of the debris-clogged gullies that crisscross the capital."""""""
-\end{tabular}
-
-~\\
-~\\
+%\tabitem  & \small A few reliable forecast models bring Sandy close enough to the coast to 
+%   produce heavy rains, strong winds and beach erosion.\\
+%\small
+%\tabitem  & \small Max winds are 65 mph with strengthening to a hurricane expected in the next
+%   12 to 18 hours.\\
+%\small
+%\tabitem & \small The two international airports, as well as schools and businesses, 
+%         will remain closed today until the hurricane warning, which is now in
+%         effect for the island, is lifted. \\
+%\hline
+%%at 11 a.m., sandy was in the caribbean about 65 miles south of kingston , jamaica , moving north at 13 mph with sustained winds of 80 mph.
+%%"""""""we heard on the radio that the hurricane was coming this way,"""""""" he said in the poor kingston community of standpipe, situated next to one of the debris-clogged gullies that crisscross the capital."""""""
+%\end{tabular}
+%
+%~\\
+%~\\
 \begin{tabular}{|l m{15cm}|}

### 2015-07-22 acl-slides/msr.tex
Acl slides added.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%   PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}

\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usepackage[utf8]{inputenc}
%\usefonttheme{professionalfonts} 
%\usefonttheme{serif}    
%\usepackage{tgbonum}
%\usepackage[T1]{fontenc}

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{svg}
\newcommand{\light}[1]{\textcolor{gray}{#1}}
%----------------------------------------------------------------------------------------
%   TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Disaster Summarization]{Predicting Salient Updates for Disaster Summarization}
\author[Chris Kedzie]{Chris Kedzie, Columbia University\\
Kathy McKeown, Columbia University\\
Fernando Diaz, Microsoft Research\\}

\institute[Columbia U.] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
%Computer Science Department \\
%Columbia University \\ % Your institution for the title page
\medskip
\textit{kedzie@cs.columbia.edu} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\newcommand{\tikzmark}[1]{\tikz[overlay,remember picture] \node (#1) {};}
\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

%\begin{frame}
%\frametitle{Overview} % Table of contents slide, comment this block out to remove it
%\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
%\end{frame}

%----------------------------------------------------------------------------------------
%   PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

\begin{frame}
\frametitle{It's October 28th, 2012...}
\vspace{10pt}
\includegraphics<2>[scale=0.4]{png/news_anim1}
\includegraphics<3>[scale=0.4]{png/news_anim2}
\includegraphics<4>[scale=0.4]{png/news_anim3}
\includegraphics<5>[scale=0.4]{png/news_anim4}
\includegraphics<6>[scale=0.4]{png/news_anim5}
\end{frame}

\begin{frame}
\frametitle{Information During Disaster}
\begin{itemize}
\item<1-> People turn to the web and mobile services
\item[]
\item<2-> Realtime curation of data requires significant manpower! 
\begin{itemize}
\item<3-> summarizing official information/resources
\item<4-> summarizing eye witness information
\item<5-> correcting/updating erroneous information
\item<6-> predicting likely outcomes
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Information During Disaster}
\begin{itemize}
\item \light{People turn to the web and mobile services}
\item[]
\item \light{Realtime curation of data requires significant manpower!}
\begin{itemize}
\item \textbf{summarizing official information/resources}
\item \light{summarizing eye witness information}
\item \light{correcting/updating erroneous information}
\item \light{predicting likely outcomes}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Update Summarization}
\textbf{Given an event:} \pause ``guatemala earthquake''
\begin{itemize}
    \item{Monitor a stream of news}
    \item{Identify relevant information}
    \item{Update the user}

\end{itemize}
    \pause
    \textbf{CAVEAT:} Updates should be timely and novel! 
\end{frame}

\begin{frame}
    \includegraphics{images/anim1.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim2.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim3.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim4.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim5.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim6.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim7.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim8.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim9.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim10.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim11.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim12.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim13.pdf}
\end{frame}
\begin{frame}
    \includegraphics{images/anim14.pdf}
\end{frame}

\begin{frame}
\frametitle{Characteristics of a good summary of disaster}

$\pmb{+}$ \textbf{Salience} \\
Updates should contain the most important information.\\
~\\

$\pmb{-}$ \textbf{Latency} \\
New information should be delivered as quickly as possible.\\
~\\

$\pmb{-}$ \textbf{Redundancy} \\
Updates should focus on new or updated information
that has not yet been presented to the user.\\
~\\

\end{frame}

\begin{frame}
    \frametitle{Online Update Summarization}
    \textbf{Summarization as Search}\\
    \begin{itemize}
        \item learning 2 search
        \item each search state is a multiclass classification problem 
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Update Summarization Approach}
~\\
At time $t$:
\begin{enumerate}
\item Predict salience for input sentences. 
\pause
\begin{itemize}
\item Domain-specific features for predicting salience
\end{itemize}
\pause
\item Select exemplar sentences for $t$ that are both representative and
    salient.
\pause
%\begin{itemize}
%\item Incorporate salience predictions in this decision
%\end{itemize}
\item Remove redundant sentences.
%\pause
\end{enumerate}
\end{frame}

\begin{frame}
    \frametitle{How do we measure salience?}

    \pause

    event: 2012 Guatemala Earthquake
    \tikzmark{nuggets}{
    \begin{itemize}
        \pause
        \item epicenter was located in the Pacific Ocean
        \pause
    \item \alert<8>{tsunami warnings 100-200 miles from epicenter}
        \pause
        \item felt in El Salvador and parts of Mexico
        \item[] $\;\;\;\;\;\;\;\;\;\vdots \;\;\;\;\;\;\;\;\;\;\;\;\;\;\vdots\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\vdots$

    \end{itemize}
    }
    \pause
    \tikz[overlay,remember picture]{
        \draw[draw=red,thick,double,fill opacity=0.2] ($(nuggets)+(-5.6,-0.2)$) rectangle ($(nuggets)+(2.8,-2.7)$);
        \node[draw=red,thick,double,fill=white] at ($(nuggets)+(1.7,-2.7)$) {nuggets}; 
    }

\uncover<7->{\vspace{16pt}
\pause
\textit{Nicaragua's disaster management said it had issued a 
\alert<8>{local tsunami alert.}}\\
\uncover<9->{
~\\
Given a sentence $s$ and a set of gold nuggets $N$, we define the salience 
of a sentence as
$$\operatorname{salience}(s) = \max_{n \in N} \operatorname{sim}(s,n).$$
    }
}
\end{frame}

\begin{frame}
    \frametitle{How do we measure salience?}

    Given a sentence $s$ and a set of gold nuggets $N$, we define the salience 
    of a sentence as
    $$\operatorname{salience}(s) = \max_{n \in N} \operatorname{sim}(s,n).$$

    \pause
    ~\\

    \textbf{Sentence Similarity}\\
    We use the sentence similarity model of \textit{Guo and Diab, (2012)}\\
    \begin{itemize}
        \item model is trained on domain related Wikipedia abstracts
        \item reduces sentences to a low $k$-dimensional latent vector ($k=100$
            in our experiments)
        \item similarity of two sentences is the cosine similarity of their 
            latent vectors
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{How do we measure salience?}

    Given a sentence $s \in S$ and a set of gold nuggets $N$, we define the 
    salience of a sentence as
    $$\operatorname{salience}(s) = \max_{n \in N} \operatorname{sim}(s,n).$$

    ~\\
    $$\uncover<2->{s_1 }\uncover<6->{,\;\; s_2} 
         \uncover<9->{,\;\; s_3} 
         \uncover<12->{,\;\; \ldots,\;\; s_k}\uncover<2->{ \sim S}$$
    ~\\
    \begin{center}
    \uncover<3->{
    \begin{tabular}{c | c }
        $y$ & $X$\\[2pt]
    \hline
    \uncover<4->{$\operatorname{salience}(s_1)$} & \uncover<5->{\alert<17>{$\phi(s_1)$}}\\[2pt]
    \uncover<7->{$\operatorname{salience}(s_2)$} & \uncover<8->{\alert<17>{$\phi(s_2)$}}\\[2pt] 
    \uncover<10->{$\operatorname{salience}(s_3)$} & \uncover<11->{\alert<17>{$\phi(s_3)$}}\\[2pt]
    \uncover<13->{$\vdots$} & \uncover<14->{$\vdots$}\\[2pt]    
    \uncover<15->{$\operatorname{salience}(s_k)$} & \uncover<16->{\alert<17>{$\phi(s_k)$}}\\[2pt]
    \end{tabular}
    }
    \end{center}
\end{frame}
\begin{frame}
\frametitle{Predicting Salience: Model Features}
\begin{itemize}
\item Basic sentence level features
\begin{itemize}
\item sentence length
\item punctuation count
\item \alert<2-4>{number of capitalized words}
\item sentence position in document
\end{itemize}
\end{itemize}

\pause
\textbf{High Salience}\\ 
\alert<2>{Nicaragua's} disaster management said it had issued a local tsunami alert.\\
\textbf{Medium Salience} \\
\alert<3>{People} streamed out of homes, schools and office buildings as far north as \alert<3>{Mexico City.}\\

\textbf{Low Salience} \\
\alert<4>{Add} to \alert<4>{Digg Add} to del.icio.us \alert<4>{Add} to \alert<4>{Facebook Add} to \alert<4>{Myspace}

\end{frame}

\begin{frame}
\frametitle{Predicting Salience: Model Features}
\begin{itemize}
\item Basic sentence level features
\item Query features
\begin{itemize}
\item query term matches
\item event type synonym, \alert<2>{hypernym}, and hyponym matches
\end{itemize}
\end{itemize}

\textbf{High Salience}\\ 
Nicaragua's \alert<2>{disaster} management said it had issued a local tsunami alert.\\
\textbf{Medium Salience} \\
People streamed out of homes, schools and office buildings as far north as 
Mexico City.\\

\textbf{Low Salience} \\
Add to Digg Add to del.icio.us Add to Facebook Add to Myspace

\end{frame}

\begin{frame}
\frametitle{Predicting Salience: Model Features}
\begin{itemize}
\item Basic sentence level features
\item Query features
\item Language Models (5-gram interpolated Kneser-Ney model)
\begin{itemize}
\item \alert<2>{generic news corpus (10 years of AP and NY Times articles)}
\item \alert<4>{domain specific corpus (domain related Wikipedia articles)}
\end{itemize}
\end{itemize}

\pause
\textbf{High Salience}\\ 
\alert<2,4>{Nicaragua's disaster management said it had issued a local tsunami alert.}\\
\textbf{Medium Salience} \\
\alert<2>{People streamed out of homes, schools and office buildings as far north as Mexico City.}\\
\textbf{Low Salience} \\
Add to Digg Add to del.icio.us Add to Facebook Add to Myspace

\end{frame}

\begin{frame}
\frametitle{Predicting Salience: Model Features}
\begin{itemize}
\item Basic sentence level features
\item Query features
\item Language Models
\item Geographic Features
\begin{itemize}
\item input sentences tagged with Named-Entity tagger
\item coordinates are retrieved for each location in a sentence
\item mean distance to event
\end{itemize}
\end{itemize}

\pause
\textbf{High Salience}\\ 
\alert<2>{Nicaragua's} disaster management said it had issued a local tsunami alert.\\
\textbf{Medium Salience} \\
People streamed out of homes, schools and office buildings as far north as \alert<2>{Mexico City.}\\
\textbf{Low Salience} \\
Add to Digg Add to del.icio.us Add to Facebook Add to Myspace

\end{frame}

\begin{frame}
\frametitle{Predicting Salience: Model Features}
\begin{itemize}
\item Basic sentence level features
\item Query features
\item Language Models (5-gram interpolated Kneser-Ney model)
\item Geographic Features
\item Temporal Features
\begin{itemize}
\item average tf-idf at the current time $t_0$
\item difference of average tf-idf at time $t_0$ and $t_{-1}$
\item $\;\;\;\;\;\;\;\;\;\vdots$
\item difference of average tf-idf at time $t_0$ and $t_{-24}$
\item hours since event start time
\end{itemize}
\end{itemize}

\pause
\textbf{High Salience}\\ 
\alert<2>{Nicaragua's disaster management} said it had issued a local \alert<2>{tsunami} alert.\\
\textbf{Medium Salience} \\
People streamed out of homes, schools and office buildings as far north as Mexico City.\\
\textbf{Low Salience} \\
\alert<3>{Add to Digg Add to del.icio.us Add to Facebook Add to Myspace}

\end{frame}

\begin{frame}
\frametitle{Affinity Propagation Clustering}
\begin{itemize}
\pause
\item assigns each data point to an exemplar data point \\
     (these assignments
      define a cluster)
\pause
\item exemplars are actual data points that best represent their cluster
\pause
\item exemplar selection is determined by a similarity matrix $S$ and 
      exemplar \textit{preference}
\pause
\begin{itemize}
\pause 
\item similarities $S$ -- element $s(i,j)$ is the semantic similarity of sentence $i$ to sentence $j$
\pause
\item \textit{preference} -- scalar value that represents 
      apriori how suited a data point is to be an exemplar
\pause
\item \alert<+->{preference $=$ salience}
\pause
\item \alert<+->{exemplar $=$ update}
\end{itemize}

\end{itemize}
%\begin{columns}[T]
%\column{0.5\textwidth}
%$$r(i,k) \leftarrow s(i,k) - \max_{k^\prime\ne k} a(i,k^\prime) 
%+ s(i,k^\prime) $$

%\column{0.5\textwidth}
%    Update 2
%\end{columns}

\end{frame}

\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim1.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim2.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim3.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim4.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim5.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim6.pdf}
\end{frame}
\begin{frame}
\frametitle{Affinity Propagation Clustering}
    \includegraphics{images/cluster_anim7.pdf}
\end{frame}

\begin{frame}
\frametitle{Trec Temporal Summarization Track}
\begin{itemize}
\pause
\item TREC KBA Stream Corpus
\pause
\begin{itemize}
\item hourly web crawl
\item October 2011 - February 2013
\item 16.1TB!
\end{itemize}
\pause
\item[]
\item Training Data
\begin{itemize}
\item Training Events
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Trec TS Events}
\begin{itemize}
    \item{Natural Disasters}
        \begin{itemize}
            \item{\textbf{2012 Guatemalan Earthquake}}
            \item{Hurricane Sandy\\}
            $\;\;\;\;\;\;\;\;\vdots$
        \end{itemize}
    \item{Man-made Disasters}
        \begin{itemize}
            \item{2012 Pakistan Garment Factory Fires}
            \item{2012 Buenos Aires Rail Disaster\\}
            $\;\;\;\;\;\;\;\;\vdots$
        \end{itemize}
    \item{Terrorism/Violence}
    \item{Social Unrest}
\end{itemize}
\end{frame}
%\begin{frame}
%put pic of doc frequency over time

\begin{frame}
\frametitle{Trec Temporal Summarization Track}
\begin{itemize}
\item TREC KBA Stream Corpus
\begin{itemize}
\item hourly web crawl
\item October 2011 - February 2013
\item 16.1TB!
\end{itemize}
\item[]
\item Training Data
\begin{itemize}
\item Training Events
\pause
\item Gold Nugget Information
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Gold Nugget Information}
\begin{itemize}
\item epicenter was located in the Pacific Ocean 
%\item[]
\item tsunami warnings 100-200 miles from epicenter
%\item[]
\item felt in El Salvador and parts of Mexico
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-07-19\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_1}
\end{frame}
\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-07-20\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_2}
\end{frame}
\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-08-05\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_3}
\end{frame}
\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-08-05\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_4}
\end{frame}

\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-08-05\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_3}
\end{frame}

\begin{frame}
\frametitle{Extracting Gold Nuggets from Wikipedia}
%SYSTEM TIME: 2012-11-08-05\\
\vspace{5pt}
\includegraphics[width=325pt,height=200pt]{png/wp_evo_5}
\end{frame}

\begin{frame}
\frametitle{Trec Temporal Summarization Track}
\begin{itemize}
\item TREC KBA Stream Corpus
\begin{itemize}
\item hourly web crawl
\item October 2011 - February 2013
\item 16.1TB!
\end{itemize}
\item[]
\item Training Data
\begin{itemize}
\item Training Events
\item Gold Nugget Information
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Experiments}

    \begin{itemize}
        \pause
        \item leave one out evaluation on 21 TREC TS events
        \pause
        \item 3 events held out for tuning similarity threshold parameters
        \pause
        \item trained salience models for each event
        \pause
        \begin{itemize}
            \item 1000 sentences sampled for each event
            \item at test time, salience prediction is the average of the 
                23 other models
        \end{itemize}
    \end{itemize}

\end{frame}

\begin{frame}
    \frametitle{Evaluation Metrics}
    \begin{itemize}
        \pause
        \item ROUGE 
        \pause
        \begin{itemize}
            \item model summary generated by concatenating event nuggets
        \end{itemize}
        \pause
    \item Expected Gain and Comprehensiveness
        $$\mathbb{E}[\mathrm{Gain}] = \frac{|S_n|}{|S|}\;\;\;\;\;\;\;\;\;
            \textrm{Comprehensiveness} = \frac{|S_n|}{|N|}$$
            where \begin{itemize}
                    \pause
            \item[] $S$ is the set of system updates
                    \pause
                \item[] $S_n$ is the set of nuggets mapped to updates in $S$
                    \pause
                \item[] $N$ is the set of nuggets for the event
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Baselines}
    \begin{itemize}
    \item AP+Salience -- full model
    \item AP -- affinity propagation clustering with no salience
    \item HAC -- hierarchical agglomerative clustering
    \item RS -- rank by salience
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{ROUGE}
\begin{table}[h]
\centering
% centering table
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-1}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Recall}$ & $\mathrm{Prec.}$ & $\mathrm{F}_1$\\
[0.5ex]
\hline
AP+Salience & $\mathbf{0.282}$ & $\mathbf{0.344}$ & $\mathbf{0.306}$\\
AP          & $0.245$ & $0.285$ & $0.263$ \\
RS          & $0.230$ & $0.271$ & $0.247$ \\
HAC         & $0.169$ & $0.230$ & $0.186$ \\
\hline % inserts single-line
\end{tabular}
~\\[1ex]
~\\
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-2}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Recall}$ & $\mathrm{Prec.}$ & $\mathrm{F}_1$\\[0.5ex]
\hline
\textsc{AP+Salience} & $\mathbf{0.045}$ & $\mathbf{0.056}$ & $\mathbf{0.049}$\\
\textsc{AP}          & $0.033$ & $0.038$ & $0.035$ \\
\textsc{RS}          & $0.031$ & $0.037$ & $0.034$ \\
\textsc{HAC}         & $0.017$ & $0.024$ & $0.019$ \\
\hline % inserts single-line
\end{tabular}
\end{table}
\end{frame}

\begin{frame}
\frametitle{ROUGE-1 over time}
\begin{center}
\includegraphics[]{images/rouge-time.eps}
\end{center}
\end{frame}

\begin{frame}
\frametitle{$\mathbb{E}[\mathrm{Gain}]$ and $\mathrm{Comprehensiveness}$}
\begin{center}
\includegraphics[scale=.7]{images/nuggets-metrics2.eps}
\end{center}
\end{frame}

\section{Conclusion}
\begin{frame}
\frametitle{Future Work}

\begin{itemize}
\item Online update summarization
\pause
\item[]
\item Improved redundancy
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Links}

\begin{itemize}
\item Trec Temporal Summarization Track \\ 
    \begin{center}\url{http://www.trec-ts.org}\end{center}
\item[]
\item ACL paper\\\begin{center} \url{http://www.columbia.edu/~crk2130}\end{center}
\end{itemize}

\end{frame}

\end{document}

### 2015-07-22 acl-slides/images/make-cluster-animation.py

@@ -1,0 +1,55 @@
+import re
+import os
+
+
+def make_invisible(layers, xml):
+    for layer in layers:
+        print layer
+        m = re.search(r'(<g[^>]+?id="' + layer + r'"[^>]+>)', xml, 
+            re.DOTALL)
+        layer_tag = m.groups()[0]
+        layer_tag_inv = layer_tag.replace("display:inline", "display:none")
+        xml = xml.replace(layer_tag, layer_tag_inv)
+    return xml        
+
+def make_visible(layers, xml):
+    for layer in layers:
+        m = re.search(r'(<g[^>]+?id="' + layer + r'"[^>]+>)', xml, 
+            re.DOTALL)
+        layer_tag = m.groups()[0]
+        layer_tag_vis = layer_tag.replace("display:none", "display:inline")
+        print layer_tag_vis
+        xml = xml.replace(layer_tag, layer_tag_vis)
+    return xml        
+
+
+def main():
+    with open("clustering.svg", "r") as f:
+        xml = f.read()
+
+    layers = ["unbiased_assignments", "biased_assignments",
+              "unbiased_clusters", "ex_label", "label", 
+              "salient_regions"]
+    img_layers = [["label", ],
+        ["label",  "unbiased_assignments",],
+        ["label", "points", "unbiased_assignments", "ex_label"],
+        ["label", "points", "unbiased_assignments", "ex_label", 
+            "unbiased_clusters"],
+        ["label", "points"],
+        ["label", "points", "salient_regions",],
+        ["label", "points", "salient_regions", "biased_assignments"],]
+
+    for i, img_layer in enumerate(img_layers, 1):
+        print i, img_layer
+        xml = make_invisible(layers, xml)    
+        xml = make_visible(img_layer, xml) 
+        
+        fname = "cluster_anim{}.svg".format(i)
+        dest = "cluster_anim{}.pdf".format(i)
+        with open(fname, "w") as f:
+            f.write(xml)
+        os.system("inkscape -z -C -f {} -A {}".format(fname, dest))
+        os.remove(fname)
+
+if __name__ == "__main__":
+    main()

### 2015-07-22 acl-slides/images/make-stream.sh

@@ -1,0 +1,14 @@
+hour1,51.754418,27.446495,125.64625,29.04
+hour1art1,55.976527,63.27952,192.55498,241.94488
+hour1art2,93.979178,91.422723,192.55497,153.27724
+hour1art3,147.12878,73.706187,192.55498,241.94487
+hour1art4,217.99492,79.968209,192.55498,241.94487
+hour1blur,-1.8163636,-0.35544285,456.91332,341.97493
+hour1updates,12.050875,136.159,391.53604,124.84863
+hour2bg,-1.8163635,0,455.35968,341.61951
+hour2,51.754418,27.446495,128.64625,29.04
+hour2art1,55.976527,63.27952,192.55498,241.94488
+hour2art2,93.979178,91.422723,192.55497,153.27724
+hour2art3,147.12878,73.706187,192.55498,241.94487
+hour2art4,217.99492,79.968209,192.55498,241.94487
+hour2blur,12.050875,136.159,391.53603,179.97972

### 2015-07-22 acl-slides/images/make-svg-animation.py

@@ -1,0 +1,69 @@
+import re
+import os
+
+
+def make_invisible(layers, xml):
+    for layer in layers:
+        m = re.search(r'(<g[^>]+?id="' + layer + r'"[^>]+>)', xml, 
+            re.DOTALL)
+        layer_tag = m.groups()[0]
+        layer_tag_inv = layer_tag.replace("display:inline", "display:none")
+        xml = xml.replace(layer_tag, layer_tag_inv)
+    return xml        
+
+def make_visible(layers, xml):
+    for layer in layers:
+        m = re.search(r'(<g[^>]+?id="' + layer + r'"[^>]+>)', xml, 
+            re.DOTALL)
+        layer_tag = m.groups()[0]
+        layer_tag_vis = layer_tag.replace("display:none", "display:inline")
+        print layer_tag_vis
+        xml = xml.replace(layer_tag, layer_tag_vis)
+    return xml        
+
+
+def main():
+    with open("stream-images.svg", "r") as f:
+        xml = f.read()
+
+    layers = ["clock", "hour1art1", "hour1art2", "hour1art3", "hour1art4", 
+        "hour1blur", "hour1updates", "hour2bg", "hour2", "hour2art1", 
+        "hour2art2", "hour2art3", "hour2art4", "hour2blur", "hour2updates"]
+
+    clock_layers = ["15_minutes", "quarterhour", "30_minutes", "halfhour",
+        "45_minutes", "three_quarter", "1hour", "four_quarter", "zero_minutes",
+        "2hour"]
+    img_layers = [layers[:1] + ["zero_minutes"],
+        layers[:2] + ["15_minutes", "quarterhour"],
+        layers[:3] + ["30_minutes", "halfhour"],
+        layers[:4] + ["45_minutes", "three_quarter"],
+        layers[:5] + ["zero_minutes", "1hour", "four_quarter"],
+        layers[:6] + ["zero_minutes", "1hour", "four_quarter"],
+        layers[:7] + ["zero_minutes", "1hour", "four_quarter"],
+        #layers[7:8] + ["zero_minutes", "1hour", "clock"],
+        ["zero_minutes", "1hour", "clock"],
+        [layers[9]] + ["1hour", "15_minutes", "quarterhour", "clock"],
+        layers[9:11] + ["1hour", "30_minutes", "halfhour", "clock"],
+        layers[9:12] + ["1hour", "45_minutes", "three_quarter", "clock"],
[…]
